### 2022-01-19 README.Rmd
add first few thoughts

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

Preregistration is becoming increasingly popular in the psychological sciences.
At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
Should they surrender and discard the idea of preregistration altogether? Not at all.
We argue for the benefits of preregistration beyond strictly confirmatory studies.
Grounded in Bayesian philosophy of science, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigor.
Our argument rests on the relevance of “theoretical risk”, a generalization of type-I error rate, for judging theories.
Our proposal is that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of limiting the type-I error rate.
We argue that this reformulation extends the utility of preregistration and is closer aligned with researchers intuition.
If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
This uncertainty about the theoretical risk does us a disservice by decreasing the evidence boost.

A successful prediction should lend more credibility to a theory than a postdiction, all else being held equal.
Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis generating and hypothesis testing and so forth (see XXXX).
To honor this distinction, many psychological studies are preregistered to produce results that are considered confirmatory.
This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
We leverage this reasoning to outline two paradigms that rationalize the use of preregistration.

<!-- Caroline does not see the connection to type I error rate/theoretical risk -->

Let us formalize the judgment about a hypothesis (H) given some evidence (E) as the conditional probability $p(H|E)$.

Using Bayes' rule we arrive at:

$$
p(H|E) = \frac{p(H)p(E|H)}{p(H)p(E|H) + p(¬H)p(E|¬H)}
$$

Note the connection to null hypothesis testing.
If H represents the alternative Hypothesis and ¬H represents the Null hypothesis, then $p(E|H)$ is the power, while $p(E|¬H)$ represents the type I/alpha error.
However, generally a theory is richer in content, than equating it with the alternaive hypothesis in Null hypothesis testing.

<!-- continue to justify the importance of theorethical risk -->

The first paradigm is embraced by the open science community and conceptualizes preregistration as a tool to reduce the type I error rate (XXXX).
As we will later show, reducing the type I error rate is indeed a powerful leaver to gather persuasive evidence.
In practice, however, researchers can not reduce the type I error rate at will.
They face three limiting factors.
First, psychological theories tend to be vague.
Some decisions are just not derivable from theory, which makes it extraordinarily difficult to maintain a type I error rate near zero.
Second, researchers are constrained by resource limitations.
One way to counter the first problem is conducting exploratory pilot studies, but running extensive pilot studies is expensive.
Doubling the resources that go into investigating a research question is often just not tenable and mostly beyond the individual researcher to decide.
Third, by reducing the type I error rate the researcher directly increases the likelihood that their study will result in disappointing null results.
A study that fails to deliver results can be crippling to the researchers career.
Depending on individual circumstances, e.g. the theory in question, resources allocated to a study, the career stage of the researcher, a researcher may simply feel unable to reduce the type I error rate to the threshold necessary for a confirmatory study.
Constrained by these factors researchers tasked with writing a preregistration are incentiviced to either give up or to muddle through by being intentionally vague. The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".

### 2022-02-16 README.Rmd
begin with the main argument

@@
 Grounded in Bayesian philosophy of science, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigor.
-Our argument rests on the relevance of “theoretical risk”, a generalization of type-I error rate, for judging theories.
-Our proposal is that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of limiting the type-I error rate.
+Our argument rests on the relevance of “theoretical risk”, a generalization of type-I error rate, for judging the evidential support for theories.
+Our proposal is that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
 We argue that this reformulation extends the utility of preregistration and is closer aligned with researchers intuition.
 If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
-This uncertainty about the theoretical risk does us a disservice by decreasing the evidence boost.
+This uncertainty severely limits the utility of the researchers' evidence.
 
-A successful prediction should lend more credibility to a theory than a postdiction, all else being held equal.
+A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
 Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis generating and hypothesis testing and so forth (see XXXX).
 To honor this distinction, many psychological studies are preregistered to produce results that are considered confirmatory.
+Rigorous application of preregistration prevents researcher from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
+The scientific community, widely agrees that actively concealing such "trying for results" is misconduct.
+But what exactly is such concealment cheating the scientific community of?
+We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
+These misunderstandings are interesting from a theoretical perspective but also lead to decreased efficiency of the scientific process if researcher  misguidedly decide that preregistration is not applicable for their intended study.
+
+One reason why researchers are hesitant to preregister is that they feel too limited by preregistration.
+Even researchers who have been persuaded by its advantages sometimes apply strategies to skirt the self imposed restrictions.
+One strategy is to write a loose preregistration to begin with, another is to simply deviate from the preregistration afterward.
+To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating good looking results no matter the nature of the phenomenon under study.
+To summarize, the conundrum is that current conceptualization of preregistration deem a study "properly" preregistrated even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
+This, of course, goes against common researchers intuition about what is considered exploratory and confirmatory.
+We point these issues out as symptoms of a misunderstanding.#change
+We do want not to discredit preregistration as a useful practice for advancing science, but challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
+
+Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
+The aforementioned issues arise by delegating this judgment to a simple decision rule, that is, if something is preregistered it is confirmatory and if it is not preregistered it is exploratory.
+Preregistering an analysis that is inherently exploratory (like testing dozens of relations) does not make it confirmatory nor will a careful confirmatory study become exploratory if researcher deviate from the preregistration in minor details.
+To uphold the simple decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
+Under these conditions the decision rule holds, but such restricted use makes preregistration a niche solution not able to match the greater problem of replicability in psychology and elsewhere.
+
+We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
+To that end, we first introduce some basic tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theorethical risk", that is inversely related to type I error rate.
+We than propose two interpretations of how theorethical risk is influenced by preregistration.
+The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
+The second interpretation, is our main contribution, and shows broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
+Under this interpretation theoretical risk is not necessarily directly maximized by preregistration but the uncertainty in judging the theoretical risk is minimized.
+This changed viewpoint is crucial for the following reason.
+Directly maximizing theoretical risk does, as the name implies, increase the risk that a study does fail to deliver results in favor of the theory.
+Minimizing the uncertainty around the theoretical risk, as we will later show, shares the beneficial properties but without increasing likelihood of a "failed" study.
+Let us note, that a study that is not able to find evidence that a researcher is hoping for, is still immensely valuable for the scientific process.
+Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privalige to operate outside of these incentives.
+We therefore propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
+
+## The epistemic value of studies from a bayesian perspective
+
+Let us start by defining what we mean by expected epistemic value.
+If researchers plan to conduct a study they usually hope it will change their assessment of the verisimilitude (truthlikeness) of some theory.
+In other words, they hope to learn something from conducting the study.
+The amount of knowledge researchers gain from a particular study in relation to a specific theory is what we call epistemic value.
+While researcher can not know what exactly they will learn from a study, they can form an expectation.
+This expectation is what we call expected epistemic value.
+We assume three things about this process of estimation and how it relates to choosing a study to conduct.
+
+1. Researchers judge evidence for or against a hypothesis rationally.
+2. They expect other researchers to apply the same rational process.
+2. All else equal, researchers try to increase the expected epistemic value for other researchers.
+
+The first assumption leads to our adoption of a Bayesian framework.
+Our rational is as follows.
+Researchers that decide to conduct a study are akin to choosing a study to bet on.
+They have to "place the bet" by conducting the study therefore invest resources and stand to gain epistemic value with some probability.
+This conceptualization of choosing a study to conduct as a betting problem, lets us apply a "Dutch Book" argument.
+This argument, states that any better must follow the axioms of probability if they want to avoid being "irrational", i.e., accept bets with sure losses.
+Fully developing a Dutch book argument for this problem requires a careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in choosing a study.
+Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
+The axioms of probability are all that is required to derive the famous Bayes formula, on which we will heavily rely for our arguments.
+Please note that our decision to adopt a Bayesian philosophy of science does not imply anything about the statistical methods researcher use.
+In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+Our conclusions can be applied to bayesian and frequentist methods alike.
+Of course, frequentist methods are traditionally viewed from a poperian philosophy of science, but our conclusions are compatible with it and notable later advancements like the error statistical view.
+However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make.
+
+## Relating Riskiness to epistemic value
+
+The first point is that risky predictions create persuasive evidence, if they turn out to be correct.
+This point is important because we attribute much of the appeal of preregistration to this fact.
+
+Let us make some simplifying assumptions and define notation.
+We restrict ourselfs to consider evidence of a binary nature (either exists or does not), since continues evidence would led to some quite involved derivations.
+We denote the probability of a theory before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+The probability of observing evidence under some hypothesis is $P(E|H)$.
+The total probability of evidence, $P(E)$, is given by $P(H)P(E|H) + P(¬H)P(E|¬H)$.
+From these probabilities we can calculate the probability of the hypothesis after observing the evidence via:
+
+$$
+P(H|E) = \frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}
+$$
+
+From the three probabilities involved in the calculation $P(E|¬H)$ is the one our argument rests on.
+
+The other two assumptions are later necessary to connect individual decisions that effect epistemic value to the research community as a whole.
+
+Observe, that since $P(E|¬H)$
+
+Please note, that we adopt a Bayesian rational for the meta scientific process of preregistration but that this does not imply any ties to the methods a researcher uses.
+
+However, we aim to show that preregistration is indispensable for adequately judging a much broader range of studies.
+To that end, we first rationalize the appreciation for conformation using Bayesian reasoning.
+Equipped with some of the basic tools of Bayesian Philosophy of Science, we can move beyond a simple dichotomy of exploration and confirmation.
+
+To that end we show that the appreciation for the distinction between exploration and confirmatory neatly falls in line with Bayesian reasoning connected through a quantity we call *theorethical risk*.
+Than we proceed to show that preregistration impacts the theorethical risk via two pathways.
+Only considering the one path rationalizes recommendation to employ preregistration only for confirmatory studies, taking both paths together into consideration leads to a much wider applicability of preregistration.
+
+But what exactly have you been cheated of?
 This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
-We leverage this reasoning to outline two paradigms that rationalize the use of preregistration.
+<!-- We leverage this reasoning to outline two paradigms that rationalize the use of preregistration. -->
+In order to apply Bayesian reasoning we have to cast conformation and exploration in terms of probability.
+Consider the hypothesis that people who regularly engage in fitness activities are healthier.
+A confirmatory story might operationalize regularly as weekly, fitness activity as light jogging and healthy as blood cholesterol level.
+A more exploratory study might consider several definitions of "regularly", "fitness activities", and "healthy", e.g.
+"monthly", "weekly", "daily", or "marathon", "light jogging", "intense cleaning", or "cholesterol", "self rated health", "ability to play tag".
+Both studies have some probability to turn up with positive results even in a world where fitness activities do not lead to healthier people, because there are other plausible explanations for the results.
+However, for the second study the number of explanations that could lead to evidence in favor, aside from the theory is much greater.
+Suppose a researcher actually conducted the second study but is reporting it as if they conducted the first study.
+You would feel cheated.
+But what exactly have you been cheated of?
+In fact, you have been cheated out of the information about all the possibilities that could have brought about this evidence.
+We summarize these possibilities as a probability.
+Therefore, we argue that confirmatory studies have a low probability of observing evidence in favor of their theory if we assume the theory to be false.
+Exploratory studies on the other hand are characterized by a higher probability that they find evidence, even when their hypothesis turns out to be wrong.
+
+We connect the question of exploration vs confirmation to Bayesian reasoning via a quantity we call theoretical risk (we borrow the term from Meehl).
+
+Theoretical risk is the inverse of the probability that one observes evidence in favor of a theory when we assume that the theory does not hold.
+We assume that confirmatory studies take a high theoretical risk, compared to exploratory studies which take a low theoretical risk.
+The tighter your definition of "evidence in favor" is, the more the probability of observing the evidence is tied to your theory holding.
 

### 2022-02-17 README.Rmd
introduce bayes

@@
 The scientific community, widely agrees that actively concealing such "trying for results" is misconduct.
-But what exactly is such concealment cheating the scientific community of?
+But how is such concealment hurting the scientific process?
 We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
@@
 This, of course, goes against common researchers intuition about what is considered exploratory and confirmatory.
-We point these issues out as symptoms of a misunderstanding.#change
-We do want not to discredit preregistration as a useful practice for advancing science, but challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
+We take these issues as signals of a misunderstanding about how preregistration contributes to the scientific progress.
+In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
 
@@
 The aforementioned issues arise by delegating this judgment to a simple decision rule, that is, if something is preregistered it is confirmatory and if it is not preregistered it is exploratory.
+This rule is often assumed implicitly as a rule of thumb, other times it is stated as the explicit goal of preregistration, but always leads to confusion when researchers try to apply it outside specific bounds.
 Preregistering an analysis that is inherently exploratory (like testing dozens of relations) does not make it confirmatory nor will a careful confirmatory study become exploratory if researcher deviate from the preregistration in minor details.
-To uphold the simple decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
-Under these conditions the decision rule holds, but such restricted use makes preregistration a niche solution not able to match the greater problem of replicability in psychology and elsewhere.
+To uphold the simplilicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
+Under these conditions the rule holds, but such restricted use makes preregistration a niche solution not able to match the greater problem of replicability in psychology and elsewhere.
 
@@
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study in relation to a specific theory is what we call epistemic value.
+The amount of knowledge researchers gain from a particular study in relation to the versimilitude of a specific theory is what we call epistemic value.
 While researcher can not know what exactly they will learn from a study, they can form an expectation.
@@
 The axioms of probability are all that is required to derive the famous Bayes formula, on which we will heavily rely for our arguments.
-Please note that our decision to adopt a Bayesian philosophy of science does not imply anything about the statistical methods researcher use.
+Please note that our decision to adopt this aspect of Bayesian philosophy of science does not imply anything about the statistical methods researcher use.
 In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 Our conclusions can be applied to bayesian and frequentist methods alike.
-Of course, frequentist methods are traditionally viewed from a poperian philosophy of science, but our conclusions are compatible with it and notable later advancements like the error statistical view.
-However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make.
+Of course, frequentist methods are traditionally viewed from a poperian philosophy of science, but our conclusions are compatible with this view and and notable later advancements like the error statistical view.
+However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make about the value of preregistration.
 
@@
 The probability of observing evidence under some hypothesis is $P(E|H)$.
-The total probability of evidence, $P(E)$, is given by $P(H)P(E|H) + P(¬H)P(E|¬H)$.
-From these probabilities we can calculate the probability of the hypothesis after observing the evidence via:
+We can calculate the probability of the hypothesis after observing the evidence with help from the Bayes formula:
+
+$$
+P(H|E) = \frac{P(H)P(E|H)}{P(E)}
+$$
+
+There we can replace the total probability of evidence, $P(E)$ with $P(H)P(E|H) + P(¬H)P(E|¬H)$ and arrive at:
 
@@
 
-From the three probabilities involved in the calculation $P(E|¬H)$ is the one our argument rests on.
+```{r, include=FALSE}
+bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
+```
+
+From the three probabilities involved in the calculation, $P(E|¬H)$ is the one our argument rests on.
+It tells use how likely it is that we find evidence, when in fact the theory is not true.
+How $P(E|¬H)$ relates to $P(H|E)$ is of great relevance, since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In carnaps tradition, in its direct use, it is called corroboration as firmness, in its relation to $P(H)$ it is called increase in firmness.
+We refrain from discussing specific measures of corroboration, since there is no measure that show universally better properties than others.
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+
+In short, we want to increase $P(H|E)$.
+Increases in $P(H|E)$ are associated with increases epistemic value, of which we want to maximize the expectation.
+So how can we increase $P(H|E)$?
+The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E|¬H)$.
+Before we continue, let us note that $P(E|H)$ and $P(E|¬H)$ are not complements (they do not necessarily sum to 1).
+The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
+The second option is similar commonsense, that is an increase $P(E|H)$ leads to higher $P(H|E)$.
+Researcher therefore should make sure that they design their studies in a way that in case their hypothesis is true, they can actually find evidence for it.
+While this truism is strongly related to study design, it is unrelated to preregistration.
+The third option, while mathematically equally straightforward, is not quite as intuitive.
+It concerns $P(E|¬H)$, therefore the probability of observing evidence for the theory, when in fact the theory is wrong.
+Before we discuss what this probability means and how it relates to preregistration, let us show how it impacts $P(H|E)$.
+First, increasing $P(E|¬H)$ does decrease $P(H|E)$.
+
+## How statistical method contribute to $P(E|¬H)$
+
+For the overly simplistic case that the hypothesis is a statistical nill-hypothesis, $P(E|¬H)$ becomes the type I error rate.
+Researchers who choose a smaller type I error rate, can therefore be surer of their results.
+However, the research hypothesis is seldomly equal to the statistical null hypothesis.
+$P(E|¬H)$ encompasses factors also outside of the statistical realm, like the studies design.
+Statistical methods have the great advantage that we have a large toolbox of assessing and controlling their contribution to $P(E|¬H)$.
+Examples of our ability to exert control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²) or cross validation in machine learning.
+These are all tools that help us to factor in the tendency of statistical methods to show results even when in fact their are none and are therefore closely related to the probability of observing evidence when in fact there is none.
 

### 2022-02-19 README.Rmd
save

@@
 ---
-output: github_document
+title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
+output:
+  bookdown::pdf_document2:
+    latex_engine: "xelatex"
+    toc: false
+linestretch: 2
+abstract: |
+  Preregistration is becoming increasingly popular in the psychological sciences.
+  At the same time, many researchers struggle to make precise predictions and feel   overwhelmed by the need to prespecify every detail of their analysis plan.
+  Should they surrender and discard the idea of preregistration altogether?
+  Not at all.
+  We argue for the benefits of preregistration beyond strictly confirmatory studies.
+  Grounded in Bayesian reasoning, we define a formal objective for   preregistration that neither declares changes to a preregistration to be sinful   nor punishes rigor.
+  This objective rests on the relevance of “theoretical risk”, a generalization of   type-I error rate, for judging the evidential support for theories.
+  Our proposal is that preregistration should aim to reduce the uncertainty in   judging the theoretical risk instead of directly limiting the type-I error rate.
+  We argue that this reformulation extends the utility of preregistration and is   closer aligned with researchers intuition.
+  If we do not know what a researcher did, we have great uncertainty about the   theoretical risk they took.
+  This uncertainty severely limits the utility of the researchers' evidence.
 ---
@@
 
-# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.
-
-Preregistration is becoming increasingly popular in the psychological sciences.
-At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
-Should they surrender and discard the idea of preregistration altogether? Not at all.
-We argue for the benefits of preregistration beyond strictly confirmatory studies.
-Grounded in Bayesian philosophy of science, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigor.
-Our argument rests on the relevance of “theoretical risk”, a generalization of type-I error rate, for judging the evidential support for theories.
-Our proposal is that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-We argue that this reformulation extends the utility of preregistration and is closer aligned with researchers intuition.
-If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
-This uncertainty severely limits the utility of the researchers' evidence.
-
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
@@
 Rigorous application of preregistration prevents researcher from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
-The scientific community, widely agrees that actively concealing such "trying for results" is misconduct.
-But how is such concealment hurting the scientific process?
+The scientific community, widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
+We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
 We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
@@
 
-Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
+Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
 The aforementioned issues arise by delegating this judgment to a simple decision rule, that is, if something is preregistered it is confirmatory and if it is not preregistered it is exploratory.
-This rule is often assumed implicitly as a rule of thumb, other times it is stated as the explicit goal of preregistration, but always leads to confusion when researchers try to apply it outside specific bounds.
-Preregistering an analysis that is inherently exploratory (like testing dozens of relations) does not make it confirmatory nor will a careful confirmatory study become exploratory if researcher deviate from the preregistration in minor details.
-To uphold the simplilicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
+This rule is often assumed implicitly as a rule of thumb, other times it is stated as the explicit goal of preregistration.
+In both cases it leads to confusion when applied outside specific bounds.
+Preregistering an analysis that is inherently exploratory (like testing dozens of relations) does not make it confirmatory nor will a carefully conducted confirmatory study become exploratory if researcher deviate from the preregistration in minor details.
+To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
 Under these conditions the rule holds, but such restricted use makes preregistration a niche solution not able to match the greater problem of replicability in psychology and elsewhere.
@@
 Let us note, that a study that is not able to find evidence that a researcher is hoping for, is still immensely valuable for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privalige to operate outside of these incentives.
+Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
 We therefore propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
 
-## The epistemic value of studies from a bayesian perspective
+# The epistemic value of studies from a bayesian perspective
 
@@
 This conceptualization of choosing a study to conduct as a betting problem, lets us apply a "Dutch Book" argument.
-This argument, states that any better must follow the axioms of probability if they want to avoid being "irrational", i.e., accept bets with sure losses.
+This argument, states that any better must follow the axioms of probability if they want to avoid being "irrational", i.e., by accepting bets which lead to sure losses.
 Fully developing a Dutch book argument for this problem requires a careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in choosing a study.
@@
 
-## Relating Riskiness to epistemic value
+# Theorethical Risk and Epistemic Value
 
@@
 We restrict ourselfs to consider evidence of a binary nature (either exists or does not), since continues evidence would led to some quite involved derivations.
-We denote the probability of a theory before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
 The probability of observing evidence under some hypothesis is $P(E|H)$.
@@
 
-There we can replace the total probability of evidence, $P(E)$ with $P(H)P(E|H) + P(¬H)P(E|¬H)$ and arrive at:
+$P(H|E)$ is of great relevance, since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In carnaps tradition, in its direct use, it is called corroboration as firmness, in its relation to $P(H)$ it is called increase in firmness.
+We refrain from discussing specific measures of corroboration, since there is no measure shows universally better properties than others.
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+
+In short, we want to increase $P(H|E)$.
+Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
+So how can we increase $P(H|E)$?
+The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
+The second option is similar commonsense, that is an increase $P(E|H)$ leads to higher $P(H|E)$.
+<!--this sentence is shit-->
+Researcher therefore should make sure that they design their studies in a way that in case their hypothesis is true, they can actually find evidence for it.
+While this truism is strongly related to study design, it is unrelated to preregistration.
+So $P(E)$ remains to consider.
+Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$.
+So the unlikelier it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+In other words high risk, high reword.
+
+If we equate riskiness with low probability of evidence, the bayesian rational is perfectly in line with the observation that risky prediction do lead to persuasive evidence.
+This tension between high risk leading to high reward is central to our consideration of preregistration.
+A high risk, high reword strategy, is bound to result in many losses that are eventually absorbed by the high gains.
+Sustaining many "failed" studies is not exactly aligned with the incentive structure many researchers operate under.
+Therefore researchers have an incentive to appear to take more risks than they actually do, which misleads their colleges to give their claims more credence than they deserve.
+It is here, where preregistrations make their impact.
+They try to ensure a proper judgment of the riskyness of a study, by making tempering with the appearance of $P(E)$ difficult.
+
+To gain a better understanding on how preregistrations can do that, let us take a closer look at what contributes to $P(E)$.
+Using the law of total probability we can split $P(E)$ in two terms:
 
 $$
-P(H|E) = \frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}
+P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
 $$
+
+We already have noted that there is not much to do about $P(H)$ (and hence its counterprobability $P(¬H)$) and that is common sense to increase $P(E|H)$.
+The real lever to pull is therefore $P(E|¬H)$.
+This probability tells use how likely it is that we find evidence, when in fact the theory is not true.
+Its counterprobability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theorethical risk", because it is the risk a theory takes on in making a prediction about the occurrence of particular evidence in its favor.
 
@@
 
-From the three probabilities involved in the calculation, $P(E|¬H)$ is the one our argument rests on.
-It tells use how likely it is that we find evidence, when in fact the theory is not true.
-How $P(E|¬H)$ relates to $P(H|E)$ is of great relevance, since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In carnaps tradition, in its direct use, it is called corroboration as firmness, in its relation to $P(H)$ it is called increase in firmness.
-We refrain from discussing specific measures of corroboration, since there is no measure that show universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+Let us note some interesting properties about theoretical risk $P(¬H|¬E)$.
+First, increasing theorethical risk leads to higher $P(H|E)$ (our objective).
+Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrees when observing the evidence.
+Third, if the theorethical risk equals zero, than $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
+In other words observing a sure fact does not lead credence to a hypothesis.
 
-In short, we want to increase $P(H|E)$.
-Increases in $P(H|E)$ are associated with increases epistemic value, of which we want to maximize the expectation.
-So how can we increase $P(H|E)$?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E|¬H)$.
-Before we continue, let us note that $P(E|H)$ and $P(E|¬H)$ are not complements (they do not necessarily sum to 1).
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsense, that is an increase $P(E|H)$ leads to higher $P(H|E)$.
-Researcher therefore should make sure that they design their studies in a way that in case their hypothesis is true, they can actually find evidence for it.
-While this truism is strongly related to study design, it is unrelated to preregistration.
-The third option, while mathematically equally straightforward, is not quite as intuitive.
-It concerns $P(E|¬H)$, therefore the probability of observing evidence for the theory, when in fact the theory is wrong.
-Before we discuss what this probability means and how it relates to preregistration, let us show how it impacts $P(H|E)$.
-First, increasing $P(E|¬H)$ does decrease $P(H|E)$.
+This sounds like a truism but is directly related to Poppers famous criterion of demarcation.
+He stated, that if it is impossible to proof a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis (Popper 1962, 39).
+We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly hold views on riskiness and epistemic value.
 
-## How statistical method contribute to $P(E|¬H)$
+# Should Preregistration Increase the Theorethical Risk?
 
-For the overly simplistic case that the hypothesis is a statistical nill-hypothesis, $P(E|¬H)$ becomes the type I error rate.
+Now after we discussed that increasing the theoretical risk will increase the epistemic value it is only natural to task preregistration with maximizing theoretical risk.
+
+# How Statistical Methods Influence the Theorethical Risk
+
+Preregistration is often discussed in its relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+If you consider the overly simplistic case where the research hypothesis is equal to the statistical nill-hypothesis, than the alternative Hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
 Researchers who choose a smaller type I error rate, can therefore be surer of their results.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-$P(E|¬H)$ encompasses factors also outside of the statistical realm, like the studies design.
-Statistical methods have the great advantage that we have a large toolbox of assessing and controlling their contribution to $P(E|¬H)$.
-Examples of our ability to exert control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²) or cross validation in machine learning.
-These are all tools that help us to factor in the tendency of statistical methods to show results even when in fact their are none and are therefore closely related to the probability of observing evidence when in fact there is none.
+We argue that $P(E|¬H)$ also encompasses factors outside of the statistical realm most notably the study design and broader analytical strategies.
+Beside fitting a statistical model researchers
+
+Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
+Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross validation in machine learning.
+These are all tools that help us to factor in the tendency of statistical methods to show results even when in fact their are none and are therefore closely related to $P(E|¬H)$  (the probability of observing evidence when in fact there is none).
+
+The point is, the contribution of statistical methods to $P(E|¬H)$ can be formally assesst.
+For many statistical models $P(E|¬H)$ can be analytically computed under some assumptions.
+For those models or assumptions where this is not possible, one can employ monte carlo simulation to estimate $P(E|¬H)$.
+The precision with which researcher can discuss $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+We can not hope to resolve this uncertainty, however, we may discuss its implications.
+
+If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
+Take for example a t test.
+Under ideal circumstances (assumption of independence, normality and so forth), it stays true to its type I error rate.
+However, researchers might do a number of very reasonable things in the broader context of the study.
+They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
+Even if the t test itself does perfectly maintain its type I error rate these factors have to be added to $P(E|¬H)$.
+Statistical properties, therefore, only guarantee a lower bound to the quantity we seek to minimize.
+Put simply, the only thing we can conclude with certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
+
+# What to do with uncertainty about the theorethical risk?
 
@@
 
-Observe, that since $P(E|¬H)$
+Observe, that since $P(E|¬H)$.
 

### 2022-03-02 README.Rmd
save

--- 2022-02-16 README.Rmd
@@
 ---
-output: github_document
+title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
+output:
+  bookdown::pdf_document2:
+    latex_engine: "xelatex"
+    toc: false
+linestretch: 2
+abstract: |
+  Preregistration is becoming increasingly popular in the psychological sciences.
+  At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
+  Should they surrender and discard the idea of preregistration altogether?
+  Not at all.
+  We argue for the benefits of preregistration beyond strictly confirmatory studies.
+  Grounded in Bayesian reasoning, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigour.
+  This objective rests on the relevance of "theoretical risk", a generalization of type-I error rate, for judging the evidential support for theories.
+  We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
+  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
+  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
+  This uncertainty severely limits the utility of the researchers' evidence.
 ---
@@
 
-# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.
-
-Preregistration is becoming increasingly popular in the psychological sciences.
-At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
-Should they surrender and discard the idea of preregistration altogether? Not at all.
-We argue for the benefits of preregistration beyond strictly confirmatory studies.
-Grounded in Bayesian philosophy of science, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigor.
-Our argument rests on the relevance of “theoretical risk”, a generalization of type-I error rate, for judging the evidential support for theories.
-Our proposal is that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-We argue that this reformulation extends the utility of preregistration and is closer aligned with researchers intuition.
-If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
-This uncertainty severely limits the utility of the researchers' evidence.
-
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis generating and hypothesis testing and so forth (see XXXX).
-To honor this distinction, many psychological studies are preregistered to produce results that are considered confirmatory.
-Rigorous application of preregistration prevents researcher from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
-The scientific community, widely agrees that actively concealing such "trying for results" is misconduct.
-But what exactly is such concealment cheating the scientific community of?
+Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth (see XXXX).
+To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory.
+Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
+The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
 We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
-These misunderstandings are interesting from a theoretical perspective but also lead to decreased efficiency of the scientific process if researcher  misguidedly decide that preregistration is not applicable for their intended study.
+These misunderstandings are interesting from a theoretical perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
 
 One reason why researchers are hesitant to preregister is that they feel too limited by preregistration.
-Even researchers who have been persuaded by its advantages sometimes apply strategies to skirt the self imposed restrictions.
-One strategy is to write a loose preregistration to begin with, another is to simply deviate from the preregistration afterward.
+Even researchers who have been persuaded by its advantages sometimes apply strategies to skirt the self-imposed restrictions.
+One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
 To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating good looking results no matter the nature of the phenomenon under study.
-To summarize, the conundrum is that current conceptualization of preregistration deem a study "properly" preregistrated even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
+To summarize, the conundrum is that the current conceptualization of preregistration deem a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
 This, of course, goes against common researchers intuition about what is considered exploratory and confirmatory.
-We point these issues out as symptoms of a misunderstanding.#change
-We do want not to discredit preregistration as a useful practice for advancing science, but challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
+We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
+In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
 
-Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
-The aforementioned issues arise by delegating this judgment to a simple decision rule, that is, if something is preregistered it is confirmatory and if it is not preregistered it is exploratory.
-Preregistering an analysis that is inherently exploratory (like testing dozens of relations) does not make it confirmatory nor will a careful confirmatory study become exploratory if researcher deviate from the preregistration in minor details.
-To uphold the simple decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analysed without changes no matter what.
-Under these conditions the decision rule holds, but such restricted use makes preregistration a niche solution not able to match the greater problem of replicability in psychology and elsewhere.
+Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
+The issues above arise by delegating this judgment to a simple decision rule.
+That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
+This rule is often assumed implicitly as a rule of thumb.
+Other times it is stated as the explicit goal of preregistration.
+In both cases, it leads to confusion when applied outside specific bounds.
+Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
+To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what.
+Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
 
 We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
-To that end, we first introduce some basic tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theorethical risk", that is inversely related to type I error rate.
-We than propose two interpretations of how theorethical risk is influenced by preregistration.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk", which is inversely related to type I error rate.
+We then propose two interpretations of how theoretical risk is influenced by preregistration.
 The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
-The second interpretation, is our main contribution, and shows broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
-Under this interpretation theoretical risk is not necessarily directly maximized by preregistration but the uncertainty in judging the theoretical risk is minimized.
+The second interpretation is our main contribution and shows the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
+Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
 This changed viewpoint is crucial for the following reason.
-Directly maximizing theoretical risk does, as the name implies, increase the risk that a study does fail to deliver results in favor of the theory.
-Minimizing the uncertainty around the theoretical risk, as we will later show, shares the beneficial properties but without increasing likelihood of a "failed" study.
-Let us note, that a study that is not able to find evidence that a researcher is hoping for, is still immensely valuable for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privalige to operate outside of these incentives.
-We therefore propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
+As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
+As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
+Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
+Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
+We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
 
-## The epistemic value of studies from a bayesian perspective
+# Epistemic value and the Bayesian rationale
 
 Let us start by defining what we mean by expected epistemic value.
-If researchers plan to conduct a study they usually hope it will change their assessment of the verisimilitude (truthlikeness) of some theory.
+If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study in relation to a specific theory is what we call epistemic value.
-While researcher can not know what exactly they will learn from a study, they can form an expectation.
+The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
+While researchers can not know what exactly they will learn from a study, they can form an expectation.
 This expectation is what we call expected epistemic value.
-We assume three things about this process of estimation and how it relates to choosing a study to conduct.
+We assume three things about this estimation process and how it relates to choosing a study to conduct.
 
-1. Researchers judge evidence for or against a hypothesis rationally.
+1. Researchers judge the evidence for or against a hypothesis rationally.
 2. They expect other researchers to apply the same rational process.
@@
 The first assumption leads to our adoption of a Bayesian framework.
-Our rational is as follows.
+Our rationale is as follows.
 Researchers that decide to conduct a study are akin to choosing a study to bet on.
-They have to "place the bet" by conducting the study therefore invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study to conduct as a betting problem, lets us apply a "Dutch Book" argument.
-This argument, states that any better must follow the axioms of probability if they want to avoid being "irrational", i.e., accept bets with sure losses.
-Fully developing a Dutch book argument for this problem requires a careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in choosing a study.
+They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
+This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument.
+This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
+Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
 Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
-The axioms of probability are all that is required to derive the famous Bayes formula, on which we will heavily rely for our arguments.
-Please note that our decision to adopt a Bayesian philosophy of science does not imply anything about the statistical methods researcher use.
+The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
+Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
 In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 Our conclusions can be applied to bayesian and frequentist methods alike.
-Of course, frequentist methods are traditionally viewed from a poperian philosophy of science, but our conclusions are compatible with it and notable later advancements like the error statistical view.
-However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make.
+Of course, frequentist methods are traditionally viewed from a Popperian philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
+However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make about the value of preregistration.
 
-## Relating Riskiness to epistemic value
+# Epistemic Value and Theoretical Risk
 
-The first point is that risky predictions create persuasive evidence, if they turn out to be correct.
-This point is important because we attribute much of the appeal of preregistration to this fact.
+The first point is that risky predictions create persuasive evidence if they turn out to be correct.
+This point is crucial because we attribute much of the appeal of preregistration to this fact.
 
 Let us make some simplifying assumptions and define notation.
-We restrict ourselfs to consider evidence of a binary nature (either exists or does not), since continues evidence would led to some quite involved derivations.
-We denote the probability of a theory before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+We restrict ourselves to consider evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
 The probability of observing evidence under some hypothesis is $P(E|H)$.
-The total probability of evidence, $P(E)$, is given by $P(H)P(E|H) + P(¬H)P(E|¬H)$.
-From these probabilities we can calculate the probability of the hypothesis after observing the evidence via:
+We can calculate the probability of the hypothesis after observing the evidence with help from the Bayes formula:
 
 $$
-P(H|E) = \frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}
+P(H|E) = \frac{P(H)P(E|H)}{P(E)}
 $$
 
-From the three probabilities involved in the calculation $P(E|¬H)$ is the one our argument rests on.
+$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
+We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
 
-The other two assumptions are later necessary to connect individual decisions that effect epistemic value to the research community as a whole.
+In short, we want to increase $P(H|E)$.
+Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
+So how can we increase $P(H|E)$?
+The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
+The second option is similar commonsense, that is an increase in $P(E|H)$ leads to higher $P(H|E)$.
+Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis in case it is true.
+While this truism is strongly related to study design, it is unrelated to preregistration.
+So $P(E)$ remains to consider.
+Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$.
+So the unlikelier it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+In other words, high risk, high reward.
 
-Observe, that since $P(E|¬H)$
+If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
+This tension between high risk leading to high reward is central to our consideration of preregistration.
+A high risk, high reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
+Sustaining many "failed" studies is not exactly aligned with the incentive structure many researchers operate under.
+Therefore researchers have an incentive to appear to take more risks than they actually do, which misleads their colleges to give their claims more credence than they deserve.
+It is here where preregistrations make their impact.
+They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult.
+
+To better understand how preregistrations can do that, let us take a closer look at what contributes to $P(E)$.
+Using the law of total probability, we can split $P(E)$ into two terms:
+
+$$
+P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
+$$
+
+We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
+The real lever to pull is, therefore, $P(E|¬H)$.
+This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
+Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+
+```{r, include=FALSE}
+bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
+```
+
+Let us note some interesting properties about theoretical risk $P(¬H|¬E)$.
+First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
+Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
+Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
+In other words, observing a sure fact does not lend credence to a hypothesis.
+
+This sounds like a truism but is directly related to Poppers famous criterion of demarcation.
+He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis (Popper 1962, 39).
+We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+
+# Preregistration Tasked with Increasing the Theoretical Risk
+
+After we discussed that increasing the theoretical risk will increase the epistemic value, it is only natural to task preregistration with maximizing theoretical risk.
+In fact, limiting type I error rate is a commonly stated as a goal of preregistration.
+While we think it is often beneficial to increase the theorethical risk but it is comes with a cost.
+
+# Theorethical Risk and Statistical Methods
+
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+If you consider the overly simplistic case where the research hypothesis is equal to the statistical nill-hypothesis, then the alternative hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
+Researchers who choose a smaller type I error rate can hence be surer of their results.
+However, the research hypothesis is seldomly equal to the statistical null hypothesis.
+We argue that $P(E|¬H)$ also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+
+Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
+Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
+These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to $P(E|¬H)$  (the probability of observing evidence when in fact, there is none).
+
+The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
+For many statistical models, $P(E|¬H)$ can be analytically computed under some assumptions.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate $P(E|¬H)$.
+The precision with which researchers can discuss $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+We can not hope to resolve this uncertainty; however, we may discuss its implications.
+
+If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
+Take, for example, a t-test.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
+However, researchers might do many very reasonable things in the broader context of the study.
+They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors have to be added to $P(E|¬H)$.
+Statistical properties, therefore, only guarantee a lower bound to the quantity we seek to minimize.
+Put simply, the only thing we can conclude with certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
+
+# Preregistration Tasked with Decreasing Uncertainty about the Theoretical Risk
+
+The other two assumptions are later necessary to connect individual decisions that affect epistemic value to the research community as a whole.
+
+Observe, that since $P(E|¬H)$.
 

### 2022-05-18 README.Rmd
add problems
small changes

@@
     toc: false
-linestretch: 2
+linestretch: 1
 abstract: |
@@
   We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
-  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took.
-  This uncertainty severely limits the utility of the researchers' evidence.
+  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took which severely limits the utility of the researchers' evidence.
+  Preregistration effectively reduces this uncertainty and therefore generates more persuasiveness evidence.
 ---
@@
 ```
+
+<!--problems to solve:
+* I feel like I am describing a ruby goldberg machine here and expect the reader to understand what happens when you move the spoon
+* what about falsificationists, every time I write a paragraph relating to I delete it and try to move it a discussion
+* references!!!
+-->
 
@@
 That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
-This rule is often assumed implicitly as a rule of thumb.
+This rule of thumb is often assumed implicitly.
 Other times it is stated as the explicit goal of preregistration.
@@
 We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk", which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" (a term borrowed from Meehl), which is inversely related to type I error rate.
 We then propose two interpretations of how theoretical risk is influenced by preregistration.
@@
 In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-Our conclusions can be applied to bayesian and frequentist methods alike.
-Of course, frequentist methods are traditionally viewed from a Popperian philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
-However, those philosophies of science require a much richer set of assumptions that are not quite necessary for the points we want to make about the value of preregistration.
 
@@
 $P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
+In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.[^corroboration]
+<!--I think I get myself in hot water here, one the one hand I claim compatibility of hypo-deductive methods and the other hand I appeal to carnap!!!-->
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
 However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+
+[^corroboration]: This move to assume measures of corroboration is a leap of faith and not covered by the Dutch book argument we make earlier, so a falsificationist can rightfully reject it.
 

### 2022-05-18 README.Rmd
structure following sections

@@
 Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
+
+<!--this does not belong to the introduction-->
 This changed viewpoint is crucial for the following reason.
@@
 We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
+
+To arrive at this interpretation we rely on three arguments.
+The first is that theoretical risk is important for judging theories.
+This argument is not self evident for a conformationist, so we make a lengthy argument for why an inductionist should care.
+We assume that a falsificationist is generally interested in the probability of finding evidence against a theory if it is wrong.
+The second argument is that the theoretical risk for a given study is generally uncertain.
+The third and last assumption is that this uncertainty is reduced by applying preregistration.
+
+When we put these arguments into the Bayes' formula, we conclude that decreasing uncertainty about the theoretical risk increases our expectation to gain evidence for or against a theory.
+We will later discuss how our argument relates to a frequentist philosophy of science.
+For now we stick to a Bayesian rational since it is straightforward to incorporate uncertainty; a move that would require considerable thought otherwise.
 
@@
 
-# Preregistration Tasked with Increasing the Theoretical Risk
+# Preregistration as a Means to Increase Theoretical Risk?
 
 After we discussed that increasing the theoretical risk will increase the epistemic value, it is only natural to task preregistration with maximizing theoretical risk.
-In fact, limiting type I error rate is a commonly stated as a goal of preregistration.
-While we think it is often beneficial to increase the theorethical risk but it is comes with a cost.
+In fact, limiting type I error rate is commonly stated as a goal of preregistration.
+We argue that while such conclusion it plausible, we must first consider at least two constrains that put an upper bound to the theoretical risk.
 
-# Theorethical Risk and Statistical Methods
+First, the theory itself.
+Some theories simply do not make risky predictions and preregistration will not change that.
+It would show a great deal of arrogance to advise researchers to stop investigating such theories.
+Consider the case of a researcher contemplating the relation between two sets of variables.
+Suppose each set considered alone is well studied and strong theories tell the researcher how the variables within the set relate.
+But our imaginary researcher considers the relation of these two sets and for lack of a better theory they assume that some relation between any variables of the two sets exists.
+This is not a risky prediction to make in psychology even without statistical issues like alpha inflation (cite crud factor).
+However, we would consider it an success if the researcher would use the evidence to develop a more precise (and hence risky) theory, e.g.
+by proposing which variables from the one set relate to which from the other set to what extend, direction, functional shape etc.
+We will later show that preregistration increases the expected epistemic value of the further specified theory, though it remains low till substantiated by testing it again.
+The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
+Changing what theories to test is not on our agenda.
+
+The second upper bound to theoretical risk is resources.
+Increasing $P(¬E|¬H)$ will usually decease $P(E|H)$ unless more resources are invested.
+In other words, one can not increase power, while maintaining the same type I error rate, without increasing the invested resources.
+Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade off.
+As we alluded to earlier the resource and incentive structure of research also does not make this easier.
+
+# Uncertainty about Theorethical Risk
+
+We established that higher theoretical risk leads to more credible evidence.
+Nevertheless there are trait offs for increasing it.
+In other words we have reconstructed the interpretation that preregistrations restrict the researchers which increases theoretical risk (or equivalently limit type I error rate) and thereby creates more impactfull evidence.
+However, employing a mathematical framework allows us to navigate the trade offs more effectively and move towards a second, more favorable, interpretation.
+To that end, we incorporate uncertainty into the our mathematical framework.
+
+## Statistical Methods
 
@@
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to $P(E|¬H)$  (the probability of observing evidence when in fact, there is none).
+These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to theorethical risk.
 
 The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
-For many statistical models, $P(E|¬H)$ can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate $P(E|¬H)$.
-The precision with which researchers can discuss $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution $P(E|¬H)$.
+The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
 We can not hope to resolve this uncertainty; however, we may discuss its implications.
@@
 Even if the t-test itself perfectly maintains its type I error rate, these factors have to be added to $P(E|¬H)$.
-Statistical properties, therefore, only guarantee a lower bound to the quantity we seek to minimize.
-Put simply, the only thing we can conclude with certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
+While in theory these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that in practice this is not the case.
+If researchers want to or not, except under strict blinding, they continuously observe information about how the study is going, which at best does not effect them in any way.
+We therefore conclude, that statistical properties, only guarantee a lower bound to the quantity we seek to minimize.
+Put simply, the only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
 
-# Preregistration Tasked with Decreasing Uncertainty about the Theoretical Risk
+## The effects of uncertainty
+
+Before we ask how preregistration is influencing uncertainty about theorethical risk, we must consider
+
+# Preregistration as a Means to Decreasing Uncertainty about the Theoretical Risk
+
+We hope to have persuaded you to accept two arguments: First, theoretical risk is important for judging evidential support for theories.
+Second, the theorethical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
+
+# Old Material
+
+If you are a follow a hypo-deductive rational this is self evident, if you are positivist we hope our
 
@@
 The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".
+
+Our conclusions can be applied to bayesian and frequentist methods alike.
+Of course, frequentist methods are traditionally viewed from a Popperian (or hypo-deductive) philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
+However, error statistical view marries a statistical philosophy with specific statistical models and procedures.
+A marriage which leaves us with richer set of assumptions than necessary for the points we want to make about the value of preregistration.

### 2022-05-20 README.Rmd
send to ylva

@@
   We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
-  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took which severely limits the utility of the researchers' evidence.
-  Preregistration effectively reduces this uncertainty and therefore generates more persuasiveness evidence.
+  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers' intuition.
+  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took, which severely limits the utility of the researchers' evidence.
+  Preregistration effectively reduces this uncertainty and therefore generates more persuasive evidence.
 ---
@@
 knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
+  echo = FALSE
+)
+
+if(!requireNamespace("pacman"))install.packages("pacman")
+pacman::p_load("tidyverse", "ggplot2", "ggthemes", "patchwork")
+```
+
+```{r tikz-setup, include=FALSE}
+if(!requireNamespace("tikzDevice"))install.packages("tikzDevice")
+
+options(
+  tikzDefaultEngine = "xetex",
+  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}")
 )
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth (see XXXX).
+Much has been theorized on this vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth (see XXXX).
 To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory.
-Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
+Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error (XXXX).
 The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
 We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
-These misunderstandings are interesting from a theoretical perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
+These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
 
 One reason why researchers are hesitant to preregister is that they feel too limited by preregistration.
-Even researchers who have been persuaded by its advantages sometimes apply strategies to skirt the self-imposed restrictions.
+Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt the self-imposed restrictions.
 One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
 To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating good looking results no matter the nature of the phenomenon under study.
-To summarize, the conundrum is that the current conceptualization of preregistration deem a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
-This, of course, goes against common researchers intuition about what is considered exploratory and confirmatory.
+To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
+This, of course, goes against common researchers' intuition about what is considered exploratory and confirmatory.
 We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
@@
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what.
+To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what (XXX).
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
@@
 
-<!--this does not belong to the introduction-->
-This changed viewpoint is crucial for the following reason.
-As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
-As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
-We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
-
-To arrive at this interpretation we rely on three arguments.
-The first is that theoretical risk is important for judging theories.
-This argument is not self evident for a conformationist, so we make a lengthy argument for why an inductionist should care.
+To arrive at this interpretation, we rely on three arguments.
+The first is that theoretical risk is vital for judging theories.
+This argument is not self-evident for a conformationist, so we make a lengthy argument for why an inductionist should care.
 We assume that a falsificationist is generally interested in the probability of finding evidence against a theory if it is wrong.
@@
 
-When we put these arguments into the Bayes' formula, we conclude that decreasing uncertainty about the theoretical risk increases our expectation to gain evidence for or against a theory.
-We will later discuss how our argument relates to a frequentist philosophy of science.
-For now we stick to a Bayesian rational since it is straightforward to incorporate uncertainty; a move that would require considerable thought otherwise.
+When we put these arguments into the Bayes' formula, we conclude that because preregistration decrease the uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration are warranted for any study.
+We will later discuss how our argument relates to a (Neo-)Popperian philosophy of science.
+For now, we stick to a Bayesian rationale since quantifying uncertainty is crucial for our argument, which would require considerable thought in a falsificationist framework.
 
@@
 The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-While researchers can not know what exactly they will learn from a study, they can form an expectation.
+While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
 This expectation is what we call expected epistemic value.
@@
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument.
+This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument (XXX).
 This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
@@
 The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
+The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
 Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
@@
 
-The first point is that risky predictions create persuasive evidence if they turn out to be correct.
+Our first argument is that risky predictions create persuasive evidence if they turn out to be correct.
 This point is crucial because we attribute much of the appeal of preregistration to this fact.
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to consider evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
@@
 $P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.[^corroboration]
-<!--I think I get myself in hot water here, one the one hand I claim compatibility of hypo-deductive methods and the other hand I appeal to carnap!!!-->
+In Carnaps (xxx) tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
 However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
-
-[^corroboration]: This move to assume measures of corroboration is a leap of faith and not covered by the Dutch book argument we make earlier, so a falsificationist can rightfully reject it.
 
@@
 The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsense, that is an increase in $P(E|H)$ leads to higher $P(H|E)$.
-Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis in case it is true.
+The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
+Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
 While this truism is strongly related to study design, it is unrelated to preregistration.
@@
 
-Let us note some interesting properties about theoretical risk $P(¬H|¬E)$.
+Let us note some interesting properties of theoretical risk $P(¬H|¬E)$.
 First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
@@
 
-This sounds like a truism but is directly related to Poppers famous criterion of demarcation.
+This sounds like a truism but is directly related to Popper's famous criterion of demarcation.
 He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis (Popper 1962, 39).
@@
 In fact, limiting type I error rate is commonly stated as a goal of preregistration.
-We argue that while such conclusion it plausible, we must first consider at least two constrains that put an upper bound to the theoretical risk.
+We argue that while such a conclusion is plausible, we must first consider at least two constraints that put an upper bound on the theoretical risk.
 
-First, the theory itself.
-Some theories simply do not make risky predictions and preregistration will not change that.
+First is the theory itself.
+Some theories simply do not make risky predictions, and preregistration will not change that.
 It would show a great deal of arrogance to advise researchers to stop investigating such theories.
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set considered alone is well studied and strong theories tell the researcher how the variables within the set relate.
-But our imaginary researcher considers the relation of these two sets and for lack of a better theory they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology even without statistical issues like alpha inflation (cite crud factor).
-However, we would consider it an success if the researcher would use the evidence to develop a more precise (and hence risky) theory, e.g.
-by proposing which variables from the one set relate to which from the other set to what extend, direction, functional shape etc.
+Suppose each set considered alone is well studied, and strong theories tell the researcher how the variables within the set relate.
+But our imaginary researcher considers the relation between these two sets.
+For lack of a better theory, they assume that some relation between any variables of the two sets exists.
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation (cite crud factor XXX).
+However, we would consider it a success if the researcher would use the evidence to develop a more precise (and hence risky) theory, e.g.
+by proposing which variables from the one set relate to which from the other set to what extent, direction, functional shape etc.
 We will later show that preregistration increases the expected epistemic value of the further specified theory, though it remains low till substantiated by testing it again.
@@
 The second upper bound to theoretical risk is resources.
-Increasing $P(¬E|¬H)$ will usually decease $P(E|H)$ unless more resources are invested.
-In other words, one can not increase power, while maintaining the same type I error rate, without increasing the invested resources.
-Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade off.
-As we alluded to earlier the resource and incentive structure of research also does not make this easier.
+Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
+In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
+Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
 
-# Uncertainty about Theorethical Risk
+At last, we can not help but speculate about a sociological factor at play.
+If preregistrations aim to increase theoretical risk, then even researchers who are sold on the merits may deem the associated "high risk; high reward"-strategy as unfit for the incentive structure they operate under.
+
+# Uncertainty about Theoretical Risk
 
 We established that higher theoretical risk leads to more credible evidence.
-Nevertheless there are trait offs for increasing it.
-In other words we have reconstructed the interpretation that preregistrations restrict the researchers which increases theoretical risk (or equivalently limit type I error rate) and thereby creates more impactfull evidence.
-However, employing a mathematical framework allows us to navigate the trade offs more effectively and move towards a second, more favorable, interpretation.
-To that end, we incorporate uncertainty into the our mathematical framework.
+Nevertheless, there are trait offs for increasing it.
+In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.
+However, employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
+To that end, we incorporate uncertainty into our framework.
 
@@
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to theorethical risk.
+These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to theoretical risk.
 
@@
 
+## Causes of Uncertainty
+
+As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
+However, other factors in the broader context of the study are much harder to quantify.
 If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
@@
 They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors have to be added to $P(E|¬H)$.
-While in theory these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that in practice this is not the case.
+Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
+While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 If researchers want to or not, except under strict blinding, they continuously observe information about how the study is going, which at best does not effect them in any way.
-We therefore conclude, that statistical properties, only guarantee a lower bound to the quantity we seek to minimize.
-Put simply, the only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
+We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
+The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
+
+Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of the resulting article (or, more generally, the consumer of the research product).
+We have to remember that the process of weighing evidence for or against a theory extends to the original authors and all the people they hope to pursue.
+While the authors may have a deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
+In particular, they never know what relevant factors the authors, intentionally or not, failed to lay open.
+From an ultimate sceptic perspective, they may claim maximum uncertainty.
 
@@
 
-Before we ask how preregistration is influencing uncertainty about theorethical risk, we must consider
+Before we ask how preregistration is influencing uncertainty about theoretical risk, we must consider the implications of being uncertain.
+Within the Bayesian framework, this is both straightforward and insightful.
+To get an expectation, we express uncertainty as a probability distribution and then integrate over it:
+
+$$
+\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H))
+$$
+
+To illustrate the effect of uncertainty, let $p(E|H) = .8$ (e.g., power of 80%) and $p(H) = .1$ and assume a uniform distribution for $p(E|\neg{}H)$ of the form:
+
+$$
+f(x)=\begin{cases}
+  \frac{1}{1 - \alpha} & \mathrm{for}\ \alpha \le x \le 1,\\
+  0 & \mathrm{for}\ x<\alpha\ \mathrm{or}\ x>1
+  \end{cases}
+$$
+
+Where $\alpha$ is, e.g., the statistical model's nominal type I error rate.
+
+We chose this uniform distribution to capture our statement that a statistical model only guarantees a lower bound to $p(E|\neg{}H)$ (and since its a probability the upper bound is 1) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework (https://arxiv.org/pdf/0708.1593.pdf XXX).
+
+```{r, include=FALSE, cache=TRUE}
+bayes <- function(h1, power, alpha, ...) {
+  (h1 * power) / (h1 * power + (1 - h1) * alpha)
+}
+uncertain_bayes <- function(h1, power, alpha, alpha_upper = 1, ...) {
+  mean(bayes(h1, power, runif(1e6, alpha, alpha_upper)))
+}
+
+step <- 0.01
+vary_alpha <-
+  tidyr::expand_grid(
+    h1 = .1,
+    power = .8,
+    alpha = seq(step, 1 - step, step)
+  ) %>% 
+  dplyr::mutate(.,
+    certain = purrr::pmap_dbl(., bayes),
+    uncertain = purrr::pmap_dbl(., uncertain_bayes)
+  )
+
+vary_alpha_long <- vary_alpha %>% 
+  pivot_longer(c(certain, uncertain),
+               names_to = "certainty",
+               values_to = "h2")
+```
+
+```{r, dev = "tikz", warning=FALSE}
+vary_alpha_long %>%
+  ggplot() +
+  geom_line(aes(alpha, h2, linetype = certainty)) +
+  lims(x = c(0, 1), y = 0:1) +
+  geom_ribbon(data = vary_alpha, aes(x = alpha, ymin = uncertain, ymax = certain), alpha = .1) +
+  labs(caption = "$p(E|H) = .8$, $p(H) = .1$",
+       y = "$\\mathbb{E}\\big(p(H|E)\\big)$",
+       x = "$\\alpha$") +
+  theme_tufte() +
+  theme(legend.position = c(.85, .8),
+        legend.title = element_blank())
+```
+
+```{r, include=FALSE}
+# h1 = prior
+# eh = p(E|H)
+# enh = p(E|\neg{}H)
+
+bayes <- function(h1, eh, enh, ...) {
+  (h1 * eh) / (h1 * eh + (1 - h1) * enh)
+}
+# all formulas taken from Bayesian Philosophy of Science, p. 51
+# first letter = function name in Bayesian Philosophy of Science
+# optional letter = p for prime
+# second letter = m for measure
+measures <- list(
+  dm = list(
+    name = "Difference Measure",
+    formula = "d(H, E) = p(H|E) - p(H)",
+    fun = function(h1, eh, enh, ...) {
+      he <- bayes(h1, eh, enh)
+      he - h1
+    }
+  ),
+  rm = list(
+    name = "Log-Ratio Measure",
+    formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
+    fun = function(h1, eh, enh, ...) {
+      he <- bayes(h1, eh, enh)
+      log(he / h1)
+    }
+  ),
+  lm = list(
+    name = "Log-Likelihood Measure",
+    formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
+    fun = function(h1, eh, enh, ...) {
+      he <- bayes(h1, eh, enh)
+      log(he / enh)
+    }
+  ),
+  km = list(
+    name = "Kemeny–Oppenheim Measure",
+    formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
+    fun = function(h1, eh, enh, ...) {
+      (eh - enh) / (eh + enh)
+    }
+  ),
+  zm = list(
+    name = "Generalized Entailment Measure",
+    formula =
+      "z(H, E) =\\begin{cases}
+  \\frac{p(E|H) - p(H)}{1 - p(H)}, & \\text{if } p(H|E) \\geq{} p(H) \\\\
+  \\frac{p(E|H) - p(H)}{p(H)}, & \\text{if } p(H|E) < p(H)
+\\end{cases}",
+fun = function(h1, eh, enh, ...) {
+  he <- bayes(h1, eh, enh)
+  ifelse(he >= h1, (he - h1)/(1 - h1), (he - h1)/h1)
+}
+  ),
+sm = list(
+  name = "Christensen–Joyce Measure",
+  formula = "s(H, E) = p(H|E) - p(H|\\neg{}E)",
+  fun = function(h1, eh, enh, ...) {
+    he <- bayes(h1, eh, enh)
+    e <- h1 * eh + (1 - h1) * enh
+    hne <- (h1 * (1 - eh)) / (1 - e)
+    he - hne
+  }
+),
+cpm = list(
+  name = "Carnap’s Relevance Measure",
+  formula = "c'(H, E) = p(E)(p(H|E) - p(H))",
+  fun = function(h1, eh, enh, ...) {
+    he <- bayes(h1, eh, enh)
+    e <- h1 * eh + (1 - h1) * enh
+    e * (he - h1)
+  }
+),
+rpm = list(
+  name = "Rips Measure",
+  formula = "r'(H, E) = 1 - \\frac{p(\\neg{}H|E)}{p(\\neg{}H)}",
+  fun = function(h1, eh, enh, ...) {
+    he <- bayes(h1, eh, enh)
+    1 - (1 - he) / (1 - h1)
+  }
+)
+)
+
+measures <- tibble(!!!transpose(measures)) %>% 
+  mutate(name = simplify(name, "chr"),
+         formula = simplify(formula, "chr"))
+
+uncertainly <- function(fun, n = 1e5, seed = NULL){
+  function(h1, eh, enh){
+    if(!is.null(seed)){
+      if(exists(".Random.seed"))on.exit(set.seed(.Random.seed))
+      set.seed(seed)
+    }
+    enh_samples <- runif(n, enh, 1)
+    mean(fun(h1, eh, enh_samples))
+  }
+}
+```
+
+```{r, cache=TRUE, echo=FALSE}
+step <- 0.01
+
+grid <- tidyr::expand_grid(
+    h1 = .1,
+    eh = .8,
+    enh = seq(step, 1 - step, step)
+  )
+apply_on_grid <- function(f, grid) {
+  mutate(grid,
+         certain = pmap_dbl(grid, f),
+         uncertain = pmap_dbl(grid, uncertainly(f, seed = .Random.seed)))
+}
+
+results <- mutate(measures, results = map(fun, apply_on_grid, grid)) %>% 
+  select(-fun) %>% 
+  unnest(results) %>% 
+  pivot_longer(c(certain, uncertain),
+               names_to = "certainty",
+               values_to = "value")
+```
+
+```{r, dev = "tikz", cache=TRUE}
+results %>%
+  mutate(name = str_remove(name, "Measure")) %>% 
+  split(., .$name) %>%
+  imap(
+    ~ ggplot(.x, aes(enh, value)) +
+      geom_line(aes(linetype = certainty)) + 
+      labs(title = .x$name, y = element_blank(), x = element_blank())
+  ) %>% 
+  reduce(`+`) +
+  plot_layout(ncol = 2, guides = 'collect') &
+  theme_tufte() &
+  theme(legend.title = element_blank(),
+        legend.position = "bottom",
+        plot.title = element_text(size = 7),
+        axis.text = element_text(size = 6))
+```
+
+Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall in.
+Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
+Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
+Generally, we expect that uncertainty (and hence more entropic distributions) lead to a decreased expected epistemic value.
 
@@
 
-We hope to have persuaded you to accept two arguments: First, theoretical risk is important for judging evidential support for theories.
-Second, the theorethical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
+We hope to have persuaded you to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
+Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
+The last argument is that preregistrations reduce this uncertainty.
 
 # Old Material
+
+This changed viewpoint is crucial for the following reason.
+As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
+As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
+Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
+Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
+We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
 

### 2022-05-30 manuscript.Rmd
rename to manuscript

--- 2022-05-20 README.Rmd
@@
 linestretch: 1
+bibliography: references.bib
 abstract: |
@@
   We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers' intuition.
+  We argue that this reformulation extends the utility of preregistration, provides principled justification for preregistration and is more closely aligned with researchers' intuition.
   If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took, which severely limits the utility of the researchers' evidence.
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on this vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth (see XXXX).
-To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory.
-Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error (XXXX).
+Much has been theorized on this vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory [@nosekPreregistrationRevolution2018].
+Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error [@wagenmakersAgendaPurelyConfirmatory2012].
 The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
@@
 These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
+Asking researcher to invest time and resources into preregistration demands a principled justification for its utility.
 
@@
 This rule of thumb is often assumed implicitly.
-Other times it is stated as the explicit goal of preregistration.
+Other times it is stated as the explicit goal of preregistration [@mellorEasyPreregistrationWill2018].
 In both cases, it leads to confusion when applied outside specific bounds.
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what (XXX).
+To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
@@
 We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" (a term borrowed from Meehl), which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
 We then propose two interpretations of how theoretical risk is influenced by preregistration.
@@
 The first is that theoretical risk is vital for judging theories.
-This argument is not self-evident for a conformationist, so we make a lengthy argument for why an inductionist should care.
-We assume that a falsificationist is generally interested in the probability of finding evidence against a theory if it is wrong.
+This argument comes natural to a (Neo-)Popperian or falsificationist, but  is not self-evident for a conformationist.
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 
-When we put these arguments into the Bayes' formula, we conclude that because preregistration decrease the uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration are warranted for any study.
+When we put these arguments into the Bayes' formula, we conclude that because preregistration decrease the uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
 We will later discuss how our argument relates to a (Neo-)Popperian philosophy of science.
@@
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument (XXX).
+This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
 This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
@@
 
-Our first argument is that risky predictions create persuasive evidence if they turn out to be correct.
+Our first argument is that theorethical risk is crucial for judging theories.
+Put simply, risky predictions create persuasive evidence if they turn out to be correct.
 This point is crucial because we attribute much of the appeal of preregistration to this fact.
@@
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation (cite crud factor XXX).
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and hence risky) theory, e.g.
@@
 
-At last, we can not help but speculate about a sociological factor at play.
-If preregistrations aim to increase theoretical risk, then even researchers who are sold on the merits may deem the associated "high risk; high reward"-strategy as unfit for the incentive structure they operate under.
+This objective also raises the problem of changing the plan after data collection started.
+One could argue that allowing deviating from the preregistration would allow researchers the same freedoms as having no preregistration at all.
+Hypothethically at least, researchers might try any possible deviation and pick the most favorably for their theory.
+Hence, allowing deviation would void any otherwise guaranteed lower limits to the theoretical risk.
+It follows that allowing deviations is wholly incompatible with the goal to maximize theoretical risk.
 
@@
 We established that higher theoretical risk leads to more credible evidence.
-Nevertheless, there are trait offs for increasing it.
 In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.
+Nevertheless, there are trait offs for increasing theoretical risk.
 However, employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
@@
 Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
-If you consider the overly simplistic case where the research hypothesis is equal to the statistical nill-hypothesis, then the alternative hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be surer of their results.
+If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
+Researchers who choose a smaller type I error rate can hence be surer of their results, if significant.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
@@
 For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution $P(E|¬H)$.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
 The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
@@
 Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
-However, researchers might do many very reasonable things in the broader context of the study.
+However, researchers might do many very reasonable things in the broader context of the study that effect theoretical risk.
 They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
@@
 
-Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of the resulting article (or, more generally, the consumer of the research product).
-We have to remember that the process of weighing evidence for or against a theory extends to the original authors and all the people they hope to pursue.
-While the authors may have a deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
-In particular, they never know what relevant factors the authors, intentionally or not, failed to lay open.
-From an ultimate sceptic perspective, they may claim maximum uncertainty.
-
 ## The effects of uncertainty
 
-Before we ask how preregistration is influencing uncertainty about theoretical risk, we must consider the implications of being uncertain.
+Before we ask how preregistration is influencing this uncertainty, we must consider the implications of being uncertain about the theoretical risk.
 Within the Bayesian framework, this is both straightforward and insightful.
@@
 
-We chose this uniform distribution to capture our statement that a statistical model only guarantees a lower bound to $p(E|\neg{}H)$ (and since its a probability the upper bound is 1) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework (https://arxiv.org/pdf/0708.1593.pdf XXX).
+We chose this uniform distribution to capture our statement that a statistical model only guarantees a lower bound to $p(E|\neg{}H)$ (and since its a probability the upper bound is 1) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
@@
 Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-Generally, we expect that uncertainty (and hence more entropic distributions) lead to a decreased expected epistemic value.
+Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
 
@@
 
+Recollect, our three assumptions:
+
+1. Researchers judge the evidence for or against a hypothesis rationally.
+2. They expect other researchers to apply the same rational process.
+2. All else equal, researchers try to increase the expected epistemic value for other researchers.
+
+The point we make with these assumption is that the authors aim to persuade other researchers, their readers.
+Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of the resulting article (or, more generally, the consumer of the research product).
+We have to remember that the process of weighing evidence for or against a theory extends beyond the original authors to all the people they hope to persuade.
+While the authors may have a deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
+In particular, they never know what relevant factors the authors, intentionally or not, failed to lay open.
+From an ultimate sceptic perspective, they may claim maximum uncertainty.
+
+Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
+Preregistration are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
+The authors could have decided to exclude a number of analytic strategies they tried out etc.
+That is not to say any not preregistered study was p hacked.
+The point is, we can not exclude this and a myriad of other possibilities and, hence, are uncertain.
+If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete picture of the situation.
+They still might be uncertain about the actual theoretical risk the authors took but much less so.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
+In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+
+# Discussion
+
+We started out with the observation that preregistrations do no always cleanly divide preregistration from conformation.
+Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
+No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
+However, as the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
+As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
+Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
+Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
+
+The advantage of this perspective that puts uncertainty about theorethical risk front and center is twofold.
+First, aiming to reduce uncertainty is beneficial across a wide range of theoretical risks.
+That is, researchers benefit from preregistrating their study, regardless weather or not the study is "exploratory".
+If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
+Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
+Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and than select the most beneficial for their theory.
+This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
+However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
+
+> As far as I am aware, Mayo’s severity argument currently provides one of the few philosophies of science that allows for a coherent conceptual analysis of the value of preregistration.
+
 # Old Material
 
-This changed viewpoint is crucial for the following reason.
-As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
-As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
 We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
@@
 A marriage which leaves us with richer set of assumptions than necessary for the points we want to make about the value of preregistration.
+
+# References

### 2022-05-31 manuscript.Rmd
move abstract to own file

@@
 bibliography: references.bib
-abstract: |
-  Preregistration is becoming increasingly popular in the psychological sciences.
-  At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
-  Should they surrender and discard the idea of preregistration altogether?
-  Not at all.
-  We argue for the benefits of preregistration beyond strictly confirmatory studies.
-  Grounded in Bayesian reasoning, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigour.
-  This objective rests on the relevance of "theoretical risk", a generalization of type-I error rate, for judging the evidential support for theories.
-  We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-  We argue that this reformulation extends the utility of preregistration, provides principled justification for preregistration and is more closely aligned with researchers' intuition.
-  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took, which severely limits the utility of the researchers' evidence.
-  Preregistration effectively reduces this uncertainty and therefore generates more persuasive evidence.
+abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 

### 2022-05-31 README.Rmd
add actual readme

--- 2022-05-18 README.Rmd
@@
 ---
-title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
-output:
-  bookdown::pdf_document2:
-    latex_engine: "xelatex"
-    toc: false
-linestretch: 1
-abstract: |
-  Preregistration is becoming increasingly popular in the psychological sciences.
-  At the same time, many researchers struggle to make precise predictions and feel overwhelmed by the need to prespecify every detail of their analysis plan.
-  Should they surrender and discard the idea of preregistration altogether?
-  Not at all.
-  We argue for the benefits of preregistration beyond strictly confirmatory studies.
-  Grounded in Bayesian reasoning, we define a formal objective for preregistration that neither declares changes to a preregistration to be sinful nor punishes rigour.
-  This objective rests on the relevance of "theoretical risk", a generalization of type-I error rate, for judging the evidential support for theories.
-  We propose that preregistration should aim to reduce the uncertainty in judging the theoretical risk instead of directly limiting the type-I error rate.
-  We argue that this reformulation extends the utility of preregistration and is more closely aligned with researchers intuition.
-  If we do not know what a researcher did, we have great uncertainty about the theoretical risk they took which severely limits the utility of the researchers' evidence.
-  Preregistration effectively reduces this uncertainty and therefore generates more persuasiveness evidence.
+output: github_document
 ---
@@
 
-<!--problems to solve:
-* I feel like I am describing a ruby goldberg machine here and expect the reader to understand what happens when you move the spoon
-* what about falsificationists, every time I write a paragraph relating to I delete it and try to move it a discussion
-* references!!!
--->
+# Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.
 
-A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on this important distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth (see XXXX).
-To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory.
-Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results that were produced by an arduous process of try and error.
-The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
-We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
-These misunderstandings are interesting from a theoretical perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
-
-One reason why researchers are hesitant to preregister is that they feel too limited by preregistration.
-Even researchers who have been persuaded by its advantages sometimes apply strategies to skirt the self-imposed restrictions.
-One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
-To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating good looking results no matter the nature of the phenomenon under study.
-To summarize, the conundrum is that the current conceptualization of preregistration deem a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
-This, of course, goes against common researchers intuition about what is considered exploratory and confirmatory.
-We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
-In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
-
-Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
-The issues above arise by delegating this judgment to a simple decision rule.
-That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
-This rule of thumb is often assumed implicitly.
-Other times it is stated as the explicit goal of preregistration.
-In both cases, it leads to confusion when applied outside specific bounds.
-Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-To uphold the simplicity of the decision rule, it is routinely suggested to only preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what.
-Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
-
-We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" (a term borrowed from Meehl), which is inversely related to type I error rate.
-We then propose two interpretations of how theoretical risk is influenced by preregistration.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
-The second interpretation is our main contribution and shows the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
-Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
-This changed viewpoint is crucial for the following reason.
-As the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favouring the theory.
-As we will later show, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcome and few, if any, have the privilege to operate outside of these incentives.
-We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
-
-# Epistemic value and the Bayesian rationale
-
-Let us start by defining what we mean by expected epistemic value.
-If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
-In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-While researchers can not know what exactly they will learn from a study, they can form an expectation.
-This expectation is what we call expected epistemic value.
-We assume three things about this estimation process and how it relates to choosing a study to conduct.
-
-1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else equal, researchers try to increase the expected epistemic value for other researchers.
-
-The first assumption leads to our adoption of a Bayesian framework.
-Our rationale is as follows.
-Researchers that decide to conduct a study are akin to choosing a study to bet on.
-They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument.
-This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
-Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
-The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
-Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-
-# Epistemic Value and Theoretical Risk
-
-The first point is that risky predictions create persuasive evidence if they turn out to be correct.
-This point is crucial because we attribute much of the appeal of preregistration to this fact.
-
-Let us make some simplifying assumptions and define notation.
-We restrict ourselves to consider evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
-We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
-The probability of observing evidence under some hypothesis is $P(E|H)$.
-We can calculate the probability of the hypothesis after observing the evidence with help from the Bayes formula:
-
-$$
-P(H|E) = \frac{P(H)P(E|H)}{P(E)}
-$$
-
-$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In carnaps tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.[^corroboration]
-<!--I think I get myself in hot water here, one the one hand I claim compatibility of hypo-deductive methods and the other hand I appeal to carnap!!!-->
-We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
-
-[^corroboration]: This move to assume measures of corroboration is a leap of faith and not covered by the Dutch book argument we make earlier, so a falsificationist can rightfully reject it.
-
-In short, we want to increase $P(H|E)$.
-Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
-So how can we increase $P(H|E)$?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsense, that is an increase in $P(E|H)$ leads to higher $P(H|E)$.
-Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis in case it is true.
-While this truism is strongly related to study design, it is unrelated to preregistration.
-So $P(E)$ remains to consider.
-Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$.
-So the unlikelier it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high reward.
-
-If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
-This tension between high risk leading to high reward is central to our consideration of preregistration.
-A high risk, high reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
-Sustaining many "failed" studies is not exactly aligned with the incentive structure many researchers operate under.
-Therefore researchers have an incentive to appear to take more risks than they actually do, which misleads their colleges to give their claims more credence than they deserve.
-It is here where preregistrations make their impact.
-They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult.
-
-To better understand how preregistrations can do that, let us take a closer look at what contributes to $P(E)$.
-Using the law of total probability, we can split $P(E)$ into two terms:
-
-$$
-P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
-$$
-
-We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
-The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
-Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-
-```{r, include=FALSE}
-bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
+```{r, echo=FALSE, results='asis'}
+cat(readLines('abstract.md'))
 ```
-
-Let us note some interesting properties about theoretical risk $P(¬H|¬E)$.
-First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
-Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
-Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
-In other words, observing a sure fact does not lend credence to a hypothesis.
-
-This sounds like a truism but is directly related to Poppers famous criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis (Popper 1962, 39).
-We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
-
-# Preregistration Tasked with Increasing the Theoretical Risk
-
-After we discussed that increasing the theoretical risk will increase the epistemic value, it is only natural to task preregistration with maximizing theoretical risk.
-In fact, limiting type I error rate is a commonly stated as a goal of preregistration.
-While we think it is often beneficial to increase the theorethical risk but it is comes with a cost.
-
-# Theorethical Risk and Statistical Methods
-
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
-If you consider the overly simplistic case where the research hypothesis is equal to the statistical nill-hypothesis, then the alternative hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be surer of their results.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
-Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to $P(E|¬H)$  (the probability of observing evidence when in fact, there is none).
-
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
-For many statistical models, $P(E|¬H)$ can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate $P(E|¬H)$.
-The precision with which researchers can discuss $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
-We can not hope to resolve this uncertainty; however, we may discuss its implications.
-
-If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
-However, researchers might do many very reasonable things in the broader context of the study.
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors have to be added to $P(E|¬H)$.
-Statistical properties, therefore, only guarantee a lower bound to the quantity we seek to minimize.
-Put simply, the only thing we can conclude with certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees.
-
-# Preregistration Tasked with Decreasing Uncertainty about the Theoretical Risk
-
-The other two assumptions are later necessary to connect individual decisions that affect epistemic value to the research community as a whole.
-
-Observe, that since $P(E|¬H)$.
-
-Please note, that we adopt a Bayesian rational for the meta scientific process of preregistration but that this does not imply any ties to the methods a researcher uses.
-
-However, we aim to show that preregistration is indispensable for adequately judging a much broader range of studies.
-To that end, we first rationalize the appreciation for conformation using Bayesian reasoning.
-Equipped with some of the basic tools of Bayesian Philosophy of Science, we can move beyond a simple dichotomy of exploration and confirmation.
-
-To that end we show that the appreciation for the distinction between exploration and confirmatory neatly falls in line with Bayesian reasoning connected through a quantity we call *theorethical risk*.
-Than we proceed to show that preregistration impacts the theorethical risk via two pathways.
-Only considering the one path rationalizes recommendation to employ preregistration only for confirmatory studies, taking both paths together into consideration leads to a much wider applicability of preregistration.
-
-But what exactly have you been cheated of?
-This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
-<!-- We leverage this reasoning to outline two paradigms that rationalize the use of preregistration. -->
-In order to apply Bayesian reasoning we have to cast conformation and exploration in terms of probability.
-Consider the hypothesis that people who regularly engage in fitness activities are healthier.
-A confirmatory story might operationalize regularly as weekly, fitness activity as light jogging and healthy as blood cholesterol level.
-A more exploratory study might consider several definitions of "regularly", "fitness activities", and "healthy", e.g.
-"monthly", "weekly", "daily", or "marathon", "light jogging", "intense cleaning", or "cholesterol", "self rated health", "ability to play tag".
-Both studies have some probability to turn up with positive results even in a world where fitness activities do not lead to healthier people, because there are other plausible explanations for the results.
-However, for the second study the number of explanations that could lead to evidence in favor, aside from the theory is much greater.
-Suppose a researcher actually conducted the second study but is reporting it as if they conducted the first study.
-You would feel cheated.
-But what exactly have you been cheated of?
-In fact, you have been cheated out of the information about all the possibilities that could have brought about this evidence.
-We summarize these possibilities as a probability.
-Therefore, we argue that confirmatory studies have a low probability of observing evidence in favor of their theory if we assume the theory to be false.
-Exploratory studies on the other hand are characterized by a higher probability that they find evidence, even when their hypothesis turns out to be wrong.
-
-We connect the question of exploration vs confirmation to Bayesian reasoning via a quantity we call theoretical risk (we borrow the term from Meehl).
-
-Theoretical risk is the inverse of the probability that one observes evidence in favor of a theory when we assume that the theory does not hold.
-We assume that confirmatory studies take a high theoretical risk, compared to exploratory studies which take a low theoretical risk.
-The tighter your definition of "evidence in favor" is, the more the probability of observing the evidence is tied to your theory holding.
-
-<!-- Caroline does not see the connection to type I error rate/theoretical risk -->
-
-Let us formalize the judgment about a hypothesis (H) given some evidence (E) as the conditional probability $p(H|E)$.
-
-Using Bayes' rule we arrive at:
-
-$$
-p(H|E) = \frac{p(H)p(E|H)}{p(H)p(E|H) + p(¬H)p(E|¬H)}
-$$
-
-Note the connection to null hypothesis testing.
-If H represents the alternative Hypothesis and ¬H represents the Null hypothesis, then $p(E|H)$ is the power, while $p(E|¬H)$ represents the type I/alpha error.
-However, generally a theory is richer in content, than equating it with the alternaive hypothesis in Null hypothesis testing.
-
-<!-- continue to justify the importance of theorethical risk -->
-
-The first paradigm is embraced by the open science community and conceptualizes preregistration as a tool to reduce the type I error rate (XXXX).
-As we will later show, reducing the type I error rate is indeed a powerful leaver to gather persuasive evidence.
-In practice, however, researchers can not reduce the type I error rate at will.
-They face three limiting factors.
-First, psychological theories tend to be vague.
-Some decisions are just not derivable from theory, which makes it extraordinarily difficult to maintain a type I error rate near zero.
-Second, researchers are constrained by resource limitations.
-One way to counter the first problem is conducting exploratory pilot studies, but running extensive pilot studies is expensive.
-Doubling the resources that go into investigating a research question is often just not tenable and mostly beyond the individual researcher to decide.
-Third, by reducing the type I error rate the researcher directly increases the likelihood that their study will result in disappointing null results.
-A study that fails to deliver results can be crippling to the researchers career.
-Depending on individual circumstances, e.g.
-the theory in question, resources allocated to a study, the career stage of the researcher, a researcher may simply feel unable to reduce the type I error rate to the threshold necessary for a confirmatory study.
-Constrained by these factors researchers tasked with writing a preregistration are incentiviced to either give up or to muddle through by being intentionally vague.
-The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".

### 2022-05-31 manuscript.Rmd
non breaking hyphon

@@
 The first is that theoretical risk is vital for judging theories.
-This argument comes natural to a (Neo-)Popperian or falsificationist, but  is not self-evident for a conformationist.
+This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 When we put these arguments into the Bayes' formula, we conclude that because preregistration decrease the uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
-We will later discuss how our argument relates to a (Neo-)Popperian philosophy of science.
+We will later discuss how our argument relates to a (Neo‑)Popperian philosophy of science.
 For now, we stick to a Bayesian rationale since quantifying uncertainty is crucial for our argument, which would require considerable thought in a falsificationist framework.

### 2022-05-31 manuscript.Rmd
cite popper

@@
 This sounds like a truism but is directly related to Popper's famous criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis (Popper 1962, 39).
+He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
+18].
 We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.

### 2022-05-31 manuscript.Rmd
change font

@@
     toc: false
-linestretch: 1
+header-includes:
+   - \usepackage{libertine}
+fontsize: 11pt
+linestretch: 1.2
+geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
+papersize: a4
 bibliography: references.bib
@@
   tikzDefaultEngine = "xetex",
-  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}")
+  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}", "\\usepackage{libertine}")
 )
@@
 
-```{r, dev = "tikz", warning=FALSE}
+```{r, dev='tikz'}
+#| fig.cap: "Corroboration as firmness ($p(E|H)$) as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
 vary_alpha_long %>%

### 2022-05-31 manuscript.Rmd
try to load encoding

@@
     toc: false
+    number_sections: false
 header-includes:
@@
   tikzDefaultEngine = "xetex",
-  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}", "\\usepackage{libertine}")
+  tikzXelatexPackages = c(
+    getOption("tikzXelatexPackages"),
+    "\\usepackage{amsfonts}",
+    "\\usepackage[utf8]{inputenc}",
+    "\\usepackage{libertine}"
+  )
 )
@@
 
-```{r, dev = "tikz", cache=TRUE}
+```{r, dev = "tikz", cache=TRUE, warning=FALSE}
+#| fig.cap: "Several measures for corroboration as increase in firmness as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
+
 results %>%
@@
       geom_line(aes(linetype = certainty)) +
-      labs(title = .x$name, y = element_blank(), x = element_blank())
+      labs(
+        title = .x$name,
+        y = element_blank(),
+        x = element_blank()
+      )
   ) %>%
@@
   theme_tufte() &
-  theme(legend.title = element_blank(),
+  theme(
+    legend.title = element_blank(),
     legend.position = "bottom",
     plot.title = element_text(size = 7),
-        axis.text = element_text(size = 6))
+    axis.text = element_text(size = 6)
+  )
 ```

### 2022-05-31 manuscript.Rmd
Revert "try to load encoding"
This reverts commit 054312fd65f3212acfa70dc598322ce81056cc79.

@@
   tikzDefaultEngine = "xetex",
-  tikzXelatexPackages = c(
-    getOption("tikzXelatexPackages"),
-    "\\usepackage{amsfonts}",
-    "\\usepackage[utf8]{inputenc}",
-    "\\usepackage{libertine}"
-  )
+  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}", "\\usepackage{libertine}")
 )

### 2022-06-01 manuscript.Rmd
Update manuscript.Rmd

@@
 Much has been theorized on this vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-To honour this distinction, many researchers preregister their studies to produce results that are considered confirmatory [@nosekPreregistrationRevolution2018].
-Rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error [@wagenmakersAgendaPurelyConfirmatory2012].
+To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
+Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error [@wagenmakersAgendaPurelyConfirmatory2012].
 The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
-We believe that a rigorous answer to this question enables us to clear some common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
+We believe that a rigorous answer to this question enables us to clarify common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
 These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
-Asking researcher to invest time and resources into preregistration demands a principled justification for its utility.
+Asking researchers to invest time and resources into preregistration demands a principled justification of its utility.
 
-One reason why researchers are hesitant to preregister is that they feel too limited by preregistration.
-Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt the self-imposed restrictions.
+One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration.
+Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt these self-imposed restrictions.
 One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
-To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating good looking results no matter the nature of the phenomenon under study.
+To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
 To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
-This, of course, goes against common researchers' intuition about what is considered exploratory and confirmatory.
+This, of course, goes against researchers' common intuition about what is considered exploratory and confirmatory.
 We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
@@
 Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
-The issues above arise by delegating this judgment to a simple decision rule.
-That is, if something is preregistered, it is confirmatory, and if it is not preregistered, it is exploratory.
-This rule of thumb is often assumed implicitly.
+The issues outlined above arise by delegating this judgment to a simple decision rule.
+That is, if something is preregistered, it is considered confirmatory, and if it is not preregistered, it is considered exploratory.
+This rule of thumb is often applied implicitly.
 Other times it is stated as the explicit goal of preregistration [@mellorEasyPreregistrationWill2018].
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
-We then propose two interpretations of how theoretical risk is influenced by preregistration.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently limiting type I error.
-The second interpretation is our main contribution and shows the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes.
-Under this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but the uncertainty in judging the theoretical risk is minimized.
+We then outline two interpretations of how theoretical risk is impacted by preregistration.
+The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type I error.
+The second interpretation is our main contribution and demonstrates the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
+Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
 
@@
 The first is that theoretical risk is vital for judging theories.
-This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
+This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist. (WHY; A BIT MORE MORE ELABORATION)
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 
-When we put these arguments into the Bayes' formula, we conclude that because preregistration decrease the uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
+When we put these arguments into the Bayes' formula, we conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
 We will later discuss how our argument relates to a (Neo‑)Popperian philosophy of science.
@@
 
-Let us start by defining what we mean by expected epistemic value.
+Let us start by defining what we mean by expected epistemic value. (METADISCOURSE UNNECESSARY)
 If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
@@
 While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
-This expectation is what we call expected epistemic value.
+This expectation is what we term expected epistemic value.
 We assume three things about this estimation process and how it relates to choosing a study to conduct.
@@
 
-The first assumption leads to our adoption of a Bayesian framework.
+The first assumption leads to our adoption of a Bayesian framework. (WHY)
 Our rationale is as follows.
@@
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem lets us apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
+This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
 This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
@@
 Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
-The axioms of probability are sufficient to derive the famous Bayes formula, on which we will heavily rely for our arguments.
+The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
 The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
@@
 
-Let us make some simplifying assumptions and define notation.
+Let us make some simplifying assumptions and define notation. (METADISCOURSE UNNECESSARY)
 We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
@@
 $P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In Carnaps (xxx) tradition, in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
+In the tradition of Carnap (xxx), in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness. (SENTENCE STRUCTURE)
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
@@
 The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$, but the a priori probability of a hypothesis is nothing our study design can change.
+The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$. However, the a priori probability of a hypothesis is nothing our study design can change.
 The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
@@
 While this truism is strongly related to study design, it is unrelated to preregistration.
-So $P(E)$ remains to consider.
-Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$.
-So the unlikelier it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high reward.
+Thus, $P(E)$ remains to be considered.
+Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+(REPHRASE SENTENCE)
+In other words, high risk, high reward. (REPHRASE SENTENCE)
 
@@
 A high risk, high reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
-Sustaining many "failed" studies is not exactly aligned with the incentive structure many researchers operate under.
-Therefore researchers have an incentive to appear to take more risks than they actually do, which misleads their colleges to give their claims more credence than they deserve.
-It is here where preregistrations make their impact.
-They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult.
+Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many if not most researchers operate.
+Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their peers to give their claims more credence than they deserve.
+It is at this juncture that the practice and mispractice of preregistration comes into play.
+They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult. (REPHRASE; THIS SENTENCE IS NON-SENSICAL)
 
@@
 
-Let us note some interesting properties of theoretical risk $P(¬H|¬E)$.
+Let us note some interesting properties of theoretical risk $P(¬H|¬E)$. (WHY? // THIS IS NOT WELL INTRODUCED)
 First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
@@
 
-This sounds like a truism but is directly related to Popper's famous criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false, therefore $P(E|¬H) = 1$ or conversely $P(¬E|¬H) = 0$, it can not be a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
-18].
+This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
+He stated that if it is impossible to prove a hypothesis false ($P(E|¬H) = 1$ or $P(¬E|¬H) = 0$), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
 We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
@@
 
-After we discussed that increasing the theoretical risk will increase the epistemic value, it is only natural to task preregistration with maximizing theoretical risk.
-In fact, limiting type I error rate is commonly stated as a goal of preregistration.
-We argue that while such a conclusion is plausible, we must first consider at least two constraints that put an upper bound on the theoretical risk.
+After we discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
+Indeed, limiting type I error rate is commonly stated as a goal of preregistration.
+We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
 
-First is the theory itself.
+First, the theory itself limits theoretical risk:
 Some theories simply do not make risky predictions, and preregistration will not change that.
-It would show a great deal of arrogance to advise researchers to stop investigating such theories.
+It would show a great deal of arrogance to advise researchers to stop investigating such theories. (META-DISCOURSE; UNNECESSARY LINE OF ARGUMENTATION?)
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set considered alone is well studied, and strong theories tell the researcher how the variables within the set relate.
-But our imaginary researcher considers the relation between these two sets.
+Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
+However, our imaginary researcher considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
 This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
-However, we would consider it a success if the researcher would use the evidence to develop a more precise (and hence risky) theory, e.g.
-by proposing which variables from the one set relate to which from the other set to what extent, direction, functional shape etc.
+However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
+by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
 We will later show that preregistration increases the expected epistemic value of the further specified theory, though it remains low till substantiated by testing it again.
 The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
-Changing what theories to test is not on our agenda.
+It is outside the scope of our agenda to changing or place demands on what theories should be tested through empirical research.
 
-The second upper bound to theoretical risk is resources.
+Second, available resources limit theoretical risk.
 Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
@@
 
-This objective also raises the problem of changing the plan after data collection started.
-One could argue that allowing deviating from the preregistration would allow researchers the same freedoms as having no preregistration at all.
-Hypothethically at least, researchers might try any possible deviation and pick the most favorably for their theory.
+This objective also raises the problem of changing the plan after data collection started. (WHY? THIS IS NOT WELL INTRODUCED)
+One could argue that allowing deviation from the preregistration would allow researchers the same freedoms as having no preregistration at all.
+Hypothethically at least, researchers might try any number of possible deviations and pick the one that is most favorable to their theory.
 Hence, allowing deviation would void any otherwise guaranteed lower limits to the theoretical risk.
-It follows that allowing deviations is wholly incompatible with the goal to maximize theoretical risk.
+It follows that allowing deviations is wholly incompatible with the goal of maximizing theoretical risk.
 
@@
 In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.
-Nevertheless, there are trait offs for increasing theoretical risk.
+Nevertheless, there are trade-offs for increasing theoretical risk.
 However, employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
-To that end, we incorporate uncertainty into our framework.
+To that end, we incorporate uncertainty into our framework. (WELL DONE)
 
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods. (UNNECESSARY META-DISCOURSE)
 If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be surer of their results, if significant.
+Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
@@
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These are all tools that help us factor in the tendency of statistical methods to show results even when there are none and are therefore closely related to theoretical risk.
+These are all tools that help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
+They are thus  closely related to theoretical risk.
 
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
+The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed. (META-DISCOURSE AVOID)
 For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
@@
 The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
-We can not hope to resolve this uncertainty; however, we may discuss its implications.
+We can not hope to resolve this uncertainty; but we have to be aware of its implications. These are relayed in the following.
 
@@
 Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
-However, researchers might do many very reasonable things in the broader context of the study that effect theoretical risk.
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaire, impute missing values, or any number of other things, that all carry a small risk that they obtain evidence despite their hypothesis being false.
+However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
+They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
 Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
 While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
-If researchers want to or not, except under strict blinding, they continuously observe information about how the study is going, which at best does not effect them in any way.
+Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
+While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
 We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
@@
 
-Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall in.
+Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall .
 Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
@@
 Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
-The last argument is that preregistrations reduce this uncertainty.
+The last argument is that preregistrations reduces this uncertainty.
 
@@
 We have to remember that the process of weighing evidence for or against a theory extends beyond the original authors to all the people they hope to persuade.
-While the authors may have a deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
-In particular, they never know what relevant factors the authors, intentionally or not, failed to lay open.
-From an ultimate sceptic perspective, they may claim maximum uncertainty.
+While the authors may have deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
+In particular, they never know what relevant factors the authors, intentionally or not, failed to disclose.
+From the perspective of an ultimate sceptic, they may claim maximum uncertainty.
 
 Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
-Preregistration are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
+Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
 The authors could have decided to exclude a number of analytic strategies they tried out etc.
-That is not to say any not preregistered study was p hacked.
-The point is, we can not exclude this and a myriad of other possibilities and, hence, are uncertain.
-If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete picture of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so.
+That is not to say that any study that was not preregistered was subjected to practices of p hacking.
+The point is, we can not exclude this and a myriad of other possibilities and, hence, are left uncertain.
+If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
+They still might be uncertain about the actual theoretical risk the authors took but much less so. (REITERATE)
 Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
@@
 No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as the name implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
+However, as the term implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
 As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
@@
 The advantage of this perspective that puts uncertainty about theorethical risk front and center is twofold.
-First, aiming to reduce uncertainty is beneficial across a wide range of theoretical risks.
-That is, researchers benefit from preregistrating their study, regardless weather or not the study is "exploratory".
+First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
+That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
@@
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
+However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount. (RESTATE)
 

### 2022-06-03 manuscript.Rmd
cite concerns with preregistration

@@
 
-One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration.
+One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration [@goldin-meadowWhyPreregistrationMakes2016].
 Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt these self-imposed restrictions.
@@
 To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
-This, of course, goes against researchers' common intuition about what is considered exploratory and confirmatory.
+This, of course, goes against most researchers' intuition about what is considered exploratory and confirmatory.
 We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
@@
 The first is that theoretical risk is vital for judging theories.
-This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist. (WHY; A BIT MORE MORE ELABORATION)
+This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 
-Let us start by defining what we mean by expected epistemic value. (METADISCOURSE UNNECESSARY)
+Let us start by defining what we mean by expected epistemic value.
 If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
@@
 
-The first assumption leads to our adoption of a Bayesian framework. (WHY)
+The first assumption leads to our adoption of a Bayesian framework.
 Our rationale is as follows.
@@
 
-Let us make some simplifying assumptions and define notation. (METADISCOURSE UNNECESSARY)
+Let us make some simplifying assumptions and define notation.
 We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
@@
 $P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (xxx), in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness. (SENTENCE STRUCTURE)
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
@@
 The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$. However, the a priori probability of a hypothesis is nothing our study design can change.
+The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$.
+However, the a priori probability of a hypothesis is nothing our study design can change.
 The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
@@
 Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-(REPHRASE SENTENCE)
-In other words, high risk, high reward. (REPHRASE SENTENCE)
+In other words, high risk, high reward.
 
@@
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many if not most researchers operate.
-Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their peers to give their claims more credence than they deserve.
+Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
 It is at this juncture that the practice and mispractice of preregistration comes into play.
-They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult. (REPHRASE; THIS SENTENCE IS NON-SENSICAL)
+They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult.
 
@@
 
-Let us note some interesting properties of theoretical risk $P(¬H|¬E)$. (WHY? // THIS IS NOT WELL INTRODUCED)
+Let us note some interesting properties of theoretical risk $P(¬H|¬E)$.
 First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
@@
 Some theories simply do not make risky predictions, and preregistration will not change that.
-It would show a great deal of arrogance to advise researchers to stop investigating such theories. (META-DISCOURSE; UNNECESSARY LINE OF ARGUMENTATION?)
+It would show a great deal of arrogance to advise researchers to stop investigating such theories.
 Consider the case of a researcher contemplating the relation between two sets of variables.
@@
 
-This objective also raises the problem of changing the plan after data collection started. (WHY? THIS IS NOT WELL INTRODUCED)
+This objective also raises the problem of changing the plan after data collection started.
 One could argue that allowing deviation from the preregistration would allow researchers the same freedoms as having no preregistration at all.
@@
 However, employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
-To that end, we incorporate uncertainty into our framework. (WELL DONE)
+To that end, we incorporate uncertainty into our framework.
 
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods. (UNNECESSARY META-DISCOURSE)
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
 If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
@@
 
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed. (META-DISCOURSE AVOID)
+The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
 For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
@@
 The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
-We can not hope to resolve this uncertainty; but we have to be aware of its implications. These are relayed in the following.
+We can not hope to resolve this uncertainty; but we have to be aware of its implications.
+These are relayed in the following.
 
@@
 
-Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall .
+Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall in.
 Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
@@
 If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so. (REITERATE)
+They still might be uncertain about the actual theoretical risk the authors took but much less so.
 Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
@@
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount. (RESTATE)
+However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
 

### 2022-06-03 manuscript.Rmd
adopt ylvas style suggestions

@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on this vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+Much has been theorized on the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
 To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
-Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of try and error [@wagenmakersAgendaPurelyConfirmatory2012].
+Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
 The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
@@
 
-Deciding if a study is confirmatory or exploratory is a judgment made by the researcher assessing it.
+Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
 The issues outlined above arise by delegating this judgment to a simple decision rule.
@@
 
-Let us start by defining what we mean by expected epistemic value.
+Let us start by defining what we call expected epistemic value.
 If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
@@
 2. They expect other researchers to apply the same rational process.
-2. All else equal, researchers try to increase the expected epistemic value for other researchers.
+2. All else equal, researchers try to maximize the expected epistemic value for other researchers.
 
@@
 Our rationale is as follows.
-Researchers that decide to conduct a study are akin to choosing a study to bet on.
+Researchers who decide to conduct a study are akin to choosing a study to bet on.
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.

### 2022-06-03 manuscript.Rmd
closes #11
closes #12

@@
 It is at this juncture that the practice and mispractice of preregistration comes into play.
-They try to ensure a proper judgment of the riskiness of a study by making tempering with the appearance of $P(E)$ difficult.
+We argue that the main function of preregistration is to ensure proper judgment of the riskiness ($P(E)$) a study.
 
@@
 Some theories simply do not make risky predictions, and preregistration will not change that.
-It would show a great deal of arrogance to advise researchers to stop investigating such theories.
 Consider the case of a researcher contemplating the relation between two sets of variables.
@@
 The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
-It is outside the scope of our agenda to changing or place demands on what theories should be tested through empirical research.
 

### 2022-06-05 manuscript.Rmd
correct typo
wording
correct theorethical risk

@@
 It is at this juncture that the practice and mispractice of preregistration comes into play.
-We argue that the main function of preregistration is to ensure proper judgment of the riskiness ($P(E)$) a study.
+We argue that the main function of preregistration is to enable proper judgment of the riskiness of a study.
 
@@
 
-Let us note some interesting properties of theoretical risk $P(¬H|¬E)$.
+Let us note some interesting properties of theoretical risk $P(¬E|¬H)$.
 First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).

### 2022-06-06 manuscript.Rmd
clarify terminology

@@
 by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the expected epistemic value of the further specified theory, though it remains low till substantiated by testing it again.
+We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
 The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
@@
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
@@
 The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
-For many statistical models, their $P(E|¬H)$ can be analytically computed under some assumptions.
+For many statistical models it can be analytically computed under some assumptions.
 For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.

### 2022-06-06 manuscript.Rmd
clarify terminology
add a more motivating introduction

@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-Much has been theorized on the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+Many researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
-The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
-We believe that a rigorous answer to this question enables us to clarify common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
-These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
-Asking researchers to invest time and resources into preregistration demands a principled justification of its utility.
+However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
+Those without this intuition ask the fundamental question of why should a study be preregistered at all?
+Nevertheless, even researchers convinced of its advantages have some problems with an intuitive approach.
+Before data collection, they must make difficult decisions about what to include in the preregistration and in how much detail.
+After data collection, they might be uncertain what conditions warrant deviating from the preregistration.
+If the research community remains confused by the purpose of preregistrations, it decreases the efficiency of the scientific endeavour. 
+We propose a principled justification for preregistration with a formal objective against which researcher can evaluate their trade-offs.
 
-One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration [@goldin-meadowWhyPreregistrationMakes2016].
-Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt these self-imposed restrictions.
+Preregistrations trade-off researchers' degrees of freedom (XXX) to obtain more confirmatory results.
+However, balancing this trade-off is difficult.
+Being too restricted by the boundaries set in the preregistration make some researchers hesitant to adopt the practice [@goldin-meadowWhyPreregistrationMakes2016].
+Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt the self-imposed restrictions.
 One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
-To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
+Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
+The conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
 This, of course, goes against most researchers' intuition about what is considered exploratory and confirmatory.
@@
 
-We established that higher theoretical risk leads to more credible evidence.
+We established that higher theoretical risk leads to more persuasive evidence.
 In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.

### 2022-06-06 manuscript.Rmd
remove early mention of deviation closes #19

--- 2022-06-06 manuscript.Rmd
@@
 
-This objective also raises the problem of changing the plan after data collection started.
-One could argue that allowing deviation from the preregistration would allow researchers the same freedoms as having no preregistration at all.
-Hypothethically at least, researchers might try any number of possible deviations and pick the one that is most favorable to their theory.
-Hence, allowing deviation would void any otherwise guaranteed lower limits to the theoretical risk.
-It follows that allowing deviations is wholly incompatible with the goal of maximizing theoretical risk.
-
 # Uncertainty about Theoretical Risk
 
-We established that higher theoretical risk leads to more credible evidence.
+We established that higher theoretical risk leads to more persuasive evidence.
 In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.

### 2022-06-07 manuscript.Rmd
summarise why this tradeoff is difficult #10
move quote to old material
sharpen the criticism of preregistration

--- 2022-06-06 manuscript.Rmd
@@
 However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
-Those without this intuition ask the fundamental question of why should a study be preregistered at all?
-Nevertheless, even researchers convinced of its advantages have some problems with an intuitive approach.
-Before data collection, they must make difficult decisions about what to include in the preregistration and in how much detail.
-After data collection, they might be uncertain what conditions warrant deviating from the preregistration.
-If the research community remains confused by the purpose of preregistrations, it decreases the efficiency of the scientific endeavour. 
+Skeptics question why a study should be preregistered at all.
+Even researchers convinced of the advantages have some problems when left with nothing more than an intuitive appeal.
+Their intuition must answer difficult questions, such as what to include, in how much detail, in the preregistration, or what conditions warrant deviating from the preregistration.
+If the research community remains confused about the purpose of preregistrations, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
 We propose a principled justification for preregistration with a formal objective against which researcher can evaluate their trade-offs.
 
-Preregistrations trade-off researchers' degrees of freedom (XXX) to obtain more confirmatory results.
-However, balancing this trade-off is difficult.
-Being too restricted by the boundaries set in the preregistration make some researchers hesitant to adopt the practice [@goldin-meadowWhyPreregistrationMakes2016].
-Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt the self-imposed restrictions.
+The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018, @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
+Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
+However, this postulate is indefensible.
+Researchers can conduct confirmatory research without preregisteration.
+Wether they can convince other researchers of the confirmatory nature of their research is a seperate issue.
+Preregistered but not strictly confirmatory studies are also commonplace.
+Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
 One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-The conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
-This, of course, goes against most researchers' intuition about what is considered exploratory and confirmatory.
-We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
-In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
+Insisting on equating preregistration and confirmation has led to criticism that preregistration is actually negative sum, and neither sufficent nor neccesary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
-Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
-The issues outlined above arise by delegating this judgment to a simple decision rule.
-That is, if something is preregistered, it is considered confirmatory, and if it is not preregistered, it is considered exploratory.
-This rule of thumb is often applied implicitly.
-Other times it is stated as the explicit goal of preregistration [@mellorEasyPreregistrationWill2018].
-In both cases, it leads to confusion when applied outside specific bounds.
+These issues arise from a fundamental confusion about the objective of preregistration.
+Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
 
-We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
+We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
@@
 The first is that theoretical risk is vital for judging theories.
-This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
 The second argument is that the theoretical risk for a given study is generally uncertain.
 The third and last assumption is that this uncertainty is reduced by applying preregistration.
-
-When we put these arguments into the Bayes' formula, we conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
-We will later discuss how our argument relates to a (Neo‑)Popperian philosophy of science.
-For now, we stick to a Bayesian rationale since quantifying uncertainty is crucial for our argument, which would require considerable thought in a falsificationist framework.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
 

### 2022-06-07 manuscript.Rmd
stylistic changes

@@
 However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
-Skeptics question why a study should be preregistered at all.
-Even researchers convinced of the advantages have some problems when left with nothing more than an intuitive appeal.
-Their intuition must answer difficult questions, such as what to include, in how much detail, in the preregistration, or what conditions warrant deviating from the preregistration.
-If the research community remains confused about the purpose of preregistrations, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
-We propose a principled justification for preregistration with a formal objective against which researcher can evaluate their trade-offs.
+These skeptics question why a study should be preregistered at all.
+Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
+Specifially, an intuitive convinction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.
+If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
+We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
 
-The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018, @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
+The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
 Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate is indefensible.
+However, this postulate not warrented.
 Researchers can conduct confirmatory research without preregisteration.
-Wether they can convince other researchers of the confirmatory nature of their research is a seperate issue.
+Weather they can convince other researchers of the confirmatory nature of their research is a separate issue.
 Preregistered but not strictly confirmatory studies are also commonplace.
@@
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has led to criticism that preregistration is actually negative sum, and neither sufficent nor neccesary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually negative sum, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
@@
 The second argument is that the theoretical risk for a given study is generally uncertain.
-The third and last assumption is that this uncertainty is reduced by applying preregistration.
+The third and last argument is that this uncertainty is reduced by applying preregistration.
 We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
@@
 This expectation is what we term expected epistemic value.
-We assume three things about this estimation process and how it relates to choosing a study to conduct.
+To make our three arguments, we must assume three things about this estimation process and how it relates to choosing a study to conduct.
 
@@
 
-The first assumption leads to our adoption of a Bayesian framework.
+The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
 Our rationale is as follows.

### 2022-06-07 manuscript.Rmd
cite replication crisis

@@
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-Many researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+Confusing exploratory findings with confirmed theories has lead to a crisis of confidence of in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
@@
 Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate not warrented.
+However, this postulate is not warrented.
 Researchers can conduct confirmatory research without preregisteration.
-Weather they can convince other researchers of the confirmatory nature of their research is a separate issue.
+Convincing other researchers of the confirmatory nature of their research is a separate issue.
 Preregistered but not strictly confirmatory studies are also commonplace.

### 2022-06-07 manuscript.Rmd

@@
 Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
-
-This objective also raises the problem of changing the plan after data collection started.
-One could argue that allowing deviation from the preregistration would allow researchers the same freedoms as having no preregistration at all.
-Hypothethically at least, researchers might try any number of possible deviations and pick the one that is most favorable to their theory.
-Hence, allowing deviation would void any otherwise guaranteed lower limits to the theoretical risk.
-It follows that allowing deviations is wholly incompatible with the goal of maximizing theoretical risk.
+Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
 
@@
 
+# Old Material
+
 > As far as I am aware, Mayo’s severity argument currently provides one of the few philosophies of science that allows for a coherent conceptual analysis of the value of preregistration.
-
-# Old Material
+> Borsboom
 

### 2022-06-07 manuscript.Rmd
minor changes

@@
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-Confusing exploratory findings with confirmed theories has lead to a crisis of confidence of in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
@@
 However, this postulate is not warrented.
-Researchers can conduct confirmatory research without preregisteration.
-Convincing other researchers of the confirmatory nature of their research is a separate issue.
-Preregistered but not strictly confirmatory studies are also commonplace.
+Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
+The converse, therefore preregistered but not strictly confirmatory studies are also commonplace.
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
@@
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually negative sum, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmfull all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 

### 2022-06-08 manuscript.Rmd
Update manuscript.Rmd

@@
 
-A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
+A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
@@
 Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifially, an intuitive convinction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.
+Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.<!-- Message unclear, if meant like I think: "...such as what to include in the preregistraion and how detailed this should be"-->
 If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
@@
 Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate is not warrented.
+However, this postulate is not warranted.
 Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The converse, therefore preregistered but not strictly confirmatory studies are also commonplace.
+The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace.
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
-One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
+One strategy is to write a loose preregistration to begin with; another is to deviate from the preregistration afterwards.
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
@@
 Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
+Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
 
@@
 2. They expect other researchers to apply the same rational process.
-2. All else equal, researchers try to maximize the expected epistemic value for other researchers.
+2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.
 
@@
 Nevertheless, there are trade-offs for increasing theoretical risk.
-However, employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
+Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
 To that end, we incorporate uncertainty into our framework.
@@
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These are all tools that help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
 They are thus closely related to theoretical risk.
@@
 
-We started out with the observation that preregistrations do no always cleanly divide preregistration from conformation.
+We started out with the observation that preregistrations do no always cleanly divide preregistration from confirmation.
 Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
@@
 
-The advantage of this perspective that puts uncertainty about theorethical risk front and center is twofold.
+The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
 First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
@@
 Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
-Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and than select the most beneficial for their theory.
+Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
 However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
+<!-- Last sentence is a bit confusing -->
 

### 2022-06-08 manuscript.Rmd

@@
 Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.<!-- Message unclear, if meant like I think: "...such as what to include in the preregistraion and how detailed this should be"-->
+Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistraion and how detailed this should be, or about the conditions that warrant deviating from it.
 If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
@@
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
-<!-- Last sentence is a bit confusing -->
+However, if the authors offer a convincing argument for their deviation, their readers might believe it to be a likely explanation for the deviation.
+If the readers find the reason for the deviation likely to be something other than decreasing the theorethical risk, their uncertainty about the theorethical risk should only increase modestly.
 

### 2022-06-09 manuscript.Rmd
use apa citation style

--- 2022-06-07 manuscript.Rmd
@@
 abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
+csl: apa.csl
 ---
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+Much has been theorized on the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
-However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
-These skeptics question why a study should be preregistered at all.
-Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifially, an intuitive convinction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.
-If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
-We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
+The scientific community widely agrees that actively concealing such "trying for results" is misconduct --- but how exactly is it impeding scientific progress?
+We believe that a rigorous answer to this question enables us to clarify common misunderstandings that lead researchers to deem preregistration as unfit for their purposes.
+These misunderstandings are interesting from a conceptual perspective but also lead to decreased efficiency of the scientific process if researchers misguidedly decide that preregistration is not applicable for their intended study.
+Asking researchers to invest time and resources into preregistration demands a principled justification of its utility.
 
-The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate is not warrented.
-Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The converse, therefore preregistered but not strictly confirmatory studies are also commonplace.
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
+One reason why researchers are hesitant to preregister is that they feel restricted by the boundaries specified in their preregistration [@goldin-meadowWhyPreregistrationMakes2016].
+Even researchers who have been persuaded of its advantages sometimes apply strategies to skirt these self-imposed restrictions.
 One strategy is to write a loose preregistration, to begin with; another is to deviate from the preregistration afterwards.
-Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmfull all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+To complicate matters further, both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
+To summarize, the conundrum is that the current conceptualization of preregistration deems a study "properly" preregistered even if it contains a recipe for p-hacking before data collection but "improper" if the researcher did technically deviate afterwards.
+This, of course, goes against most researchers' intuition about what is considered exploratory and confirmatory.
+We take these issues as signals of a misunderstanding about how preregistration contributes to scientific progress.
+In particular, we want to challenge the common view that preregistration somehow cleanly separates confirmatory from exploratory.
 
-These issues arise from a fundamental confusion about the objective of preregistration.
-Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
+Deciding if a study is confirmatory or exploratory is a judgment made by the researcher.
+The issues outlined above arise by delegating this judgment to a simple decision rule.
+That is, if something is preregistered, it is considered confirmatory, and if it is not preregistered, it is considered exploratory.
+This rule of thumb is often applied implicitly.
+Other times it is stated as the explicit goal of preregistration [@mellorEasyPreregistrationWill2018].
+In both cases, it leads to confusion when applied outside specific bounds.
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+To uphold the simplicity of the decision rule, it is routinely suggested only to preregister when a study is expected to have a low type I error rate by design and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
 
-We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
+We show that this simple decision rule is a special case of a more general conceptualization under Bayesian reasoning.
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
@@
 The first is that theoretical risk is vital for judging theories.
+This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
 The second argument is that the theoretical risk for a given study is generally uncertain.
-The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
+The third and last assumption is that this uncertainty is reduced by applying preregistration.
+
+When we put these arguments into the Bayes' formula, we conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
+We will later discuss how our argument relates to a (Neo‑)Popperian philosophy of science.
+For now, we stick to a Bayesian rationale since quantifying uncertainty is crucial for our argument, which would require considerable thought in a falsificationist framework.
 
@@
 This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to choosing a study to conduct.
+We assume three things about this estimation process and how it relates to choosing a study to conduct.
 
@@
 
-The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
+The first assumption leads to our adoption of a Bayesian framework.
 Our rationale is as follows.

### 2022-06-09 manuscript.Rmd
first set of comments

@@
 abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
-csl: apa.csl
 ---
@@
 
+<!--AB: this is a pretty steep start - what is the context of prediction here - do you think that all researchers subscribe to the importance of prediction in the first place? What is "successful"?-->
 A successful prediction should lend more credibility to a theory than a postdiction, all else equal.
 Much has been theorized on the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+<!--AB add ref to Shmueli, 2010-->
 To honour this distinction and produce results that are considered confirmatory, many researchers opt to preregister their studies [@nosekPreregistrationRevolution2018].
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
+<!--AB how exactly is it borrowed?
+Did Meehl use it differently?
+If so, what is different?-->
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 To arrive at this interpretation, we rely on three arguments.
-The first is that theoretical risk is vital for judging theories.
+The first is that theoretical risk is vital for judging theories. <!--AB judging what? the verisimilitude/trueness/plausibility?-->
 This argument comes natural to a (Neo‑)Popperian or falsificationist, but  is not self-evident for a conformationist.
+<!--AB: do all readers know what these philosophical stances mean?
+Depends on the target audience and journal you aim for-->
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 
-Our first argument is that theorethical risk is crucial for judging theories.
+Our first argument is that theorethical risk is crucial for judging theories.<!-- again: judging what?-->
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
@@
 We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+<!--AB: What about measurement error? Is evidence thought to be measurement-free her? -->
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+<!--AB: Not all readers may understand what the probability of a hypothesis is. Elaborate more on the probability interpretation of Bayesians as credibility.-->
 The probability of observing evidence under some hypothesis is $P(E|H)$.
@@
 So how can we increase $P(H|E)$?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
 The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$.
@@
 We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
+<!--AB: from the perspective of a rational agent/researcher, of course, it makes sense to make P(E|H) large but isn't psychology facing a widespread problem of underpowered studies?
+I would make this very clear that this essentially relates to power (right?) and that historically we haven't been doing well on that account but that this problem is not pertinent to preregistration as such-->
 The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
+This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
@@
 
+<!--AB can we add an example that explains this whole idea once more but with a concrete example?
+I think that the probability of evidence given NOT H is just difficult to wrap one's head around.
+It would be ideal to illustrate this with an example that presents both a weak and a strong P(E given not H).
+
+I have a hard time coming up with an easy example that relates this simplified case
+to a problem of psychological research.
+
+H: a depression treatment is effective
+not H: the treatment is not effective
+
+In a weak risk case, the prediction is fuzzy, P(E | not H) is large, theoretical risk is small:
+
+E: people have some depression score (other than zero) after the treatment
+not E: people have exactly zero depression scores
+
+In a strong risk case, the prediction is precise, P(E | not H) is small, theoretical risk is large
+E: people improve at least 10 points on the depression scale
+not E: people improve less or equal to 10 points on the depression scale
+-->
+
 Let us note some interesting properties of theoretical risk $P(¬E|¬H)$.
-First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
+<!--AB: I rephrased this sentence. Is credibility equal to epistemic value?-->
+First, if the objective is to increase the credibility of a theory $P(H|E)$, this can
+be achieved by increasing theoretical risk.
 Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
+<!--AB: explain more what this means. When is this condition true in reality?-->
 Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
@@
 
+<!--Is this the word "this" refering to the third point only or all three?-->
 This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
@@
 After we discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
-Indeed, limiting type I error rate is commonly stated as a goal of preregistration.
+Indeed, limiting the false positive (or, type I error rate) is commonly stated as a goal of preregistration. <!--AB: citation needed; maybe verbose quote?-->
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
 
+<!--AB: the kind of more concrete example that you give here is needed much earlier when you define the different conditional probabilities; maybe even in the very beginning, it could be helpful to give examples of risky vs non-risky predictions in psychology from published literature?!-->
 First, the theory itself limits theoretical risk:
@@
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
-However, our imaginary researcher considers the relation between these two sets.
+Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within each of the sets relate to each other.
+Now, our imaginary researcher considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation due to multiple testing [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
 by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
+We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low until substantiated by testing it again.
+<!--AB: "till" is never used in formal writing-->
+The point, however, is that we want to show that preregistration increases the expected epistemic value independent of the theory being tested.
 
@@
 Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
+<!-- I recommend giving specific names to each of the conditional probabilities and then refer to those names and have the math notation in parentheses; then, this becomes easier to read and understand, e.g.,
+
+Increasing theoretical risk ($P(¬E|¬H)$) will usually decrease the probability of observing the evidence under the hypothesis ($P(E|H)$).
+
+Is there a term, we can introduce for $P(E|H)$, like power?
+
+-->
 In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
@@
 Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
+In other words, the probability that a given significant finding reflects a true effect is larger if researchers lower their type I error rates.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
 We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+<!--AB: can you elaborate more?!-->
 
@@
 If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
+Take, for example, a t-test of mean differences.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its promised type I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled, translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
 While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.

### 2022-06-17 manuscript.Rmd
add sources for preregistered but exploratory studies closes #37

--- 2022-06-08 manuscript.Rmd
@@
 Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace.
+The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
-One strategy is to write a loose preregistration to begin with; another is to deviate from the preregistration afterwards.
+One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.

### 2022-06-17 manuscript.Rmd
most probabilities are referered to by name closes #35

@@
 
-$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
+The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+However, it is generally expected that any measure of corroboration (and even increases monotonically with an increase in posterior probability ($P(H|E)$).
 
-In short, we want to increase $P(H|E)$.
-Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
-So how can we increase $P(H|E)$?
+In short, we want to increase posterior probability ($P(H|E)$).
+Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, of which we want to maximize the expectation.
+So how can we increase posterior probability?
 The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$.
-However, the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
+The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
+However, the prior probability of a hypothesis is nothing our study design can change.
+The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
+$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
+In other words, how probable is it that we "detect" that a theory holds, we therefore call it "detectability".
 Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-While this truism is strongly related to study design, it is unrelated to preregistration.
+When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type II error rate).
+However, while detectability is of great importance for study design, it is not directly relevant for the objective of preregistration.
 Thus, $P(E)$ remains to be considered.
-Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
 In other words, high risk, high reward.
@@
 
-We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
+We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
 The real lever to pull is, therefore, $P(E|¬H)$.
@@
 
-Let us note some interesting properties of theoretical risk $P(¬E|¬H)$.
-First, increasing theoretical risk leads to higher $P(H|E)$ (our objective).
-Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
-Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
+Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
+First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
+Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
+If detectability exceeds theorethical risk, the evidence is less likely under the theory than it is when the theory does not hold.
+Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
 In other words, observing a sure fact does not lend credence to a hypothesis.
@@
 This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false ($P(E|¬H) = 1$ or $P(¬E|¬H) = 0$), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
+He stated that if it is impossible to prove a hypothesis false ($P(¬E|¬H) = 0$, theoretical risk is zero), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
+18].
 We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
@@
 Second, available resources limit theoretical risk.
-Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
+Increasing theorethical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
 In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ (the inverse of theorethical risk) and statistical methods.
 If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
@@
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+We argue that $P(E|¬H)$ (and hence theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 

### 2022-06-17 manuscript.Rmd
add a concrete example #34

@@
 
-```{r, include=FALSE}
-bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
-```
-
 Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
@@
 We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+
+Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+To make these concepts less abstract, consider the following illustrative example of a theory and different experiments that test it.
+Suppose the theory is about the cognitive phenomenon of "insight", and allow us to define it somewhat hand-waving as an cognitive abstraction that allows agents to consistently solve a well defined class of problems.
+We pose the hypothesis that the following problem belongs to such class of insight problems:
+
+> Use five matches (Ⅰ Ⅰ Ⅰ Ⅰ Ⅰ) to form the number eight.
+
+We propose three experiments that differ in theoretical risk and detectability.
+All take a sample of ten psychology students and present them the problem for brief span of time.
+After that:
+
+1. the experimenter explains that with roman-numerals the problem is easy to solve; if all students come up with the solution, she records it as evidence for the hypothesis.
+2. the experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
+3. the experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
+
+We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
+If "insight" has nothing to do with solving the problem ($¬H$), than presenting the insight that roman literals might be used, should not lead to all students solving the problem ($¬E_1$); the experiment has therefore high theoretical risk ($P(¬E_1|¬H)$)
+Conversly, if insight is required to solve the problem ($H$), than it is probable to help all students to solve the problem ($E_1$); the experiment has therefore high detectability ($P(E_1|H)$).
+The experiment 2 on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
+Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students can simply copy the solution, without any insight.
+With regard to detectability experiment 1 and 2 differ in no obvious way.
+Experiment 3, however, has low detectability.
+It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); the experiment 3 has therefore low detectability ($P(E_3|H)$).
+The theoretical risk, however, is also low in absolute terms but high compared to the theoretical risk.
+In the unlikely event that all 10 students lay the matches down in the form of the roman numeral Ⅷ ($E_3$) it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
+Of course, it comes to mind to allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
+We discuss the relation between statistical methods and theoretical risk in the section [Statistical Methods].
+
+```{r, include=FALSE}
+bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
+```
 

### 2022-06-19 manuscript.Rmd
cite Shmueli 2010 closes #30

--- 2022-06-17 manuscript.Rmd
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
 Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].

### 2022-06-20 manuscript.Rmd
cite meehls paper more widely
closes #31

@@
 We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
+In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
+However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ is theoretical risk.
 

### 2022-06-24 manuscript.Rmd

@@
 As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
+
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 To arrive at this interpretation, we rely on three arguments.
+
 The first is that theoretical risk is vital for judging theories.
+
 The second argument is that the theoretical risk for a given study is generally uncertain.

### 2022-06-24 manuscript.Rmd
judge what? closes #32

--- 2022-06-20 manuscript.Rmd
@@
 To arrive at this interpretation, we rely on three arguments.
-The first is that theoretical risk is vital for judging theories.
+The first is that theoretical risk is vital for judging evidential support for theories.
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
+This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
@@
 by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
+We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low until substantiated by testing it again.
+The point, however, is that we want to show that preregistration increases the expected epistemic value independent of the theory being tested.
 
@@
 Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
+In other words, the probability that a given significant finding reflects a true effect is larger if researchers lower their type I error rates.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.

### 2022-06-24 manuscript.Rmd
give all probabilities a name

--- 2022-06-20 manuscript.Rmd
@@
 Nevertheless, there are trade-offs for increasing theoretical risk.
-Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
+Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favorable interpretation.
 To that end, we incorporate uncertainty into our framework.
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
-If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+Theoretical risk is deeply connected with statistical methods, because its the inverse of $P(E|¬H)$.
+$P(E|¬H)$ is equivalent to the type I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill-hypothesis is $¬H$.
+Because many researchers are familiar with type I error rate, it can be helpful to remember this connection to theorethical risk.
+Researchers who choose a smaller type I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
+However, the research hypothesis is seldomly equal to the statistical null hypothesis, and therefore, the relation between statistical type I error rate and theoretical risk should not be over interpreted.
+We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
+Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to theoretical risk.
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
-They are thus closely related to theoretical risk.
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
 
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
+The point is that the contribution of statistical methods to theoretical risk can be formally assessed.
 For many statistical models it can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
-The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to theoretical risk.
+The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
 We can not hope to resolve this uncertainty; but we have to be aware of its implications.
@@
 However, other factors in the broader context of the study are much harder to quantify.
-If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
+If one chooses to focus only on the contribution of statistical methods to theorethical risk, one is bound to overestimate it.
 Take, for example, a t-test.
@@
 All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
-While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
+Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$ and, hence, be subtracted from theoretical risk.
+While, in theory, these factors may leave theoretical risk unaffected or even increase it, we argue that this is not the case in practice.
 Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
 While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
-The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
+The only thing we can conclude with some certainty is that theorethical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
 

### 2022-06-24 manuscript.Rmd
graphics show theorethical risk instead of 𝑃(𝐸|¬𝐻) closes #22

@@
 
-Our first argument is that theorethical risk is crucial for judging theories.
+Our first argument is that theorethical risk is crucial for judging evidential support for theories.
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
@@
 f(x)=\begin{cases}
-  \frac{1}{1 - \alpha} & \mathrm{for}\ \alpha \le x \le 1,\\
-  0 & \mathrm{for}\ x<\alpha\ \mathrm{or}\ x>1
+  \frac{1}{\tau} & \mathrm{for}\ 0 \le x \le \tau,\\
+  0 & \mathrm{for}\ x<0\ \mathrm{or}\ x>\tau
   \end{cases}
@@
 
-Where $\alpha$ is, e.g., the statistical model's nominal type I error rate.
+Where $\tau{}$ is the upper bound of theoretical risk (e.g., .95 for a statistical model with a nominal type I error rate of 5%).
 
-We chose this uniform distribution to capture our statement that a statistical model only guarantees a lower bound to $p(E|\neg{}H)$ (and since its a probability the upper bound is 1) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theorethical rsik (and since its a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
-```{r, include=FALSE, cache=TRUE}
-bayes <- function(h1, power, alpha, ...) {
-  (h1 * power) / (h1 * power + (1 - h1) * alpha)
-}
-uncertain_bayes <- function(h1, power, alpha, alpha_upper = 1, ...) {
-  mean(bayes(h1, power, runif(1e6, alpha, alpha_upper)))
+```{r, include=FALSE}
+forbid_additional_args <- function(...)stopifnot(length(list(...)) == 0L)
+
+#' Apply the Bayes Theorem
+#'
+#' @param h The prior probability of the hypothesis \eqn{P(H)}.
+#' @param detectability The probability of detecting evidence, if hypothesis holds \eqn{P(E|H)}.
+#' @param theoretical_risk The probability of not detecting evidence, if hypothesis does not hold \eqn{P(¬E|¬H)}.
+#' @param tau Same as `theoretical_risk` for compatibility.
+#' @return The posterior probability of the hypothesis after observing evidence \eqn{P(H|E)}.
+bayes <- function(h, detectability, theoretical_risk, ...) {
+  forbid_additional_args(...)
+  true_positive <- h * detectability
+  false_positive <- (1 - h) * (1 - theoretical_risk)
+  true_positive / (true_positive + false_positive)
 }
 
-step <- 0.01
-vary_alpha <-
-  tidyr::expand_grid(
-    h1 = .1,
-    power = .8,
-    alpha = seq(step, 1 - step, step)
-  ) %>% 
-  dplyr::mutate(.,
-    certain = purrr::pmap_dbl(., bayes),
-    uncertain = purrr::pmap_dbl(., uncertain_bayes)
-  )
-
-vary_alpha_long <- vary_alpha %>% 
-  pivot_longer(c(certain, uncertain),
-               names_to = "certainty",
-               values_to = "h2")
-```
-
-```{r, dev='tikz'}
-#| fig.cap: "Corroboration as firmness ($p(E|H)$) as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
-vary_alpha_long %>%
-  ggplot() +
-  geom_line(aes(alpha, h2, linetype = certainty)) +
-  lims(x = c(0, 1), y = 0:1) +
-  geom_ribbon(data = vary_alpha, aes(x = alpha, ymin = uncertain, ymax = certain), alpha = .1) +
-  labs(caption = "$p(E|H) = .8$, $p(H) = .1$",
-       y = "$\\mathbb{E}\\big(p(H|E)\\big)$",
-       x = "$\\alpha$") +
-  theme_tufte() +
-  theme(legend.position = c(.85, .8),
-        legend.title = element_blank())
-```
-
-```{r, include=FALSE}
-# h1 = prior
-# eh = p(E|H)
-# enh = p(E|\neg{}H)
-
-bayes <- function(h1, eh, enh, ...) {
-  (h1 * eh) / (h1 * eh + (1 - h1) * enh)
-}
 # all formulas taken from Bayesian Philosophy of Science, p. 51
@@
 measures <- list(
+  pm = list(
+    name = "Posterior",
+    formula = "\\frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}",
+    fun = bayes
+  ),
   dm = list(
@@
     formula = "d(H, E) = p(H|E) - p(H)",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
-      he - h1
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      he - h
     }
@@
     formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
-      log(he / h1)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      log(he / h)
     }
@@
     formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      enh <- 1 - theoretical_risk
       log(he / enh)
@@
     formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
-    fun = function(h1, eh, enh, ...) {
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      eh <- detectability
+      enh <- 1 - theoretical_risk
       (eh - enh) / (eh + enh)
@@
 \\end{cases}",
-fun = function(h1, eh, enh, ...) {
-  he <- bayes(h1, eh, enh)
-  ifelse(he >= h1, (he - h1)/(1 - h1), (he - h1)/h1)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      ifelse(he >= h, (he - h) / (1 - h), (he - h) / h)
     }
@@
     formula = "s(H, E) = p(H|E) - p(H|\\neg{}E)",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    e <- h1 * eh + (1 - h1) * enh
-    hne <- (h1 * (1 - eh)) / (1 - e)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      enh <- 1 - theoretical_risk
+      eh <- detectability
+      e <- h * eh + (1 - h) * enh
+      hne <- (h * (1 - eh)) / (1 - e)
       he - hne
@@
     formula = "c'(H, E) = p(E)(p(H|E) - p(H))",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    e <- h1 * eh + (1 - h1) * enh
-    e * (he - h1)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      enh <- 1 - theoretical_risk
+      e <- h * detectability + (1 - h) * enh
+      e * (he - h)
     }
@@
     formula = "r'(H, E) = 1 - \\frac{p(\\neg{}H|E)}{p(\\neg{}H)}",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    1 - (1 - he) / (1 - h1)
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      1 - (1 - he) / (1 - h)
     }
@@
 
-uncertainly <- function(fun, n = 1e5, seed = NULL){
-  function(h1, eh, enh){
+#' Integrate over uncertain theoretical risk
+#' 
+#' If theorethical risk is maximally uncertain, we represent this uncertainty
+#' as a uniform distribution from zero to tau from which we sample and average the measures.
+#'
+#' @param tau The guaranteed upper bound of theoretical risk.
+#' @param n Number of samples from the uniform distribution.
+#' @param seed Optional seed to set before execution. Is reset to global state on exit.
+#' @return A function similar to the original
+uncertainly <- function(measure, tau, n = 1e5, seed = NULL){
+  function(h, detectability, ...){
+    forbid_additional_args(...)
     if(!is.null(seed)){
@@
     }
-    enh_samples <- runif(n, enh, 1)
-    mean(fun(h1, eh, enh_samples))
+    theoretical_risk_samples <- runif(n, 0, tau)
+    mean(measure(h, detectability, theoretical_risk_samples))
+  }
+}
+certainly <- function(measure, tau){
+  function(h, detectability, ...){
+    forbid_additional_args(...)
+    measure(h, detectability, tau)
   }
@@
 
-grid <- tidyr::expand_grid(
-    h1 = .1,
-    eh = .8,
-    enh = seq(step, 1 - step, step)
+grid <-
+  tidyr::expand_grid(
+    h = .1,
+    detectability = .8,
+    tau = seq(step, 1 - step, step)
   )
-apply_on_grid <- function(f, grid) {
+apply_on_grid <- function(f, grid, seed = NULL, n = 1e4) {
+  certain <- function(h, detectability, tau, ...){
+    forbid_additional_args(...)
+    certainly(f, tau)(h, detectability)
+  }
+  uncertain <- function(h, detectability, tau, ...){
+    forbid_additional_args(...)
+    uncertainly(f, tau, seed = seed, n = n)(h, detectability)
+  }
   mutate(grid,
-         certain = pmap_dbl(grid, f),
-         uncertain = pmap_dbl(grid, uncertainly(f, seed = .Random.seed)))
+         certain = pmap_dbl(grid, certain),
+         uncertain = pmap_dbl(grid, uncertain))
 }
@@
 
-```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Several measures for corroboration as increase in firmness as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
-
-results %>%
+```{r, cache = TRUE}
+plots <- results %>%
   mutate(name = str_remove(name, "Measure")) %>%
@@
   imap(
-    ~ ggplot(.x, aes(enh, value)) +
-      geom_line(aes(linetype = certainty)) +
-      labs(
-        title = .x$name,
+    ~ ggplot() +
+      geom_line(aes(tau, value, linetype = certainty), data = .x) +
+      labs(title = .y,
            y = element_blank(),
-        x = element_blank()
+           x = "$\\tau$") +
+      geom_ribbon(
+        data = pivot_wider(.x, names_from = certainty, values_from = value),
+        aes(x = tau, ymin = uncertain, ymax = certain),
+        alpha = .1
+      ) +
+      theme_tufte() +
+      NULL
   )
-  ) %>%
+```
+
+```{r, dev = "tikz", cache=TRUE, warning=FALSE}
+#| fig.cap: "Posterior (corroboration as firmness) as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
+
+plots[["Posterior"]] +
+  labs(caption = "$p(E|H) = .8$, $p(H) = .1$",
+       y = "$\\mathbb{E}\\big(p(H|E)\\big)$") +
+  theme(legend.position = c(.2, .8),
+        legend.title = element_blank())
+```
+
+```{r, dev = "tikz", cache=TRUE, warning=FALSE}
+#| fig.cap: "Several measures for corroboration as increase in firmness as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
+
+plots[!(names(plots) %in% "Posterior")] %>% 
   reduce(`+`) +

### 2022-06-24 manuscript.Rmd

--- 2022-06-24 manuscript.Rmd
@@
 Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
+Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychological research or elsewhere.
 
@@
 To arrive at this interpretation, we rely on three arguments.
-The first is that theoretical risk is vital for judging evidential support for theories.
+The first is that theoretical risk is vital for judging theories.
 The second argument is that the theoretical risk for a given study is generally uncertain.
@@
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
+The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory (which itself is an ontologocial concept) is what we call epistemic value.
+Researchers cannot know what they will learn from a study before they have run it.
+However, they can develop an expectation that helps them to decide which study to conduct <!--AB in the sense of maximum information gain?
+What does "which study" mean?
+I guess you mean just any possible property of an study in the planning phase that can be altered?-->.
 This expectation is what we term expected epistemic value.
@@
 Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+In fact, this conceptualization is intentionally reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 
@@
 We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
+<!--AB: from the perspective of a rational agent/researcher, of course, it makes sense to make P(E|H) large but isn't psychology facing a widespread problem of underpowered studies?
+I would make this very clear that this essentially relates to power (right?) and that historically we haven't been doing well on that account but that this problem is not pertinent to preregistration as such-->
 The real lever to pull is, therefore, $P(E|¬H)$.
@@
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
-However, our imaginary researcher considers the relation between these two sets.
+Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within each of the sets relate to each other.
+Now, our imaginary researcher considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation due to multiple testing [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
@@
 We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low until substantiated by testing it again.
+<!--AB: "till" is never used in formal writing-->
 The point, however, is that we want to show that preregistration increases the expected epistemic value independent of the theory being tested.
@@
 We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+<!--AB: can you elaborate more?!-->
 

### 2022-06-24 manuscript.Rmd
reintroduce some of andreas changes

--- 2022-06-24 manuscript.Rmd
@@
 Nevertheless, there are trade-offs for increasing theoretical risk.
-Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favorable interpretation.
+Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
 To that end, we incorporate uncertainty into our framework.
@@
 
-Theoretical risk is deeply connected with statistical methods, because its the inverse of $P(E|¬H)$.
-$P(E|¬H)$ is equivalent to the type I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill-hypothesis is $¬H$.
-Because many researchers are familiar with type I error rate, it can be helpful to remember this connection to theorethical risk.
-Researchers who choose a smaller type I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis, and therefore, the relation between statistical type I error rate and theoretical risk should not be over interpreted.
-We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
+Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
+In other words, the probability that a given significant finding reflects a true effect is larger if researchers lower their type I error rates.
+However, the research hypothesis is seldomly equal to the statistical null hypothesis.
+We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to theoretical risk.
+Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
+They are thus closely related to theoretical risk.
 
-The point is that the contribution of statistical methods to theoretical risk can be formally assessed.
+The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
 For many statistical models it can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to theoretical risk.
-The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
+The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
 We can not hope to resolve this uncertainty; but we have to be aware of its implications.
@@
 However, other factors in the broader context of the study are much harder to quantify.
-If one chooses to focus only on the contribution of statistical methods to theorethical risk, one is bound to overestimate it.
-Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
+If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
+Take, for example, a t-test of mean differences.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its promised type I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$ and, hence, be subtracted from theoretical risk.
-While, in theory, these factors may leave theoretical risk unaffected or even increase it, we argue that this is not the case in practice.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled, translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
+While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
 While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-The only thing we can conclude with some certainty is that theorethical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
+We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
+The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
 

### 2022-06-24 manuscript.Rmd
Update manuscript.Rmd

--- 2022-06-24 manuscript.Rmd
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+We restrict ourselves to evidence of a binary nature (either exists or does not exist) <!--this seems like an ontological level but here again, aren't we on an epistemic level, so shouldn't it be: was observed or not?
+--> since continuous evidence would lead to some quite involved derivations.
 <!--AB: What about measurement error? Is evidence thought to be measurement-free her? -->
-We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+We denote the probability <!--AB: personally, I think that a lot of confusion with Bayesian results is that the 'term' probability makes it sound objective and truthful; would you be okay with replacing probability with credibility to underline the subjective nature of this approach?
+Or would this be even more confusing?--> of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
 <!--AB: Not all readers may understand what the probability of a hypothesis is. Elaborate more on the probability interpretation of Bayesians as credibility.-->
@@
 $P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to $P(H)$, it is called increase in firmness.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the prior believe in the hypothesis, $P(H)$, it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
@@
 
-In short, we want to increase $P(H|E)$.
+In short, to support our theory as good as possible, we want to increase $P(H|E)$. <!-- or: to corroborate or to confirm or...?-->
 Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
+<!--Why the expectation?
+Because you take the expectation of an entire field or a scientist who does multiple experiments over her/his career?
+-->
 So how can we increase $P(H|E)$?
 The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher $P(H)$ leads to higher $P(H|E)$.
-However, the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
-Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
+The first option leads us to the unsurprising conclusion that a higher prior believe in a theory, $P(H)$, leads to higher $P(H|E)$.
+If a hypothesis is more credible to begin with, the same increase in corroboration will lead to a higher credibility than if the study was less credible to begin with.
+<!--AB: not sure - state the obvious explicitly?-->
+Anyhow, the a priori probability of a hypothesis is nothing our study design can change.
+The second option is similar unsurprising; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
+In other words, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
 While this truism is strongly related to study design, it is unrelated to preregistration.
@@
 In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
-However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ is theoretical risk.
+However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ can be related to or even equated to theoretical risk.
 
@@
 
-Let us note some interesting properties of theoretical risk $P(¬E|¬H)$.
+Let us note some interesting properties of theoretical risk, $P(¬E|¬H)$.
 <!--AB: I rephrased this sentence. Is credibility equal to epistemic value?-->
@@
 
-## Causes of Uncertainty
+## Sources of Uncertainty <!--AB really causes?-->
 
 As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
-However, other factors in the broader context of the study are much harder to quantify.
+However, other factors in the broader context of a given study are much harder to quantify.
 If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test of mean differences.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its promised type I error rate.
+Take, for example, a t-test of mean differences in two samples.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it will keep its promised type I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled, translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying hypothesis being false.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled (thus jeopardizing distributional assumptions<!-- AB: is this what you meant?-->), translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
 Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
@@
 Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
-While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
+While one can hope that processing this information does not affect their decision making either way, this cannot be ruled out.
 We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
-The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
+The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about other factors that play a role.
 
@@
 Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
-Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
+Uncertainty about theoretical risk is expressed through the entropy of the distribution. Under Gaussian assumptions, this is proportional to the variance of the distribution.
 Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
+
+<!--AB: I am confused that a new section heading is following but I haven't still understood or seen what follows from the expectation that you just gave.
+I assume the problem is that you created Figures 1 & 2 to illustrate the consequences of the choices you made and the expectation you presented but you do not refer to these Figures from the text, o the reader is lost and may never read the figures...
+please refer to the figures and explain what is going on in the text.
+Currently, I do not know what to take from Figure 2 other than that there are different ways to formalize this.
+Please also cite all sources for these measures in the text.-->
 
@@
 
-We hope to have persuaded you to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
-Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
+We hope to have persuaded the reader to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
+Second, the theoretical risk is inherently uncertain and the degree of uncertainty diminishes the persuasiveness of the gathered evidence. <!--AB my rewrite correct?-->
 The last argument is that preregistrations reduces this uncertainty.
@@
 
-Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
+Communicating clearly what the authors did to arrive at the evidence (<!--AB: "to arrive at the evidence" is unclear to me, do you mean a conditional probability or really the evidence itself, say, in numeric form in a data table)?
+--> is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.

### 2022-06-24 manuscript.Rmd
Update manuscript.Rmd

@@
 The authors could have decided to exclude a number of analytic strategies they tried out etc.
-That is not to say that any study that was not preregistered was subjected to practices of p hacking.
-The point is, we can not exclude this and a myriad of other possibilities and, hence, are left uncertain.
-If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
+That is not to say that a study that was not preregistered was subjected to practices of p hacking or other questionable research practices.
+The point is, as a sceptic, we cannot exclude this and a myriad of other possibilities that may bias the conclusions drawn from the presented evidence and, hence, are left in a state of uncertainty.
+If the researchers do describe what they intend to do beforehand, and then report they did exactly that, this uncertainty is drastically reduced.
+They still might be uncertain about the concrete theoretical risk the authors took but much less so.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analysies, etc.
 In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
@@
 
-We started out with the observation that preregistrations do no always cleanly divide preregistration from confirmation.
+We started out with the observation that preregistrations do no always cleanly distinguish exploration from confirmation. 
 Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
 No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as the term implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
-As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
+However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
+As we further argued, minimizing the uncertainty about theoretical risk is similarly beneficial without increasing the likelihood of a "failed" study.
+Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
+<!--AB: is my rephrasing correct?
+Please elaborate more.
+Why is it helpful?
+In the sense of a meta-analysis?
+-->
+Under the current incentive structure in academia, however, many researchers have reason to avoid such negative outcomes and few, if any, have the privilege to operate outside of these incentives.
+<!--AB: but see registered reports, journals that increasingly publish null findings, Bayesian analyses etc-->
 

### 2022-06-24 manuscript.Rmd
Update manuscript.Rmd

@@
 However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
-As we further argued, minimizing the uncertainty about theoretical risk is similarly beneficial without increasing the likelihood of a "failed" study.
+As we further argued, minimizing the uncertainty about theoretical risk (without necessarily changing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
+<!--AB: is my rephrasing correct?-->
 Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
@@
 Under the current incentive structure in academia, however, many researchers have reason to avoid such negative outcomes and few, if any, have the privilege to operate outside of these incentives.
-<!--AB: but see registered reports, journals that increasingly publish null findings, Bayesian analyses etc-->
+<!--AB: but see registered reports, journals that increasingly publish null findings, Bayesian analyses etc; I am not sure whether we need these last two sentences?
+-->
 
-The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
-First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
+In our view, the advantage of underlining uncertainty reduction about theoretical risk front and center is two-fold.
+First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks. <!--AB: do you refer to the risk functions in Fig2? I am not sure; or type of studies?!-->
 That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
+<!--AB: this point, I still do not get; the more exploratory a study, the less I can reduce uncertainty to the extreme that I have full uncertainty; but I assume your argument is that whatever reduction in uncertainty is possible is beneficial; but in practice: as a reader: how can I really judge this continuum of uncertainty from a given manuscript??-->
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
+<!--AB to repeat: this is still a claim that yet has to stand the test of reality.
+how should we do this really?
+I think we need to stress that this is a philosophical perspective rather than a tutorial on how to do this in practice.
+But...
+can we give practical advise?-->
 Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
 Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
+<!--AB: can you connect this to the probability terms, e.g., thus they reduce their actual theoretical risk through QRPs?-->
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.

### 2022-06-24 manuscript.Rmd
Update manuscript.Rmd

--- 2022-06-24 manuscript.Rmd
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
 Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
-
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
@@
 Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychological research or elsewhere.
+Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
 
 We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
-
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory (which itself is an ontologocial concept) is what we call epistemic value.
-Researchers cannot know what they will learn from a study before they have run it.
-However, they can develop an expectation that helps them to decide which study to conduct <!--AB in the sense of maximum information gain?
-What does "which study" mean?
-I guess you mean just any possible property of an study in the planning phase that can be altered?-->.
+The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
+While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
 This expectation is what we term expected epistemic value.
@@
 Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is intentionally reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 
@@
 
-Our first argument is that theorethical risk is crucial for judging theories.<!-- again: judging what?-->
+Our first argument is that theorethical risk is crucial for judging theories.
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either exists or does not exist) <!--this seems like an ontological level but here again, aren't we on an epistemic level, so shouldn't it be: was observed or not?
---> since continuous evidence would lead to some quite involved derivations.
-<!--AB: What about measurement error? Is evidence thought to be measurement-free her? -->
-We denote the probability <!--AB: personally, I think that a lot of confusion with Bayesian results is that the 'term' probability makes it sound objective and truthful; would you be okay with replacing probability with credibility to underline the subjective nature of this approach?
-Or would this be even more confusing?--> of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
-<!--AB: Not all readers may understand what the probability of a hypothesis is. Elaborate more on the probability interpretation of Bayesians as credibility.-->
+We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
 The probability of observing evidence under some hypothesis is $P(E|H)$.
@@
 
-$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the prior believe in the hypothesis, $P(H)$, it is called increase in firmness.
+The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
+However, it is generally expected that any measure of corroboration (and even increases monotonically with an increase in posterior probability ($P(H|E)$).
 
-In short, to support our theory as good as possible, we want to increase $P(H|E)$. <!-- or: to corroborate or to confirm or...?-->
-Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
-<!--Why the expectation?
-Because you take the expectation of an entire field or a scientist who does multiple experiments over her/his career?
--->
-So how can we increase $P(H|E)$?
-The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that a higher prior believe in a theory, $P(H)$, leads to higher $P(H|E)$.
-If a hypothesis is more credible to begin with, the same increase in corroboration will lead to a higher credibility than if the study was less credible to begin with.
-<!--AB: not sure - state the obvious explicitly?-->
-Anyhow, the a priori probability of a hypothesis is nothing our study design can change.
-The second option is similar unsurprising; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
-In other words, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-While this truism is strongly related to study design, it is unrelated to preregistration.
+In short, we want to increase posterior probability ($P(H|E)$).
+Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, of which we want to maximize the expectation.
+So how can we increase posterior probability?
+The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
+However, the prior probability of a hypothesis is nothing our study design can change.
+The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
+$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
+In other words, how probable is it that we "detect" that a theory holds, we therefore call it "detectability".
+Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
+When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type II error rate).
+However, while detectability is of great importance for study design, it is not directly relevant for the objective of preregistration.
 Thus, $P(E)$ remains to be considered.
-Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
 In other words, high risk, high reward.
@@
 
-To better understand how preregistrations can do that, let us take a closer look at what contributes to $P(E)$.
+To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$.
+<!--AB: is this expectedness (Pap, 1962) as quoted on p.818 in Meehl's risk paper?-->
 Using the law of total probability, we can split $P(E)$ into two terms:
@@
 
-We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
-<!--AB: from the perspective of a rational agent/researcher, of course, it makes sense to make P(E|H) large but isn't psychology facing a widespread problem of underpowered studies?
-I would make this very clear that this essentially relates to power (right?) and that historically we haven't been doing well on that account but that this problem is not pertinent to preregistration as such-->
+We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
 The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
+This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
-In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
-However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ can be related to or even equated to theoretical risk.
+
+Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
+First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
+Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
+If detectability exceeds theorethical risk, the evidence is less likely under the theory than it is when the theory does not hold.
+Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
+In other words, observing a sure fact does not lend credence to a hypothesis.
+
+This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
+He stated that if it is impossible to prove a hypothesis false ($P(¬E|¬H) = 0$, theoretical risk is zero), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
+18].
+We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+
+Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
+The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
+Suppose the primary theory is about the cognitive phenomenon of "insight".
+For the purpose of illustration, we define it somewhat hand-wavinlyg as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
+We pose the hypothesis that the following problem belongs to such class of insight problems:
+
+> Use five matches (Ⅰ Ⅰ Ⅰ Ⅰ Ⅰ) to form the number eight.
+
+We propose three experiments that differ in theoretical risk and detectability.
+All experiments take a sample of ten psychology students.
+We present the students the problem for a brief span of time.
+After that, the three experiments differ as follows:
+
+1. the experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
+2. the experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
+3. the experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
+
+We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
+If "insight" has nothing to do with solving the problem ($¬H$), than presenting the insight that roman literals might be used, should not lead to all students solving the problem ($¬E_1$); the experiment has therefore high theoretical risk ($P(¬E_1|¬H)$)
+Conversly, if insight is required to solve the problem ($H$), than it is probable to help all students to solve the problem ($E_1$); the experiment has therefore high detectability ($P(E_1|H)$).
+The second experiment, on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
+Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
+With regard to detectability experiment 1 and 2 differ in no obvious way.
+Experiment 3, however, also has low detectability.
+It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); the experiment 3 has therefore low detectability ($P(E_3|H)$).
+The theoretical risk, however, is also low in absolute terms but high compared to the theoretical risk.
+In the unlikely event that all 10 students lay the matches down in the form of the roman numeral Ⅷ ($E_3$) it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
+Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
+As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
+We discuss the relation between statistical methods and theoretical risk in the section [Statistical Methods].
 
@@
 
-<!--AB can we add an example that explains this whole idea once more but with a concrete example?
-I think that the probability of evidence given NOT H is just difficult to wrap one's head around.
-It would be ideal to illustrate this with an example that presents both a weak and a strong P(E given not H).
-
-I have a hard time coming up with an easy example that relates this simplified case
-to a problem of psychological research.
-
-H: a depression treatment is effective
-not H: the treatment is not effective
-
-In a weak risk case, the prediction is fuzzy, P(E | not H) is large, theoretical risk is small:
-
-E: people have some depression score (other than zero) after the treatment
-not E: people have exactly zero depression scores
-
-In a strong risk case, the prediction is precise, P(E | not H) is small, theoretical risk is large
-E: people improve at least 10 points on the depression scale
-not E: people improve less or equal to 10 points on the depression scale
--->
-
-Let us note some interesting properties of theoretical risk, $P(¬E|¬H)$.
-<!--AB: I rephrased this sentence. Is credibility equal to epistemic value?-->
-First, if the objective is to increase the credibility of a theory $P(H|E)$, this can
-be achieved by increasing theoretical risk.
-Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
-<!--AB: explain more what this means. When is this condition true in reality?-->
-Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
-In other words, observing a sure fact does not lend credence to a hypothesis.
-
-<!--Is this the word "this" refering to the third point only or all three?-->
-This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false ($P(E|¬H) = 1$ or $P(¬E|¬H) = 0$), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
-We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
-
 # Preregistration as a Means to Increase Theoretical Risk?
@@
 After we discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
-Indeed, limiting the false positive (or, type I error rate) is commonly stated as a goal of preregistration. <!--AB: citation needed; maybe verbose quote?-->
+Indeed, limiting type I error rate is commonly stated as a goal of preregistration.
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
 
-<!--AB: the kind of more concrete example that you give here is needed much earlier when you define the different conditional probabilities; maybe even in the very beginning, it could be helpful to give examples of risky vs non-risky predictions in psychology from published literature?!-->
 First, the theory itself limits theoretical risk:
@@
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within each of the sets relate to each other.
-Now, our imaginary researcher considers the relation between these two sets.
+Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
+However, our imaginary researcher considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation due to multiple testing [@orbenCrudReDefined2020].
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
 by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low until substantiated by testing it again.
-<!--AB: "till" is never used in formal writing-->
-The point, however, is that we want to show that preregistration increases the expected epistemic value independent of the theory being tested.
+We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
+The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
 
 Second, available resources limit theoretical risk.
-Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
-<!-- I recommend giving specific names to each of the conditional probabilities and then refer to those names and have the math notation in parentheses; then, this becomes easier to read and understand, e.g.,
-
-Increasing theoretical risk ($P(¬E|¬H)$) will usually decrease the probability of observing the evidence under the hypothesis ($P(E|H)$).
-
-Is there a term, we can introduce for $P(E|H)$, like power?
-
--->
+Increasing theorethical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
 In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ (the inverse of theorethical risk) and statistical methods.
 If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
 Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
-In other words, the probability that a given significant finding reflects a true effect is larger if researchers lower their type I error rates.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-<!--AB: can you elaborate more?!-->
+We argue that $P(E|¬H)$ (and hence theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
@@
 
-## Sources of Uncertainty <!--AB really causes?-->
+## Causes of Uncertainty
 
 As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
-However, other factors in the broader context of a given study are much harder to quantify.
+However, other factors in the broader context of the study are much harder to quantify.
 If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test of mean differences in two samples.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it will keep its promised type I error rate.
+Take, for example, a t-test.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled (thus jeopardizing distributional assumptions<!-- AB: is this what you meant?-->), translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
+They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
 While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
-While one can hope that processing this information does not affect their decision making either way, this cannot be ruled out.
+While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
 We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
-The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about other factors that play a role.
+The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
 
@@
 Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
-Uncertainty about theoretical risk is expressed through the entropy of the distribution. Under Gaussian assumptions, this is proportional to the variance of the distribution.
+Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
 Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
-
-<!--AB: I am confused that a new section heading is following but I haven't still understood or seen what follows from the expectation that you just gave.
-I assume the problem is that you created Figures 1 & 2 to illustrate the consequences of the choices you made and the expectation you presented but you do not refer to these Figures from the text, o the reader is lost and may never read the figures...
-please refer to the figures and explain what is going on in the text.
-Currently, I do not know what to take from Figure 2 other than that there are different ways to formalize this.
-Please also cite all sources for these measures in the text.-->
 
@@
 
-We hope to have persuaded the reader to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
-Second, the theoretical risk is inherently uncertain and the degree of uncertainty diminishes the persuasiveness of the gathered evidence. <!--AB my rewrite correct?-->
+We hope to have persuaded you to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
+Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
 The last argument is that preregistrations reduces this uncertainty.
@@
 
-Communicating clearly what the authors did to arrive at the evidence (<!--AB: "to arrive at the evidence" is unclear to me, do you mean a conditional probability or really the evidence itself, say, in numeric form in a data table)?
---> is crucial for judging the theoretical risk they took.
+Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.

### 2022-06-24 manuscript.Rmd
Test PR Preview

@@
 
-```{r, include = FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE
-)
-
-if(!requireNamespace("pacman"))install.packages("pacman")
-pacman::p_load("tidyverse", "ggplot2", "ggthemes", "patchwork")
-```
-
-```{r tikz-setup, include=FALSE}
-if(!requireNamespace("tikzDevice"))install.packages("tikzDevice")
-
-options(
-  tikzDefaultEngine = "xetex",
-  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}", "\\usepackage{libertine}")
-)
-```
-
-<!--problems to solve:
-* I feel like I am describing a ruby goldberg machine here and expect the reader to understand what happens when you move the spoon
-* what about falsificationists, every time I write a paragraph relating to I delete it and try to move it a discussion
-* references!!!
--->
-
-A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
-Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
-Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
-However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
-These skeptics question why a study should be preregistered at all.
-Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistraion and how detailed this should be, or about the conditions that warrant deviating from it.
-If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
-We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
-
-The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate is not warranted.
-Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
-One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
-Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmfull all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
-
-These issues arise from a fundamental confusion about the objective of preregistration.
-Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
-Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
-
-We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
-We then outline two interpretations of how theoretical risk is impacted by preregistration.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type I error.
-The second interpretation is our main contribution and demonstrates the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
-Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
-
-To arrive at this interpretation, we rely on three arguments.
-The first is that theoretical risk is vital for judging theories.
-The second argument is that the theoretical risk for a given study is generally uncertain.
-The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
-
-# Epistemic value and the Bayesian rationale
-
-Let us start by defining what we call expected epistemic value.
-If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
-In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
-This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to choosing a study to conduct.
-
-1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.
-
-The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
-Our rationale is as follows.
-Researchers who decide to conduct a study are akin to choosing a study to bet on.
-They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
-This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
-Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
-The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
-The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
-Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is purposefully reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-
-# Epistemic Value and Theoretical Risk
-
-Our first argument is that theorethical risk is crucial for judging theories.
-Put simply, risky predictions create persuasive evidence if they turn out to be correct.
-This point is crucial because we attribute much of the appeal of preregistration to this fact.
-
-Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
-We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
-The probability of observing evidence under some hypothesis is $P(E|H)$.
-We can calculate the probability of the hypothesis after observing the evidence with help from the Bayes formula:
-
-$$
-P(H|E) = \frac{P(H)P(E|H)}{P(E)}
-$$
-
-The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
-We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration (and even increases monotonically with an increase in posterior probability ($P(H|E)$).
-
-In short, we want to increase posterior probability ($P(H|E)$).
-Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, of which we want to maximize the expectation.
-So how can we increase posterior probability?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
-However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
-$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-In other words, how probable is it that we "detect" that a theory holds, we therefore call it "detectability".
-Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant for the objective of preregistration.
-Thus, $P(E)$ remains to be considered.
-Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high reward.
-
-If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
-This tension between high risk leading to high reward is central to our consideration of preregistration.
-A high risk, high reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
-Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many if not most researchers operate.
-Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
-It is at this juncture that the practice and mispractice of preregistration comes into play.
-We argue that the main function of preregistration is to enable proper judgment of the riskiness of a study.
-
-To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$.
-<!--AB: is this expectedness (Pap, 1962) as quoted on p.818 in Meehl's risk paper?-->
-Using the law of total probability, we can split $P(E)$ into two terms:
-
-$$
-P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
-$$
-
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
-The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
-Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-
-Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
-First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
-Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theorethical risk, the evidence is less likely under the theory than it is when the theory does not hold.
-Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
-In other words, observing a sure fact does not lend credence to a hypothesis.
-
-This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false ($P(¬E|¬H) = 0$, theoretical risk is zero), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
-18].
-We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
-
-Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
-To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
-The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
-Suppose the primary theory is about the cognitive phenomenon of "insight".
-For the purpose of illustration, we define it somewhat hand-wavinlyg as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
-We pose the hypothesis that the following problem belongs to such class of insight problems:
-
-> Use five matches (Ⅰ Ⅰ Ⅰ Ⅰ Ⅰ) to form the number eight.
-
-We propose three experiments that differ in theoretical risk and detectability.
-All experiments take a sample of ten psychology students.
-We present the students the problem for a brief span of time.
-After that, the three experiments differ as follows:
-
-1. the experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-2. the experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
-3. the experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
-
-We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($¬H$), than presenting the insight that roman literals might be used, should not lead to all students solving the problem ($¬E_1$); the experiment has therefore high theoretical risk ($P(¬E_1|¬H)$)
-Conversly, if insight is required to solve the problem ($H$), than it is probable to help all students to solve the problem ($E_1$); the experiment has therefore high detectability ($P(E_1|H)$).
-The second experiment, on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
-Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
-With regard to detectability experiment 1 and 2 differ in no obvious way.
-Experiment 3, however, also has low detectability.
-It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); the experiment 3 has therefore low detectability ($P(E_3|H)$).
-The theoretical risk, however, is also low in absolute terms but high compared to the theoretical risk.
-In the unlikely event that all 10 students lay the matches down in the form of the roman numeral Ⅷ ($E_3$) it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
-Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
-As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
-We discuss the relation between statistical methods and theoretical risk in the section [Statistical Methods].
-
-```{r, include=FALSE}
-bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
-```
-
-# Preregistration as a Means to Increase Theoretical Risk?
-
-After we discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
-Indeed, limiting type I error rate is commonly stated as a goal of preregistration.
-We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
-
-First, the theory itself limits theoretical risk:
-Some theories simply do not make risky predictions, and preregistration will not change that.
-Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
-However, our imaginary researcher considers the relation between these two sets.
-For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
-However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
-by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
-
-Second, available resources limit theoretical risk.
-Increasing theorethical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
-In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
-Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
-Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-
-# Uncertainty about Theoretical Risk
-
-We established that higher theoretical risk leads to more persuasive evidence.
-In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.
-Nevertheless, there are trade-offs for increasing theoretical risk.
-Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favourable interpretation.
-To that end, we incorporate uncertainty into our framework.
-
-## Statistical Methods
-
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ (the inverse of theorethical risk) and statistical methods.
-If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
-Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
-They are thus closely related to theoretical risk.
-
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
-For many statistical models it can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
-The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
-We can not hope to resolve this uncertainty; but we have to be aware of its implications.
-These are relayed in the following.
-
-## Causes of Uncertainty
-
-As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
-However, other factors in the broader context of the study are much harder to quantify.
-If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
-However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
-While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
-Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
-While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
-The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
-
-## The effects of uncertainty
-
-Before we ask how preregistration is influencing this uncertainty, we must consider the implications of being uncertain about the theoretical risk.
-Within the Bayesian framework, this is both straightforward and insightful.
-To get an expectation, we express uncertainty as a probability distribution and then integrate over it:
-
-$$
-\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H))
-$$
-
-To illustrate the effect of uncertainty, let $p(E|H) = .8$ (e.g., power of 80%) and $p(H) = .1$ and assume a uniform distribution for $p(E|\neg{}H)$ of the form:
-
-$$
-f(x)=\begin{cases}
-  \frac{1}{1 - \alpha} & \mathrm{for}\ \alpha \le x \le 1,\\
-  0 & \mathrm{for}\ x<\alpha\ \mathrm{or}\ x>1
-  \end{cases}
-$$
-
-Where $\alpha$ is, e.g., the statistical model's nominal type I error rate.
-
-We chose this uniform distribution to capture our statement that a statistical model only guarantees a lower bound to $p(E|\neg{}H)$ (and since its a probability the upper bound is 1) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
-
-```{r, include=FALSE, cache=TRUE}
-bayes <- function(h1, power, alpha, ...) {
-  (h1 * power) / (h1 * power + (1 - h1) * alpha)
-}
-uncertain_bayes <- function(h1, power, alpha, alpha_upper = 1, ...) {
-  mean(bayes(h1, power, runif(1e6, alpha, alpha_upper)))
-}
-
-step <- 0.01
-vary_alpha <-
-  tidyr::expand_grid(
-    h1 = .1,
-    power = .8,
-    alpha = seq(step, 1 - step, step)
-  ) %>% 
-  dplyr::mutate(.,
-    certain = purrr::pmap_dbl(., bayes),
-    uncertain = purrr::pmap_dbl(., uncertain_bayes)
-  )
-
-vary_alpha_long <- vary_alpha %>% 
-  pivot_longer(c(certain, uncertain),
-               names_to = "certainty",
-               values_to = "h2")
-```
-
-```{r, dev='tikz'}
-#| fig.cap: "Corroboration as firmness ($p(E|H)$) as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
-vary_alpha_long %>%
-  ggplot() +
-  geom_line(aes(alpha, h2, linetype = certainty)) +
-  lims(x = c(0, 1), y = 0:1) +
-  geom_ribbon(data = vary_alpha, aes(x = alpha, ymin = uncertain, ymax = certain), alpha = .1) +
-  labs(caption = "$p(E|H) = .8$, $p(H) = .1$",
-       y = "$\\mathbb{E}\\big(p(H|E)\\big)$",
-       x = "$\\alpha$") +
-  theme_tufte() +
-  theme(legend.position = c(.85, .8),
-        legend.title = element_blank())
-```
-
-```{r, include=FALSE}
-# h1 = prior
-# eh = p(E|H)
-# enh = p(E|\neg{}H)
-
-bayes <- function(h1, eh, enh, ...) {
-  (h1 * eh) / (h1 * eh + (1 - h1) * enh)
-}
-# all formulas taken from Bayesian Philosophy of Science, p. 51
-# first letter = function name in Bayesian Philosophy of Science
-# optional letter = p for prime
-# second letter = m for measure
-measures <- list(
-  dm = list(
-    name = "Difference Measure",
-    formula = "d(H, E) = p(H|E) - p(H)",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
-      he - h1
-    }
-  ),
-  rm = list(
-    name = "Log-Ratio Measure",
-    formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
-      log(he / h1)
-    }
-  ),
-  lm = list(
-    name = "Log-Likelihood Measure",
-    formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
-    fun = function(h1, eh, enh, ...) {
-      he <- bayes(h1, eh, enh)
-      log(he / enh)
-    }
-  ),
-  km = list(
-    name = "Kemeny–Oppenheim Measure",
-    formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
-    fun = function(h1, eh, enh, ...) {
-      (eh - enh) / (eh + enh)
-    }
-  ),
-  zm = list(
-    name = "Generalized Entailment Measure",
-    formula =
-      "z(H, E) =\\begin{cases}
-  \\frac{p(E|H) - p(H)}{1 - p(H)}, & \\text{if } p(H|E) \\geq{} p(H) \\\\
-  \\frac{p(E|H) - p(H)}{p(H)}, & \\text{if } p(H|E) < p(H)
-\\end{cases}",
-fun = function(h1, eh, enh, ...) {
-  he <- bayes(h1, eh, enh)
-  ifelse(he >= h1, (he - h1)/(1 - h1), (he - h1)/h1)
-}
-  ),
-sm = list(
-  name = "Christensen–Joyce Measure",
-  formula = "s(H, E) = p(H|E) - p(H|\\neg{}E)",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    e <- h1 * eh + (1 - h1) * enh
-    hne <- (h1 * (1 - eh)) / (1 - e)
-    he - hne
-  }
-),
-cpm = list(
-  name = "Carnap’s Relevance Measure",
-  formula = "c'(H, E) = p(E)(p(H|E) - p(H))",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    e <- h1 * eh + (1 - h1) * enh
-    e * (he - h1)
-  }
-),
-rpm = list(
-  name = "Rips Measure",
-  formula = "r'(H, E) = 1 - \\frac{p(\\neg{}H|E)}{p(\\neg{}H)}",
-  fun = function(h1, eh, enh, ...) {
-    he <- bayes(h1, eh, enh)
-    1 - (1 - he) / (1 - h1)
-  }
-)
-)
-
-measures <- tibble(!!!transpose(measures)) %>% 
-  mutate(name = simplify(name, "chr"),
-         formula = simplify(formula, "chr"))
-
-uncertainly <- function(fun, n = 1e5, seed = NULL){
-  function(h1, eh, enh){
-    if(!is.null(seed)){
-      if(exists(".Random.seed"))on.exit(set.seed(.Random.seed))
-      set.seed(seed)
-    }
-    enh_samples <- runif(n, enh, 1)
-    mean(fun(h1, eh, enh_samples))
-  }
-}
-```
-
-```{r, cache=TRUE, echo=FALSE}
-step <- 0.01
-
-grid <- tidyr::expand_grid(
-    h1 = .1,
-    eh = .8,
-    enh = seq(step, 1 - step, step)
-  )
-apply_on_grid <- function(f, grid) {
-  mutate(grid,
-         certain = pmap_dbl(grid, f),
-         uncertain = pmap_dbl(grid, uncertainly(f, seed = .Random.seed)))
-}
-
-results <- mutate(measures, results = map(fun, apply_on_grid, grid)) %>% 
-  select(-fun) %>% 
-  unnest(results) %>% 
-  pivot_longer(c(certain, uncertain),
-               names_to = "certainty",
-               values_to = "value")
-```
-
-```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Several measures for corroboration as increase in firmness as a function of α, where α is either certain (solid line) or maximally uncertain (dotted line)."
-
-results %>%
-  mutate(name = str_remove(name, "Measure")) %>%
-  split(., .$name) %>%
-  imap(
-    ~ ggplot(.x, aes(enh, value)) +
-      geom_line(aes(linetype = certainty)) +
-      labs(
-        title = .x$name,
-        y = element_blank(),
-        x = element_blank()
-      )
-  ) %>%
-  reduce(`+`) +
-  plot_layout(ncol = 2, guides = 'collect') &
-  theme_tufte() &
-  theme(
-    legend.title = element_blank(),
-    legend.position = "bottom",
-    plot.title = element_text(size = 7),
-    axis.text = element_text(size = 6)
-  )
-```
-
-Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall in.
-Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
-Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
-
-# Preregistration as a Means to Decreasing Uncertainty about the Theoretical Risk
-
-We hope to have persuaded you to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
-Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
-The last argument is that preregistrations reduces this uncertainty.
-
-Recollect, our three assumptions:
-
-1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else equal, researchers try to increase the expected epistemic value for other researchers.
-
-The point we make with these assumption is that the authors aim to persuade other researchers, their readers.
-Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of the resulting article (or, more generally, the consumer of the research product).
-We have to remember that the process of weighing evidence for or against a theory extends beyond the original authors to all the people they hope to persuade.
-While the authors may have deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
-In particular, they never know what relevant factors the authors, intentionally or not, failed to disclose.
-From the perspective of an ultimate sceptic, they may claim maximum uncertainty.
-
-Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
-Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
-The authors could have decided to exclude a number of analytic strategies they tried out etc.
-That is not to say that any study that was not preregistered was subjected to practices of p hacking.
-The point is, we can not exclude this and a myriad of other possibilities and, hence, are left uncertain.
-If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
-In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
-
-# Discussion
-
-We started out with the observation that preregistrations do no always cleanly divide preregistration from confirmation.
-Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
-No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as the term implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
-As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
-
-The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
-First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
-That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
-If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
-Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
-Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
-This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might believe it to be a likely explanation for the deviation.
-If the readers find the reason for the deviation likely to be something other than decreasing the theorethical risk, their uncertainty about the theorethical risk should only increase modestly.
-
-# Old Material
-
-> As far as I am aware, Mayo’s severity argument currently provides one of the few philosophies of science that allows for a coherent conceptual analysis of the value of preregistration.
-> Borsboom
-
-We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
-
-If you are a follow a hypo-deductive rational this is self evident, if you are positivist we hope our
-
-The other two assumptions are later necessary to connect individual decisions that affect epistemic value to the research community as a whole.
-
-Observe, that since $P(E|¬H)$.
-
-Please note, that we adopt a Bayesian rational for the meta scientific process of preregistration but that this does not imply any ties to the methods a researcher uses.
-
-However, we aim to show that preregistration is indispensable for adequately judging a much broader range of studies.
-To that end, we first rationalize the appreciation for conformation using Bayesian reasoning.
-Equipped with some of the basic tools of Bayesian Philosophy of Science, we can move beyond a simple dichotomy of exploration and confirmation.
-
-To that end we show that the appreciation for the distinction between exploration and confirmatory neatly falls in line with Bayesian reasoning connected through a quantity we call *theorethical risk*.
-Than we proceed to show that preregistration impacts the theorethical risk via two pathways.
-Only considering the one path rationalizes recommendation to employ preregistration only for confirmatory studies, taking both paths together into consideration leads to a much wider applicability of preregistration.
-
-But what exactly have you been cheated of?
-This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
-<!-- We leverage this reasoning to outline two paradigms that rationalize the use of preregistration. -->
-In order to apply Bayesian reasoning we have to cast conformation and exploration in terms of probability.
-Consider the hypothesis that people who regularly engage in fitness activities are healthier.
-A confirmatory story might operationalize regularly as weekly, fitness activity as light jogging and healthy as blood cholesterol level.
-A more exploratory study might consider several definitions of "regularly", "fitness activities", and "healthy", e.g.
-"monthly", "weekly", "daily", or "marathon", "light jogging", "intense cleaning", or "cholesterol", "self rated health", "ability to play tag".
-Both studies have some probability to turn up with positive results even in a world where fitness activities do not lead to healthier people, because there are other plausible explanations for the results.
-However, for the second study the number of explanations that could lead to evidence in favor, aside from the theory is much greater.
-Suppose a researcher actually conducted the second study but is reporting it as if they conducted the first study.
-You would feel cheated.
-But what exactly have you been cheated of?
-In fact, you have been cheated out of the information about all the possibilities that could have brought about this evidence.
-We summarize these possibilities as a probability.
-Therefore, we argue that confirmatory studies have a low probability of observing evidence in favor of their theory if we assume the theory to be false.
-Exploratory studies on the other hand are characterized by a higher probability that they find evidence, even when their hypothesis turns out to be wrong.
-
-We connect the question of exploration vs confirmation to Bayesian reasoning via a quantity we call theoretical risk (we borrow the term from Meehl).
-
-Theoretical risk is the inverse of the probability that one observes evidence in favor of a theory when we assume that the theory does not hold.
-We assume that confirmatory studies take a high theoretical risk, compared to exploratory studies which take a low theoretical risk.
-The tighter your definition of "evidence in favor" is, the more the probability of observing the evidence is tied to your theory holding.
-
-<!-- Caroline does not see the connection to type I error rate/theoretical risk -->
-
-Let us formalize the judgment about a hypothesis (H) given some evidence (E) as the conditional probability $p(H|E)$.
-
-Using Bayes' rule we arrive at:
-
-$$
-p(H|E) = \frac{p(H)p(E|H)}{p(H)p(E|H) + p(¬H)p(E|¬H)}
-$$
-
-Note the connection to null hypothesis testing.
-If H represents the alternative Hypothesis and ¬H represents the Null hypothesis, then $p(E|H)$ is the power, while $p(E|¬H)$ represents the type I/alpha error.
-However, generally a theory is richer in content, than equating it with the alternaive hypothesis in Null hypothesis testing.
-
-<!-- continue to justify the importance of theorethical risk -->
-
-The first paradigm is embraced by the open science community and conceptualizes preregistration as a tool to reduce the type I error rate (XXXX).
-As we will later show, reducing the type I error rate is indeed a powerful leaver to gather persuasive evidence.
-In practice, however, researchers can not reduce the type I error rate at will.
-They face three limiting factors.
-First, psychological theories tend to be vague.
-Some decisions are just not derivable from theory, which makes it extraordinarily difficult to maintain a type I error rate near zero.
-Second, researchers are constrained by resource limitations.
-One way to counter the first problem is conducting exploratory pilot studies, but running extensive pilot studies is expensive.
-Doubling the resources that go into investigating a research question is often just not tenable and mostly beyond the individual researcher to decide.
-Third, by reducing the type I error rate the researcher directly increases the likelihood that their study will result in disappointing null results.
-A study that fails to deliver results can be crippling to the researchers career.
-Depending on individual circumstances, e.g.
-the theory in question, resources allocated to a study, the career stage of the researcher, a researcher may simply feel unable to reduce the type I error rate to the threshold necessary for a confirmatory study.
-Constrained by these factors researchers tasked with writing a preregistration are incentiviced to either give up or to muddle through by being intentionally vague.
-The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".
-
-Our conclusions can be applied to bayesian and frequentist methods alike.
-Of course, frequentist methods are traditionally viewed from a Popperian (or hypo-deductive) philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
-However, error statistical view marries a statistical philosophy with specific statistical models and procedures.
-A marriage which leaves us with richer set of assumptions than necessary for the points we want to make about the value of preregistration.
-
-# References
+Test PR Preview

### 2022-06-24 manuscript.Rmd
Update manuscript.Rmd

--- 2022-06-24 manuscript.Rmd
@@
 A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
 Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
@@
 We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
+
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration (and even increases monotonically with an increase in posterior probability ($P(H|E)$).
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
 
@@
 So how can we increase posterior probability?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
 The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
@@
 To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$. 
-<!--AB: is this expectedness (Pap, 1962) as quoted on p.818 in Meehl's risk paper?-->
 Using the law of total probability, we can split $P(E)$ into two terms:
@@
 The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
+This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
+In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
+However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ is theoretical risk.
 
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ (the inverse of theorethical risk) and statistical methods.
-If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
-Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+Theoretical risk is deeply connected with statistical methods, because its the inverse of $P(E|¬H)$.
+$P(E|¬H)$ is equivalent to the type I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill-hypothesis is $¬H$.
+Because many researchers are familiar with type I error rate, it can be helpful to remember this connection to theorethical risk.
+Researchers who choose a smaller type I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
+However, the research hypothesis is seldomly equal to the statistical null hypothesis, and therefore, the relation between statistical type I error rate and theoretical risk should not be over interpreted.
+We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to $P(E|¬H)$.
+Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to theoretical risk.
 Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
-They are thus closely related to theoretical risk.
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
 
-The point is that the contribution of statistical methods to $P(E|¬H)$ can be formally assessed.
+The point is that the contribution of statistical methods to theoretical risk can be formally assessed.
 For many statistical models it can be analytically computed under some assumptions.
-For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to $P(E|¬H)$.
-The precision with which researchers can discuss statistical contributions to $P(E|¬H)$ has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to theoretical risk.
+The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
 We can not hope to resolve this uncertainty; but we have to be aware of its implications.

### 2022-06-30 manuscript.Rmd
number equations closes #46
Update manuscript.Rmd

--- 2022-06-24 manuscript.Rmd
@@
 
+<!--AB: i feel that the intro needs a slower start and take the reader by the hand -->
+Imagine three researchers named Ben, Holly and the Wise Old Elf. <!--AB you may want to change these names--> 
+Based on his theoretical considerations, Ben predicts that personality traits are useful predictors of job success. 
+Based on her own theoretical considerations Holly specifically predicts that agreeableness is a good predictor of job success and she specifically predicts a correlation of 0.5.
+The Wise Old Elf who did not have a theory to begin with waits until Ben and Holly collected their data and after examining all correlations of job success and personality traits postdicts that indeed agreeableness is correlated with job success (which he claims to have known all along).
+Which theory should we attribute the largest credibility?
+
+Voice from off:  could we discuss precision of a theory and the exploration/confirmation distinction with an example like this?? And probably a lot more... let's discuss!
+-->
+
 A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
 Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
 As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+
 Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
@@
 Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
+Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychological research or elsewhere.
 
 We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [we borrow the term from @meehlTheoreticalRisksTabular1978], which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
+
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
-We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+We restrict ourselves to evidence of a binary nature (either exists or does not exist) <!--this seems like an ontological level but here again, aren't we on an epistemic level, so shouldn't it be: was observed or not?
+--> since continuous evidence would lead to some quite involved derivations.
+<!--AB: What about measurement error? Is evidence thought to be measurement-free her? -->
+We denote the probability <!--AB: personally, I think that a lot of confusion with Bayesian results is that the 'term' probability makes it sound objective and truthful; would you be okay with replacing probability with credibility to underline the subjective nature of this approach?
+Or would this be even more confusing?--> of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+<!--AB: Not all readers may understand what the probability of a hypothesis is. Elaborate more on the probability interpretation of Bayesians as credibility.-->
 The probability of observing evidence under some hypothesis is $P(E|H)$.
@@
 
-The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
+$P(H|E)$ is of great relevance since $P(H|E)$ is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the prior believe in the hypothesis, $P(H)$, it is called increase in firmness.
 We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration (and even increases monotonically with an increase in posterior probability ($P(H|E)$).
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in $P(H|E)$.
 
-In short, we want to increase posterior probability ($P(H|E)$).
-Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, of which we want to maximize the expectation.
-So how can we increase posterior probability?
-The Bayes formula gives us three options to investigate, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
-However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
-$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-In other words, how probable is it that we "detect" that a theory holds, we therefore call it "detectability".
-Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant for the objective of preregistration.
+In short, to support our theory as good as possible, we want to increase $P(H|E)$. <!-- or: to corroborate or to confirm or...?-->
+Increases in $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
+<!--Why the expectation?
+Because you take the expectation of an entire field or a scientist who does multiple experiments over her/his career?
+-->
+So how can we increase $P(H|E)$?
+The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The first option leads us to the unsurprising conclusion that a higher prior believe in a theory, $P(H)$, leads to higher $P(H|E)$.
+If a hypothesis is more credible to begin with, the same increase in corroboration will lead to a higher credibility than if the study was less credible to begin with.
+<!--AB: not sure - state the obvious explicitly?-->
+Anyhow, the a priori probability of a hypothesis is nothing our study design can change.
+The second option is similar unsurprising; that is, an increase in $P(E|H)$ leads to higher $P(H|E)$.
+In other words, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
+While this truism is strongly related to study design, it is unrelated to preregistration.
 Thus, $P(E)$ remains to be considered.
-Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+Since $P(E)$ is the denominator, increasing it will decrease $P(H|E)$: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
 In other words, high risk, high reward.
@@
 
-To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$.
-<!--AB: is this expectedness (Pap, 1962) as quoted on p.818 in Meehl's risk paper?-->
+To better understand how preregistrations can do that, let us take a closer look at what contributes to $P(E)$.
 Using the law of total probability, we can split $P(E)$ into two terms:
@@
 
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
+We already have noted that there is not much to do about $P(H)$ (and hence its counter probability $P(¬H)$), and that it is common sense to increase $P(E|H)$.
+<!--AB: from the perspective of a rational agent/researcher, of course, it makes sense to make P(E|H) large but isn't psychology facing a widespread problem of underpowered studies?
+I would make this very clear that this essentially relates to power (right?) and that historically we haven't been doing well on that account but that this problem is not pertinent to preregistration as such-->
 The real lever to pull is, therefore, $P(E|¬H)$.
-This probability tells us how likely it is that we find evidence when in fact, the theory is not true.
+This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-
-Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
-First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
-Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theorethical risk, the evidence is less likely under the theory than it is when the theory does not hold.
-Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
-In other words, observing a sure fact does not lend credence to a hypothesis.
-
-This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false ($P(¬E|¬H) = 0$, theoretical risk is zero), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
-18].
-We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
-
-Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
-To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
-The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
-Suppose the primary theory is about the cognitive phenomenon of "insight".
-For the purpose of illustration, we define it somewhat hand-wavinlyg as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
-We pose the hypothesis that the following problem belongs to such class of insight problems:
-
-> Use five matches (Ⅰ Ⅰ Ⅰ Ⅰ Ⅰ) to form the number eight.
-
-We propose three experiments that differ in theoretical risk and detectability.
-All experiments take a sample of ten psychology students.
-We present the students the problem for a brief span of time.
-After that, the three experiments differ as follows:
-
-1. the experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-2. the experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
-3. the experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
-
-We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($¬H$), than presenting the insight that roman literals might be used, should not lead to all students solving the problem ($¬E_1$); the experiment has therefore high theoretical risk ($P(¬E_1|¬H)$)
-Conversly, if insight is required to solve the problem ($H$), than it is probable to help all students to solve the problem ($E_1$); the experiment has therefore high detectability ($P(E_1|H)$).
-The second experiment, on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
-Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
-With regard to detectability experiment 1 and 2 differ in no obvious way.
-Experiment 3, however, also has low detectability.
-It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); the experiment 3 has therefore low detectability ($P(E_3|H)$).
-The theoretical risk, however, is also low in absolute terms but high compared to the theoretical risk.
-In the unlikely event that all 10 students lay the matches down in the form of the roman numeral Ⅷ ($E_3$) it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
-Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
-As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
-We discuss the relation between statistical methods and theoretical risk in the section [Statistical Methods].
+We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
+In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
+However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ can be related to or even equated to theoretical risk.
 
@@
 
+<!--AB can we add an example that explains this whole idea once more but with a concrete example?
+I think that the probability of evidence given NOT H is just difficult to wrap one's head around.
+It would be ideal to illustrate this with an example that presents both a weak and a strong P(E given not H).
+
+I have a hard time coming up with an easy example that relates this simplified case
+to a problem of psychological research.
+
+H: a depression treatment is effective
+not H: the treatment is not effective
+
+In a weak risk case, the prediction is fuzzy, P(E | not H) is large, theoretical risk is small:
+
+E: people have some depression score (other than zero) after the treatment
+not E: people have exactly zero depression scores
+
+In a strong risk case, the prediction is precise, P(E | not H) is small, theoretical risk is large
+E: people improve at least 10 points on the depression scale
+not E: people improve less or equal to 10 points on the depression scale
+-->
+
+Let us note some interesting properties of theoretical risk, $P(¬E|¬H)$.
+<!--AB: I rephrased this sentence. Is credibility equal to epistemic value?-->
+First, if the objective is to increase the credibility of a theory $P(H|E)$, this can
+be achieved by increasing theoretical risk.
+Second, if the theoretical risk is smaller than $P(E|H)$ it follows that $P(H|E)$ must decrease when observing the evidence.
+<!--AB: explain more what this means. When is this condition true in reality?-->
+Third, if the theoretical risk equals zero, then $P(H|E)$ can at best be equal to $P(H)$ but only if $P(H|E)$ is one.
+In other words, observing a sure fact does not lend credence to a hypothesis.
+
+<!--Is this the word "this" refering to the third point only or all three?-->
+This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
+He stated that if it is impossible to prove a hypothesis false ($P(E|¬H) = 1$ or $P(¬E|¬H) = 0$), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p. 18].
+We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+
 # Preregistration as a Means to Increase Theoretical Risk?
@@
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
-However, our imaginary researcher considers the relation between these two sets.
+Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within each of the sets relate to each other.
+Now, our imaginary researcher considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
+This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation due to multiple testing [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
 by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
+We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low until substantiated by testing it again.
+<!--AB: "till" is never used in formal writing-->
+The point, however, is that we want to show that preregistration increases the expected epistemic value independent of the theory being tested.
 
 Second, available resources limit theoretical risk.
-Increasing theorethical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
+Increasing $P(¬E|¬H)$ will usually decrease $P(E|H)$ unless more resources are invested.
+<!-- I recommend giving specific names to each of the conditional probabilities and then refer to those names and have the math notation in parentheses; then, this becomes easier to read and understand, e.g.,
+
+Increasing theoretical risk ($P(¬E|¬H)$) will usually decrease the probability of observing the evidence under the hypothesis ($P(E|H)$).
+
+Is there a term, we can introduce for $P(E|H)$, like power?
+
+-->
 In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
@@
 
-Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ (the inverse of theorethical risk) and statistical methods.
+Preregistration is often discussed in relation to statistical methods, so let us note some relations between $P(E|¬H)$ and statistical methods.
 If you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, then the nill-hypothesis is $¬H$, and therefore $P(E|¬H)$ is equivalent to the type I error rate.
 Researchers who choose a smaller type I error rate can hence be more sure of their results, if significant.
+In other words, the probability that a given significant finding reflects a true effect is larger if researchers lower their type I error rates.
 However, the research hypothesis is seldomly equal to the statistical null hypothesis.
-We argue that $P(E|¬H)$ (and hence theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+We argue that $P(E|¬H)$ (and hence its inverse, theorethical risk) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+<!--AB: can you elaborate more?!-->
 
@@
 
-## Causes of Uncertainty
+## Sources of Uncertainty <!--AB really causes?-->
 
 As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
-However, other factors in the broader context of the study are much harder to quantify.
+However, other factors in the broader context of a given study are much harder to quantify.
 If one chooses to focus only on the contribution of statistical methods to $P(E|¬H)$, one is bound to underestimate it.
-Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
+Take, for example, a t-test of mean differences in two samples.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it will keep its promised type I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled (thus jeopardizing distributional assumptions<!-- AB: is this what you meant?-->), translate their questionnaires into a different language without validating it, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
 While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
-While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
+While one can hope that processing this information does not affect their decision making either way, this cannot be ruled out.
 We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
-The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about the other factors at play.
+The only thing we can conclude with some certainty is that $P(E|¬H)$ is not lower than what the statistical model guarantees without knowledge about other factors that play a role.
 
@@
 
-Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
+Communicating clearly what the authors did to arrive at the evidence (<!--AB: "to arrive at the evidence" is unclear to me, do you mean a conditional probability or really the evidence itself, say, in numeric form in a data table)?
+--> is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
 The authors could have decided to exclude a number of analytic strategies they tried out etc.
-That is not to say that any study that was not preregistered was subjected to practices of p hacking.
-The point is, we can not exclude this and a myriad of other possibilities and, hence, are left uncertain.
-If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
+That is not to say that a study that was not preregistered was subjected to practices of p hacking or other questionable research practices.
+The point is, as a sceptic, we cannot exclude this and a myriad of other possibilities that may bias the conclusions drawn from the presented evidence and, hence, are left in a state of uncertainty.
+If the researchers do describe what they intend to do beforehand, and then report they did exactly that, this uncertainty is drastically reduced.
+They still might be uncertain about the concrete theoretical risk the authors took but much less so.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analysies, etc.
 In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
@@
 
-We started out with the observation that preregistrations do no always cleanly divide preregistration from confirmation.
+We started out with the observation that preregistrations do no always cleanly distinguish exploration from confirmation. 
 Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
 No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as the term implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
-As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
+However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
+As we further argued, minimizing the uncertainty about theoretical risk (without necessarily changing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
+<!--AB: is my rephrasing correct?-->
+Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
+<!--AB: is my rephrasing correct?
+Please elaborate more.
+Why is it helpful?
+In the sense of a meta-analysis?
+-->
+Under the current incentive structure in academia, however, many researchers have reason to avoid such negative outcomes and few, if any, have the privilege to operate outside of these incentives.
+<!--AB: but see registered reports, journals that increasingly publish null findings, Bayesian analyses etc; I am not sure whether we need these last two sentences?
+-->
 
-The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
-First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
+In our view, the advantage of underlining uncertainty reduction about theoretical risk front and center is two-fold.
+First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks. <!--AB: do you refer to the risk functions in Fig2? I am not sure; or type of studies?!-->
 That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
+<!--AB: this point, I still do not get; the more exploratory a study, the less I can reduce uncertainty to the extreme that I have full uncertainty; but I assume your argument is that whatever reduction in uncertainty is possible is beneficial; but in practice: as a reader: how can I really judge this continuum of uncertainty from a given manuscript??-->
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
+<!--AB to repeat: this is still a claim that yet has to stand the test of reality.
+how should we do this really?
+I think we need to stress that this is a philosophical perspective rather than a tutorial on how to do this in practice.
+But...
+can we give practical advise?-->
 Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
 Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
+<!--AB: can you connect this to the probability terms, e.g., thus they reduce their actual theoretical risk through QRPs?-->
 This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.

### 2022-07-11 manuscript.Rmd
remove old materials

--- 2022-06-24 manuscript.Rmd
@@
 
-$$
-P(H|E) = \frac{P(H)P(E|H)}{P(E)}
-$$
+\begin{equation}
+P(H|E) = \frac{P(H)P(E|H)}{P(E)} (\#eq:bayes)
+\end{equation}
 
@@
 
-$$
-P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
-$$
+\begin{equation}
+P(E) = P(H)P(E|H) + P(¬H)P(E|¬H) (\#eq:total-expectedness)
+\end{equation}
 
@@
 Suppose the primary theory is about the cognitive phenomenon of "insight".
-For the purpose of illustration, we define it somewhat hand-wavinlyg as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
+For the purpose of illustration, we define it somewhat hand-wavingly as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
 We pose the hypothesis that the following problem belongs to such class of insight problems:
@@
 
-$$
-\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H))
-$$
+\begin{equation}
+\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H)) (\#eq:integral)
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 f(x)=\begin{cases}
@@
   \end{cases}
-$$
+\end{equation}
 
@@
 
-# Old Material
-
-> As far as I am aware, Mayo’s severity argument currently provides one of the few philosophies of science that allows for a coherent conceptual analysis of the value of preregistration.
-> Borsboom
-
-We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.
-
-If you are a follow a hypo-deductive rational this is self evident, if you are positivist we hope our
-
-The other two assumptions are later necessary to connect individual decisions that affect epistemic value to the research community as a whole.
-
-Observe, that since $P(E|¬H)$.
-
-Please note, that we adopt a Bayesian rational for the meta scientific process of preregistration but that this does not imply any ties to the methods a researcher uses.
-
-However, we aim to show that preregistration is indispensable for adequately judging a much broader range of studies.
-To that end, we first rationalize the appreciation for conformation using Bayesian reasoning.
-Equipped with some of the basic tools of Bayesian Philosophy of Science, we can move beyond a simple dichotomy of exploration and confirmation.
-
-To that end we show that the appreciation for the distinction between exploration and confirmatory neatly falls in line with Bayesian reasoning connected through a quantity we call *theorethical risk*.
-Than we proceed to show that preregistration impacts the theorethical risk via two pathways.
-Only considering the one path rationalizes recommendation to employ preregistration only for confirmatory studies, taking both paths together into consideration leads to a much wider applicability of preregistration.
-
-But what exactly have you been cheated of?
-This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
-<!-- We leverage this reasoning to outline two paradigms that rationalize the use of preregistration. -->
-In order to apply Bayesian reasoning we have to cast conformation and exploration in terms of probability.
-Consider the hypothesis that people who regularly engage in fitness activities are healthier.
-A confirmatory story might operationalize regularly as weekly, fitness activity as light jogging and healthy as blood cholesterol level.
-A more exploratory study might consider several definitions of "regularly", "fitness activities", and "healthy", e.g.
-"monthly", "weekly", "daily", or "marathon", "light jogging", "intense cleaning", or "cholesterol", "self rated health", "ability to play tag".
-Both studies have some probability to turn up with positive results even in a world where fitness activities do not lead to healthier people, because there are other plausible explanations for the results.
-However, for the second study the number of explanations that could lead to evidence in favor, aside from the theory is much greater.
-Suppose a researcher actually conducted the second study but is reporting it as if they conducted the first study.
-You would feel cheated.
-But what exactly have you been cheated of?
-In fact, you have been cheated out of the information about all the possibilities that could have brought about this evidence.
-We summarize these possibilities as a probability.
-Therefore, we argue that confirmatory studies have a low probability of observing evidence in favor of their theory if we assume the theory to be false.
-Exploratory studies on the other hand are characterized by a higher probability that they find evidence, even when their hypothesis turns out to be wrong.
-
-We connect the question of exploration vs confirmation to Bayesian reasoning via a quantity we call theoretical risk (we borrow the term from Meehl).
-
-Theoretical risk is the inverse of the probability that one observes evidence in favor of a theory when we assume that the theory does not hold.
-We assume that confirmatory studies take a high theoretical risk, compared to exploratory studies which take a low theoretical risk.
-The tighter your definition of "evidence in favor" is, the more the probability of observing the evidence is tied to your theory holding.
-
-<!-- Caroline does not see the connection to type I error rate/theoretical risk -->
-
-Let us formalize the judgment about a hypothesis (H) given some evidence (E) as the conditional probability $p(H|E)$.
-
-Using Bayes' rule we arrive at:
-
-$$
-p(H|E) = \frac{p(H)p(E|H)}{p(H)p(E|H) + p(¬H)p(E|¬H)}
-$$
-
-Note the connection to null hypothesis testing.
-If H represents the alternative Hypothesis and ¬H represents the Null hypothesis, then $p(E|H)$ is the power, while $p(E|¬H)$ represents the type I/alpha error.
-However, generally a theory is richer in content, than equating it with the alternaive hypothesis in Null hypothesis testing.
-
-<!-- continue to justify the importance of theorethical risk -->
-
-The first paradigm is embraced by the open science community and conceptualizes preregistration as a tool to reduce the type I error rate (XXXX).
-As we will later show, reducing the type I error rate is indeed a powerful leaver to gather persuasive evidence.
-In practice, however, researchers can not reduce the type I error rate at will.
-They face three limiting factors.
-First, psychological theories tend to be vague.
-Some decisions are just not derivable from theory, which makes it extraordinarily difficult to maintain a type I error rate near zero.
-Second, researchers are constrained by resource limitations.
-One way to counter the first problem is conducting exploratory pilot studies, but running extensive pilot studies is expensive.
-Doubling the resources that go into investigating a research question is often just not tenable and mostly beyond the individual researcher to decide.
-Third, by reducing the type I error rate the researcher directly increases the likelihood that their study will result in disappointing null results.
-A study that fails to deliver results can be crippling to the researchers career.
-Depending on individual circumstances, e.g.
-the theory in question, resources allocated to a study, the career stage of the researcher, a researcher may simply feel unable to reduce the type I error rate to the threshold necessary for a confirmatory study.
-Constrained by these factors researchers tasked with writing a preregistration are incentiviced to either give up or to muddle through by being intentionally vague.
-The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".
-
-Our conclusions can be applied to bayesian and frequentist methods alike.
-Of course, frequentist methods are traditionally viewed from a Popperian (or hypo-deductive) philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
-However, error statistical view marries a statistical philosophy with specific statistical models and procedures.
-A marriage which leaves us with richer set of assumptions than necessary for the points we want to make about the value of preregistration.
-
 # References

### 2022-07-14 manuscript.Rmd
actual double spaceing

--- 2022-07-11 manuscript.Rmd
@@
 fontsize: 11pt
-linestretch: 1.2
+linestretch: 1.6
 geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"

### 2022-07-14 manuscript.Rmd
no indention in abstract

@@
 bibliography: references.bib
-abstract: "`r paste0(readLines('abstract.md'), collapse = ' ')`"
+abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
 csl: apa.csl

### 2022-07-14 manuscript.Rmd
remove comments
incorporate julias suggestions

@@
 
-<!--problems to solve:
-* I feel like I am describing a ruby goldberg machine here and expect the reader to understand what happens when you move the spoon
-* what about falsificationists, every time I write a paragraph relating to I delete it and try to move it a discussion
-* references!!!
--->
-
-A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
+A successful prediction should lend more credibility to a theory than a "postdiction," everything else being equal.
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
-Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
+Confusing exploratory findings with confirmed theories has led to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honor this distinction and produce results that are considered confirmatory.
 
@@
 Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistraion and how detailed this should be, or about the conditions that warrant deviating from it.
-If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
+Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistration and how detailed this information should be, or about the conditions that warrant deviations.
+If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavor. 
 We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
@@
 However, this postulate is not warranted.
-Researchers can conduct confirmatory research without preregisteration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
+Researchers can conduct confirmatory research without preregistration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
 The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
@@
 One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
-Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmfull all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter what the nature of the phenomenon under study.
+Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmful all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
@@
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-Equating confirmatory research with preregistered research is only possible for studies that expect low type I error rate and will run/be analyzed without changes no matter what [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule upholds, but such restricted use makes preregistration a niche solution unable to match the greater problem of replicability in psychology and elsewhere.
+Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without changes no matter whatin any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of lacking replicability in psychology and elsewhere.
 
 We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978 but assigned to the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to type I error rate.
+To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
 
 We then outline two interpretations of how theoretical risk is impacted by preregistration.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type I error.
-The second interpretation is our main contribution and demonstrates the broad applicability of preregistration for both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
+The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error.
+The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
 Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
@@
 Let us start by defining what we call expected epistemic value.
-If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude (truthlikeness).
+If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude.
 In other words, they hope to learn something from conducting the study.
@@
 This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to choosing a study to conduct.
+To make our three arguments, we must assume three things about this estimation process and how it relates to what studies to conduct.
 
@@
 Our rationale is as follows.
-Researchers who decide to conduct a study are akin to choosing a study to bet on.
+Researchers who decide to conduct a certain study are actually choosing a study to bet on.
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
 This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
-This argument states that any better must follow the axioms of probability to avoid being "irrational", i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in choosing a study.
+This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
+Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
 Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
@@
 
-# Epistemic Value and Theoretical Risk
+# Epistemic value and theoretical risk
 
-Our first argument is that theorethical risk is crucial for judging evidential support for theories.
+Our first argument is that theoretical risk is crucial for judging evidential support for theories.
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either exists or does not) since continuous evidence would lead to some quite involved derivations.
+We restrict ourselves to evidence of a binary nature (either it exists or does not) since continuous evidence would lead to some quite involved derivations.
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
 The probability of observing evidence under some hypothesis is $P(E|H)$.
-We can calculate the probability of the hypothesis after observing the evidence with help from the Bayes formula:
+We can calculate the probability of the hypothesis after observing the evidence with the help of the Bayes formula:
 
@@
 In short, we want to increase posterior probability ($P(H|E)$).
-Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, of which we want to maximize the expectation.
+Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, for which we want to maximize the expectation.
 So how can we increase posterior probability?
@@
 However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similar commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
+The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
 $P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-In other words, how probable is it that we "detect" that a theory holds, we therefore call it "detectability".
+We call this probability of detecting evidence given that the theory holds "detectability."
 Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant for the objective of preregistration.
+When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type-II error rate).
+However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
 Thus, $P(E)$ remains to be considered.
@@
 This tension between high risk leading to high reward is central to our consideration of preregistration.
-A high risk, high reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
-Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many if not most researchers operate.
+A high-risk, high-reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
+Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
 Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
@@
 Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theorethical risk, the evidence is less likely under the theory than it is when the theory does not hold.
+If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
@@
 This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove a hypothesis false ($P(¬E|¬H) = 0$, theoretical risk is zero), it can not be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
+He stated that if it is impossible to prove that a hypothesis is false ($P(¬E|¬H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
 18].
@@
 
-Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+Both theoretical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
 To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
 The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
-Suppose the primary theory is about the cognitive phenomenon of "insight".
-For the purpose of illustration, we define it somewhat hand-wavingly as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
-We pose the hypothesis that the following problem belongs to such class of insight problems:
+Suppose the primary theory is about the cognitive phenomenon of "insight."
+For the purpose of illustration, we defin it, with quite some hand-waving, as a cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
+We present the hypothesis that the following problem belongs to such a class of insight problems:
 
@@
 All experiments take a sample of ten psychology students.
-We present the students the problem for a brief span of time.
+We present the students with the problem for a brief span of time.
 After that, the three experiments differ as follows:
 
-1. the experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-2. the experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
-3. the experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
+1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
+2. The experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
+3. The experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
 
 We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($¬H$), than presenting the insight that roman literals might be used, should not lead to all students solving the problem ($¬E_1$); the experiment has therefore high theoretical risk ($P(¬E_1|¬H)$)
-Conversly, if insight is required to solve the problem ($H$), than it is probable to help all students to solve the problem ($E_1$); the experiment has therefore high detectability ($P(E_1|H)$).
+If "insight" has nothing to do with solving the problem ($¬H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($¬E_1$); the experiment therefore has high theoretical risk ($P(¬E_1|¬H)$)
+Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability ($P(E_1|H)$).
 The second experiment, on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
 Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
-With regard to detectability experiment 1 and 2 differ in no obvious way.
+With regard to detectability, experiments 1 and 2 differ in no obvious way.
 Experiment 3, however, also has low detectability.
-It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); the experiment 3 has therefore low detectability ($P(E_3|H)$).
-The theoretical risk, however, is also low in absolute terms but high compared to the theoretical risk.
-In the unlikely event that all 10 students lay the matches down in the form of the roman numeral Ⅷ ($E_3$) it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
+It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$) experiment 3 therefore has low detectability ($P(E_3|H)$).
+The theoretical risk, however, is also low in absolute terms, but high compared to the detectability.
+In the unlikely event that all 10 students place their matches to form the Roman numeral Ⅷ ($E_3$), it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
+
 As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
-We discuss the relation between statistical methods and theoretical risk in the section [Statistical Methods].
+We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.
 
@@
 
-# Preregistration as a Means to Increase Theoretical Risk?
+# Preregistration as a means to increase theoretical risk?
 
-After we discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
-Indeed, limiting type I error rate is commonly stated as a goal of preregistration.
+Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
+Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration.
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
@@
 Consider the case of a researcher contemplating the relation between two sets of variables.
-Suppose each set by itself is well studied, and strong theories tell the researcher how the variables within the set relate.
-However, our imaginary researcher considers the relation between these two sets.
+Suppose each set is separately well studied, and strong theories tell the researcher how the variables within the set relate.
+However, our imaginary researcher now considers the relation between these two sets.
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
 This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
-However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g.
-by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape etc.
-We will later show that preregistration increases the belief one can stake in the further specified theory, though it remains low till substantiated by testing it again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value without regard to the theory being tested.
+However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
+We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing it again.
+The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.
 
 Second, available resources limit theoretical risk.
-Increasing theorethical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
-In other words, one can not increase power while maintaining the same type I error rate without increasing the invested resources.
+Increasing theoretical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
+In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
 Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
@@
 
-# Uncertainty about Theoretical Risk
+# Uncertainty about theoretical risk
 
-We established that higher theoretical risk leads to more persuasive evidence.
-In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limit type I error rate) and thereby creates more persuasive evidence.
+We have established that higher theoretical risk leads to more persuasive evidence.
+In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limits the type-I error rate) and thereby creates more compelling evidence.
 Nevertheless, there are trade-offs for increasing theoretical risk.
@@
 
-## Statistical Methods
+## Statistical methods
 
-Theoretical risk is deeply connected with statistical methods, because its the inverse of $P(E|¬H)$.
-$P(E|¬H)$ is equivalent to the type I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill-hypothesis is $¬H$.
-Because many researchers are familiar with type I error rate, it can be helpful to remember this connection to theorethical risk.
-Researchers who choose a smaller type I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
-However, the research hypothesis is seldomly equal to the statistical null hypothesis, and therefore, the relation between statistical type I error rate and theoretical risk should not be over interpreted.
+Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|¬H)$.
+$P(E|¬H)$ is equivalent to the type-I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $¬H$.
+Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
+Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
+However, the research hypothesis seldomly equals  the statistical null hypothesis, and therefore, the relation between statistical type-I error rate and theoretical risk should not be overinterpreted.
 We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
@@
 Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to theoretical risk.
-Examples of our ability to exert this control are the setting of type I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria or cross-validation in machine learning.
+Examples of our ability to exert this control are the setting of type-I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
 These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
@@
 The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
-We can not hope to resolve this uncertainty; but we have to be aware of its implications.
-These are relayed in the following.
+We cannot hope to resolve this uncertainty; but we have to be aware of its implications.
+These are presented in the following.
 
-## Causes of Uncertainty
+## Causes of uncertainty
 
@@
 However, other factors in the broader context of the study are much harder to quantify.
-If one chooses to focus only on the contribution of statistical methods to theorethical risk, one is bound to overestimate it.
+If one chooses to focus only on the contribution of statistical methods to theoretical risk, one is bound to overestimate it.
 Take, for example, a t-test.
-Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type I error rate.
+Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type-I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
 They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelhood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors must be added to $P(E|¬H)$ and, hence, be subtracted from theoretical risk.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying hypothesis being false.
+Even if the t-test itself perfectly maintains its type-I error rate, these factors must be added to $P(E|¬H)$ and, hence, be subtracted from theoretical risk.
 While, in theory, these factors may leave theoretical risk unaffected or even increase it, we argue that this is not the case in practice.
-Whether researchers want to or not, except under strict blinding, they continuously process information about how the study is going.
+Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
 While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-The only thing we can conclude with some certainty is that theorethical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
+The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
 
@@
 
-Before we ask how preregistration is influencing this uncertainty, we must consider the implications of being uncertain about the theoretical risk.
+Before we ask how preregistration influences this uncertainty, we must consider the implications of being uncertain about the theoretical risk.
 Within the Bayesian framework, this is both straightforward and insightful.
@@
 
-To illustrate the effect of uncertainty, let $p(E|H) = .8$ (e.g., power of 80%) and $p(H) = .1$ and assume a uniform distribution for $p(E|\neg{}H)$ of the form:
+To illustrate the effect of uncertainty, let $p(E|H) = 0.8$ (e.g., power of 80%) and $p(H) = 0.1$ and assume a uniform distribution for $p(E|\neg{}H)$ of the form:
 
@@
 
-Where $\tau{}$ is the upper bound of theoretical risk (e.g., .95 for a statistical model with a nominal type I error rate of 5%).
+Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
 
-We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theorethical rsik (and since its a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore with our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and thus conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
@@
   pm = list(
-    name = "Posterior",
+    name = "Posterior Probability",
     formula = "\\frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}",
@@
 #' 
-#' If theorethical risk is maximally uncertain, we represent this uncertainty
+#' If theoretical risk is maximally uncertain, we represent this uncertainty
 #' as a uniform distribution from zero to tau from which we sample and average the measures.
@@
 ```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Posterior (corroboration as firmness) as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "Posterior Probability (corroboration as firmness) as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
 
-plots[["Posterior"]] +
-  labs(caption = "$p(E|H) = .8$, $p(H) = .1$",
+plots[["Posterior Probability"]] +
+  labs(caption = "$p(E|H) = 0.8$, $p(H) = 0.1$",
        y = "$\\mathbb{E}\\big(p(H|E)\\big)$") +
@@
 
-plots[!(names(plots) %in% "Posterior")] %>% 
+plots[!(names(plots) %in% "Posterior Probability")] %>% 
   reduce(`+`) +
@@
 
-Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions that all realistic conditions fall in.
-Depending on the distribution that expresses our state of knowledge about the theoretical risk a study took on, our expectation for what to gain varies considerably.
-Uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
+Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions into which all realistic conditions fall.
+Depending on how uncertain we are about the theoretical risk a study took on, our expectation the gained epistemic value varies considerably.
+Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
 Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
 
-# Preregistration as a Means to Decreasing Uncertainty about the Theoretical Risk
+# Preregistration as a means to decrease uncertainty about the theoretical risk
 
@@
 Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
-The last argument is that preregistrations reduces this uncertainty.
+The third and last argument is that preregistrations reduce this uncertainty.
 
@@
 2. They expect other researchers to apply the same rational process.
-2. All else equal, researchers try to increase the expected epistemic value for other researchers.
+2. All else being equal, researchers try to increase the expected epistemic value for other researchers.
 
-The point we make with these assumption is that the authors aim to persuade other researchers, their readers.
-Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of the resulting article (or, more generally, the consumer of the research product).
+The point we make with these assumptions is that researchers aim to persuade other researchers, the readers of their articles.
 We have to remember that the process of weighing evidence for or against a theory extends beyond the original authors to all the people they hope to persuade.
+Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of a resulting article (or, more generally, the consumer of the research product).
 While the authors may have deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
-In particular, they never know what relevant factors the authors, intentionally or not, failed to disclose.
+In particular, they never know which relevant factors the authors of a given article failed to disclose, intentionally or not.
 From the perspective of an ultimate sceptic, they may claim maximum uncertainty.
 
-Communicating clearly what the authors did to arrive at the evidence is crucial for judging the theoretical risk they took.
+Communicating clearly what authors did to arrive at the evidence they present is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
-The authors could have decided to exclude a number of analytic strategies they tried out etc.
-That is not to say that any study that was not preregistered was subjected to practices of p hacking.
-The point is, we can not exclude this and a myriad of other possibilities and, hence, are left uncertain.
-If the researcher do describe what they intend to do beforehand, and then report they did exactly that, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took but much less so.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses etc.
-In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+For instance, the authors could have decided to exclude a number of analytic strategies they tried out.
+That is not to say that every study that was not-preregistered was subjected to practices of p hacking.
+The point is, that we cannot exclude this and innumerable possibilities and, hence, are left uncertain.
+If the researchers have described what they intended to do beforehand, and then report that they did exactly that, the readers can be certain, that they have received a complete account of the situation.
+They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extend than if the study would not have been preregistered.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses, etc.
+In any case, a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
 
@@
 
-We started out with the observation that preregistrations do no always cleanly divide preregistration from confirmation.
-Enforcing this distinction would mean to use preregistrations as a means to increase the theorethical risk researcher take on.
+We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
+Enforcing this distinction would mean using preregistrations as a means to increase the theoretical risk researchers take on.
 No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
@@
 As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still precious for the scientific process.
+Let us note that a study that does not result in the evidence a researcher is hoping for is still valuable for the scientific process.
 Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
@@
 First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
-That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
+That is, researchers benefit from preregistering their study, regardless of whether the study is "exploratory" or not.
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
-Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
-Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
-This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might believe it to be a likely explanation for the deviation.
-If the readers find the reason for the deviation likely to be something other than decreasing the theorethical risk, their uncertainty about the theorethical risk should only increase modestly.
+Second, if researchers realize after the fact that it would be unwise to follow their preregistration in certain details, they may change their strategy.
+Allowing researchers to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
+In its fullest consequence, this argument might result in the same uncertainty that we argued plagues for not-preregistered studies.
+However, if the authors offer a convincing argument for their deviation, their readers might may judge it to be a reasonable explanation for the deviation.
+In fact, using the here presented framework, we can distinguish reasonable from unreasonable deviations.
+If a change increases detectability enough to offset the increase in uncertainty in theoretical risk, it is worthwhile.
 

### 2022-07-27 manuscript.Rmd
typo

@@
 Suppose the primary theory is about the cognitive phenomenon of "insight."
-For the purpose of illustration, we defin it, with quite some hand-waving, as a cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
+For the purpose of illustration, we define it, with quite some hand-waving, as a cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
 We present the hypothesis that the following problem belongs to such a class of insight problems:

### 2022-08-09 manuscript.Rmd

@@
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without changes no matter whatin any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without in any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of lacking replicability in psychology and elsewhere.
@@
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-While researchers can not know what exactly they will learn from a study, they can form an expectation that helps them decide which study to conduct.
+The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory (which itself is an ontologocial concept) is what we call epistemic value.
+Researchers cannot know what exactly they will learn from a study before they have run it.
+However, they can develop, they can form an expectation that helps them decide which study to conduct in what manner.
 This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to what studies to conduct.
+To make our three arguments, we must assume three things about this estimation process and how it relates to what studies (preregistered vs not preregistered) to conduct.
 
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either it exists or does not) since continuous evidence would lead to some quite involved derivations.
+We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
@@
 The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
+If a hypothesis is more probable to begin with, the same increase in corroboration will lead to a higher probability than if the study was less probable to begin with.
 However, the prior probability of a hypothesis is nothing our study design can change.
@@
 
-This sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
+The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
 He stated that if it is impossible to prove that a hypothesis is false ($P(¬E|¬H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
@@
 
-## Causes of uncertainty
+## Sources of Uncertainty
 
@@
 If one chooses to focus only on the contribution of statistical methods to theoretical risk, one is bound to overestimate it.
-Take, for example, a t-test.
+Take, for example, a t-test  of mean differences in two samples.
 Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type-I error rate.
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item, enlarge their definition of the population to be sampled, translate their questionnaires, impute missing values, or any number of other things.
-All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying hypothesis being false.
-Even if the t-test itself perfectly maintains its type-I error rate, these factors must be added to $P(E|¬H)$ and, hence, be subtracted from theoretical risk.
-While, in theory, these factors may leave theoretical risk unaffected or even increase it, we argue that this is not the case in practice.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, or any number of other things.
+All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
+Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
+While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
 While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
+We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
 The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
@@
 
-We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and thus conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
@@
 
-We hope to have persuaded you to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
-Second, the theoretical risk is inherently uncertain, which diminishes the persuasiveness of the gathered evidence.
+We hope to have persuaded the reader to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
+Second, the theoretical risk is inherently uncertain and the degree of uncertainty diminishes the persuasiveness of the gathered evidence.
 The third and last argument is that preregistrations reduce this uncertainty.
@@
 
-Communicating clearly what authors did to arrive at the evidence they present is crucial for judging the theoretical risk they took.
+Communicating clearly how the authors gathered the data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
@@
 The point is, that we cannot exclude this and innumerable possibilities and, hence, are left uncertain.
-If the researchers have described what they intended to do beforehand, and then report that they did exactly that, the readers can be certain, that they have received a complete account of the situation.
+This uncertainty is drastically reduced, if the researchers have described what they intended to do beforehand, and then report that they did exactly that.
+In that case, the readers can be certain, that they have received a complete account of the situation.
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extend than if the study would not have been preregistered.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms, the probability of plain error in analyses, etc.
-In any case, a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analysies, etc.
+In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
 
@@
 No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as the term implies, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory.
-As we showed, minimizing the uncertainty around the theoretical risk shares the beneficial properties without increasing the likelihood of a "failed" study.
-Let us note that a study that does not result in the evidence a researcher is hoping for is still valuable for the scientific process.
+However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
+As we further argued, minimizing the uncertainty about theoretical risk (without necessarily increasing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
+Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
 Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
 
-The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
+In our view, the advantage of uncertainty reduction about theoretical risk front and center is two-fold.
 First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
-That is, researchers benefit from preregistering their study, regardless of whether the study is "exploratory" or not.
+That is, researchers benefit from preregistering their study, regardless of whether or not the study is "exploratory".
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.

### 2022-08-09 manuscript.Rmd
remove double spacing

@@
 fontsize: 11pt
-linestretch: 1.6
+linestretch: 1.2
 geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"

### 2022-08-09 manuscript.Rmd
Update manuscript.Rmd

--- 2022-08-09 manuscript.Rmd
@@
 A successful prediction should lend more credibility to a theory than a "postdiction," everything else being equal.
+<!-- ME: what is a postdiction?
+I think this word is to difficult if it is not a technical term -->
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
@@
 
-Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
+Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
 However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
@@
 If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavor.
+<!-- what does efficiency mean in this case?
+Maybe there is a better word -->
 We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
@@
 The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
+Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered. <!-- I dont think it does. I think it implies research can be 
 However, this postulate is not warranted.

### 2022-08-09 manuscript.Rmd
Update manuscript.Rmd

@@
 The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered. <!-- I dont think it does. I think it implies research can be 
+Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
+<!-- This sounds too much like a strawman to me.
+I assume it implies more that research can be recognized as confirmatory.
+-->
 However, this postulate is not warranted.
@@
 The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
-One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
+Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
 Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter what the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmful all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
-These issues arise from a fundamental confusion about the objective of preregistration.
-Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
+These issues arise from a fundamental missunderstanding of the objective of preregistration.
+Researchers are bound to be misled when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
+<!-- I dont like "confused" -->
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
 Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without in any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+<!-- I dont understand this sentence -->
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of lacking replicability in psychology and elsewhere.

### 2022-08-19 manuscript.Rmd
Update manuscript.Rmd

@@
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
-Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter what the nature of the phenomenon under study.
+Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
 Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
+<!-- why is it called theoretical risk when it is about proving a hypothesis wrong?
+-->
 
@@
 
-To arrive at this interpretation, we rely on three arguments.
-The first is that theoretical risk is vital for judging evidential support for theories.
-The second argument is that the theoretical risk for a given study is generally uncertain.
-The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
+<!-- you don't rely on all of those, I would not state this here -->
 
@@
 Researchers cannot know what exactly they will learn from a study before they have run it.
-However, they can develop, they can form an expectation that helps them decide which study to conduct in what manner.
+However, they can develop an expectation that helps them to decide which study to conduct in what manner.
 This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to what studies (preregistered vs not preregistered) to conduct.
-
-1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.
-
-The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
-Our rationale is as follows.
-Researchers who decide to conduct a certain study are actually choosing a study to bet on.
-They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
-This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
-Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
-The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
-The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
-Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is intentionally reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+<!-- not compelling, deleted -->
 
@@
 Let us make some simplifying assumptions and define notation.
-We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
+To avoid unnessesary technicality, we restrict ourselves to evidence of a binary nature.
+<!-- I am not shure if this is good.
+Even in NHS, there is no "binary evidence".
+In fact, I can't think of any example except for a single coinflip.
+Perhaps you mean non-randomized statistical tests.
+-->
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
@@
 
-The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
-We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
+The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness.
+We do not discuss specific measures of corroboration since it is generally expected that any measure of corroboration increases monotonically with an increase in the posterior probability $P(H|E)$.
+<!-- citation missing -->
 
-In short, we want to increase posterior probability ($P(H|E)$).
-Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, for which we want to maximize the expectation.
+In short, we want to increase the posterior probability $P(H|E)$.
+Increases in the posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation. <!-- very complicated sentence -->
 So how can we increase posterior probability?
@@
 Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high reward.
 
 If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
+<!-- I believe this is the wrong conceptualization.
+Riskiness refers to the probability of obtaining the evidence given the Hypothesis is false, not to the total probability of the evidence.
+I would swap Eq.
+1 for a version that already entails Eq.
+2 -->
 This tension between high risk leading to high reward is central to our consideration of preregistration.
 A high-risk, high-reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
+<!-- I think this is not true.
+A high-risk strategy means conducting studies where it is unlikely to observe the evidence, given the hypothesis is wrong.
+However, if the hypothesis is true, and P(E|H) = 1, there are no losses, even in risky studies.
+-->
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.

### 2022-08-19 manuscript.Rmd
Update manuscript.Rmd

@@
 
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
+We already have noted that there is not much to do about prior probability ($P(H)$ and that it is common sense to increase detectability ($P(E|H)$).
+<!-- the question is not whether it is common sense, but whether it has something to do with preregistration -->
 The real lever to pull is, therefore, $P(E|¬H)$.
@@
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+<!-- I am not shure if thats why we call it theoretical risk.
+I am also not shure if this is gramatically sensible.
+-->
 We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
@@
 First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
-Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
+Second, if the theoretical risk is smaller than the detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
 If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
-In other words, observing a sure fact does not lend credence to a hypothesis.
+In other words, observing a fact that would also be observed if the hypotheses was wrong does not lend credence to the hypothesis.
 
@@
 Both theoretical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+<!-- I dont understand this sentence -->
 To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it.
-The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
+These examples were created to illustrate how experiments may differ in their theoretical risk and detectability.
 Suppose the primary theory is about the cognitive phenomenon of "insight."
@@
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
-
+<!-- I think this example needs to be swapped for a better one.
+For example: H: A coin is fair.
+E1: Observe if the person throwing the coin is betwenn 0 and 10 meters in height (high detectability, no theoretical risk).
+E2: Throw the coin 2 times and observe if it lands one time on heads and one time on
 As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.

### 2022-08-19 manuscript.Rmd
Update manuscript.Rmd

@@
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
-<!-- I think this example needs to be swapped for a better one.
+<!-- I think this example needs to be swapped for a simpler one.
 For example: H: A coin is fair.
 E1: Observe if the person throwing the coin is betwenn 0 and 10 meters in height (high detectability, no theoretical risk).
-E2: Throw the coin 2 times and observe if it lands one time on heads and one time on
+E2: Throw the coin 2 times and observe if it lands one time on heads and one time on tails (detectability: P(E|H) = 0.5, theoretical risk: P(E|p != 0.5) = 1/3) E3: throw the coin 200 times, compute #heads/#200, construct a test (high power -> high detectability, high theoretical risk).
+-->
 As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
+<!-- what is non-probabilistic evidence supposed to be?
+Evidence is always a random variable -->
 We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.

### 2022-08-19 manuscript.Rmd
Update manuscript.Rmd

@@
 The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.
+<!-- I think the example is too theoretical and i dont understand the conclusions drawn from it -->
 
@@
 However, the research hypothesis seldomly equals  the statistical null hypothesis, and therefore, the relation between statistical type-I error rate and theoretical risk should not be overinterpreted.
-We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+<!-- what is it now, the null or the alternative?
+-->
+We argue that theoretical risk also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
@@
 Examples of our ability to exert this control are the setting of type-I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none.
 
@@
 The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
+<!-- why are there no factors that increase theoretical risk?
+-->
 

### 2022-08-24 manuscript.Rmd
remove term "postdiction" closes #15

--- 2022-08-19 manuscript.Rmd
@@
 
-A successful prediction should lend more credibility to a theory than a "postdiction," everything else being equal.
-<!-- ME: what is a postdiction?
-I think this word is to difficult if it is not a technical term -->
+A theory able to generate successful prediction should be more credible than a theory build around the facts afterwards.
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
@@
 
-Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
+Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
 However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
@@
 If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavor. 
-<!-- what does efficiency mean in this case?
-Maybe there is a better word -->
 We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
@@
 Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-<!-- This sounds too much like a strawman to me.
-I assume it implies more that research can be recognized as confirmatory.
--->
 However, this postulate is not warranted.
@@
 The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
-Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
-Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
+One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
+Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter what the nature of the phenomenon under study.
+Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmful all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
-These issues arise from a fundamental missunderstanding of the objective of preregistration.
-Researchers are bound to be misled when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
-<!-- I dont like "confused" -->
+These issues arise from a fundamental confusion about the objective of preregistration.
+Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
 Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without in any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-<!-- I dont understand this sentence -->
 Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of lacking replicability in psychology and elsewhere.
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
-<!-- why is it called theoretical risk when it is about proving a hypothesis wrong?
--->
 
@@
 
-<!-- you don't rely on all of those, I would not state this here -->
+To arrive at this interpretation, we rely on three arguments.
+The first is that theoretical risk is vital for judging evidential support for theories.
+The second argument is that the theoretical risk for a given study is generally uncertain.
+The third and last argument is that this uncertainty is reduced by applying preregistration.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
 
@@
 Researchers cannot know what exactly they will learn from a study before they have run it.
-However, they can develop an expectation that helps them to decide which study to conduct in what manner.
+However, they can develop, they can form an expectation that helps them decide which study to conduct in what manner.
 This expectation is what we term expected epistemic value.
-<!-- not compelling, deleted -->
+To make our three arguments, we must assume three things about this estimation process and how it relates to what studies (preregistered vs not preregistered) to conduct.
+
+1. Researchers judge the evidence for or against a hypothesis rationally.
+2. They expect other researchers to apply the same rational process.
+2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.
+
+The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
+Our rationale is as follows.
+Researchers who decide to conduct a certain study are actually choosing a study to bet on.
+They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
+This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
+This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
+Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
+Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
+The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
+The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
+Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
+In fact, this conceptualization is intentionally reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 
@@
 Let us make some simplifying assumptions and define notation.
-To avoid unnessesary technicality, we restrict ourselves to evidence of a binary nature.
-<!-- I am not shure if this is good.
-Even in NHS, there is no "binary evidence".
-In fact, I can't think of any example except for a single coinflip.
-Perhaps you mean non-randomized statistical tests.
--->
+We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
@@
 
-The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness.
-We do not discuss specific measures of corroboration since it is generally expected that any measure of corroboration increases monotonically with an increase in the posterior probability $P(H|E)$.
-<!-- citation missing -->
+The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
+We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
 
-In short, we want to increase the posterior probability $P(H|E)$.
-Increases in the posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation. <!-- very complicated sentence -->
+In short, we want to increase posterior probability ($P(H|E)$).
+Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, for which we want to maximize the expectation.
 So how can we increase posterior probability?
@@
 Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+In other words, high risk, high reward.
 
 If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
-<!-- I believe this is the wrong conceptualization.
-Riskiness refers to the probability of obtaining the evidence given the Hypothesis is false, not to the total probability of the evidence.
-I would swap Eq.
-1 for a version that already entails Eq.
-2 -->
 This tension between high risk leading to high reward is central to our consideration of preregistration.
 A high-risk, high-reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
-<!-- I think this is not true.
-A high-risk strategy means conducting studies where it is unlikely to observe the evidence, given the hypothesis is wrong.
-However, if the hypothesis is true, and P(E|H) = 1, there are no losses, even in risky studies.
--->
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
@@
 
-We already have noted that there is not much to do about prior probability ($P(H)$ and that it is common sense to increase detectability ($P(E|H)$).
-<!-- the question is not whether it is common sense, but whether it has something to do with preregistration -->
+We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
 The real lever to pull is, therefore, $P(E|¬H)$.
@@
 Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-<!-- I am not shure if thats why we call it theoretical risk.
-I am also not shure if this is gramatically sensible.
--->
 We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
@@
 First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
-Second, if the theoretical risk is smaller than the detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
+Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
 If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
-In other words, observing a fact that would also be observed if the hypotheses was wrong does not lend credence to the hypothesis.
+In other words, observing a sure fact does not lend credence to a hypothesis.
 
@@
 Both theoretical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
-<!-- I dont understand this sentence -->
 To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
-These examples were created to illustrate how experiments may differ in their theoretical risk and detectability.
+The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
 Suppose the primary theory is about the cognitive phenomenon of "insight."
@@
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
-<!-- I think this example needs to be swapped for a better one.
-For example: H: A coin is fair.
-E1: Observe if the person throwing the coin is betwenn 0 and 10 meters in height (high detectability, no theoretical risk).
-E2: Throw the coin 2 times and observe if it lands one time on heads and one time on
+
 As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.

### 2022-09-11 manuscript.Rmd
use repro to generate dockerfile

@@
 csl: apa.csl
+repro:
+  files: abstract.md
+  packages:
+    - tidyverse
+    - ggplot2
+    - ggthemes
+    - patchwork
+    - tikzDevice
+  tlmgr:
+    - pgf
+    - preview
+    - libertine
+  apt:
+    - git
+    - rsync
+    - curl
+    - inkscape
 ---
@@
 Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistration and how detailed this information should be, or about the conditions that warrant deviations.
-If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavor. 
+If the research community remains confused about the purpose of preregistration, it is bound to misapply it and waste resources on writing preregistrations that do not address the underlying problem.
 We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
@@
 The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-Taken at face value, this objective implies that research is confirmatory if, and only if, it is preregistered.
-However, this postulate is not warranted.
-Researchers can conduct confirmatory research without preregistration, though it might be difficult to convince other researchers of the confirmatory nature of their research.
+However, it is not principally warranted to suggest a close coupling of the categories preregistered vs. non-preregistered and confirmatory vs. exploratory research.
+In fact, researchers can conduct confirmatory research without preregistration (though it might be difficult to convince other researchers of the confirmatory nature of their research).
 The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations.
-One strategy is to write a loose preregistration to begin with [@stefanBigLittleLies2022]; another is to deviate from the preregistration afterwards.
-Both strategies may be used with compelling scientific reasoning or with the self-serving intent of generating desirable results no matter what the nature of the phenomenon under study.
-Insisting on equating preregistration and confirmation has, hence, led to criticism that preregistration is actually harmful all things considered, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
+Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
+Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
 
-These issues arise from a fundamental confusion about the objective of preregistration.
-Researchers are bound to be confused when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
+These issues arise from a fundamental misunderstanding about the objective of preregistration.
+Researchers are bound to be misled when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
 Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
+To force preregistrations to generate results that are rightfully considered confirmatory additional restrictions on the types of analyses allowed have to be imposed.
+First, the analysis must have a low type-I error rate. second, 
 Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without in any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].

### 2022-09-12 manuscript.Rmd
move plot to own script

@@
 repro:
-  files: abstract.md
+  scripts:
+    - R/plots.R
+  files:
+    - abstract.md
+    - data/simulation.rds
   packages:
@@
 )
+knitr::read_chunk("R/plots.R")
 
@@
 
-```{r, include=FALSE}
-bayes <- function(h, eh, enh)(h * eh)/((h * eh) + ((1 - h) * enh))
-```
-
 # Preregistration as a means to increase theoretical risk?
@@
 
-```{r, include=FALSE}
-forbid_additional_args <- function(...)stopifnot(length(list(...)) == 0L)
-
-#' Apply the Bayes Theorem
-#'
-#' @param h The prior probability of the hypothesis \eqn{P(H)}.
-#' @param detectability The probability of detecting evidence, if hypothesis holds \eqn{P(E|H)}.
-#' @param theoretical_risk The probability of not detecting evidence, if hypothesis does not hold \eqn{P(¬E|¬H)}.
-#' @param tau Same as `theoretical_risk` for compatibility.
-#' @return The posterior probability of the hypothesis after observing evidence \eqn{P(H|E)}.
-bayes <- function(h, detectability, theoretical_risk, ...) {
-  forbid_additional_args(...)
-  true_positive <- h * detectability
-  false_positive <- (1 - h) * (1 - theoretical_risk)
-  true_positive / (true_positive + false_positive)
-}
-
-# all formulas taken from Bayesian Philosophy of Science, p. 51
-# first letter = function name in Bayesian Philosophy of Science
-# optional letter = p for prime
-# second letter = m for measure
-measures <- list(
-  pm = list(
-    name = "Posterior Probability",
-    formula = "\\frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}",
-    fun = bayes
-  ),
-  dm = list(
-    name = "Difference Measure",
-    formula = "d(H, E) = p(H|E) - p(H)",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      he - h
-    }
-  ),
-  rm = list(
-    name = "Log-Ratio Measure",
-    formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      log(he / h)
-    }
-  ),
-  lm = list(
-    name = "Log-Likelihood Measure",
-    formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      enh <- 1 - theoretical_risk
-      log(he / enh)
-    }
-  ),
-  km = list(
-    name = "Kemeny–Oppenheim Measure",
-    formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      eh <- detectability
-      enh <- 1 - theoretical_risk
-      (eh - enh) / (eh + enh)
-    }
-  ),
-  zm = list(
-    name = "Generalized Entailment Measure",
-    formula =
-      "z(H, E) =\\begin{cases}
-  \\frac{p(E|H) - p(H)}{1 - p(H)}, & \\text{if } p(H|E) \\geq{} p(H) \\\\
-  \\frac{p(E|H) - p(H)}{p(H)}, & \\text{if } p(H|E) < p(H)
-\\end{cases}",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      ifelse(he >= h, (he - h) / (1 - h), (he - h) / h)
-    }
-  ),
-  sm = list(
-    name = "Christensen–Joyce Measure",
-    formula = "s(H, E) = p(H|E) - p(H|\\neg{}E)",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      enh <- 1 - theoretical_risk
-      eh <- detectability
-      e <- h * eh + (1 - h) * enh
-      hne <- (h * (1 - eh)) / (1 - e)
-      he - hne
-    }
-  ),
-  cpm = list(
-    name = "Carnap’s Relevance Measure",
-    formula = "c'(H, E) = p(E)(p(H|E) - p(H))",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      enh <- 1 - theoretical_risk
-      e <- h * detectability + (1 - h) * enh
-      e * (he - h)
-    }
-  ),
-  rpm = list(
-    name = "Rips Measure",
-    formula = "r'(H, E) = 1 - \\frac{p(\\neg{}H|E)}{p(\\neg{}H)}",
-    fun = function(h, detectability, theoretical_risk, ...) {
-      forbid_additional_args(...)
-      he <- bayes(h, detectability, theoretical_risk)
-      1 - (1 - he) / (1 - h)
-    }
-  )
-)
-
-measures <- tibble(!!!transpose(measures)) %>% 
-  mutate(name = simplify(name, "chr"),
-         formula = simplify(formula, "chr"))
-
-#' Integrate over uncertain theoretical risk
-#' 
-#' If theoretical risk is maximally uncertain, we represent this uncertainty
-#' as a uniform distribution from zero to tau from which we sample and average the measures.
-#'
-#' @param tau The guaranteed upper bound of theoretical risk.
-#' @param n Number of samples from the uniform distribution.
-#' @param seed Optional seed to set before execution. Is reset to global state on exit.
-#' @return A function similar to the original
-uncertainly <- function(measure, tau, n = 1e5, seed = NULL){
-  function(h, detectability, ...){
-    forbid_additional_args(...)
-    if(!is.null(seed)){
-      if(exists(".Random.seed"))on.exit(set.seed(.Random.seed))
-      set.seed(seed)
-    }
-    theoretical_risk_samples <- runif(n, 0, tau)
-    mean(measure(h, detectability, theoretical_risk_samples))
-  }
-}
-certainly <- function(measure, tau){
-  function(h, detectability, ...){
-    forbid_additional_args(...)
-    measure(h, detectability, tau)
-  }
-}
+```{r}
+results <- readr::read_rds(here::here("data", "simulation.rds"))
 ```
 
-```{r, cache=TRUE, echo=FALSE}
-step <- 0.01
-
-grid <-
-  tidyr::expand_grid(
-    h = .1,
-    detectability = .8,
-    tau = seq(step, 1 - step, step)
-  )
-apply_on_grid <- function(f, grid, seed = NULL, n = 1e4) {
-  certain <- function(h, detectability, tau, ...){
-    forbid_additional_args(...)
-    certainly(f, tau)(h, detectability)
-  }
-  uncertain <- function(h, detectability, tau, ...){
-    forbid_additional_args(...)
-    uncertainly(f, tau, seed = seed, n = n)(h, detectability)
-  }
-  mutate(grid,
-         certain = pmap_dbl(grid, certain),
-         uncertain = pmap_dbl(grid, uncertain))
-}
-
-results <- mutate(measures, results = map(fun, apply_on_grid, grid)) %>% 
-  select(-fun) %>% 
-  unnest(results) %>% 
-  pivot_longer(c(certain, uncertain),
-               names_to = "certainty",
-               values_to = "value")
-```
-
-```{r, cache = TRUE}
-plots <- results %>%
-  mutate(name = str_remove(name, "Measure")) %>%
-  split(., .$name) %>%
-  imap(
-    ~ ggplot() +
-      geom_line(aes(tau, value, linetype = certainty), data = .x) +
-      labs(title = .y,
-           y = element_blank(),
-           x = "$\\tau$") +
-      geom_ribbon(
-        data = pivot_wider(.x, names_from = certainty, values_from = value),
-        aes(x = tau, ymin = uncertain, ymax = certain),
-        alpha = .1
-      ) +
-      theme_tufte() +
-      NULL
-  )
+```{r measure-plots}
+# see R/plots.R
 ```

### 2022-09-12 R/plots.R

@@ -1,0 +1,21 @@
+library(tidyverse)
+source(here::here("simulate.R"))
+
+#----measure-plots----
+plots <- results %>%
+  mutate(name = str_remove(name, "Measure")) %>%
+  split(., .$name) %>%
+  imap(
+    ~ ggplot() +
+      geom_line(aes(tau, value, linetype = certainty), data = .x) +
+      labs(title = .y,
+           y = element_blank(),
+           x = "$\\tau$") +
+      geom_ribbon(
+        data = pivot_wider(.x, names_from = certainty, values_from = value),
+        aes(x = tau, ymin = uncertain, ymax = certain),
+        alpha = .1
+      ) +
+      theme_tufte() +
+      NULL
+  )

### 2022-09-12 R/simulate.R

@@ -1,0 +1,122 @@
+library("tidyverse")
+source(here::here("R", "funs.R"))
+#----measures----
+# all formulas taken from Bayesian Philosophy of Science, p. 51
+# first letter = function name in Bayesian Philosophy of Science
+# optional letter = p for prime
+# second letter = m for measure
+measures <- list(
+  pm = list(
+    name = "Posterior Probability",
+    formula = "\\frac{P(H)P(E|H)}{P(H)P(E|H) + P(¬H)P(E|¬H)}",
+    fun = bayes
+  ),
+  dm = list(
+    name = "Difference Measure",
+    formula = "d(H, E) = p(H|E) - p(H)",
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      he - h
+    }
+  ),
+  rm = list(
+    name = "Log-Ratio Measure",
+    formula = "r(H, E) = log(\\frac{p(H|E)}{p(H)})",
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      log(he / h)
+    }
+  ),
+  lm = list(
+    name = "Log-Likelihood Measure",
+    formula = "l(H, E) = log(\\frac{p(H|E)}{p(E|\\neg{}H)})",
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      he <- bayes(h, detectability, theoretical_risk)
+      enh <- 1 - theoretical_risk
+      log(he / enh)
+    }
+  ),
+  km = list(
+    name = "Kemeny–Oppenheim Measure",
+    formula = "k(H, E) = \\frac{p(E|H) - p(E|\\neg{}H)}{p(E|H) + p(E|\\neg{}H)}",
+    fun = function(h, detectability, theoretical_risk, ...) {
+      forbid_additional_args(...)
+      eh <- detectability
+      enh <- 1 - theoretical_risk
+      (eh - enh) / (eh + enh)
+    }
+  ),
+  zm = list(
+    name = "Generalized Entailment Measure",
+    formula =
+      "z(H, E) =\\begin{cases}
+  \\frac{p(E|H) - p(H)}{1 - p(H)}, & \\text{if } p(H|E) \\geq{} p(H) \\\\
[…]

### 2022-09-12 manuscript.Rmd
add here package

--- 2022-09-11 manuscript.Rmd
@@
   packages:
+    - here
     - tidyverse

### 2022-09-12 manuscript.Rmd
seperate theme

--- 2022-09-12 manuscript.Rmd
@@
        y = "$\\mathbb{E}\\big(p(H|E)\\big)$") +
+  theme_tufte() +
   theme(legend.position = c(.2, .8),
@@
 plots[!(names(plots) %in% "Posterior Probability")] %>% 
+  map(~ . +   theme_tufte()) %>% 
   reduce(`+`) +

### 2022-09-12 R/plots.R

@@ -16,6 +16,5 @@ plots <- results %>%
         aes(x = tau, ymin = uncertain, ymax = certain),
         alpha = .1
       ) +
-      theme_tufte() +
       NULL
   )

### 2022-09-12 manuscript.Rmd
add apa.csl to dependencies

@@
     - data/simulation.rds
+    - apa.csl
   packages:

### 2022-09-12 README.Rmd
add link to pdf to readme closes #2
cite meehls paper more widely
closes #31

@@
 output: github_document
+bibliography: references.bib
 ---
@@
 
-```{r, echo=FALSE, results='asis'}
+```{r abstract, echo=FALSE, results='asis'}
 cat(readLines('abstract.md'))
 ```
+
+## [🗎 PDF of draft 🗎](https://aaronpeikert.github.io/bayes-prereg/manuscript.pdf)
+
+## [Slides for Talk](https://aaronpeikert.github.io/bayes-prereg/presentation.html)
+
+## References

### 2022-09-12 manuscript.Rmd

--- 2022-09-12 manuscript.Rmd
@@
     - data/simulation.rds
+    - apa.csl
   packages:
+    - here
     - tidyverse

### 2022-09-12 README.Rmd
update readme

@@
 
+## [🗎 Manuscript draft 🗎](https://aaronpeikert.github.io/bayes-prereg/manuscript.pdf) / [Talk slides](https://aaronpeikert.github.io/bayes-prereg/presentation.html)
+
 ```{r abstract, echo=FALSE, results='asis'}
@@
 
-## [🗎 PDF of draft 🗎](https://aaronpeikert.github.io/bayes-prereg/manuscript.pdf)
-
-## [Slides for Talk](https://aaronpeikert.github.io/bayes-prereg/presentation.html)
-
 ## References

### 2022-12-05 manuscript.Rmd
streamline intro. what problem is solved?

@@
 
+<!--introduce confirmatory vs. exploratory-->
 A theory able to generate successful prediction should be more credible than a theory build around the facts afterwards.
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
-Confusing exploratory findings with confirmed theories has led to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honor this distinction and produce results that are considered confirmatory.
+Mistaking exploratory findings for empirically confirmed results has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to observe this distinction and produce results that are considered confirmatory.
+Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
+This ability to provide a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
+Still an overwhelming majority of empirical researchers do not routinely preregister their studies because they do not find that the theoretical advantages outweigh the practical hurdles.
+This suggests a demand for explicating what the objective of preregistration is and what such objective implies in practice.
 
-Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
-However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
-These skeptics question why a study should be preregistered at all.
-Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
-Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistration and how detailed this information should be, or about the conditions that warrant deviations.
-If the research community remains confused about the purpose of preregistration, it is bound to misapply it and waste resources on writing preregistrations that do not address the underlying problem.
-We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.
+<!--the current goal of preregistration is confirmation-->
+One objective of preregistration that has received widespread attention, is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+In such narrative, preregistration is justified by a confirmatory research agenda.
+However, two problems emerge under closer inspection.
+First, many researcher do not subscribe to a confirmatory research agenda.
+Second, it is not principally warranted to suggest a close coupling of the categories preregistered vs. non-preregistered and confirmatory vs. exploratory research.
 
-The scientific community asserts that the objective of preregistration is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012].
-However, it is not principally warranted to suggest a close coupling of the categories preregistered vs. non-preregistered and confirmatory vs. exploratory research.
-In fact, researchers can conduct confirmatory research without preregistration (though it might be difficult to convince other researchers of the confirmatory nature of their research).
+<!---false dichotomy--->
+In fact, researchers can conduct confirmatory research without preregistration—though it might be difficult to convince other researchers of the confirmatory nature of their research.
 The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
@@
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
-Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary to establish confirmation [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021]
+Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
 
-These issues arise from a fundamental misunderstanding about the objective of preregistration.
-Researchers are bound to be misled when they delegate an important scientific judgment to a simple decision rule, that only takes into account the existence of a preregistration.
-Preregistering an inherently exploratory analysis (like testing dozens of relations) does not make it confirmatory, nor will a carefully conducted confirmatory study become exploratory if the researcher deviates from the preregistration in minor details.
-To force preregistrations to generate results that are rightfully considered confirmatory additional restrictions on the types of analyses allowed have to be imposed.
-First, the analysis must have a low type-I error rate. second, 
-Equating confirmatory research with preregistered research is only possible for studies that expect a low type-I error rate and will run/be analyzed without in any case [@simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-Under these conditions, the rule holds, but such restricted use makes preregistration a niche solution unable to match the greater problem of lacking replicability in psychology and elsewhere.
+<!---many researcher do not subscribe to confirmatory agenda--->
+We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
+When researcher criticize preregistration as too inflexible to fit their research question they often simply acknowledge that their research goals are not strictly confirmatory.
+Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
+However reasonable such move is, changing core believes of large community is much harder than convincing them that a method is well justified.
+We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
 
-We show that the simple decision rule is just a special case of a more general conceptualization under Bayesian reasoning. 
+<!---outline-structure--->
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.

### 2022-12-20 manuscript.Rmd
correct subscript

@@
 The theoretical risk, however, is also low in absolute terms, but high compared to the detectability.
-In the unlikely event that all 10 students place their matches to form the Roman numeral Ⅷ ($E_3$), it is probably due to insight ($H$) and not by chance $P(¬E_2|¬H)$).
+In the unlikely event that all 10 students place their matches to form the Roman numeral Ⅷ ($E_3$), it is probably due to insight ($H$) and not by chance $P(¬E_3|¬H)$).
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.

### 2023-01-11 manuscript.Rmd
fix sentance
expand on statistical relevancy condition

@@
 Researchers cannot know what exactly they will learn from a study before they have run it.
-However, they can develop, they can form an expectation that helps them decide which study to conduct in what manner.
+However, they can develop an expectation that helps them decide which study to conduct in what manner.
 This expectation is what we term expected epistemic value.
@@
 The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
-The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of probability; that remains a leap of faith.
+The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" (XXX).
+That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.
+
 Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
@@
 In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
-We refrain from discussing specific measures of corroboration since no measure shows universally better properties than others.
+As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
 However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
+We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|¬H)$.
+This assumption is sometimes called statistical relevance condition (XXX).
 

### 2023-01-11 manuscript.Rmd
typo

@@
 The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
-If a hypothesis is more probable to begin with, the same increase in corroboration will lead to a higher probability than if the study was less probable to begin with.
+If a hypothesis is more probable to begin with, observing evidence in favor will result a more corroborated hypothesis, all else equal.
 However, the prior probability of a hypothesis is nothing our study design can change.
@@
 We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($¬H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($¬E_1$); the experiment therefore has high theoretical risk ($P(¬E_1|¬H)$)
+If "insight" has nothing to do with solving the problem ($¬H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($¬E_1$); the experiment therefore has high theoretical risk ($P(¬E_1|¬H)$).
 Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability ($P(E_1|H)$).
@@
 Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
-Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration.
+Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration (XXX).
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
@@
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extend than if the study would not have been preregistered.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analysies, etc.
+Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
 In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.

### 2023-01-19 manuscript.Rmd
Update manuscript.Rmd
some more changes.

@@
 <!--introduce confirmatory vs. exploratory-->
-A theory able to generate successful prediction should be more credible than a theory build around the facts afterwards.
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
-Mistaking exploratory findings for empirically confirmed results has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to observe this distinction and produce results that are considered confirmatory.
-Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
+<!--AB: I am not sure about whether this is a good first sentence; most would subscribe to a view that both exploration and confirmation are necessary for progress but the distinction is important to keep the confirmation step valid; A theory able to generate successful predictions is often considered more valuable than a theory generated by postdiction, that is, one that is built to explain why observations came about.-->
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing, or prediction and postdiction [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010, @nosekPreregistrationRevolution2018].
+There is a broad consensus that both are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far it has been corroborated.
+Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation to be true if it is false.
+A variety of practices, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
+Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As a response to the crisis, more and more researchers preregister their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
+They do so to stress the predictive nature of their registered statistical analyses and to produce results that were clearly and transparently to be labeled confirmatory.
+Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping promised false-positive rates.
 This ability to provide a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
-Still an overwhelming majority of empirical researchers do not routinely preregister their studies because they do not find that the theoretical advantages outweigh the practical hurdles.
-This suggests a demand for explicating what the objective of preregistration is and what such objective implies in practice.
+Still, the majority of empirical researchers do not routinely preregister their studies.
+One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
+We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that leave data-dependent decisions unspecified.
 
 <!--the current goal of preregistration is confirmation-->
-One objective of preregistration that has received widespread attention, is to distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+One goal of preregistration that has received widespread attention, is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 In such narrative, preregistration is justified by a confirmatory research agenda.
 However, two problems emerge under closer inspection.
-First, many researcher do not subscribe to a confirmatory research agenda.
-Second, it is not principally warranted to suggest a close coupling of the categories preregistered vs. non-preregistered and confirmatory vs. exploratory research.
+First, many researcher do not subscribe to a purely confirmatory research agenda.
+Second, it is not necessary to strictly map the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
 
 <!---false dichotomy--->
-In fact, researchers can conduct confirmatory research without preregistration—though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+Obviously, researchers can conduct confirmatory research without preregistration — though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
+The opposite would be preregistered but not strictly confirmatory studies and those are possible as well [e.g., @dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
-Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
+Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
 
@@
 We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
-When researcher criticize preregistration as too inflexible to fit their research question they often simply acknowledge that their research goals are not strictly confirmatory.
+When researcher criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
 Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
-However reasonable such move is, changing core believes of large community is much harder than convincing them that a method is well justified.
+However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
 We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
+While preregistration is especially useful for dividing confirmatory and exploratory research, we argue that preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
 
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
+Theoretical risk serves as a general measure of juding how much evidence supports a given theory.
 
-We then outline two interpretations of how theoretical risk is impacted by preregistration.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error.
+We then outline two interpretations of how preregistration impacts theoretical risk.
+The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
 The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
-Following this interpretation, the theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
+We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and such.
+Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
 
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is warranted for any study.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is applicable and potentially useful for any kind of study, whether it is one of prediction, postdiction, or a mixture of both.
 
@@
 If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude.
+<!--AB: reference needed, where this term is defined; whose definition do we use?-->
 In other words, they hope to learn something from conducting the study.
@@
 We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
-Enforcing this distinction would mean using preregistrations as a means to increase the theoretical risk researchers take on.
-No doubt this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
+<--AB: can you rewrite this and be more clear?
+Isn't this more about researchers who do not distinguish these concepts?
+-->
+Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.
+Undoubtedly, this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
 However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
+<!--AB: maybe add "everything else being the same" or the specific conditions?-->
 As we further argued, minimizing the uncertainty about theoretical risk (without necessarily increasing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
@@
 
-In our view, the advantage of uncertainty reduction about theoretical risk front and center is two-fold.
-First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
-That is, researchers benefit from preregistering their study, regardless of whether or not the study is "exploratory".
+In our view, there are two advantages to focusing on uncertainty over theoretical risk itself.
+First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks one encounters as researcher.
+That is, researchers benefit from preregistering their study, regardless of whether or not the study is tagged "exploratory".
 If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
-Second, if researchers realize after the fact that it would be unwise to follow their preregistration in certain details, they may change their strategy.
-Allowing researchers to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
-In its fullest consequence, this argument might result in the same uncertainty that we argued plagues for not-preregistered studies.
-However, if the authors offer a convincing argument for their deviation, their readers might may judge it to be a reasonable explanation for the deviation.
+Second, if researchers realize after the fact that a deviation from their preregistration in certain details is sensible (and sometimes, the only reasonable way forward), they are explicitly allowed to change their strategy.
+<!--AB: this sounds like a new claim but it is accepted that deviations are OK, isn't it?
+maybe stress that this view is natural under the proposed framework but less so in stricter views of preregistration-->
+Allowing researchers to deviate from the preregistration is controversial (XXX) <!--AB: ref missing--> because the authors might sift through all the possibilities and then select the most beneficial for their theory.
+In its full consequence, this argument might result in the same uncertainty that we argued plagues not-preregistered studies.
+However, if authors of study offer a convincing argument for their deviation, their readers might may judge it to be a reasonable explanation for the deviation.
 In fact, using the here presented framework, we can distinguish reasonable from unreasonable deviations.
-If a change increases detectability enough to offset the increase in uncertainty in theoretical risk, it is worthwhile.
+If a deviation increases detectability enough to offset the increase in uncertainty in theoretical risk, it is a deviation worth doing.
 

### 2023-01-26 manuscript.Rmd
Revert "Update manuscript.Rmd"
This reverts commit 3a9a85cd92e5855e0bf17695ef0a0373872b1c67.

--- 2023-01-11 manuscript.Rmd
@@
 <!--introduce confirmatory vs. exploratory-->
-A theory able to generate successful prediction should be more credible than a theory build around the facts afterwards.
+A theory able to generate successful predictions is often considered more valuable than a theory generated by postdiction, that is, one that is built to explain why observations came about.
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010].
-Mistaking exploratory findings for empirically confirmed results has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to observe this distinction and produce results that are considered confirmatory.
-Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012].
+Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation to be true if it is false.
+This has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
+As a response to the crisis, more and more researchers preregister their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
+They do so to stress the predictive nature of their registered statistical analyses and to produce results that are considered confirmatory.
+Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping promised false-positive rates.
 This ability to provide a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
-Still an overwhelming majority of empirical researchers do not routinely preregister their studies because they do not find that the theoretical advantages outweigh the practical hurdles.
+Still, the majority of empirical researchers do not routinely preregister their studies because they do not find that the theoretical advantages outweigh the practical hurdles.
 This suggests a demand for explicating what the objective of preregistration is and what such objective implies in practice.
@@
 However, two problems emerge under closer inspection.
-First, many researcher do not subscribe to a confirmatory research agenda.
+First, many researcher do not subscribe to a purely confirmatory research agenda.
 Second, it is not principally warranted to suggest a close coupling of the categories preregistered vs. non-preregistered and confirmatory vs. exploratory research.
@@
 <!---false dichotomy--->
-In fact, researchers can conduct confirmatory research without preregistration—though it might be difficult to convince other researchers of the confirmatory nature of their research.
-The exact opposite, preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+Obviously, researchers can conduct confirmatory research without preregistration — though it might be difficult to convince other researchers of the confirmatory nature of their research.
+The opposite would be preregistered but not strictly confirmatory studies and those are possible as well [e.g., @dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
-Thus, insisting on equating preregistration and confirmation has led to criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
+Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
 
@@
 We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
-When researcher criticize preregistration as too inflexible to fit their research question they often simply acknowledge that their research goals are not strictly confirmatory.
+When researcher criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
 Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
-However reasonable such move is, changing core believes of large community is much harder than convincing them that a method is well justified.
-We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
+However reasonable such move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
+We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science and argue that preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
 

### 2023-01-31 manuscript.Rmd

--- 2023-01-19 manuscript.Rmd
@@
 <!--introduce confirmatory vs. exploratory-->
-<!--AB: I am not sure about whether this is a good first sentence; most would subscribe to a view that both exploration and confirmation are necessary for progress but the distinction is important to keep the confirmation step valid; A theory able to generate successful predictions is often considered more valuable than a theory generated by postdiction, that is, one that is built to explain why observations came about.-->
 The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing, or prediction and postdiction [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010, @nosekPreregistrationRevolution2018].
-There is a broad consensus that both are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far it has been corroborated.
+There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far it has been corroborated.
 Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation to be true if it is false.
@@
 <!---false dichotomy--->
-Obviously, researchers can conduct confirmatory research without preregistration — though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
-The opposite would be preregistered but not strictly confirmatory studies and those are possible as well [e.g., @dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+Obviously, researchers can conduct confirmatory research without preregistration— hough it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
+The opposite would be preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
@@
 To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
-Theoretical risk serves as a general measure of juding how much evidence supports a given theory.
 
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is applicable and potentially useful for any kind of study, whether it is one of prediction, postdiction, or a mixture of both.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is potentially useful for any kind of study, no matter how exploratory.
 
@@
 We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
-<--AB: can you rewrite this and be more clear?
-Isn't this more about researchers who do not distinguish these concepts?
--->
+<--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
 Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.

### 2023-02-02 manuscript.Rmd
minor changes after reading

@@
 There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far it has been corroborated.
-Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation to be true if it is false.
+Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation even if it is false.
 A variety of practices, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
@@
 <!---false dichotomy--->
-Obviously, researchers can conduct confirmatory research without preregistration— hough it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
-The opposite would be preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
+The opposite, i.e., preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.

### 2023-02-02 manuscript.Rmd
change format to taylor and francis

@@
 title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
-output:
-  bookdown::pdf_document2:
-    latex_engine: "xelatex"
-    toc: false
-    number_sections: false
-header-includes:
-   - \usepackage{libertine}
-fontsize: 11pt
-linestretch: 1.2
-geometry: "left=3.9cm, right=3.3cm, top=2.5cm, bottom=3cm"
-papersize: a4
 bibliography: references.bib
 abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
-csl: apa.csl
+author:
+  - name: Aaron Peikert
+    affil: a, b
+    email: peikert@mpib-berlin.mpg.de
+  - name: Andreas M. Brandmaier
+    affil: a, c
+    email: brandmaier@mpib-berlin.mpg.de
+affiliation:
+  - num: a
+    address: |
+      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
+  - num: b
+    address: |
+      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
+  - num: c
+    address: |
+      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
 repro:
@@
     - tikzDevice
+    - rticles
   tlmgr:
@@
     - inkscape
+output:
+  rticles::tf_article:
+    latex_engine: xelatex
+type: ARTICLE TEMPLATE
+keywords: |
+  preregistration; confirmatory analyses; exploratory analyses; hypothesis testing; bayesian
+header-includes: |
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
 ---

### 2023-02-02 manuscript.Rmd
new font does not have roman numerals

@@
 
-> Use five matches (Ⅰ Ⅰ Ⅰ Ⅰ Ⅰ) to form the number eight.
+> Use five matches (IIII) to form the number eight.
 
@@
 1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-2. The experimenter shows the solution "Ⅷ" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
+2. The experimenter shows the solution "IX" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
 3. The experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.

### 2023-02-02 manuscript.Rmd
correct history of theorethical risk

@@
 We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
-In a response to the original paper, @kuklaClinicalStatisticalTheory1990 argued that the arguments layed out in @meehlTheoreticalRisksTabular1978 can be reconstructed in purely Bayesian framework.
+@kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in purely Bayesian framework.
 However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ can be related to or even equated to theoretical risk.

### 2023-02-02 manuscript.Rmd
small formulation changes

@@
 They do so to stress the predictive nature of their registered statistical analyses and to produce results that were clearly and transparently to be labeled confirmatory.
-Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping promised false-positive rates.
+Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
 This ability to provide a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
@@
 One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
-We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that leave data-dependent decisions unspecified.
+We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.
 
@@
 We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
-While preregistration is especially useful for dividing confirmatory and exploratory research, we argue that preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
+It might be that psychology needs more confirmatory studies but preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
 

### 2023-02-06 manuscript.Rmd
remove UTF8 chars

@@
 We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
-We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(¬H) = 1 - P(H)$.
+We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(\neg{}H) = 1 - P(H)$.
 The probability of observing evidence under some hypothesis is $P(E|H)$.
@@
 However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
-We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|¬H)$.
+We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
 This assumption is sometimes called statistical relevance condition (XXX).
@@
 \begin{equation}
-P(E) = P(H)P(E|H) + P(¬H)P(E|¬H) (\#eq:total-expectedness)
+P(E) = P(H)P(E|H) + P(\neg{}H)P(E|\neg{}H) (\#eq:total-expectedness)
 \end{equation}
 
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(¬H)$), and that it is common sense to increase detectability ($P(E|H)$).
-The real lever to pull is, therefore, $P(E|¬H)$.
+We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability ($P(E|H)$).
+The real lever to pull is, therefore, $P(E|\neg{}H)$.
 This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
-Its counter probability $P(¬E|¬H)= 1 - P(E|¬H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
-We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(¬E|¬H)$.
+Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
 @kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in purely Bayesian framework.
-However, while he did not name $P(¬E|¬H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|¬H)$ which would imply that $P(¬E|¬H)$ can be related to or even equated to theoretical risk.
+However, while he did not name $P(\neg{}E|\neg{}H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|\neg{}H)$ which would imply that $P(\neg{}E|\neg{}H)$ can be related to or even equated to theoretical risk.
 
-Let us note some interesting properties of theoretical risk ($P(¬E|¬H)$).
+Let us note some interesting properties of theoretical risk ($P(\neg{}E|\neg{}H)$).
 First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
@@
 The last statement sounds like a truism but is directly related to Popper's seminal criterion of demarcation.
-He stated that if it is impossible to prove that a hypothesis is false ($P(¬E|¬H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
+He stated that if it is impossible to prove that a hypothesis is false ($P(\neg{}E|\neg{}H) = 0$, theoretical risk is zero), it cannot be considered a scientific hypothesis [@popperLogicScientificDiscovery2002, p.
 18].
@@
 
-Both theoretical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+Both theoretical risk ($P(\neg{}E|\neg{}H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
 To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
@@
 
-We argue that experiment 1 has high theoretical risk ($P(¬E_1|¬H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($¬H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($¬E_1$); the experiment therefore has high theoretical risk ($P(¬E_1|¬H)$).
+We argue that experiment 1 has high theoretical risk ($P(\neg{}E_1|\neg{}H)$) and high detectability ($P(E_1|H)$).
+If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk ($P(\neg{}E_1|\neg{}H)$).
 Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability ($P(E_1|H)$).
-The second experiment, on the other hand, has low theoretical risk ($P(¬E_2|¬H)$).
-Even if "insight" has nothing to do with solving the problem ($¬H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
+The second experiment, on the other hand, has low theoretical risk ($P(\neg{}E_2|\neg{}H)$).
+Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
 With regard to detectability, experiments 1 and 2 differ in no obvious way.
@@
 The theoretical risk, however, is also low in absolute terms, but high compared to the detectability.
-In the unlikely event that all 10 students place their matches to form the Roman numeral Ⅷ ($E_3$), it is probably due to insight ($H$) and not by chance $P(¬E_3|¬H)$).
+In the unlikely event that all 10 students place their matches to form the Roman numeral VIII ($E_3$), it is probably due to insight ($H$) and not by chance $P(\neg{}E_3|\neg{}H)$).
 Of course, in practice, we would allow the evidence to be probabilistic, e.g., relax the requirement of "all students" to nine out of ten students, more than eight, and so forth.
@@
 Second, available resources limit theoretical risk.
-Increasing theoretical risk ($P(¬E|¬H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
+Increasing theoretical risk ($P(\neg{}E|\neg{}H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
 In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
@@
 
-Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|¬H)$.
-$P(E|¬H)$ is equivalent to the type-I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $¬H$.
+Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|\neg{}H)$.
+$P(E|\neg{}H)$ is equivalent to the type-I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
 Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
@@
 However, the research hypothesis seldomly equals  the statistical null hypothesis, and therefore, the relation between statistical type-I error rate and theoretical risk should not be overinterpreted.
-We argue that theoretical risk (and hence its inverse, $P(E|¬H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+We argue that theoretical risk (and hence its inverse, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
@@
 Examples of our ability to exert this control are the setting of type-I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|¬H)$).
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|\neg{}H)$).
 
@@
 All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
-Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|¬H)$.
-While, in theory, these factors may leave $P(E|¬H)$ unaffected or even decrease it, we argue that this is not the case in practice.
+Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|\neg{}H)$.
+While, in theory, these factors may leave $P(E|\neg{}H)$ unaffected or even decrease it, we argue that this is not the case in practice.
 Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.

### 2023-02-06 manuscript.Rmd
remove brackets around probabilities in text looks weird

@@
 
-The posterior probability ($P(H|E)$) is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability ($P(H)$), it is called increase in firmness.
+The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
+In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness.
 As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability ($P(H|E)$).
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$.
 We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
@@
 
-In short, we want to increase posterior probability ($P(H|E)$).
-Increases in posterior probability ($P(H|E)$) are associated with increased epistemic value, for which we want to maximize the expectation.
+In short, we want to increase posterior probability $P(H|E)$.
+Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation.
 So how can we increase posterior probability?
 The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
-The first option leads us to the unsurprising conclusion that higher a priori probability ($P(H)$) leads to higher posterior probability ($P(H|E)$).
+The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
 If a hypothesis is more probable to begin with, observing evidence in favor will result a more corroborated hypothesis, all else equal.
 However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability ($P(H|E)$).
+The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability $P(H|E)$.
 $P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
@@
 
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability ($P(E|H)$).
+We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability $P(E|H)$.
 The real lever to pull is, therefore, $P(E|\neg{}H)$.
@@
 
-Let us note some interesting properties of theoretical risk ($P(\neg{}E|\neg{}H)$).
-First, increasing theoretical risk leads to higher posterior probability ($P(H|E)$, our objective).
-Second, if the theoretical risk is smaller than detectability ($P(E|H)$) it follows that the posterior probability must decrease when observing the evidence.
+Let us note some interesting properties of theoretical risk $P(\neg{}E|\neg{}H)$.
+First, increasing theoretical risk leads to higher posterior probability $P(H|E)$, our objective.
+Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
 If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
@@
 
-Both theoretical risk ($P(\neg{}E|\neg{}H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
+Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
 To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
@@
 
-We argue that experiment 1 has high theoretical risk ($P(\neg{}E_1|\neg{}H)$) and high detectability ($P(E_1|H)$).
-If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk ($P(\neg{}E_1|\neg{}H)$).
-Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability ($P(E_1|H)$).
-The second experiment, on the other hand, has low theoretical risk ($P(\neg{}E_2|\neg{}H)$).
+We argue that experiment 1 has high theoretical risk $P(\neg{}E_1|\neg{}H)$ and high detectability $P(E_1|H)$.
+If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
+Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability $P(E_1|H)$.
+The second experiment, on the other hand, has low theoretical risk $P(\neg{}E_2|\neg{}H)$.
 Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
@@
 Experiment 3, however, also has low detectability.
-It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$) experiment 3 therefore has low detectability ($P(E_3|H)$).
-The theoretical risk, however, is also low in absolute terms, but high compared to the detectability.
+It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$) experiment 3 therefore has low detectability $P(E_3|H)$.
+The theoretical risk, however, is also low in absolute terms, but high compared to the detectability (statistical relevancy condition is satisfied).
 In the unlikely event that all 10 students place their matches to form the Roman numeral VIII ($E_3$), it is probably due to insight ($H$) and not by chance $P(\neg{}E_3|\neg{}H)$).
@@
 Second, available resources limit theoretical risk.
-Increasing theoretical risk ($P(\neg{}E|\neg{}H)$) will usually decrease detectability ($P(E|H)$) unless more resources are invested.
+Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
 In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
@@
 Examples of our ability to exert this control are the setting of type-I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none ($P(E|\neg{}H)$).
+These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none $P(E|\neg{}H)$.
 

### 2023-02-06 manuscript.Rmd
add version notes

@@
 
+# Notes {-}
+
+The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
+This version was created from git commit ``r repro::current_hash(backend = "gert")``.
+The manuscript is available as preprint [@no-preprint-yet] and will be submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
+
 # References

### 2023-02-06 manuscript.Rmd
fix labels
tau is also utf8

@@
 \begin{equation}
-P(H|E) = \frac{P(H)P(E|H)}{P(E)} (\#eq:bayes)
+P(H|E) = \frac{P(H)P(E|H)}{P(E)} \label{eq:bayes}
 \end{equation}
@@
 \begin{equation}
-P(E) = P(H)P(E|H) + P(\neg{}H)P(E|\neg{}H) (\#eq:total-expectedness)
+P(E) = P(H)P(E|H) + P(\neg{}H)P(E|\neg{}H) \label{eq:total-expectedness}
 \end{equation}
@@
 \begin{equation}
-\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H)) (\#eq:integral)
+\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H)) \label{eq:integral}
 \end{equation}
@@
 ```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Posterior Probability (corroboration as firmness) as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "Posterior Probability (corroboration as firmness) as a function of $\tau$, where $\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 ```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Several measures for corroboration as increase in firmness as a function of τ, where τ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "Several measures for corroboration as increase in firmness as a function of $\tau$, where $\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 

### 2023-02-06 manuscript.Rmd
add packages for git hash version

@@
     - rticles
+    - gert
+    - aaronpeikert/repro@5075336
   tlmgr:

### 2023-02-06 manuscript.Rmd
do not render with xelatex

--- 2023-02-06 manuscript.Rmd
@@
     - inkscape
-output:
-  rticles::tf_article:
-    latex_engine: xelatex
+output: rticles::tf_article
 type: ARTICLE TEMPLATE

### 2023-02-06 manuscript.Rmd
enable cross references for article

@@
 
-```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Posterior Probability (corroboration as firmness) as a function of $\tau$, where $\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
+#| fig.cap: "\\label{fig:posterior}Posterior Probability (corroboration as firmness) as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 
-```{r, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "Several measures for corroboration as increase in firmness as a function of $\tau$, where $\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+```{r other-measures-fig, dev = "tikz", cache=TRUE, warning=FALSE}
+#| fig.cap: "\\label{fig:other-measures}Several measures for corroboration as increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 

### 2023-02-06 manuscript.Rmd
typo

@@
 We argue that such criticism is not directed against preregistration itself but against a justification through a confirmatory research agenda [@wagenmakersAgendaPurelyConfirmatory2012].
-When researcher criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
+When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
 Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
@@
 We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
-<--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
+<!--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
 Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.

### 2023-02-08 manuscript.Rmd
rewrite discussion!

@@
 <!---outline-structure--->
-To that end, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
+To form such objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
 
@@
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
-This point is crucial because we attribute much of the appeal of preregistration to this fact.
+This point is crucial because we attribute much of the appeal of a confirmatory research agenda to this fact.
 
@@
 
-> Use five matches (IIII) to form the number eight.
+> Use five matches (IIIII) to form the number eight.
 
@@
 
-Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk.
+Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
 Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration (XXX).
@@
 Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
-
 We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
@@
 ```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:posterior}Posterior Probability (corroboration as firmness) as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:posterior}Posterior Probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
   theme(legend.position = c(.2, .8),
-        legend.title = element_blank())
+        legend.title = element_blank()) +
+    scale_y_continuous(breaks = seq(0, 1, 1/4), limits = c(0,1)) +
+  NULL
 ```
@@
 
+```{r}
+confirmatory_tau <- .99
+confirmatory_uncertain <- results %>%
+  filter(certainty == "uncertain",
+         name == "Posterior Probability",
+         tau == confirmatory_tau) %>% pull(value)
+confirmatory_certain <- results %>%
+  filter(certainty == "certain",
+         name == "Posterior Probability",
+         tau == confirmatory_tau) %>% pull(value)
+some_exploration_tau <- .8
+some_exploration_certain <-  results %>%
+  filter(certainty == "certain",
+         name == "Posterior Probability",
+         tau == some_exploration_tau) %>% pull(value)
+
+some_exploration_uncertain <-  results %>%
+  filter(certainty == "uncertain",
+         name == "Posterior Probability",
+         tau == some_exploration_tau) %>% pull(value)
+```
+
+Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being fully certain (solid line) or fully uncertain (dashed line) about the theoretical risk of a study.
 Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions into which all realistic conditions fall.
@@
 Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
+
+The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
+However, if a hypothesis in a certain study has low theoretical risk, researchers cannot do much to change that.
+However, studies do not only differ by how high the theoretical risk is but also how certain the recipient is about the theoretical risk.
+A study that has a very high theoretical risk, e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed, i.e., alpha error, but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
+The other factors (detect ability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit very of course with the specifics of the study.
+Crucially, even exploratory studies benefit from preregistration, e.g., in this scenario with a $\tau$ of `r some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
 
@@
 The third and last argument is that preregistrations reduce this uncertainty.
+Following the last argument, a preregistered study is represented by the solid line (certainty about theoretical risk) and a study that was not preregistered is more similar to the dashed line (maximally uncertain about theoretical risk) in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
 
@@
 In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+A study with a perfect preregistration will, therefore, resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
+Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not the the theoretical risk is necessarily high.
 
@@
 
-We started out with the observation that preregistrations do not always cleanly divide preregistration from confirmation.
-<!--AB: can you rewrite this and be more clear? Isn't this more about researchers who do not distinguish these concepts? -->
-Subscribing to this distinction would mean to use preregistrations solely as a means to increase the theoretical risk researchers are willing to accept.
-Undoubtedly, this results in more trustworthy evidence and may even increase the effectiveness of scientific undertakings in general.
-However, as argued earlier, directly maximizing theoretical risk increases the likelihood that a study fails to deliver results favoring the theory (a special instance of this problem is the well-known trade-off of statistical power and type-I error rate).
-<!--AB: maybe add "everything else being the same" or the specific conditions?-->
-As we further argued, minimizing the uncertainty about theoretical risk (without necessarily increasing the theoretical risk itself) has benefits for the epistemic value, too, but without increasing the likelihood of producing a "failed" study.
-Note that any single study that does not produce evidence in favor of a research hypothesis is still precious for the scientific process.
-Under the current incentive structure, however, many researchers have reason to avoid such outcomes and few, if any, have the privilege to operate outside of these incentives.
+To summarize, we showed that higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
+The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal of confirmatory research.
+However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and illdefined, resources are limited, and increasing theoretical risk usually decreases detectability (a special instance of this trade off is the well known tension between alpha error and power).
+Before researchers may preregister a confirmatory study with high theoretical risk, they must ensure that they are able to conduct a confirmatory study.
+Researchers must develop theories sufficiently and then have the resources to adequately balance detectability against theoretical risk.
+A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory, but enabling confirmatory research is a task that requires action from the whole research community.
+Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted.
+Journal editors could encourage theoretical developments and be willing to accept registered reports.
+Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both.
 
-In our view, there are two advantages to focusing on uncertainty over theoretical risk itself.
-First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks one encounters as researcher.
-That is, researchers benefit from preregistering their study, regardless of whether or not the study is tagged "exploratory".
-If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
-Second, if researchers realize after the fact that a deviation from their preregistration in certain details is sensible (and sometimes, the only reasonable way forward), they are explicitly allowed to change their strategy.
-<!--AB: this sounds like a new claim but it is accepted that deviations are OK, isn't it?
-maybe stress that this view is natural under the proposed framework but less so in stricter views of preregistration-->
-Allowing researchers to deviate from the preregistration is controversial (XXX) <!--AB: ref missing--> because the authors might sift through all the possibilities and then select the most beneficial for their theory.
-In its full consequence, this argument might result in the same uncertainty that we argued plagues not-preregistered studies.
-However, if authors of study offer a convincing argument for their deviation, their readers might may judge it to be a reasonable explanation for the deviation.
-In fact, using the here presented framework, we can distinguish reasonable from unreasonable deviations.
-If a deviation increases detectability enough to offset the increase in uncertainty in theoretical risk, it is a deviation worth doing.
+Our later result, on the importance of preregistration for minimizing uncertainty, has two important implications.
+The first is, that even if all imaginable actions regarding promoting higher theoretical risk are taken, confirmatory research should be preregistered.
+Otherwise the uncertainty about the theoretical risk will diminish the advantage of confirmatory research.
+Second, even under less than ideal circumstances for confirmatory research, preregistration is beneficial.
+Preregistering exploratory studies increases the expected epistemic value by virtue of reducing uncertainty about theorethical risk.
+Nevertheless, exploratory studies will have a lower expected epistemic value than a more confirmatory study, if both are preregistered and have equal detectability.
+
+Focusing on uncertainty reduction also explains two common practices of preregistration that do not align with a confirmatory research agenda.
+First, researchers seldomly predict precise numerical outcomes, instead they use preregistrations to describe the process that generates the results.
+Precise predictions would have very high theoretical risk (they are likely incorrect if the theory is wrong).
+A statistical procedure, may have high or low theoretical risk depending on the specifics of the model used.
+Specifying the process, therefore, is in line with the rational we propose here, but is incompatible with strictly confirmatory research.
+Second, researchers often have to deviate from the preregistration and make data-dependent decisions outside of the preregistration.
+If the only goal of preregistration is to ensure confirmatory research, such changes are not justifiable.
+However, under our rational some changes may be justified.
+Any change would increase the uncertainty about the theoretical risk and may even decrease the theorethical risk.
+However, such negative outcomes may be offset by an increase in detectability due to the change.
+Consider a preregistration that failed to specify what to do with missing values where the researchers subsequently encounter missings.
+In such case, the detectability is zero, since the data cannot be analyzed without a decision about how to handle the missing data.
+Any such decision would constitute a deviation from the preregistration.
+Furthermore, a reader cannot rule out that the researchers leveraged the decision to decrease theoretical risk, i.e., picking among all options the one that delivers the most beneficial results for the theory.
+Whatever decision they make, increased uncertainty about the theoretical risk is inevitable and the expected epistemic value is decreased compared to a world where they anticipated the need to deal with missing data.
+However, it is still justified to deviate.
+After all they have not anticipated the case and are left with a detectablilty of zero.
+Any decision will increase detectability to a non-zero value, likely offsetting the increase in uncertainty.
+The researchers also may do their best to argue that the deviation was not motivated by increasing theoretical risk, thereby, decreasing the uncertainty.
+Maybe there is one decision that fits well with the theory or with the study design.
+Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision.
+
+Using reduction in uncertainty as the objective for preregistration does not only explain some existing behavior, it allows to form recommendations to improve the practice of preregistration.
+Maybe most important, we now have a measure to differentiate a good preregistration from a bad one.
+In particular, a preregistration should be specific about the procedure that is intended to generate evidence for a theory.
+Such procedure may accommodate a wide range of possible data, i.e., it may be exploratory.
+The theoretical risk, however low, must be communicated clearly.
+Parts of the process left unspecified, imply uncertainty, which is what a preregistration should reduce.
+However, specifying procedures that can be expected to not work, will lead to deviation, and uncertainty later.
+
+We have proposed a highly specific, yet practical procedure called preregistration as code (PAC) elsewhere [@citepac].
+In a PAC, researchers use computer code instead of a verbal description.
+This enables a preregistrations to accommodate many different data patterns, yet is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
+To reduce the liklihood of deviation, we suggest that a PAC is tested on simulated or pilot data.
+Another advantage is that with a PAC, deviations from the preregistration can be isolated and inspected.
+This is possible because in a PAC, preregistration and final manuscript are different versions of the same dynamic document, which can be compared via version control.
+However, while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also suited to achieve this goal.
+
+Contrary to what is widely assumed about preregistration, a preregistration is not necessarily a seal of confirmatory research.
+Confirmatory research would be much less persuasive without preregistration, but in our view preregistration simply communicates the extend of confirmation, i.e., theoretical risk, of a study.
+Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected epistemic value.
 

### 2023-02-10 manuscript.Rmd
add missing citations

@@
 The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
-However, the argument applies to any reward function that satisfies the "statistical relevancy condition" (XXX).
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
 That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.
@@
 The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap (XXX), in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness.
+In the tradition of Carnap, in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
 As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$.
+However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$ and our argument applies to those measures as well.
 We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
-This assumption is sometimes called statistical relevance condition (XXX).
+This assumption is sometimes called statistical relevance condition [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
 
@@
 Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
-Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration (XXX).
+Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration [@nosekPreregistrationRevolution2018; @oberauerPreregistrationForkingPath2019; @rubinDoesPreregistrationImprove2020].
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
@@
 A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory, but enabling confirmatory research is a task that requires action from the whole research community.
-Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted.
-Journal editors could encourage theoretical developments and be willing to accept registered reports.
-Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both.
+Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
+Journal editors could encourage theoretical developments and be willing to accept registered reports [@friedLackTheoryBuilding2020; @friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020; @vanrooijTheoryTestHow2021].
+Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both [@kooleRewardingReplicationsSure2012].
 

### 2023-02-10 manuscript.Rmd
small changes

@@
 Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
-This ability to provide a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
+This promise of a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
 Still, the majority of empirical researchers do not routinely preregister their studies.
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for or against a theory, preregistration is potentially useful for any kind of study, no matter how exploratory.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for a theory, preregistration is potentially useful for any kind of study, no matter how exploratory.
 
@@
 While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-We, therefore, conclude that statistical properties only guarantee a lower bound to the quantity we seek to minimize.
+We, therefore, conclude that statistical properties only guarantee a lower bound for theoretical risk.
 The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
@@
 
-To illustrate the effect of uncertainty, let $p(E|H) = 0.8$ (e.g., power of 80%) and $p(H) = 0.1$ and assume a uniform distribution for $p(E|\neg{}H)$ of the form:
+To illustrate the effect of uncertainty, let $p(E|H) = 0.8$ (e.g., power of 80%) and $p(H) = 0.1$ and assume a uniform distribution for theoretical risk $p(\neg{}E|\neg{}H)$ of the form:
 
@@
 A study that has a very high theoretical risk, e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed, i.e., alpha error, but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
-The other factors (detect ability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit very of course with the specifics of the study.
-Crucially, even exploratory studies benefit from preregistration, e.g., in this scenario with a $\tau$ of `r some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
+The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit very of course with the specifics of the study.
+Crucially, even exploratory studies benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
 
@@
 If the only goal of preregistration is to ensure confirmatory research, such changes are not justifiable.
+
 However, under our rational some changes may be justified.
 Any change would increase the uncertainty about the theoretical risk and may even decrease the theorethical risk.
-However, such negative outcomes may be offset by an increase in detectability due to the change.
+The changes still may be worthwhile, if the negative outcomes may be offset by an increase in detectability due to the change.
 Consider a preregistration that failed to specify what to do with missing values where the researchers subsequently encounter missings.
@@
 Maybe there is one decision that fits well with the theory or with the study design.
-Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision.
+Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision [@steegenIncreasingTransparencyMultiverse2016].
 
@@
 
-We have proposed a highly specific, yet practical procedure called preregistration as code (PAC) elsewhere [@citepac].
+We have proposed a highly specific, yet practical preregistration procedure called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code instead of a verbal description.

### 2023-02-10 manuscript.Rmd
add verisimilitude citation

@@
 keywords: |
-  preregistration; confirmatory analyses; exploratory analyses; hypothesis testing; bayesian
+  preregistration; confirmatory analyses; exploratory analyses; hypothesis testing; bayesian; open science
 header-includes: |
@@
 As a response to the crisis, more and more researchers preregister their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
-They do so to stress the predictive nature of their registered statistical analyses and to produce results that were clearly and transparently to be labeled confirmatory.
+They do so to stress the predictive nature of their registered statistical analyses and often with the hopes of obtaining a label that marks the study as "confirmatory".
 Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
@@
 However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
-We therefore attempt to disentangle the methodological goals of preregistration from the ideological goals of confirmatory science.
+We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
 It might be that psychology needs more confirmatory studies but preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
@@
 Let us start by defining what we call expected epistemic value.
-If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude.
- <!--AB: reference needed, where this term is defined; whose definition do we use?-->
+If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude [@niiniluotoVerisimilitudeThirdPeriod1998].
 In other words, they hope to learn something from conducting the study.

### 2023-02-10 manuscript.Rmd
rephare statistical relevance

@@
 1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else being equal, researchers try to maximize the expected epistemic value for other researchers.
+2. They expect other researchers to apply a similar rational process.
+2. They try to maximize the expected epistemic value for other researchers.
 
@@
 However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
-That is epistemic value is gained if evidence is observed, given that it is more likely to observe evidence when the theory holds than if the theory does not hold.
+That is, evidence only increases epistemic value for a theory, if the evidence is more likely to observe under the theory than without it.
 

### 2023-02-13 manuscript.Rmd
minor style changes

@@
 
+One factor that is known without much uncertainty is the contribution of statistical methods to theoretical risk.
 Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|\neg{}H)$.
@@
 The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit very of course with the specifics of the study.
-Crucially, even exploratory studies benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
+Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
 
@@
 Researchers must develop theories sufficiently and then have the resources to adequately balance detectability against theoretical risk.
-A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory, but enabling confirmatory research is a task that requires action from the whole research community.
+
+A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory.
+But for a confirmatory science it will not be enough to preregister.
+This undertaking requires action from the whole research community [@lishnerConciseSetCore2015].
 Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
@@
 Specifying the process, therefore, is in line with the rational we propose here, but is incompatible with strictly confirmatory research.
+
 Second, researchers often have to deviate from the preregistration and make data-dependent decisions outside of the preregistration.
 If the only goal of preregistration is to ensure confirmatory research, such changes are not justifiable.
-
 However, under our rational some changes may be justified.

### 2023-02-14 manuscript.Rmd
Update manuscript.Rmd

@@
 As a response to the crisis, more and more researchers preregister their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
-They do so to stress the predictive nature of their registered statistical analyses and often with the hopes of obtaining a label that marks the study as "confirmatory".
+They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".
 Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.

### 2023-02-14 manuscript.Rmd
Update manuscript.Rmd
Apply suggestions Andreas
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
-However, if a hypothesis in a certain study has low theoretical risk, researchers cannot do much to change that.
+However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
 However, studies do not only differ by how high the theoretical risk is but also how certain the recipient is about the theoretical risk.
@@
 
-To summarize, we showed that higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
-The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal of confirmatory research.
-However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and illdefined, resources are limited, and increasing theoretical risk usually decreases detectability (a special instance of this trade off is the well known tension between alpha error and power).
+To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
+The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal and central goal of confirmatory research.
+However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).
 Before researchers may preregister a confirmatory study with high theoretical risk, they must ensure that they are able to conduct a confirmatory study.
-Researchers must develop theories sufficiently and then have the resources to adequately balance detectability against theoretical risk.
+Specifically, this entails the development of sufficiently precise theories and having the resources (typically sample size) to adequately balance detectability against theoretical risk.
 
 A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory.
-But for a confirmatory science it will not be enough to preregister.
+But for a confirmatory science, it will not be enough to preregister all studies.
 This undertaking requires action from the whole research community [@lishnerConciseSetCore2015].
-Incentive structures must be more willing to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
+Incentive structures must be created to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
 Journal editors could encourage theoretical developments and be willing to accept registered reports [@friedLackTheoryBuilding2020; @friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020; @vanrooijTheoryTestHow2021].
@@
 
-Our later result, on the importance of preregistration for minimizing uncertainty, has two important implications.
+Our latter result, on the importance of preregistration for minimizing uncertainty, has two important implications.
 The first is, that even if all imaginable actions regarding promoting higher theoretical risk are taken, confirmatory research should be preregistered.
@@
 
-Second, researchers often have to deviate from the preregistration and make data-dependent decisions outside of the preregistration.
+Second, researchers often have to deviate from the preregistration and make data-dependent decisions after the preregistration.
 If the only goal of preregistration is to ensure confirmatory research, such changes are not justifiable.
 However, under our rational some changes may be justified.
-Any change would increase the uncertainty about the theoretical risk and may even decrease the theorethical risk.
+Any change increases the uncertainty about the theoretical risk and may even decrease the theorethical risk.
 The changes still may be worthwhile, if the negative outcomes may be offset by an increase in detectability due to the change.
-Consider a preregistration that failed to specify what to do with missing values where the researchers subsequently encounter missings.
-In such case, the detectability is zero, since the data cannot be analyzed without a decision about how to handle the missing data.
-Any such decision would constitute a deviation from the preregistration.
-Furthermore, a reader cannot rule out that the researchers leveraged the decision to decrease theoretical risk, i.e., picking among all options the one that delivers the most beneficial results for the theory.
+Consider a preregistration that failed to specify how to handle missing values, and researchers subsequently encountering missing values.
+In such case, detectability becomes zero because the data cannot be analyzed without a post-hoc decision about how to handle the missing data.
+Any such decision would constitute a deviation from the preregistration, which is perfectly waranted under our general objective.
+Note that a reader cannot rule out that the researchers leveraged the decision to decrease theoretical risk, i.e., picking among all options the one that delivers the most beneficial results for the theory (in the previous example, chosing between various options of handling missing values).
 Whatever decision they make, increased uncertainty about the theoretical risk is inevitable and the expected epistemic value is decreased compared to a world where they anticipated the need to deal with missing data.
@@
 The researchers also may do their best to argue that the deviation was not motivated by increasing theoretical risk, thereby, decreasing the uncertainty.
-Maybe there is one decision that fits well with the theory or with the study design.
+Ideally, there is a default decision that fits well with the theory or with the study design.
 Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision [@steegenIncreasingTransparencyMultiverse2016].
@@
 Parts of the process left unspecified, imply uncertainty, which is what a preregistration should reduce.
-However, specifying procedures that can be expected to not work, will lead to deviation, and uncertainty later.
+However, specifying procedures that can be expected to not work, will lead to deviation, and subsequently to larger uncertainty.
 
-We have proposed a highly specific, yet practical preregistration procedure called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+We have proposed a workflow for preregistration called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code instead of a verbal description.

### 2023-02-16 manuscript.Rmd
add my responses to comments
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 keywords: |
-  preregistration; confirmatory analyses; exploratory analyses; hypothesis testing; bayesian; open science
+preregistration; confirmation; exploration; hypothesis testing; bayesian; open science
 header-includes: |
@@
 <!---outline-structure--->
-To form such objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
 
@@
 2. They expect other researchers to apply a similar rational process.
-2. They try to maximize the expected epistemic value for other researchers.
+2. Researchers try to maximize the expected epistemic value for other researchers.
 
@@
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
-This point is crucial because we attribute much of the appeal of a confirmatory research agenda to this fact.
+This point is crucial because we attribute much of the appeal of a confirmatory research agenda to this notion.
 
@@
 As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
-However, it is generally expected that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$ and our argument applies to those measures as well.
+However, it is reasonable that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
 We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
-This assumption is sometimes called statistical relevance condition [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
+We referred to this assumption as "statistical relevance condition" earlier [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
 
@@
 Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
-Indeed, limiting the type-I error rate is commonly stated as a goal of preregistration [@nosekPreregistrationRevolution2018; @oberauerPreregistrationForkingPath2019; @rubinDoesPreregistrationImprove2020].
+Indeed, limiting the type-I error rate is commonly stated as  *the* central goal of preregistration [@nosekPreregistrationRevolution2018; @oberauerPreregistrationForkingPath2019; @rubinDoesPreregistrationImprove2020].
 We argue that while such a conclusion is plausible, we must first consider at least two constraints that place an upper bound on the theoretical risk.
@@
 Incentive structures must be created to evaluate not the outcomes of a study but the rigor with which it was conducted [@caganSanFranciscoDeclaration2013; @schonbrodtResponsibleResearchAssessment2022].
-Journal editors could encourage theoretical developments and be willing to accept registered reports [@friedLackTheoryBuilding2020; @friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020; @vanrooijTheoryTestHow2021].
+Journal editors could encourage theoretical developments that allow for precise predictions that will be tested by other researchers and be willing to accept registered reports [@friedLackTheoryBuilding2020; @friedTheoriesModelsWhat2020; @vanrooijTheoryDevelopmentRequires2020; @vanrooijTheoryTestHow2021].
 Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both [@kooleRewardingReplicationsSure2012].
@@
 
-Using reduction in uncertainty as the objective for preregistration does not only explain some existing behavior, it allows to form recommendations to improve the practice of preregistration.
-Maybe most important, we now have a measure to differentiate a good preregistration from a bad one.
+As explained above, reduction in uncertainty as the objective for preregistration does not only explain some existing behavior, that does not align with confirmation as a goal, it also allows to form recommendations to improve the practice of preregistration.
+Importantly, we now have a theoretical measure to gauge the functionality of preregistrations, which can only help increase its utility.
 In particular, a preregistration should be specific about the procedure that is intended to generate evidence for a theory.
@@
 We have proposed a workflow for preregistration called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
-In a PAC, researchers use computer code instead of a verbal description.
+In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for preregistration.
+This combination is facilitated by dynamic document generation, where the results of the code, such as numbers, figures, and tables, are inserted automatically into the document.
+The idea is that the preregistration already contains "mock results" based on simulated or pilot data, which are replaced after the actual study data becomes available.
+Using dynamic documents helps dissolve the distinction between the preregistration document and the final scientific report.
+Instead of fundamentally separate documents, preregistration, and final report are different versions of the same underlying dynamic document.
 This enables a preregistrations to accommodate many different data patterns, yet is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
 To reduce the liklihood of deviation, we suggest that a PAC is tested on simulated or pilot data.
-Another advantage is that with a PAC, deviations from the preregistration can be isolated and inspected.
+Another advantage is that with a PAC, deviations from the preregistration can be clearly (and if necessary, automatically) isolated, highlighted, and inspected.
 This is possible because in a PAC, preregistration and final manuscript are different versions of the same dynamic document, which can be compared via version control.
@@
 Contrary to what is widely assumed about preregistration, a preregistration is not necessarily a seal of confirmatory research.
-Confirmatory research would be much less persuasive without preregistration, but in our view preregistration simply communicates the extend of confirmation, i.e., theoretical risk, of a study.
+Confirmatory research would almost always be less persuasive without preregistration, but in our view, preregistration primarily communicates the extent of confirmation, i.e., theoretical risk, of a study.
 Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected epistemic value.

### 2023-02-16 manuscript.Rmd
fix indention
revise PAC section

@@
 We have proposed a workflow for preregistration called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
-In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for preregistration.
+In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.
 This combination is facilitated by dynamic document generation, where the results of the code, such as numbers, figures, and tables, are inserted automatically into the document.
 The idea is that the preregistration already contains "mock results" based on simulated or pilot data, which are replaced after the actual study data becomes available.
-Using dynamic documents helps dissolve the distinction between the preregistration document and the final scientific report.
+Such approach dissolves the distinction between the preregistration document and the final scientific report.
 Instead of fundamentally separate documents, preregistration, and final report are different versions of the same underlying dynamic document.
-This enables a preregistrations to accommodate many different data patterns, yet is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
-To reduce the liklihood of deviation, we suggest that a PAC is tested on simulated or pilot data.
-Another advantage is that with a PAC, deviations from the preregistration can be clearly (and if necessary, automatically) isolated, highlighted, and inspected.
-This is possible because in a PAC, preregistration and final manuscript are different versions of the same dynamic document, which can be compared via version control.
-However, while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also suited to achieve this goal.
+Deviations from the preregistration can therefore be clearly (and if necessary, automatically) isolated, highlighted, and inspected using version control.
+Crucially, because the preregistration contains code, it may accommodate many different data patterns, i.e., it may be exploratory.
+However, while a PAC is not the extend of exploration, it is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
+Please note that while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also able to advance this goal.
 

### 2023-02-16 manuscript.Rmd
clarify that confirmation requires more than preregistration

@@
 However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).
-Before researchers may preregister a confirmatory study with high theoretical risk, they must ensure that they are able to conduct a confirmatory study.
+While we believe that preregistration is always beneficial, it might be counterproductive to pursue high theoretical risk if the conditions for highly confirmatory research are not given.
 Specifically, this entails the development of sufficiently precise theories and having the resources (typically sample size) to adequately balance detectability against theoretical risk.
 
-A preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory.
+In terms of preparing the conditions for confirmatory research, preregistration may at most help to invest some time into developing more specific, hence riskier, implications of a theory.
 But for a confirmatory science, it will not be enough to preregister all studies.

### 2023-02-16 manuscript.Rmd
add acknowledgement
Update manuscript.Rmd
added Centre affiliation

@@
   - name: Andreas M. Brandmaier
-    affil: a, c
+    affil: a, c, d
     email: brandmaier@mpib-berlin.mpg.de
@@
       MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
+  - num: d
+    address: |
+     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
 repro:
@@
 
+# Acknowledgement {-}
+
+We thank Maximilian S.
+Ernst, Caspar van Lissa, Felix Schönbrodt, the discussants at the DGPS2022 conference and Open Science Center Munich, and many more for the insightful discussions about disentangling preregistration and confirmation.
+
 # Notes {-}

### 2023-02-16 manuscript.Rmd
Update manuscript.Rmd

@@
 <!--introduce confirmatory vs. exploratory-->
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing, or prediction and postdiction [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010, @nosekPreregistrationRevolution2018].
-There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far it has been corroborated.
-Mistaking exploratory findings for empirically confirmed results inflates the the likelihood of believing that there is evidence for a given explanation even if it is false.
-A variety of practices, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
+Despite the different names, it is fundamentally the same dichotomy that is at stake here.
+There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far they have been corroborated.
+However, mistaking exploratory findings for empirically confirmed results is dangerous.
+It inflates the likelihood of believing that there is evidence supporting a given explanation even if it is false.
+A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
 Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response to the crisis, more and more researchers preregister their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
+As a response to the crisis, more and more researchers preregister their hypotheses and their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
 They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".

### 2023-02-16 manuscript.Rmd
Update manuscript.Rmd

@@
 One goal of preregistration that has received widespread attention, is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
-In such narrative, preregistration is justified by a confirmatory research agenda.
-However, two problems emerge under closer inspection.
+In such a narrative, preregistration is justified by a confirmatory research agenda.
+However, two problems become apparent under closer inspection.
 First, many researcher do not subscribe to a purely confirmatory research agenda.
@@
 <!---false dichotomy--->
-Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, etc.
-The opposite, i.e., preregistered but not strictly confirmatory studies, are also commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, and so forth.
+The opposite, that is, preregistered but not strictly confirmatory studies, are also becoming more commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
+
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
@@
 When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
-Forcing researchers into a confirmatory research agenda does not only imply changing *how* they investigate but also *what* research questions.
+Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they ask.
 However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
 We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
-It might be that psychology needs more confirmatory studies but preregistration can be useful for any kind of study in the continuum between purely confirmatory and fully exploratory.
+It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
 

### 2023-02-16 manuscript.Rmd
Update manuscript.Rmd

@@
 <!---outline-structure--->
-To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate.
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate in statistical test theory.
 
-We then outline two interpretations of how preregistration impacts theoretical risk.
+Further, we outline two interpretations of how preregistration impacts theoretical risk.
 The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
@@
 Let us start by defining what we call expected epistemic value.
-If researchers plan to conduct a study, they usually hope it will change their assessment of some theory's verisimilitude [@niiniluotoVerisimilitudeThirdPeriod1998].
+If researchers plan to conduct a study, they usually hope that it will change their assessment of some theory's verisimilitude [@niiniluotoVerisimilitudeThirdPeriod1998].
 In other words, they hope to learn something from conducting the study.
@@
 Researchers cannot know what exactly they will learn from a study before they have run it.
-However, they can develop an expectation that helps them decide which study to conduct in what manner.
+However, they can develop an expectation that helps them decide about the specifics of a planned study.
 This expectation is what we term expected epistemic value.
@@
 Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|\neg{}H)$.
-$P(E|\neg{}H)$ is equivalent to the type-I error rate, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
+$P(E|\neg{}H)$ is equivalent to the type-I error rate in statistical hypothesis testing, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
 Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
@@
 
-Statistical methods stand out among these factors because we have a large toolbox for assessing and controlling their contribution to theoretical risk.
-Examples of our ability to exert this control are the setting of type-I error rate, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
+Statistical methods stand out among these factors because we have a large and well-understood toolbox for assessing and controlling their contribution to theoretical risk.
+Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
 These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none $P(E|\neg{}H)$.
@@
 However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, or any number of other things.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
 All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.

### 2023-02-16 manuscript.Rmd
Update manuscript.Rmd

@@
 The third and last argument is that preregistrations reduce this uncertainty.
-Following the last argument, a preregistered study is represented by the solid line (certainty about theoretical risk) and a study that was not preregistered is more similar to the dashed line (maximally uncertain about theoretical risk) in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
+Following the last argument, a preregistered study is represented by the solid line (certainty about theoretical risk) and a study that was not preregistered is more similar to the dashed line (maximally uncertain about theoretical risk) in Figure \ref{fig:posterior} and Figure \ref{fig:other-measures}.
 
-Recollect, our three assumptions:
+Let us recall our three assumptions:
 
@@
 
-The point we make with these assumptions is that researchers aim to persuade other researchers, the readers of their articles.
-We have to remember that the process of weighing evidence for or against a theory extends beyond the original authors to all the people they hope to persuade.
-Unfortunately, the case for lack of insight into the myriad of factors that influence theoretical risk is particularly strong for the reader of a resulting article (or, more generally, the consumer of the research product).
-While the authors may have deep insight into what they did and how it might influence the theoretical risk they took, their readers have much greater uncertainty about these factors.
-In particular, they never know which relevant factors the authors of a given article failed to disclose, intentionally or not.
-From the perspective of an ultimate sceptic, they may claim maximum uncertainty.
+The point we make with these assumptions is that researchers aim to persuade other researchers, for example, the readers of their articles.
+Not only the original authors are concerned with the process of weighing evidence for or against a theory but really all people the authors of a study hope to persuade.
+Unfortunately, readers of a scientific article (or, more generally, any consumer of a research product), is likely to lack insight into the various factors that influence theoretical risk.
+While the authors themselves may have a clear picture of what they did and how it might have influenced the theoretical risk they took, their readers have much greater uncertainty about these factors.
+In particular, they never know which relevant factors the authors of a given article failed to disclose, be it intentionally or not.
+From the perspective of the ultimate sceptic, they may claim maximum uncertainty.
 
-Communicating clearly how the authors gathered the data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
+Communicating clearly how authors of a scientific report gathered their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
 Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
-For instance, the authors could have decided to exclude a number of analytic strategies they tried out.
-That is not to say that every study that was not-preregistered was subjected to practices of p hacking.
-The point is, that we cannot exclude this and innumerable possibilities and, hence, are left uncertain.
+For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
+That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
+The point is, that we cannot exclude it with certainty.
 This uncertainty is drastically reduced, if the researchers have described what they intended to do beforehand, and then report that they did exactly that.

### 2023-02-16 manuscript.Rmd
Update manuscript.Rmd

@@
 In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
-A study with a perfect preregistration will, therefore, resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
-Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not the the theoretical risk is necessarily high.
+Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
+Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.
 
@@
 To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
-The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal and central goal of confirmatory research.
-However, theoretical risk is something researcher have only limited control over, e.g., theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).
-While we believe that preregistration is always beneficial, it might be counterproductive to pursue high theoretical risk if the conditions for highly confirmatory research are not given.
-Specifically, this entails the development of sufficiently precise theories and having the resources (typically sample size) to adequately balance detectability against theoretical risk.
+The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal and central goal of preregistration ofr confirmatory research agendas.
+However, theoretical risk is something researcher have only limited control over, for example, theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).
+While we believe that preregistration is always beneficial, it might be counterproductive to pursue high theoretical risk if the research context is inappropriate for strictly confirmatory research.
+Specifically, appropriateness here entails the development of precise theories and the availability of necessary resources (often, large enough sample size, but also see @brandmaier2015lifespan) to adequately balance detectability against theoretical risk.
 

### 2023-02-16 manuscript.Rmd
add center affiliation

@@
   - name: Aaron Peikert
-    affil: a, b
+    affil: a, b, c
     email: peikert@mpib-berlin.mpg.de
   - name: Andreas M. Brandmaier
-    affil: a, c, d
+    affil: a, b, c, d 
     email: brandmaier@mpib-berlin.mpg.de
@@
     address: |
-      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
+     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
   - num: d
     address: |
-     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
+      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
 repro:

### 2023-02-16 manuscript.Rmd
fix citation

@@
   - name: Andreas M. Brandmaier
-    affil: a, b, c, d 
+    affil: a, b, d 
     email: brandmaier@mpib-berlin.mpg.de
@@
     address: |
-      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
+     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
   - num: c
     address: |
-     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
+      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
   - num: d

### 2023-02-16 manuscript.Rmd
Apply suggestions from code review

@@
 <!---outline-structure--->
-To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate in statistical test theory.
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.
 
@@
 1. Researchers judge the evidence for or against a hypothesis rationally.
-2. They expect other researchers to apply the same rational process.
-2. All else being equal, researchers try to increase the expected epistemic value for other researchers.
+2. They expect other researchers to apply a similar rational process.
+2. Researchers try to maximize the expected epistemic value for other researchers.
 
@@
 To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
-The former result, that increasing theoretical risk leads to higher expected epistemic value, reconstructs the appeal and central goal of preregistration ofr confirmatory research agendas.
+The former result, that increasing theoretical risk leads to higher expected epistemic value reconstructs the appeal and central goal of preregistration of confirmatory research agendas.
 However, theoretical risk is something researcher have only limited control over, for example, theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).

### 2023-02-16 manuscript.Rmd
julias suggestions

@@
 ---
-title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization."
+title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
 bibliography: references.bib
@@
 keywords: |
-  preregistration; confirmation; exploration; hypothesis testing; bayesian; open science
+  preregistration; confirmation; exploration; hypothesis testing; Bayesian; open science
 header-includes: |
@@
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
-There is a broad consensus that both approaches are necessary for science to progress; exploration for making new discoveries and confirmation for exposing these discoveries to potential falsification and assess in how far they have been corroborated.
+There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification and assess in how far the theory has been corroborated.
 However, mistaking exploratory findings for empirically confirmed results is dangerous.
@@
 Recognizing them has led to a crisis of confidence in the empirical sciences [@ioannidisWhyMostPublished2005], and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
-As a response to the crisis, more and more researchers preregister their hypotheses and their data collection and analysis plans in advance of the study [@nosekPreregistrationRevolution2018].
+As a response to the crisis, evermore researchers preregister their hypotheses and their data collection and analysis plans in advance of their studies [@nosekPreregistrationRevolution2018].
 They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".
 Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
-This promise of a clear distinction between confirmation and exploration has obvious appeal to many who already accepted the practice.
+This promise of a clear distinction between confirmation and exploration has obvious appeal to many who have already accepted the practice.
 Still, the majority of empirical researchers do not routinely preregister their studies.
@@
 <!--the current goal of preregistration is confirmation-->
-One goal of preregistration that has received widespread attention, is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
+One goal of preregistration that has received widespread attention is to clearly distinguish confirmatory from exploratory research [@mellorEasyPreregistrationWill2018; @nosekPreregistrationRevolution2018; @wagenmakersAgendaPurelyConfirmatory2012; @simmonsPreregistrationWhyHow2021; @bakkerEnsuringQualitySpecificity2020].
 In such a narrative, preregistration is justified by a confirmatory research agenda.
 However, two problems become apparent under closer inspection.
-First, many researcher do not subscribe to a purely confirmatory research agenda.
+First, many researchers do not subscribe to a purely confirmatory research agenda.
 Second, it is not necessary to strictly map the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
@@
 When researchers criticize preregistration as being too inflexible to fit their research question, they often simply acknowledge that their research goals are not strictly confirmatory.
-Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they ask.
-However reasonable such a move is, changing core believes of a large community is much harder than convincing them that a method is well justified.
+Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they pose.
+However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
 We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
@@
 <!---outline-structure--->
-To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong, if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.
 
@@
 The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
-We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and such.
-Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather the uncertainty in judging the theoretical risk is minimized.
+We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
+Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather that the uncertainty in judging the theoretical risk is minimized.
 
@@
 They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch Book" argument [@christensenCleverBookiesCoherent1991].
+This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
 This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
@@
 
-Let us make some simplifying assumptions and define notation.
+Let us make some simplifying assumptions and define our notation.
 We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
@@
 The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
-If a hypothesis is more probable to begin with, observing evidence in favor will result a more corroborated hypothesis, all else equal.
+If a hypothesis is more probable to begin with, observing evidence in its favor will result a more corroborated hypothesis, all else being equal.
 However, the prior probability of a hypothesis is nothing our study design can change.
@@
 Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high reward.
+In other words, high risk, high gain.
 
 If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
-This tension between high risk leading to high reward is central to our consideration of preregistration.
-A high-risk, high-reward strategy is bound to result in many losses that are eventually absorbed by the high gains.
+This tension between high risk leading to high gain is central to our consideration of preregistration.
+A high-risk, high-gain strategy is bound to result in many losses that are eventually absorbed by the high gains.
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
@@
 
-To better understand how preregistrations can achieve that, let us take a closer look at what factors contribute to $P(E)$. 
+To better understand how preregistrations can achieve that, let us take a closer look at the factors contributing to $P(E)$. 
 Using the law of total probability, we can split $P(E)$ into two terms:
@@
 
-We already have noted that there is not much to do about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability $P(E|H)$.
-The real lever to pull is, therefore, $P(E|\neg{}H)$.
+We have already noted that there is not much to be done about prior probability ($P(H)$, and hence its counter probability $P(\neg{}H)$), and that it is common sense to increase detectability $P(E|H)$.
+The real lever to pull is therefore $P(E|\neg{}H)$.
 This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
@@
 We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
-@kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in purely Bayesian framework.
-However, while he did not name $P(\neg{}E|\neg{}H)$ but did suggest that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|\neg{}H)$ which would imply that $P(\neg{}E|\neg{}H)$ can be related to or even equated to theoretical risk.
+@kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in a purely Bayesian framework.
+However, while he did not mention $P(\neg{}E|\neg{}H)$ he suggested that @meehlTheoreticalRisksTabular1978 used the term "very strange coincidence" for a small $P(E|\neg{}H)$ which would imply, that $P(\neg{}E|\neg{}H)$ can be related to or even equated to theoretical risk.
 
@@
 18].
-We note these relations to underline that the Bayesian rational we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
 
-Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
-To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
+Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate countless influences, otherwise they could not model the process of evidential support for theories.
+To illustrate the concepts we have introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
 The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
@@
 1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-2. The experimenter shows the solution "IX" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
+2. The experimenter shows the solution "VIII" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
 3. The experimenter does nothing; if all students come up with the solution, she records it as evidence for the hypothesis.
@@
 We argue that experiment 1 has high theoretical risk $P(\neg{}E_1|\neg{}H)$ and high detectability $P(E_1|H)$.
-If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals could be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
+If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals can be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
 Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability $P(E_1|H)$.
@@
 Experiment 3, however, also has low detectability.
-It is unlikely that all students come up with the correct solution in a short time ($E_3$), even if insight is required ($H$) experiment 3 therefore has low detectability $P(E_3|H)$.
+It is unlikely that all students will come up with the correct solution in a short time ($E_3$), even if insight is required ($H$); experiment 3 therefore has low detectability $P(E_3|H)$.
 The theoretical risk, however, is also low in absolute terms, but high compared to the detectability (statistical relevancy condition is satisfied).
@@
 
-As argued earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
+As mentioned earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
 We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.
@@
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
-We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing it again.
+We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
 The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.
@@
 Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that increase the likelihood of signifying results even when there are none $P(E|\neg{}H)$.
+These tools help us account for biases in statistical methods that decrease theoretical risk (and its inverse, $P(E|\neg{}H$).
 
@@
 
-As we noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
-However, other factors in the broader context of the study are much harder to quantify.
+As we have noted, it is possible to quantify how statistical models affect the theoretical risk based on mathematical considerations and simulation.
+However, other factors in the broader context of a study are much harder to quantify.
 If one chooses to focus only on the contribution of statistical methods to theoretical risk, one is bound to overestimate it.
@@
 Under ideal circumstances (assumption of independence, normality of residuals, equal variance), it stays true to its type-I error rate.
-However, researchers might do many very reasonable things in the broader context of the study that affect theoretical risk:
+However, researchers may do many very reasonable things in the broader context of the study that affect theoretical risk:
 They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
-All of these decisions carry a small risk that they increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
+All of these decisions carry a small risk that they will increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
 Even if the t-test itself perfectly maintains its type I error rate, these factors influence $P(E|\neg{}H)$.
@@
 Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
-While one can hope that processing this information does not affect their decision making either way, this cannot be secured.
-We, therefore, conclude that statistical properties only guarantee a lower bound for theoretical risk.
+While one can hope that processing this information does not affect their decision making either way, this cannot be ascertained.
+Therefore, we conclude that statistical properties only guarantee a lower bound for theoretical risk.
 The only thing we can conclude with some certainty is that theoretical risk is not higher than what the statistical model guarantees without knowledge about the other factors at play.
@@
 Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
-We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and conforms therefore to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
@@
 ```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:posterior}Posterior Probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:posterior}Posterior probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 ```{r other-measures-fig, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:other-measures}Several measures for corroboration as increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:other-measures}Several measures for corroboration as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions into which all realistic conditions fall.
-Depending on how uncertain we are about the theoretical risk a study took on, our expectation the gained epistemic value varies considerably.
+Depending on how uncertain we are about the theoretical risk a study took on, our expectation of the gained epistemic value varies considerably.
 Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-Generally, we expect that increases in uncertainty (expressed as more entropic distributions) lead to a decreased expected epistemic value.
+Generally, we expect increases in uncertainty (expressed as more entropic distributions) to lead to a decreased expected epistemic value.
 
@@
 
-Communicating clearly how authors of a scientific report gathered their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
-Preregistrations are ideal to communicate just that, because any description after the fact is suspect to be incomplete.
+Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
+Preregistrations are ideal to communicate just that, because any description after the fact is prone to be incomplete.
 For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
 That is not to say that every study that was not-preregistered was subjected to practices of questionable research practices.
-The point is, that we cannot exclude it with certainty.
-This uncertainty is drastically reduced, if the researchers have described what they intended to do beforehand, and then report that they did exactly that.
+The point is that we cannot exclude it with certainty.
+This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand, and then report that they did exactly that.
 In that case, the readers can be certain, that they have received a complete account of the situation.
-They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extend than if the study would not have been preregistered.
+They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
 Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
-In any case a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+In any case, a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
 Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
@@
 Ernst, Caspar van Lissa, Felix Schönbrodt, the discussants at the DGPS2022 conference and Open Science Center Munich, and many more for the insightful discussions about disentangling preregistration and confirmation.
+We are grateful to Julia Delius for her helpful assistance in language and style editing.
 

### 2023-02-17 manuscript.Rmd

@@
 keywords: |
-  preregistration; confirmation; exploration; hypothesis testing; Bayesian; open science
+  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
 header-includes: |

### 2023-02-17 manuscript.Rmd
update releases

@@
 This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as preprint [@no-preprint-yet] and will be submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
+The manuscript is available as preprint [@peikertWhyDoesPreregistration2023] and was submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
 

### 2023-02-17 manuscript.Rmd
make link to preprint clearer

--- 2023-02-17 manuscript.Rmd
@@
 This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as preprint [@no-preprint-yet] and will be submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
+The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
 

### 2023-02-20 manuscript.Rmd
change to springer layout

--- 2023-02-17 manuscript.Rmd
@@
 title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
+titlerunning: "Preregistration increases the persuasiveness of evidence"
 bibliography: references.bib
@@
     - inkscape
-output: rticles::tf_article
-type: ARTICLE TEMPLATE
-keywords: |
-  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
+output: rticles::springer_article
+keywords:
+  - preregistration
+  - confirmation
+  - exploration
+  - hypothesis testing
+  - Bayesian
+  - Open Science
+biblio-style: spbasic
 header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
+  \usepackage{amsmath}
+  \usepackage{amssymb}
 ---
@@
 We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
-It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
+It is likely that empirical sciences need more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
 
@@
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
+This is not a risky prediction to make in empirical sciences, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
@@
 This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as preprint [@no-preprint-yet] and will be submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
+The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
 

### 2023-02-20 manuscript.Rmd
they require self citation in third person

@@
 
-We have proposed a workflow for preregistration called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+A workflow for preregistration called preregistration as code (PAC) that has several beneficial properties for uncertainty reduction was proposed in [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.

### 2023-02-20 manuscript.Rmd
anonymize

@@
 
-The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
+The materials for this article are available on GitHub [@zenodo-release].
 This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
 

### 2023-05-05 manuscript.Rmd
first half of manuscript

--- 2023-02-20 manuscript.Rmd
@@
 title: "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
-titlerunning: "Preregistration increases the persuasiveness of evidence"
 bibliography: references.bib
@@
     - inkscape
-output: rticles::springer_article
-keywords:
-  - preregistration
-  - confirmation
-  - exploration
-  - hypothesis testing
-  - Bayesian
-  - Open Science
-biblio-style: spbasic
+output: rticles::tf_article
+type: ARTICLE TEMPLATE
+keywords: |
+  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
 header-includes: |
-  \usepackage{amsmath}
-  \usepackage{amssymb}
+  \usepackage{hyperref}
+  \usepackage[utf8]{inputenc}
+  \def\tightlist{}
 ---
@@
 One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
+<!-- it thought more of an analysis than a theory -->
 We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.
@@
 <!---false dichotomy--->
-Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions whatsoever, and so forth.
+Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions, and so forth.
 The opposite, that is, preregistered but not strictly confirmatory studies, are also becoming more commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
@@
 Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
+<!-- harsh transition -->
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
@@
 However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
-We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
-It is likely that empirical sciences need more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
+We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* <!-- the word ideological may have a negative connotation. --> goals of confirmatory science.
+It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
+<!-- I think psychology does not need more confirmatory studies, and also this is not really the focus of this paper, so I would remove this.
+I believe psychology needs more confirmatory studies that are really confirmatory (and not p-hacked) and more exploratory studies that don't disguise as confirmatory.
+Using preregistration in a way that quantifies where a study is on this continuum would maybe help to encourage people making more (semi-)exploratory studies, I believe this is a point that you could make here, if you want to.
+-->
 
@@
 
-Further, we outline two interpretations of how preregistration impacts theoretical risk.
-The first interpretation corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
-The second interpretation is our main contribution and demonstrates the broad applicability of preregistration to both exploratory and confirmatory studies that are implemented as preregistered or have undergone changes after preregistration.
-We argue that the classic view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
-Importantly, we argue that theoretical risk is not necessarily directly maximized by preregistration, but rather that the uncertainty in judging the theoretical risk is minimized.
+Further, we outline two interpretations of preregistration.
+The first one, commonly found among applied scientists, is that preregistration impacts theoretical risk.
+<!-- I think your second interpretation is all about the fact that preregistration does not (mainly) impact theoretical risk, but the uncertainty about it.
+-->
+It corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
+We argue that this view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
+The second interpretation is our main contribution: We argue that, in contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk, but rather the minimization of uncertainty about the theoretical risk.
+This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.
 
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases our expectation to gain evidence for a theory, preregistration is potentially useful for any kind of study, no matter how exploratory.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases the amount of confirmation we gain from a particular study<!-- I would rephrase this...
+atm it sounds a bit like an increase in P(E|H), but you mean an increase in P(H|E) -->, preregistration is potentially useful for any kind of study, no matter how exploratory.
 
@@
 In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory (which itself is an ontologocial concept) is what we call epistemic value.
+The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
 Researchers cannot know what exactly they will learn from a study before they have run it.
@@
 2. Researchers try to maximize the expected epistemic value for other researchers.
-
+<!-- very strong assumptions -->
 The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
-Our rationale is as follows.
-Researchers who decide to conduct a certain study are actually choosing a study to bet on.
-They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
-This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
-This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
-Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
 The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
@@
 However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
-That is, evidence only increases epistemic value for a theory, if the evidence is more likely to observe under the theory than without it.
+That is, evidence only increases epistemic value for a theory, if the evidence is more likely to be observed under the theory than under the alternative.
 
 Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
-In fact, this conceptualization is intentionally reductionistic to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 
@@
 Let us make some simplifying assumptions and define our notation.
-We restrict ourselves to evidence of a binary nature (either it was observed or not) since continuous evidence would lead to some quite involved derivations.
+To keep the notation simple, we restrict ourselves to evidence of a binary nature (either it was observed or not).
 We denote the probability of a hypothesis before observing evidence as $P(H)$ and its complement as $P(\neg{}H) = 1 - P(H)$.
@@
 
-The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of corroboration of a hypothesis.
-In the tradition of Carnap, in its direct use, it is called corroboration as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
+The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
+In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
 As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
-However, it is reasonable that any measure of corroboration increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
+However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
 We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
 We referred to this assumption as "statistical relevance condition" earlier [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
+<!-- I believe it is confirmation, not corroboration.
+Also: do we really need the statistical relevance condition?
+What is the relation between the relevance condition and the monotinic increase in confirmation?
+Also I think you should cite literature that says that this monotonic increase is indeed a reasonable assumption.
+-->
 
 In short, we want to increase posterior probability $P(H|E)$.
-Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, for which we want to maximize the expectation.
+Increases in posterior probability $P(H|E)$ are associated with increased epistemic value. <!-- what is the difference between epistemic value and confirmation? -->
 So how can we increase posterior probability?
@@
 The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
-If a hypothesis is more probable to begin with, observing evidence in its favor will result a more corroborated hypothesis, all else being equal.
+If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal.
 However, the prior probability of a hypothesis is nothing our study design can change.
@@
 $P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-We call this probability of detecting evidence given that the theory holds "detectability."
+We call this probability of detecting evidence given that the theory holds "detectability." <!-- citation for the source of "detectability" -->
 Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
@@
 Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
-In other words, high risk, high gain.
+In other words, high risk, high reward.
+<!--I believe this is not true.
+P(E) is the sum of two terms (as you write below).
+Increasing the first term will increase posterior probability, increasing the second term will decrease posterior probability.
+High risk, high reward does not refer to P(E), it refers to P(H) and P(\not E| \not H): assume the researcher is wrong with his hypothesis, how probable is it that the researcher will actually be proven wrong.
+As P(E|\not H) = 1- P(\not E| \not H), increasing the riskiness while P(H) is low will decrease the second term in the denominator, thereby increasing P(H|E) -->
 
-If we equate riskiness with a low probability of obtaining evidence, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
+If we equate riskiness with a low probability of obtaining evidence under the assumption that our theory is wrong, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
+<!-- se comment before -->
 This tension between high risk leading to high gain is central to our consideration of preregistration.
 A high-risk, high-gain strategy is bound to result in many losses that are eventually absorbed by the high gains.
+<!-- Keep in mind this is only true for Hypotheses with low P(H).
+If P(H) is high, it does not really matter how risky your study is, if detectability is high, you will confirm the hypothesis (e.g.
+this corresponds to a setting where you test the certain consequence of a certain hypothesis, H: no living human is larger than 500 meters E: measure how tall every human on earth is.
+P(E|H) = 1, P(\not E| \not H) = 1, so P(H|E) = P(H) -->
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
@@
 To better understand how preregistrations can achieve that, let us take a closer look at the factors contributing to $P(E)$. 
-Using the law of total probability, we can split $P(E)$ into two terms:
+Using the law of total probability, we can split $P(E)$ into two terms: <!-- I would move this further above to beeing able to better explain riskiness -->
 
@@
 If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
-Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
+Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1). <!-- sure? -->
 In other words, observing a sure fact does not lend credence to a hypothesis.
@@
 
-1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
+1.
+The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
+<!-- I dont really yunderstand how this relates to insight...
+-->
 2. The experimenter shows the solution "VIII" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
@@
 
-As mentioned earlier, the remainder of the paper will focus on binary, non-probabilisitic evidence to keep the mathematical notation as simple as possible.
+As mentioned earlier, the we restrict ourselves to binary evidence, to keep the mathematical notation as simple as possible.
 We discuss the relation between statistical methods and theoretical risk in the [Statistical Methods] section.
@@
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in empirical sciences, even without statistical issues like alpha inflation [@orbenCrudReDefined2020].
-However, we would consider it a success if the researcher would use the evidence to develop a more precise (and therefore risky) theory, e.g., by specifying which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc.
+This is not a risky prediction to make in psychology [@orbenCrudReDefined2020]. <!-- I dont know if the citation relates to alpha-inflation or to the riskiness of such predictions -->
+However, we would consider it a success if the researcher would use the evidence from this rather exploratory study to develop a more precise (and therefore risky) theory, e.g., by using the results to specify which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc., to be able to make riskier predictions in the future.
 We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
-The point, however, is that we want to show that preregistration increases the expected epistemic value regardless of the theory being tested.
+This is because preregistration increases the expected epistemic value regardless of the theory being tested, as we will show.
 
@@
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
+<!-- This is true but I think it does not really relate to preregistration.
+With the same amount of resources, preregistration can limit researcher degrees of freedom, thereby increasing theoretical risk.
+-->
 
@@
 We have established that higher theoretical risk leads to more persuasive evidence.
-In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limits the type-I error rate) and thereby creates more compelling evidence.
+In other words, we have reconstructed the interpretation that preregistrations supposedly works by restricting the researchers, which in turn increases the theoretical risk (or equivalently limits the type-I error rate) and thereby creates more compelling evidence.
 Nevertheless, there are trade-offs for increasing theoretical risk.
 Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favorable interpretation.
-To that end, we incorporate uncertainty into our framework.
+To that end, we incorporate uncertainty about theoretical risk into our framework. <!-- you already have a lot of uncertainty in your framework, you now also consider this one -->
 
@@
 
-One factor that is known without much uncertainty is the contribution of statistical methods to theoretical risk.
-Theoretical risk is deeply connected with statistical methods, because it is the inverse of $P(E|\neg{}H)$.
-$P(E|\neg{}H)$ is equivalent to the type-I error rate in statistical hypothesis testing, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
+One widely known factor is the contribution of statistical methods to theoretical risk.
+Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$<!-- the inverse would be ^-1-->, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
+<!-- I believe this would only be true, if the occurence of a non-significant result would lead to a refutation of the theory.
+In NHST, theoretically P(\not E | \not H) may be the theoretical risk (the probability ob observing no effect under the assumption that the N-Hypothesis is true), but in practice, researchers don't count this as evidence against the alternative hypothesis - therefore, if you adapt the framework of NHST, theoretical risk is always zero and no confirmation at all takes place.
+-->
 Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
@@
 However, the research hypothesis seldomly equals  the statistical null hypothesis, and therefore, the relation between statistical type-I error rate and theoretical risk should not be overinterpreted.
+<!-- above you state this the other way around: research hypothesis = alternative hypothesis -->
 We argue that theoretical risk (and hence its inverse, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+<!-- yes!:) -->
 
@@
 Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that decrease theoretical risk (and its inverse, $P(E|\neg{}H$).
+These tools help us account for biases in statistical methods that influence theoretical risk.
 
@@
 For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to theoretical risk.
-The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
+The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods <!-- I think this is to broad...
+in other sciences, the modeling part may be more important than the statistical part.
+-->into ignoring other factors that are much more uncertain.
 We cannot hope to resolve this uncertainty; but we have to be aware of its implications.
@@
 However, researchers may do many very reasonable things in the broader context of the study that affect theoretical risk:
-They might exclude outliers, choose to drop an item before computing a sum score, broaden, enlarge their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
+They might exclude outliers, choose to drop an item before computing a sum score, broaden their definition of the population to be sampled, translate their questionnaires into a different language, impute missing values, switch between different estimators of the pooled variance, or any number of other things.
 All of these decisions carry a small risk that they will increase the likelihood of obtaining evidence despite the underlying research hypothesis being false.
@@
 Within the Bayesian framework, this is both straightforward and insightful.
-To get an expectation, we express uncertainty as a probability distribution and then integrate over it:
+Let's assume a researcher is reading a study from another lab and tries to decide whether and how much the presented results confirm the theory.
+As the researcher did not conduct the study (and the study is not preregistered), they can not be certain about the various factors influencing theoretical risk (researcher degrees of freedom).
+We therefore express this uncertainty about the theoretical risk as a probability distribution $Q$ of $P(E|\neg{}H)$ (remember that $P(E|\neg{}H)$ is related to theoretical risk by $P(E|\neg{}H)$ = 1 -$P(\neg{}E|\neg{}H)$, so it does not matter wheter we consider the distribution of theoretical risk or $P(E|\neg{}H)$).
+To get the expected value of $P(H|E)$ that follows from the researchers uncertainty about the theoretical risk, we can compute the expectation using Bayes theorem:
 
 \begin{equation}
-\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H)) \label{eq:integral}
+\mathbb{E}_Q[P(H|E)] = \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \label{eq:integral}
 \end{equation}
 
-To illustrate the effect of uncertainty, let $p(E|H) = 0.8$ (e.g., power of 80%) and $p(H) = 0.1$ and assume a uniform distribution for theoretical risk $p(\neg{}E|\neg{}H)$ of the form:
+Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a sensible example.
+Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([\tau, 1])$ where $\tau{}$ is set to 0.05, reflecting our assumption that a statistical model with a nominal type-I error rate of 5\% is used and this error rate gives a lower bound for $P(E|\neg{}H)$.
+We chose this uniform distribution as it is the maximum entropy distribution with support $[\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
-\begin{equation}
-f(x)=\begin{cases}
-  \frac{1}{\tau} & \mathrm{for}\ 0 \le x \le \tau,\\
-  0 & \mathrm{for}\ x<0\ \mathrm{or}\ x>\tau
-  \end{cases}
-\end{equation}
-
-Where $\tau{}$ is the upper bound of theoretical risk (e.g., 0.95 for a statistical model with a nominal type-I error rate of 5%).
-We chose this uniform distribution to capture our statement that a statistical model only guarantees an upper bound to theoretical risk (and since it is a probability the lower bound is 0) as it is the maximum entropy distribution under this assumption and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+With this, we derive the expected value of $P(H|E)$ as
+\begin{align}
+\mathbb{E}_Q[P(H|E)] &= \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \\
+&= \int_\mathbb{R} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} Q \\
+&= \int_{[\tau, 1]} (1-\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) x } \, \mathrm{d} x \\
+&= \frac{P(H)P(E|H)}{P(\neg{}H)(1-\tau)}\ln \left(\frac{P(H)P(E|H) + P(\neg{}H)}{P(H)P(E|H) + P(\neg{}H)\tau}\right)\label{eq:expectation}
+\end{align}
 
@@
 This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
+The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
 

### 2023-05-08 manuscript.Rmd
formula spacing
finish review

@@
 Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a sensible example.
-Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([\tau, 1])$ where $\tau{}$ is set to 0.05, reflecting our assumption that a statistical model with a nominal type-I error rate of 5\% is used and this error rate gives a lower bound for $P(E|\neg{}H)$.
+Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([\tau, 1])$ where $\tau{}$ is set to the nominal type-I error rate of the statistical model, reflecting our assumption that this error rate gives a lower bound for $P(E|\neg{}H)$.
 We chose this uniform distribution as it is the maximum entropy distribution with support $[\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
@@
 \begin{align}
-\mathbb{E}_Q[P(H|E)] &= \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \\
-&= \int_\mathbb{R} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} Q \\
-&= \int_{[\tau, 1]} (1-\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) x } \, \mathrm{d} x \\
+\mathbb{E}_Q[P(H|E)] &= \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \\[1em]
+&= \int_\mathbb{R} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} Q \\[1em]
+&= \int_{[\tau, 1]} (1-\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) x } \, \mathrm{d} x \\[1em]
 &= \frac{P(H)P(E|H)}{P(\neg{}H)(1-\tau)}\ln \left(\frac{P(H)P(E|H) + P(\neg{}H)}{P(H)P(E|H) + P(\neg{}H)\tau}\right)\label{eq:expectation}
@@
 
-Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being fully certain (solid line) or fully uncertain (dashed line) about the theoretical risk of a study.
-Neither absolute certainty nor uncertainty are realistic scenarios but represent the boundary conditions into which all realistic conditions fall.
+Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being certain (solid line) or uncertain (dashed line) about the theoretical risk of a study.
+<!-- I think the deleted sentence is not true - with "fully uncertain" you mean you use the maximum entropy distribution with support [\tau, 1], but I could for example realistically assume that in most cases, P(E|\not H) is larger than the nominal type 1 error rate, leading to a distribution that places more mass towards a higher value of P(E|\not H), leading to a lower expected value for P(H|E) -->
 Depending on how uncertain we are about the theoretical risk a study took on, our expectation of the gained epistemic value varies considerably.
 Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-Generally, we expect increases in uncertainty (expressed as more entropic distributions) to lead to a decreased expected epistemic value.
+<!-- same as before: take a low-entropy distribution that places its weight on high values of P(E|\not H) -> this leads to a lower expected value than the high entropy distributions. -->
 
@@
 However, studies do not only differ by how high the theoretical risk is but also how certain the recipient is about the theoretical risk.
-A study that has a very high theoretical risk, e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed, i.e., alpha error, but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
-The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit very of course with the specifics of the study.
-Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)`.
+A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
+The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit vary of course with the specifics of the study.
+Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)` <!-- I would prefer reporting the prior probability and the posteriors in dependence of certainty/uncertainty over report the ratio between the posteriors-->.
 
@@
 The point we make with these assumptions is that researchers aim to persuade other researchers, for example, the readers of their articles.
-Not only the original authors are concerned with the process of weighing evidence for or against a theory but really all people the authors of a study hope to persuade.
+Not only the original authors are concerned with the process of weighing evidence for or against a theory but really the whole scientific community the study authors hope to persuade.
 Unfortunately, readers of a scientific article (or, more generally, any consumer of a research product), is likely to lack insight into the various factors that influence theoretical risk.
@@
 In particular, they never know which relevant factors the authors of a given article failed to disclose, be it intentionally or not.
-From the perspective of the ultimate sceptic, they may claim maximum uncertainty.
+From the perspective of the ultimate sceptic, they may claim maximum uncertainty <!--as before: I think the ultimate sceptic would place more weight on the higher end of the distribution, resulting in a less entropic distribution -->.
 
@@
 This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand, and then report that they did exactly that.
-In that case, the readers can be certain, that they have received a complete account of the situation.
+In that case, readers can be certain they received a complete account of the situation.
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
@@
 Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
+<!-- I wonder if it would be beneficial to include a low-uncertainty distribution in the figure.
+Because there is not a linear mapping between "how uncertain" -> "expected value", so we might want to make shure that we don't arrive at a low expected value at the slightest uncertainty -->
 Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.
@@
 Precise predictions would have very high theoretical risk (they are likely incorrect if the theory is wrong).
-A statistical procedure, may have high or low theoretical risk depending on the specifics of the model used.
-Specifying the process, therefore, is in line with the rational we propose here, but is incompatible with strictly confirmatory research.
+A statistical procedure may have high or low theoretical risk depending on the specifics of the model used.
+Specifying the process, therefore, is in line with the rational we propose here, but is less reasonable when to goal of preregistration is supposed to be a strictly confirmatory research agenda.
 
@@
 However, under our rational some changes may be justified.
-Any change increases the uncertainty about the theoretical risk and may even decrease the theorethical risk.
+Any change increases the uncertainty about the theoretical risk and may even decrease the theoretical risk.
 The changes still may be worthwhile, if the negative outcomes may be offset by an increase in detectability due to the change.
@@
 In such case, detectability becomes zero because the data cannot be analyzed without a post-hoc decision about how to handle the missing data.
-Any such decision would constitute a deviation from the preregistration, which is perfectly waranted under our general objective.
+Any such decision would constitute a deviation from the preregistration, which is possible under our proposed objective.
 Note that a reader cannot rule out that the researchers leveraged the decision to decrease theoretical risk, i.e., picking among all options the one that delivers the most beneficial results for the theory (in the previous example, chosing between various options of handling missing values).
@@
 After all they have not anticipated the case and are left with a detectablilty of zero.
-Any decision will increase detectability to a non-zero value, likely offsetting the increase in uncertainty.
 The researchers also may do their best to argue that the deviation was not motivated by increasing theoretical risk, thereby, decreasing the uncertainty.
@@
 Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision [@steegenIncreasingTransparencyMultiverse2016].
+<!-- I think this paragraph is a bit long and repetitive -->
 
-As explained above, reduction in uncertainty as the objective for preregistration does not only explain some existing behavior, that does not align with confirmation as a goal, it also allows to form recommendations to improve the practice of preregistration.
+As explained above, reduction in uncertainty as the objective for preregistration does not only explain some existing practice, that does not align with confirmation as a goal, it also allows to form recommendations to improve the practice of preregistration.
 Importantly, we now have a theoretical measure to gauge the functionality of preregistrations, which can only help increase its utility.
@@
 The theoretical risk, however low, must be communicated clearly.
-Parts of the process left unspecified, imply uncertainty, which is what a preregistration should reduce.
-However, specifying procedures that can be expected to not work, will lead to deviation, and subsequently to larger uncertainty.
+Parts of the process left unspecified imply uncertainty, which preregistration should reduce.
+However, specifying procedures that can be expected to fail will lead to deviation and, subsequently, to larger uncertainty.
 
@@
 Such approach dissolves the distinction between the preregistration document and the final scientific report.
-Instead of fundamentally separate documents, preregistration, and final report are different versions of the same underlying dynamic document.
+Instead of separate documents, preregistration and final report are different versions of the same underlying dynamic document.
 Deviations from the preregistration can therefore be clearly (and if necessary, automatically) isolated, highlighted, and inspected using version control.
@@
 However, while a PAC is not the extend of exploration, it is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
+<!-- I don't understand "while a PAC is not the extend of exploration" -->
 Please note that while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also able to advance this goal.
@@
 Confirmatory research would almost always be less persuasive without preregistration, but in our view, preregistration primarily communicates the extent of confirmation, i.e., theoretical risk, of a study.
-Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected epistemic value.
+Clearly communicating theoretical risk is important because it reduces uncertainty and hence increases expected epistemic value.
 

### 2023-05-12 manuscript.Rmd

@@
 Still, the majority of empirical researchers do not routinely preregister their studies.
-One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory in advance.
-<!-- it thought more of an analysis than a theory -->
+One reason may be that some do not find that the theoretical advantages outweigh the practical hurdles, such as specifying every aspect of a theory and the corresponding analysis in advance.
 We believe that we can reach a greater acceptance of preregistration by explicating a more general objective of preregistration that benefits all kinds of studies, even those that allow data-dependent decisions.
@@
 
-Researchers may apply two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
-<!-- harsh transition -->
+This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
@@
 However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
-We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* <!-- the word ideological may have a negative connotation. --> goals of confirmatory science.
-It is likely that psychology needs more confirmatory studies to progress as a science, but quite independently of this, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
-<!-- I think psychology does not need more confirmatory studies, and also this is not really the focus of this paper, so I would remove this.
-I believe psychology needs more confirmatory studies that are really confirmatory (and not p-hacked) and more exploratory studies that don't disguise as confirmatory.
-Using preregistration in a way that quantifies where a study is on this continuum would maybe help to encourage people making more (semi-)exploratory studies, I believe this is a point that you could make here, if you want to.
--->
+We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
+It might well be the case that psychology needs more confirmatory studies to progress as a science.
+However, independently of such goal, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
 
@@
 Further, we outline two interpretations of preregistration.
-The first one, commonly found among applied scientists, is that preregistration impacts theoretical risk.
-<!-- I think your second interpretation is all about the fact that preregistration does not (mainly) impact theoretical risk, but the uncertainty about it.
--->
-It corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
+The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
 We argue that this view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
-The second interpretation is our main contribution: We argue that, in contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk, but rather the minimization of uncertainty about the theoretical risk.
+The second interpretation is our main contribution:
+We argue that, in contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk, but rather the minimization of uncertainty about the theoretical risk.
 This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases the amount of confirmation we gain from a particular study<!-- I would rephrase this...
-atm it sounds a bit like an increase in P(E|H), but you mean an increase in P(H|E) -->, preregistration is potentially useful for any kind of study, no matter how exploratory.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases the amount of knowledge we gain from a particular study, preregistration is potentially useful for any kind of study, no matter how exploratory.
 
@@
 This expectation is what we term expected epistemic value.
-To make our three arguments, we must assume three things about this estimation process and how it relates to what studies (preregistered vs not preregistered) to conduct.
+To make our three arguments, we must assume three things about what an ideal estimation process  entails and how it relates to what studies (preregistered vs not preregistered) to conduct.
 
@@
 2. Researchers try to maximize the expected epistemic value for other researchers.
-<!-- very strong assumptions -->
+
 The assumption of rationality can be connected to Bayesian reasoning and leads to our adoption of the framework.
+Our rationale is as follows.
+Researchers who decide to conduct a certain study are actually choosing a study to bet on.
+They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
+This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
+This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
+Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
+Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
 The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
@@
 However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
-We assume that any measure shows gain if, and only if, evidence is observed assuming $P(E|H) > P(E|\neg{}H)$.
-We referred to this assumption as "statistical relevance condition" earlier [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
-<!-- I believe it is confirmation, not corroboration.
-Also: do we really need the statistical relevance condition?
-What is the relation between the relevance condition and the monotinic increase in confirmation?
-Also I think you should cite literature that says that this monotonic increase is indeed a reasonable assumption.
--->
 
 In short, we want to increase posterior probability $P(H|E)$.
-Increases in posterior probability $P(H|E)$ are associated with increased epistemic value. <!-- what is the difference between epistemic value and confirmation? -->
+Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
 So how can we increase posterior probability?
@@
 $P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-We call this probability of detecting evidence given that the theory holds "detectability." <!-- citation for the source of "detectability" -->
+We call this probability of detecting evidence given that the theory holds "detectability."
 Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
-When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the inverse of type-II error rate).
+When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
 However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
 Thus, $P(E)$ remains to be considered.
-Since $P(E)$ is the denominator, increasing it will decrease the posterior probability: The more unlikely it is to observe evidence, the more it increases the probability of the hypothesis if we do observe it.
+Since $P(E)$ is the denominator, decreasing it can increase the posterior probability.
 In other words, high risk, high reward.
-<!--I believe this is not true.
-P(E) is the sum of two terms (as you write below).
-Increasing the first term will increase posterior probability, increasing the second term will decrease posterior probability.
-High risk, high reward does not refer to P(E), it refers to P(H) and P(\not E| \not H): assume the researcher is wrong with his hypothesis, how probable is it that the researcher will actually be proven wrong.
-As P(E|\not H) = 1- P(\not E| \not H), increasing the riskiness while P(H) is low will decrease the second term in the denominator, thereby increasing P(H|E) -->
 
-If we equate riskiness with a low probability of obtaining evidence under the assumption that our theory is wrong, the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
-<!-- se comment before -->
+If we equate riskiness with a low probability of obtaining evidence (when the hypothesis is false), the Bayesian rationale perfectly aligns with the observation that risky predictions lead to persuasive evidence.
 This tension between high risk leading to high gain is central to our consideration of preregistration.
 A high-risk, high-gain strategy is bound to result in many losses that are eventually absorbed by the high gains.
-<!-- Keep in mind this is only true for Hypotheses with low P(H).
-If P(H) is high, it does not really matter how risky your study is, if detectability is high, you will confirm the hypothesis (e.g.
-this corresponds to a setting where you test the certain consequence of a certain hypothesis, H: no living human is larger than 500 meters E: measure how tall every human on earth is.
-P(E|H) = 1, P(\not E| \not H) = 1, so P(H|E) = P(H) -->
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
@@
 To better understand how preregistrations can achieve that, let us take a closer look at the factors contributing to $P(E)$. 
-Using the law of total probability, we can split $P(E)$ into two terms: <!-- I would move this further above to beeing able to better explain riskiness -->
+Using the law of total probability, we can split $P(E)$ into two terms:
 
@@
 If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
-Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1). <!-- sure? -->
+Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
 In other words, observing a sure fact does not lend credence to a hypothesis.
@@
 
-1.
-The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
-<!-- I dont really yunderstand how this relates to insight...
--->
+1. The experimenter gives a hint that  the problem is easy to solve when using Roman numerals; if all students come up with the solution, she records it as evidence for the hypothesis.
 2. The experimenter shows the solution "VIII" and explains it; if all students come up with the solution, she records it as evidence for the hypothesis.
@@
 For lack of a better theory, they assume that some relation between any variables of the two sets exists.
-This is not a risky prediction to make in psychology [@orbenCrudReDefined2020]. <!-- I dont know if the citation relates to alpha-inflation or to the riskiness of such predictions -->
+This is not a risky prediction to make in psychology [@orbenCrudReDefined2020].
 However, we would consider it a success if the researcher would use the evidence from this rather exploratory study to develop a more precise (and therefore risky) theory, e.g., by using the results to specify which variables from one set relate to which variables from the other set, to what extent, in which direction, with which functional shape, etc., to be able to make riskier predictions in the future.
@@
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-<!-- This is true but I think it does not really relate to preregistration.
-With the same amount of resources, preregistration can limit researcher degrees of freedom, thereby increasing theoretical risk.
--->
 
@@
 Employing a mathematical framework allows us to navigate the trade-offs more effectively and move towards a second, more favorable interpretation.
-To that end, we incorporate uncertainty about theoretical risk into our framework. <!-- you already have a lot of uncertainty in your framework, you now also consider this one -->
+To that end, we incorporate uncertainty about theoretical risk into our framework.
 
@@
 One widely known factor is the contribution of statistical methods to theoretical risk.
-Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$<!-- the inverse would be ^-1-->, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
-<!-- I believe this would only be true, if the occurence of a non-significant result would lead to a refutation of the theory.
-In NHST, theoretically P(\not E | \not H) may be the theoretical risk (the probability ob observing no effect under the assumption that the N-Hypothesis is true), but in practice, researchers don't count this as evidence against the alternative hypothesis - therefore, if you adapt the framework of NHST, theoretical risk is always zero and no confirmation at all takes place.
--->
+Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
 Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
 Researchers who choose a smaller type-I error rate can be more sure of their results, if significant, because the theoretical risk is higher.
-However, the research hypothesis seldomly equals  the statistical null hypothesis, and therefore, the relation between statistical type-I error rate and theoretical risk should not be overinterpreted.
-<!-- above you state this the other way around: research hypothesis = alternative hypothesis -->
-We argue that theoretical risk (and hence its inverse, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-<!-- yes!:) -->
+However, this connection should not be overinterpreted for two reasons.
+First, accorindg to NHST, the absence of a significant result should not be interpreted as evidence against the hypothesis[@citation-needed].
+Second, the research hypothesis seldomly equals  the statistical alternative hypothesis.
+We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
 
@@
 Examples of our ability to exert this control are the choice of type-I error rate, adjustments for multiple testing, the use of corrected fit measures (i.e., adjusted R²), information criteria, or cross-validation in machine learning.
-These tools help us account for biases in statistical methods that influence theoretical risk.
+These tools help us account for biases in statistical methods that influence theoretical risk (and hence, $P(E|\neg{}H$).
 
@@
 For those models or assumptions where this is impossible, one can employ Monte Carlo simulation to estimate the contribution to theoretical risk.
-The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods <!-- I think this is to broad...
-in other sciences, the modeling part may be more important than the statistical part.
--->into ignoring other factors that are much more uncertain.
+The precision with which statisticians can discuss contributions to theoretical risk has lured the community concerned with research methods into ignoring other factors that are much more uncertain.
 We cannot hope to resolve this uncertainty; but we have to be aware of its implications.
@@
 Within the Bayesian framework, this is both straightforward and insightful.
-Let's assume a researcher is reading a study from another lab and tries to decide whether and how much the presented results confirm the theory.
+Let us assume a researcher is reading a study from another lab and tries to decide whether and how much the presented results confirm the theory.
 As the researcher did not conduct the study (and the study is not preregistered), they can not be certain about the various factors influencing theoretical risk (researcher degrees of freedom).
-We therefore express this uncertainty about the theoretical risk as a probability distribution $Q$ of $P(E|\neg{}H)$ (remember that $P(E|\neg{}H)$ is related to theoretical risk by $P(E|\neg{}H)$ = 1 -$P(\neg{}E|\neg{}H)$, so it does not matter wheter we consider the distribution of theoretical risk or $P(E|\neg{}H)$).
+We therefore express this uncertainty about the theoretical risk as a probability distribution $Q$ of $P(E|\neg{}H)$ (remember that $P(E|\neg{}H)$ is related to theoretical risk by $P(E|\neg{}H) = 1 - P(\neg{}E|\neg{}H)$, so it does not matter wheter we consider the distribution of theoretical risk or $P(E|\neg{}H)$).
 To get the expected value of $P(H|E)$ that follows from the researchers uncertainty about the theoretical risk, we can compute the expectation using Bayes theorem:
@@
 
-Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a sensible example.
-Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([\tau, 1])$ where $\tau{}$ is set to the nominal type-I error rate of the statistical model, reflecting our assumption that this error rate gives a lower bound for $P(E|\neg{}H)$.
+Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a example.
+Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to the nominal type-I error rate of the statistical model, reflecting our assumption that this error rate gives a lower bound for $P(E|\neg{}H)$.
 We chose this uniform distribution as it is the maximum entropy distribution with support $[\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
@@
 \mathbb{E}_Q[P(H|E)] &= \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \\[1em]
-&= \int_\mathbb{R} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} Q \\[1em]
-&= \int_{[\tau, 1]} (1-\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) x } \, \mathrm{d} x \\[1em]
-&= \frac{P(H)P(E|H)}{P(\neg{}H)(1-\tau)}\ln \left(\frac{P(H)P(E|H) + P(\neg{}H)}{P(H)P(E|H) + P(\neg{}H)\tau}\right)\label{eq:expectation}
+&= \int_{[1-\tau, 1]} (\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} P(E|\neg H) \\[1em]
+&= \frac{P(H)P(E|H)}{P(\neg{}H)\tau}\ln \left(\frac{P(H)P(E|H) + P(\neg{}H)}{P(H)P(E|H) + P(\neg{}H)(1-\tau)}\right)\label{eq:expectation}
 \end{align}
@@
 Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being certain (solid line) or uncertain (dashed line) about the theoretical risk of a study.
-<!-- I think the deleted sentence is not true - with "fully uncertain" you mean you use the maximum entropy distribution with support [\tau, 1], but I could for example realistically assume that in most cases, P(E|\not H) is larger than the nominal type 1 error rate, leading to a distribution that places more mass towards a higher value of P(E|\not H), leading to a lower expected value for P(H|E) -->
 Depending on how uncertain we are about the theoretical risk a study took on, our expectation of the gained epistemic value varies considerably.
 Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-<!-- same as before: take a low-entropy distribution that places its weight on high values of P(E|\not H) -> this leads to a lower expected value than the high entropy distributions. -->
+The increase in uncertainty (expressed as more entropic distributions) leads to a decreased expected epistemic value.
 
@@
 The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit vary of course with the specifics of the study.
-Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau`, so false positive rate of `r 1-some_exploration_tau`, moving from uncertain to certain increases the posterior by a factor of `r round(some_exploration_certain/some_exploration_uncertain, 2)` <!-- I would prefer reporting the prior probability and the posteriors in dependence of certainty/uncertainty over report the ratio between the posteriors-->.
+Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
 
@@
 In particular, they never know which relevant factors the authors of a given article failed to disclose, be it intentionally or not.
-From the perspective of the ultimate sceptic, they may claim maximum uncertainty <!--as before: I think the ultimate sceptic would place more weight on the higher end of the distribution, resulting in a less entropic distribution -->.
+From the perspective of the ultimate sceptic, they may claim maximum uncertainty.
 
@@
 Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
-<!-- I wonder if it would be beneficial to include a low-uncertainty distribution in the figure.
-Because there is not a linear mapping between "how uncertain" -> "expected value", so we might want to make shure that we don't arrive at a low expected value at the slightest uncertainty -->
 Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.
@@
 A statistical procedure may have high or low theoretical risk depending on the specifics of the model used.
-Specifying the process, therefore, is in line with the rational we propose here, but is less reasonable when to goal of preregistration is supposed to be a strictly confirmatory research agenda.
+Specifying the process, therefore, is in line with the rational we propose here, but is less reasonable when the goal of preregistration is supposed to be a strictly confirmatory research agenda.
 
@@
 After all they have not anticipated the case and are left with a detectablilty of zero.
+Any decision will increase detectability to a non-zero value offsetting the increase in uncertainty.
 The researchers also may do their best to argue that the deviation was not motivated by increasing theoretical risk, thereby, decreasing the uncertainty.
@@
 Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision [@steegenIncreasingTransparencyMultiverse2016].
-<!-- I think this paragraph is a bit long and repetitive -->
 
@@
 Crucially, because the preregistration contains code, it may accommodate many different data patterns, i.e., it may be exploratory.
-However, while a PAC is not the extend of exploration, it is very specific about exactly how probable it is to generate evidence even when the theory does not hold (theoretical risk).
-<!-- I don't understand "while a PAC is not the extend of exploration" -->
+However, while a PAC does not limit the extend of exploration, it is very specific about the probability to generate evidence even when the theory does not hold (theoretical risk).
 Please note that while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also able to advance this goal.
@@
 Confirmatory research would almost always be less persuasive without preregistration, but in our view, preregistration primarily communicates the extent of confirmation, i.e., theoretical risk, of a study.
-Clearly communicating theoretical risk is important because it reduces uncertainty and hence increases expected epistemic value.
+Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected epistemic value.
 

### 2023-05-12 manuscript.Rmd
add citation

@@
 However, this connection should not be overinterpreted for two reasons.
-First, accorindg to NHST, the absence of a significant result should not be interpreted as evidence against the hypothesis[@citation-needed].
+First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
 Second, the research hypothesis seldomly equals the statistical alternative hypothesis.

### 2023-05-12 manuscript.Rmd

@@
 Within the Bayesian framework, this is both straightforward and insightful.
-Let us assume a researcher is reading a study from another lab and tries to decide whether and how much the presented results confirm the theory.
+Let us assume a researcher is reading a study from another lab and tries to decide whether and how much the presented results confirm the hypothesis.
 As the researcher did not conduct the study (and the study is not preregistered), they can not be certain about the various factors influencing theoretical risk (researcher degrees of freedom).
@@
 Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a example.
-Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to the nominal type-I error rate of the statistical model, reflecting our assumption that this error rate gives a lower bound for $P(E|\neg{}H)$.
-We chose this uniform distribution as it is the maximum entropy distribution with support $[\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
+Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to $1 - \alpha$, reflecting our assumption that this term gives an upper bound for theoretical risk $P(\neg{}E|\neg{}H)$.
+We chose this uniform distribution as it is the maximum entropy distribution with support $[1-\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
 
@@
 \mathbb{E}_Q[P(H|E)] &= \mathbb{E}_Q \left[ \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \right] \\[1em]
-&= \int_{[1-\tau, 1]} (\tau)^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} P(E|\neg H) \\[1em]
+&= \int_{[1-\tau, 1]} \tau^{-1} \frac{P(H)P(E|H)}{P(H)P(E|H)+P(\neg H) P(E|\neg H) } \, \mathrm{d} P(E|\neg H) \\[1em]
 &= \frac{P(H)P(E|H)}{P(\neg{}H)\tau}\ln \left(\frac{P(H)P(E|H) + P(\neg{}H)}{P(H)P(E|H) + P(\neg{}H)(1-\tau)}\right)\label{eq:expectation}

### 2023-05-13 manuscript.Rmd
remove leftover "corroboration"

@@
 So how can we increase posterior probability?
-The Bayes formula yields three components that influence corroboration, namely $P(H)$, $P(E|H)$ and $P(E)$.
+The Bayes formula yields three components that influence confirmation, namely $P(H)$, $P(E|H)$ and $P(E)$.
 The first option leads us to the unsurprising conclusion that higher a priori probability $P(H)$ leads to higher posterior probability $P(H|E)$.
@@
 ```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:posterior}Posterior probability (corroboration as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:posterior}Posterior probability (confirmation as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 ```{r other-measures-fig, dev = "tikz", cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:other-measures}Several measures for corroboration as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 

### 2023-05-13 manuscript.Rmd
grammerly

@@
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
-There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification and assess in how far the theory has been corroborated.
+There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
 However, mistaking exploratory findings for empirically confirmed results is dangerous.
-It inflates the likelihood of believing that there is evidence supporting a given explanation even if it is false.
+It inflates the likelihood of believing that there is evidence supporting a given hypothesis, even if it is false.
 A variety of problems, such as researchers' degrees of freedom together with researchers' hindsight bias or naive p-hacking have led to such mistakes becoming commonplace yet unnoticed for a long time.
@@
 As a response to the crisis, evermore researchers preregister their hypotheses and their data collection and analysis plans in advance of their studies [@nosekPreregistrationRevolution2018].
-They do so to stress the predictive nature of their registered statistical analyses; often with the hopes of obtaining a label that marks the study as "confirmatory".
+They do so to stress the predictive nature of their registered statistical analyses, often with the hopes of obtaining a label that marks the study as "confirmatory".
 Indeed, rigorous application of preregistration prevents researchers from reporting a set of results produced by an arduous process of trial and error as a simple confirmatory story [@wagenmakersAgendaPurelyConfirmatory2012] while keeping low false-positive rates.
@@
 First, many researchers do not subscribe to a purely confirmatory research agenda.
-Second, it is not necessary to strictly map the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
+Second, there is strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
 
 <!---false dichotomy--->
-Obviously, researchers can conduct confirmatory research without preregistration— though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions, and so forth.
+Obviously, researchers can conduct confirmatory research without preregistration — though it might be difficult to convince other researchers of the confirmatory nature of their research, that is, that they were free of cognitive biases, made no data-dependent decisions, and so forth.
 The opposite, that is, preregistered but not strictly confirmatory studies, are also becoming more commonplace [@dwanSystematicReviewEmpirical2008; @chanEmpiricalEvidenceSelective2004; @silagyPublishingProtocolsSystematic2002].
 
-This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterwards.
+This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
-Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful, and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
+Thus, insisting on equating preregistration and confirmation has led to the criticism that, all things considered, preregistration is actually harmful and neither sufficient nor necessary for doing good science [@szollosiPreregistrationWorthwhile2020; @phamPreregistrationNeitherSufficient2021].
 
@@
 Forcing researchers into adopting a strictly confirmatory research agenda does not only imply changing *how* they investigate a phenomenon but also *what* research questions they pose.
-However reasonable such a move is, changing core beliefs of a large community is much harder than convincing them that a method is well justified.
-We therefore attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
+However reasonable such a move is, changing the core beliefs of a large community is much harder than convincing them that a method is well justified.
+We, therefore, attempt to disentangle the *methodological* goals of preregistration from the *ideological* goals of confirmatory science.
 It might well be the case that psychology needs more confirmatory studies to progress as a science.
-However, independently of such goal, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
+However, independently of such a goal, preregistration can be useful for any kind of study on the continuum between strictly confirmatory and fully exploratory.
 
@@
 Further, we outline two interpretations of preregistration.
-The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or equivalently by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
-We argue that this view on the utility of preregistration can be interpreted as maximization of theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
+The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
+We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
 The second interpretation is our main contribution:
-We argue that, in contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk, but rather the minimization of uncertainty about the theoretical risk.
+We argue that contrary to the classic view, the objective of preregistration is *not* the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk.
 This interpretation leads to a broad applicability of preregistration to both exploratory and confirmatory studies.
@@
 The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-Researchers cannot know what exactly they will learn from a study before they have run it.
+Researchers cannot know what exactly they will learn from a study before they run it.
 However, they can develop an expectation that helps them decide about the specifics of a planned study.
@@
 Researchers who decide to conduct a certain study are actually choosing a study to bet on.
-They have to "place the bet" by conducting the study, therefore, invest resources and stand to gain epistemic value with some probability.
+They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
 This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
 This argument states that any better must follow the axioms of probability to avoid being "irrational," i.e., accepting bets that lead to sure losses.
-Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modelling what liberties researchers have in what studies to conduct.
+Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
 Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
@@
 However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
-That is, evidence only increases epistemic value for a theory, if the evidence is more likely to be observed under the theory than under the alternative.
+That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
 
@@
 In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
-As noted before, we concentrate on posterior probability as a measure of epistemic value, since no measure shows universally better properties than others.
+As noted before, we concentrate on posterior probability as a measure of epistemic value since no measure shows universally better properties than others.
 However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
@@
 However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to higher posterior probability $P(H|E)$.
-$P(E|H)$ is the probability of obtaining evidence for a theory, when the theory holds.
-We call this probability of detecting evidence given that the theory holds "detectability."
+The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
+$P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
+We call this probability of detecting evidence, given that the hypothesis holds "detectability."
 Consequently, researchers should ensure that their study design allows them to find evidence for their hypothesis, in case it is true.
@@
 Sustaining many "failed" studies is not exactly aligned with the incentive structure under which many, if not most, researchers operate.
-Consequently, researchers have an incentive to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
+Consequently, researchers are incentivized to appear to take more risks than they actually do, which misleads their readers to give their claims more credence than they deserve.
 It is at this juncture that the practice and mispractice of preregistration comes into play.
@@
 This probability tells us how likely it is that we find evidence in favor of the theory when in fact, the theory is not true.
-Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favour.
+Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favor.
 We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
@@
 
-Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate countless influences, otherwise they could not model the process of evidential support for theories.
-To illustrate the concepts we have introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
+Both theoretical risk $P(\neg{}E|\neg{}H)$ and detectability $P(E|H)$ aggregate countless influences; otherwise, they could not model the process of evidential support for theories.
+To illustrate the concepts we have introduced here, consider the following example of a single theory and three experiments that may test it. 
 The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
@@
 We argue that experiment 1 has high theoretical risk $P(\neg{}E_1|\neg{}H)$ and high detectability $P(E_1|H)$.
-If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals can be used, should not lead to all students solving the problem ($\neg{}E_1$); the experiment therefore has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
-Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment therefore has high detectability $P(E_1|H)$.
+If "insight" has nothing to do with solving the problem ($\neg{}H$), then presenting the insight that Roman numerals can be used should not lead to all students solving the problem ($\neg{}E_1$); the experiment, therefore, has high theoretical risk $P(\neg{}E_1|\neg{}H)$.
+Conversely, if insight is required to solve the problem ($H$), then it is likely to help all students to solve the problem ($E_1$), the experiment, therefore, has high detectability $P(E_1|H)$.
 The second experiment, on the other hand, has low theoretical risk $P(\neg{}E_2|\neg{}H)$.
-Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution, without having any insight.
+Even if "insight" has nothing to do with solving the problem ($\neg{}H$), there are other plausible reasons for observing the evidence ($E_2$), because the students could simply copy the solution without having any insight.
 With regard to detectability, experiments 1 and 2 differ in no obvious way.
@@
 We have established that higher theoretical risk leads to more persuasive evidence.
-In other words, we have reconstructed the interpretation that preregistrations supposedly works by restricting the researchers, which in turn increases the theoretical risk (or equivalently limits the type-I error rate) and thereby creates more compelling evidence.
+In other words, we have reconstructed the interpretation that preregistrations supposedly work by restricting the researchers, which in turn increases the theoretical risk (or equivalently limits the type-I error rate) and thereby creates more compelling evidence.
 Nevertheless, there are trade-offs for increasing theoretical risk.
@@
 One widely known factor is the contribution of statistical methods to theoretical risk.
-Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative hypothesis, because then the nill hypothesis is $\neg{}H$.
+Theoretical risk $P(\neg{}E|\neg{}H)$ is deeply connected with statistical methods, because it is related to the type-I error rate in statistical hypothesis testing $P(E|\neg{}H)$ by $P(\neg{}E|\neg{}H) = 1 - P(E|\neg{}H)$, if you consider the overly simplistic case where the research hypothesis is equal to the statistical alternative-hypothesis because then the nill-hypothesis is $\neg{}H$.
 Because many researchers are familiar with the type-I error rate, it can be helpful to remember this connection to theoretical risk.
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis seldomly equals the statistical alternative hypothesis.
+Second, the research hypothesis seldomly equals the statistical alternative-hypothesis.
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
@@
 Whether researchers want to or not, they continuously process information about how the study is going, except under strict blinding.
-While one can hope that processing this information does not affect their decision making either way, this cannot be ascertained.
+While one can hope that processing this information does not affect their decision-making either way, this cannot be ascertained.
 Therefore, we conclude that statistical properties only guarantee a lower bound for theoretical risk.
@@
 As the researcher did not conduct the study (and the study is not preregistered), they can not be certain about the various factors influencing theoretical risk (researcher degrees of freedom).
-We therefore express this uncertainty about the theoretical risk as a probability distribution $Q$ of $P(E|\neg{}H)$ (remember that $P(E|\neg{}H)$ is related to theoretical risk by $P(E|\neg{}H) = 1 - P(\neg{}E|\neg{}H)$, so it does not matter wheter we consider the distribution of theoretical risk or $P(E|\neg{}H)$).
-To get the expected value of $P(H|E)$ that follows from the researchers uncertainty about the theoretical risk, we can compute the expectation using Bayes theorem:
+We therefore express this uncertainty about the theoretical risk as a probability distribution $Q$ of $P(E|\neg{}H)$ (remember that $P(E|\neg{}H)$ is related to theoretical risk by $P(E|\neg{}H) = 1 - P(\neg{}E|\neg{}H)$, so it does not matter whether we consider the distribution of theoretical risk or $P(E|\neg{}H)$).
+To get the expected value of $P(H|E)$ that follows from the researchers' uncertainty about the theoretical risk, we can compute the expectation using Bayes theorem:
 
@@
 
-Of course, the assigned probabilities and the distribution $Q$ varies from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with a example.
-Assuming a power of 80\% we assign $P(E|H) = 0.8$, we assume the tested hypotheses is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to $1 - \alpha$, reflecting our assumption that this term gives an upper bound for theoretical risk $P(\neg{}E|\neg{}H)$.
+Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with an example.
+Assuming $P(E|H) = 0.8$ (relective of the typically strived for power of 80%).
+Let us further assume that the tested hypothesis is considered unlikely to be true by the research community before the study is conducted ($P(H) = 0.1$) and assign a uniform distribution for $P(E|\neg{}H) \sim U([1-\tau, 1])$ where $\tau{}$ is set to $1 - \alpha$, reflecting our assumption that this term gives an upper bound for theoretical risk $P(\neg{}E|\neg{}H)$.
 We chose this uniform distribution as it is the maximum entropy distribution with support $[1-\tau, 1]$ and hence conforms to our Bayesian framework [@giffinUpdatingProbabilitiesData2007].
@@
 Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being certain (solid line) or uncertain (dashed line) about the theoretical risk of a study.
-Depending on how uncertain we are about the theoretical risk a study took on, our expectation of the gained epistemic value varies considerably.
+Our expectation of the gained epistemic value varies considerably depending on how uncertain we are about the theoretical risk a study took on.
 Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
@@
 However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
-However, studies do not only differ by how high the theoretical risk is but also how certain the recipient is about the theoretical risk.
+However, studies do not only differ by how high the theoretical risk is but also by how certain the recipient is about the theoretical risk.
 A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
-The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extend of the benefit vary of course with the specifics of the study.
+The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
 Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
@@
 We hope to have persuaded the reader to accept two arguments: First, the theoretical risk is important for judging evidential support for theories.
-Second, the theoretical risk is inherently uncertain and the degree of uncertainty diminishes the persuasiveness of the gathered evidence.
+Second, the theoretical risk is inherently uncertain, and the degree of uncertainty diminishes the persuasiveness of the gathered evidence.
 The third and last argument is that preregistrations reduce this uncertainty.
-Following the last argument, a preregistered study is represented by the solid line (certainty about theoretical risk) and a study that was not preregistered is more similar to the dashed line (maximally uncertain about theoretical risk) in Figure \ref{fig:posterior} and Figure \ref{fig:other-measures}.
+Following the last argument, a preregistered study is represented by the solid line (certainty about theoretical risk), and a study that was not preregistered is more similar to the dashed line (maximally uncertain about theoretical risk) in Figure \ref{fig:posterior} and Figure \ref{fig:other-measures}.
 
@@
 Not only the original authors are concerned with the process of weighing evidence for or against a theory but really the whole scientific community the study authors hope to persuade.
-Unfortunately, readers of a scientific article (or, more generally, any consumer of a research product), is likely to lack insight into the various factors that influence theoretical risk.
+Unfortunately, readers of a scientific article (or, more generally, any consumer of a research product) will likely lack insight into the various factors that influence theoretical risk.
 While the authors themselves may have a clear picture of what they did and how it might have influenced the theoretical risk they took, their readers have much greater uncertainty about these factors.
 In particular, they never know which relevant factors the authors of a given article failed to disclose, be it intentionally or not.
-From the perspective of the ultimate sceptic, they may claim maximum uncertainty.
+From the perspective of the ultimate skeptic, they may claim maximum uncertainty.
 
 Communicating clearly how authors of a scientific report collected their data and consequently analyzed it to arrive at the evidence they present is crucial for judging the theoretical risk they took.
-Preregistrations are ideal to communicate just that, because any description after the fact is prone to be incomplete.
+Preregistrations are ideal for communicating just that because any description after the fact is prone to be incomplete.
 For instance, the authors could have opted for selective reporting, that is, they decided to exclude a number of analytic strategies they tried out.
@@
 The point is that we cannot exclude it with certainty.
-This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand, and then report that they did exactly that.
+This uncertainty is drastically reduced if the researchers have described what they intended to do beforehand and then report that they did exactly that.
 In that case, readers can be certain they received a complete account of the situation.
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
-Remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
-In any case, a well written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
+The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
+In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
 Therefore, a study that perfectly adhered to its preregistration will resemble the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
@@
 To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
-The former result, that increasing theoretical risk leads to higher expected epistemic value reconstructs the appeal and central goal of preregistration of confirmatory research agendas.
-However, theoretical risk is something researcher have only limited control over, for example, theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well known tension between type-I error and statistical power).
+The former result that increasing theoretical risk leads to higher expected epistemic value reconstructs the appeal and central goal of preregistration of confirmatory research agendas.
+However, theoretical risk is something researchers have only limited control over.
+For example, theories are often vague and ill-defined, resources are limited, and increasing theoretical risk usually decreases detectability of a hypothesized effect (a special instance of this trade-off is the well-known tension between type-I error and statistical power).
 While we believe that preregistration is always beneficial, it might be counterproductive to pursue high theoretical risk if the research context is inappropriate for strictly confirmatory research.
@@
 The first is, that even if all imaginable actions regarding promoting higher theoretical risk are taken, confirmatory research should be preregistered.
-Otherwise the uncertainty about the theoretical risk will diminish the advantage of confirmatory research.
-Second, even under less than ideal circumstances for confirmatory research, preregistration is beneficial.
-Preregistering exploratory studies increases the expected epistemic value by virtue of reducing uncertainty about theorethical risk.
-Nevertheless, exploratory studies will have a lower expected epistemic value than a more confirmatory study, if both are preregistered and have equal detectability.
+Otherwise, the uncertainty about the theoretical risk will diminish the advantage of confirmatory research.
+Second, even under less-than-ideal circumstances for confirmatory research, preregistration is beneficial.
+Preregistering exploratory studies increases the expected epistemic value by virtue of reducing uncertainty about theoretical risk.
+Nevertheless, exploratory studies will have a lower expected epistemic value than a more confirmatory study if both are preregistered and have equal detectability.
 
@@
 A statistical procedure may have high or low theoretical risk depending on the specifics of the model used.
-Specifying the process, therefore, is in line with the rational we propose here, but is less reasonable when the goal of preregistration is supposed to be a strictly confirmatory research agenda.
+Specifying the process, therefore, is in line with the rationale we propose here, but is less reasonable when the goal of preregistration is supposed to be a strictly confirmatory research agenda.
 
@@
 If the only goal of preregistration is to ensure confirmatory research, such changes are not justifiable.
-However, under our rational some changes may be justified.
+However, under our rational, some changes may be justified.
 Any change increases the uncertainty about the theoretical risk and may even decrease the theoretical risk.
-The changes still may be worthwhile, if the negative outcomes may be offset by an increase in detectability due to the change.
+The changes still may be worthwhile if the negative outcomes may be offset by an increase in detectability due to the change.
 Consider a preregistration that failed to specify how to handle missing values, and researchers subsequently encountering missing values.
@@
 In particular, a preregistration should be specific about the procedure that is intended to generate evidence for a theory.
-Such procedure may accommodate a wide range of possible data, i.e., it may be exploratory.
+Such a procedure may accommodate a wide range of possible data, i.e., it may be exploratory.
 The theoretical risk, however low, must be communicated clearly.
@@
 The idea is that the preregistration already contains "mock results" based on simulated or pilot data, which are replaced after the actual study data becomes available.
-Such approach dissolves the distinction between the preregistration document and the final scientific report.
-Instead of separate documents, preregistration and final report are different versions of the same underlying dynamic document.
+Such an approach dissolves the distinction between the preregistration document and the final scientific report.
+Instead of separate documents, preregistration, and final report are different versions of the same underlying dynamic document.
 Deviations from the preregistration can therefore be clearly (and if necessary, automatically) isolated, highlighted, and inspected using version control.
 Crucially, because the preregistration contains code, it may accommodate many different data patterns, i.e., it may be exploratory.
-However, while a PAC does not limit the extend of exploration, it is very specific about the probability to generate evidence even when the theory does not hold (theoretical risk).
+However, while a PAC does not limit the extent of exploration, it is very specific about the probability to generate evidence even when the theory does not hold (theoretical risk).
 Please note that while PAC is ideally suited to reduce uncertainty about theoretical risk, other more traditional forms of preregistration are also able to advance this goal.
@@
 
-We thank Maximilian S.
-Ernst, Caspar van Lissa, Felix Schönbrodt, the discussants at the DGPS2022 conference and Open Science Center Munich, and many more for the insightful discussions about disentangling preregistration and confirmation.
+We thank Leo Richter, Caspar van Lissa, Felix Schönbrodt, the discussants at the DGPS2022 conference and Open Science Center Munich, and many more for the insightful discussions about disentangling preregistration and confirmation.
 We are grateful to Julia Delius for her helpful assistance in language and style editing.

### 2023-05-13 manuscript.Rmd
switch to apa

@@
 abstract: "`r paste0(c('\\\\noindent', readLines('abstract.md')), collapse = ' ')`"
+shorttitle: "The Objective of Preregistration"
 author: 
   - name: Aaron Peikert
-    affil: a, b, c
+    affiliation: "1,2,3"
     email: peikert@mpib-berlin.mpg.de
+    corresponding : yes    # Define only one corresponding author
+    address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
+    role:
+      - Conceptualization
+      - Writing---Original Draft Preparation
+      - Writing---Review & Editing
+      - Methodology
+      - Formal analysis
+      - Software
+      - Visualization
+      - Project administration
+  - name: Maximilian S. Ernst
+    affiliation: "3"
+    role:
+      - Writing---Review & Editing
+      - Formal analysis
+      - Validation
   - name: Andreas M. Brandmaier
-    affil: a, b, d 
-    email: brandmaier@mpib-berlin.mpg.de
+    affiliation: "1, 2, 4"
+    role:
+      - Writing---Review & Editing
+      - Supervisions
 affiliation:
-  - num: a
-    address: |
-      Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany
-  - num: b
-    address: |
-     Max Planck UCL Centre for Computational Psychiatry and Ageing Research Berlin, Germany and London, UK
-  - num: c
-    address: |
-      Department of Psychology, Humboldt-Universität zu Berlin, Unter den Linden 6, 10117 Berlin, Germany
-  - num: d
-    address: |
-      MSB Medical School Berlin, Rüdesheimer Str. 50, 14197 Berlin, Germany
+  - id: 1
+    institution: Max Planck Institute for Human Development
+  - id: 2
+    institution: Max Planck UCL Centre for Computational Psychiatry and Ageing Research
+  - id: 3
+    institution: Humboldt-Universität zu Berlin
+  - id: 4
+    institution: MSB Medical School Berlin
+keywords: |
+  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
+wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~7000')`"
+figsintext: false
+figurelist: false
+tablelist : false
+footnotelist: false
+lineno: true
+linkcolor: "blue"
+csquotes: false
+class: "man"
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
+header-includes:
+  - \raggedbottom
 repro:
@@
     - aaronpeikert/repro@5075336
+    - crsh/papaja@1c488f7
   tlmgr:
+    - apa6
     - pgf
     - preview
-    - libertine
   apt:
@@
     - inkscape
-output: rticles::tf_article
-type: ARTICLE TEMPLATE
-keywords: |
-  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
-header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
 ---
@@
 knitr::opts_chunk$set(
-  echo = FALSE
+  echo = FALSE,
+  dev = "tikz"
 )
@@
   tikzDefaultEngine = "xetex",
-  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}", "\\usepackage{libertine}")
+  tikzXelatexPackages = c(getOption("tikzXelatexPackages"), "\\usepackage{amsfonts}")
 )
@@
 
-```{r posterior-fig, dev = "tikz", cache=TRUE, warning=FALSE}
+```{r posterior-fig, cache=TRUE, warning=FALSE}
 #| fig.cap: "\\label{fig:posterior}Posterior probability (confirmation as firmness) as a function of theoretical risk $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
@@
 
-```{r other-measures-fig, dev = "tikz", cache=TRUE, warning=FALSE}
+```{r other-measures-fig, cache=TRUE, warning=FALSE}
 #| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
@@
 
+\newpage
+
 # References
+
+::: {#refs custom-style="Bibliography"}
+:::

### 2023-05-13 manuscript.Rmd
apa.csl is no longer needed

@@
     - data/simulation.rds
-    - apa.csl
   packages:

### 2023-05-13 manuscript.Rmd
remove roles

@@
     address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
-    role:
-      - Conceptualization
-      - Writing---Original Draft Preparation
-      - Writing---Review & Editing
-      - Methodology
-      - Formal analysis
-      - Software
-      - Visualization
-      - Project administration
-  - name: Maximilian S. Ernst
-    affiliation: "3"
-    role:
-      - Writing---Review & Editing
-      - Formal analysis
-      - Validation
   - name: Andreas M. Brandmaier
     affiliation: "1, 2, 4"
-    role:
-      - Writing---Review & Editing
-      - Supervisions
 affiliation:
@@
     - inkscape
+note: |
+  The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
+  This version was created from git commit ``r repro::current_hash(backend = "gert")``.
+  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Psychological Methods](https://www.apa.org/pubs/journals/).
+
 ---
@@
 
-# Notes {-}
-
-The materials for this article are available on [GitHub](https://github.com/aaronpeikert/bayes-prereg) [@zenodo-release].
-This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Philosophical Psychology](https://www.tandfonline.com/journals/cphp20).
-
 \newpage

### 2023-05-13 manuscript.Rmd
Revert "remove roles"
This reverts commit f9ba8fb760645456a0cd0e65e067ab67429be46a.

@@
     address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
+    role:
+      - Conceptualization
+      - Writing---Original Draft Preparation
+      - Writing---Review & Editing
+      - Methodology
+      - Formal analysis
+      - Software
+      - Visualization
+      - Project administration
   - name: Andreas M. Brandmaier
     affiliation: "1, 2, 4"
+    role:
+      - Writing---Review & Editing
+      - Supervisions
 affiliation:

### 2023-05-13 manuscript.Rmd
Revert "remove maximilian, he has to approve"
This reverts commit 91b5a9cc9422454737f34a2191e955a4c7785668.

@@
       - Project administration
+  - name: Maximilian S. Ernst
+    affiliation: "3"
+    role:
+      - Writing---Review & Editing
+      - Formal analysis
+      - Validation
   - name: Andreas M. Brandmaier

### 2023-05-13 manuscript.Rmd
update to apa7

--- 2023-05-13 manuscript.Rmd
@@
 class: "man"
+documentclass: "apa7"
+csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
 output:
@@
     latex_engine: xelatex
-header-includes:
-  - \raggedbottom
+header-includes: |
+  \raggedbottom
+  \makeatletter
+
+  \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+    {-1em}%
+    {\normalfont\normalsize\bfseries\typesectitle}}
+  \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+    {-\z@\relax}%
+    {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+  \makeatother
 repro:
@@
   tlmgr:
-    - apa6
+    - apa7
     - pgf

### 2023-05-13 manuscript.Rmd
has not been peer reviewed

@@
   This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Psychological Methods](https://www.apa.org/pubs/journals/).
+  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Psychological Methods](https://www.apa.org/pubs/journals/) but has not been peer reviewed.
 

### 2023-05-13 manuscript.Rmd

@@
     address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
+    role:
+      - Conceptualization
+      - Writing---Original Draft Preparation
+      - Writing---Review & Editing
+      - Methodology
+      - Formal analysis
+      - Software
+      - Visualization
+      - Project administration
+  - name: Maximilian S. Ernst
+    affiliation: "3"
+    role:
+      - Writing---Review & Editing
+      - Formal analysis
+      - Validation
   - name: Andreas M. Brandmaier
     affiliation: "1, 2, 4"
+    role:
+      - Writing---Review & Editing
+      - Supervisions
 affiliation:

### 2023-05-15 manuscript.Rmd
I hate latex
Update manuscript.Rmd
added departments and places

@@
   - id: 1
-    institution: Max Planck Institute for Human Development
+    institution: Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany
   - id: 2
-    institution: Max Planck UCL Centre for Computational Psychiatry and Ageing Research
+    institution: Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany
   - id: 3
-    institution: Humboldt-Universität zu Berlin
+    institution: Department of Psychology, Humboldt-Universität zu Berlin, Berlin, Germany
   - id: 4
-    institution: MSB Medical School Berlin
+    institution: Department of Psychology, MSB Medical School Berlin, Berlin, Germany
 keywords: |

### 2023-05-15 manuscript.Rmd
Update manuscript.Rmd
Missing word

--- 2023-05-13 manuscript.Rmd
@@
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
-There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
+ a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
 However, mistaking exploratory findings for empirically confirmed results is dangerous.
@@
 First, many researchers do not subscribe to a purely confirmatory research agenda.
-Second, there is strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
+Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
 

### 2023-05-15 manuscript.Rmd

@@
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
- a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
+There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
 However, mistaking exploratory findings for empirically confirmed results is dangerous.
@@
 First, many researchers do not subscribe to a purely confirmatory research agenda.
-Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
+Second, there is strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
 
@@
 The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
-We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which is reduced by researchers' degrees of freedom, p-hacking, and suchlike.
+We argue that this view on the utility of preregistration can be interpreted as maximizing theoretical risk, which otherwise may be reduced by researchers' degrees of freedom, p-hacking, and suchlike.
 The second interpretation is our main contribution:
@@
 The third and last argument is that this uncertainty is reduced by applying preregistration.
-We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases the amount of knowledge we gain from a particular study, preregistration is potentially useful for any kind of study, no matter how exploratory.
+We conclude that because preregistration decreases uncertainty about the theoretical risk, which in turn increases the amount of knowledge we gain from a particular study, preregistration is potentially useful for any kind of study, no matter where it falls on the exploratory-confirmatory continuum.
 
@@
 
-Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not imply anything about the statistical methods researchers use.
+Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
 In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
@@
 The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
-In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called increase in firmness [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
+In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ @carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
 As noted before, we concentrate on posterior probability as a measure of epistemic value since no measure shows universally better properties than others.
@@
 However, the prior probability of a hypothesis is nothing our study design can change.
-The second option is similarly commonsensical; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
+The second option is equally reasonable; that is, an increase in $P(E|H)$ leads to a higher posterior probability $P(H|E)$.
 $P(E|H)$ is the probability of obtaining evidence for a hypothesis when it holds.
@@
 Its counter probability $P(\neg{}E|\neg{}H)= 1 - P(E|\neg{}H)$ is what we call "theoretical risk", because it is the risk a theory takes on in predicting the occurrence of particular evidence in its favor.
-We "borrow" the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
+We borrow the term from @meehlTheoreticalRisksTabular1978, though he has not assigned it to the probability $P(\neg{}E|\neg{}H)$.
 @kuklaClinicalStatisticalTheory1990 argued that the core arguments in @meehlAppraisingAmendingTheories1990 can be reconstructed in a purely Bayesian framework.
@@
 
-We have proposed a workflow for preregistration called preregistration as code (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+We have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.

### 2023-05-16 manuscript.Rmd
Update manuscript.Rmd

--- 2023-05-15 manuscript.Rmd
@@
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
- a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
+ There is a broad consensus that both approaches are necessary for science to progress; exploration, to make new discoveries and confirmation, to expose these discoveries to potential falsification, and assess empirical support for the theory.
 However, mistaking exploratory findings for empirically confirmed results is dangerous.

### 2023-05-16 manuscript.Rmd
Update manuscript.Rmd
remove places

@@
     address: "Center for Lifespan Psychology, Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
+    role:
+      - Conceptualization
+      - Writing---Original Draft Preparation
+      - Writing---Review & Editing
+      - Methodology
+      - Formal analysis
+      - Software
+      - Visualization
+      - Project administration
+  - name: Maximilian S. Ernst
+    affiliation: "3"
+    role:
+      - Writing---Review & Editing
+      - Formal analysis
+      - Validation
   - name: Andreas M. Brandmaier
     affiliation: "1, 2, 4"
+    role:
+      - Writing---Review & Editing
+      - Supervisions
 affiliation:
   - id: 1
-    institution: Max Planck Institute for Human Development
+    institution: Center for Lifespan Psychology, Max Planck Institute for Human Development
   - id: 2
@@
   - id: 3
-    institution: Humboldt-Universität zu Berlin
+    institution: Department of Psychology, Humboldt-Universität zu Berlin
   - id: 4
-    institution: MSB Medical School Berlin
+    institution: Department of Psychology, MSB Medical School Berlin
 keywords: |

### 2023-05-16 manuscript.Rmd
maximilian is affiliated with LIP

@@
   - name: Maximilian S. Ernst
-    affiliation: "3"
+    affiliation: "1"
     role:

### 2023-10-09 manuscript.Rmd
stress subjectivity, discuss severity

@@
 <!---outline-structure--->
-To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold].
 
-Further, we outline two interpretations of preregistration.
+We are interested in why preregistrations should change researchers' evaluation of evidence.
+Applying a Bayesian framework allows us to investigate our research question most straightforwardly. 
+Specifically, it allows us to model changes in subjective degrees of belief due to preregistration or, more simply, "persuasion".
+Please note that our decision to adopt a Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
+In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we dislike "failing to disprove" too much to give it up), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
+
+We outline two interpretations of preregistration.
 The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
@@
 
-Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
-In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-
 # Epistemic value and theoretical risk
@@
 The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
-In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ @carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
-As noted before, we concentrate on posterior probability as a measure of epistemic value since no measure shows universally better properties than others.
-However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
+In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
+We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
+In reality, surely researchers differ in what function they apply to judge evidence.
+The goal is therefore to reason about the space of possible measures researchers might apply.
+However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.
 
@@
 Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
+If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (an inversion of statistical relevancy).
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence).
 
@@
 Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
-In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
+This is similar to the well known tradeoff between type-I error rate and power.
 Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
+As noted before, we want high detectability and high theorethical risk.
+Maximizing one at the expense of the other is not nessesarily beneficial for increasing epistemic value.
 
@@
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
+Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers believe.
+One may take such theoretical properties as a first starting point to form a substantive believe but surely researchers ought to take more factors into consideration.
+For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, that should decrease the theorethical risk; after all they might find an association purely on account of the confounders.
 
@@
 
-Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with an example.
+Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher (and even the measure of confirmation), but we can illustrate the effect of uncertainty with an example.
 Assuming $P(E|H) = 0.8$ (relective of the typically strived for power of 80%).
@@
 
+Theoretical risk may conceptually be related to the very fruitful conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
+However, there are crucial differences.
+First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
+Second, severity considerations are mainly post-data.
+Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
+Third, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
+However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
+
 Our latter result, on the importance of preregistration for minimizing uncertainty, has two important implications.
@@
 
+Our emphasis on transparency echoes other justifications of preregistration, especially @lakensValuePreregistrationPsychological2019's, although we come from a very different philosophical starting point.
+What we hope to contribute, however, is a rationale that more stringently deals with the issues of exploration and confirmation as well as subjective evaluations of evidence.
+We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
+In our framework, researchers can input what they assume other researchers think and calculate what their actions, be it the preregistration itself or later deviations, may have on their readers.
+We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors), researchers who suspect confounding variables (they have lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
+We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a tool to navigate the complicated questions that arise in the practice of preregistration.
+
 We have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].

### 2023-10-10 manuscript.Rmd
note that the calculus is there to balance the tradeoff

@@
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-As noted before, we want high detectability and high theorethical risk.
-Maximizing one at the expense of the other is not nessesarily beneficial for increasing epistemic value.
+As noted before, we want high detectability and high theoretical risk.
+Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
+One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.
 

### 2023-10-10 manuscript.Rmd
minimally expand discussion

@@
 Our emphasis on transparency echoes other justifications of preregistration, especially @lakensValuePreregistrationPsychological2019's, although we come from a very different philosophical starting point.
-What we hope to contribute, however, is a rationale that more stringently deals with the issues of exploration and confirmation as well as subjective evaluations of evidence.
+What we hope to contribute, however, is a rationale that more stringently deals with the issues of exploration and confirmation, post-hoc changes, as well as subjective evaluations of evidence.
 We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
@@
 
-We have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+With the goal to facilitate, rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.

### 2023-10-10 manuscript.Rmd
mention the case of null as research hypothesis closes #96

@@
 However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
+However, we latter discuss why detectability can not be disregarded from consideration.
 Thus, $P(E)$ remains to be considered.
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis seldomly equals the statistical alternative-hypothesis.
+Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then any value except zero).
+In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theortical risk and type-I/II error rate switch.
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.

### 2023-10-11 manuscript.Rmd
fix typo

@@
 Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then any value except zero).
-In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theortical risk and type-I/II error rate switch.
+In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.

### 2023-10-12 manuscript.Rmd
typo
add more examples to relate better to exploration, p hacking, and so forth

@@
 
-## Sources of Uncertainty
+## Sources of uncertainty
 
@@
 
+# Hacking, harking, and other harms
+
+The importance of distinguishing between low and highly uncertain theoretical risk becomes perhaps clearer if we consider couple of hypothetical cases.
+
+1. We know with absolute certainty that the results are hacked.
+2. The hypothesis was picked to explain the results.
+3. We can not exclude the possibility of p-hacking.
+4. The results are obtained by planned exploration.
+5. The results are obtained by unplanned exploration.
+
+In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
+If we know that no matter what, the results engineered to support the hypothesis, there is no reason to collect data.
+Case 2 has a similar problem.
+In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
+Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain and perhaps represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
+In case 4, the theoretical risk is neither zero (unless the researcher plans to do everything in their power to obtain favourable results, then we have a particular instance case of 1) nor high (that is what makes it exploratory).
+However, we can use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
+This favorable position leads us to recommend preregistration of exploratory studies.
+Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
+Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.
+
 # Discussion

### 2023-10-12 manuscript.Rmd
small changes after reading

@@
 However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we dislike "failing to disprove" too much to give it up), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
+While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with the behavior of researchers.
 
@@
 We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
-In reality, surely researchers differ in what function they apply to judge evidence.
+In reality, surely researchers differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
 The goal is therefore to reason about the space of possible measures researchers might apply.
@@
 When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
-However, we latter discuss why detectability can not be disregarded from consideration.
+However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
+However, we latter discuss how issues detectability must be considered for the issue of preregistration.
 Thus, $P(E)$ remains to be considered.
@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence).
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence beyond failing to disprove it).
 
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then any value except zero).
+Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then "any value except zero").
 In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
@@
 Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
+We find it helpful to calculate some example because of the nonlinear nature of the evidence functions.
 
@@
 Case 2 has a similar problem.
+After all, the hypothesis that it had to happen the way it did happen is irrefutable.
 In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
@@
 
-With the goal to facilitate, rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+With the goal to facilitate rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.

### 2023-10-13 manuscript.Rmd
add more references to the exploration confirmation dichotomy

@@
 <!--introduce confirmatory vs. exploratory-->
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018; @tukeyWeNeedBoth1980].
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
@@
 However, two problems become apparent under closer inspection.
-First, many researchers do not subscribe to a purely confirmatory research agenda.
+First, many researchers do not subscribe to a purely confirmatory research agenda [@tukeyExploratoryDataAnalysis1972; @baumeisterChartingFutureSocial2016; @finkelReplicabilityOtherFeatures2017].
 Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

### 2023-10-13 manuscript.Rmd
add more references on deviations after preregistration
cite computational reproducibility

@@
 This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
+The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
@@
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
+
 The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
+To further reduce the uncertainty about theoretical risk, researchers must therefore publish code and ideally data.
+After all, computational reproducibility is only possible if the data analytic procedure was communicated clearly enough to allow others to retrace the computational steps [@peikertReproducibleDataAnalysis2021].
+
 In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
@@
 First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
+This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
 Second, severity considerations are mainly post-data.
 Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
-Third, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
 However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

### 2023-10-14 manuscript.Rmd
cite diagnosticity

@@
 The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
-However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974] or as recent psychological literature come to call it "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019].
 That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.

### 2023-10-16 manuscript.Rmd
cite measures

@@
 ```{r other-measures-fig, cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
+#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line).
+Measure taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p.
+51."
 

### 2023-10-16 manuscript.Rmd
extend discussion of oberauer

@@
 
+However, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
+To our knowledge, it was not yet applied to the problem of preregistration.
+However, @oberauerAddressingTheoryCrisis2019 made use of the formalism to model the relation between theory, hypothesis, and evidence.
+In the wake of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
+Most interestingly, they are rather critical of the idea that preregistration has tangible benefits.
+Instead, they prefer multiverse analyses but contend that those could be preregistered if one fancies it.
+Their reasoning is based on two intuitions about what should *not* influence the evaluation of evidence: temporal order and the mental state of the originator.
+For one, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
+However, we agree that the information flow matters for preregistration, not the temporal order.
+For the other, we find that the mental state of the originator does matter.
+Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
+Where each $k$ has a $P(E_k | \neg H) > 0$.
+If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = \prod_{k=1}^K P(\neg E_k | \neg H)$?
+We think the latter.
+
+Whatever the difference in evaluating preregistration as a tool, maybe conceptually more profound is that @oberauerAddressingTheoryCrisis2019 conceptualizes "discovery-oriented research" differently than we do "exploratory".
+They assume the same theoretical risk ($P(\neg E| \neg H) = .05$) and detectability ($P(E| H)) = .8$) in their calculation example as we but assign different prior probabilities, namely $.06$ for discovery versus $.6$ for theory testing.
+They then conclude that discovery-oriented researcher requires a much lower type-I error rate to control false positive in light of the low prior probability.
+This runs counter to our definition of exploratory research having low theoretical risk.
+Of course, we agree that low priors require more persuasive evidence; our disagreement, therefore, lies mainly in terminology.
+They imagine discovery-oriented researchers to conduct experiments where they have low expectations that they obtain positive evidence ($.06 \cdot .8 + .94 \cdot .05 = 0.095$), but if they do, it raises the posterior significantly (from $.06$ to $.51$)
+In our view, researchers who set out to explore a data set often find "something" (due to low $P(\neg E| \neg H)$); therefore, it should only slightly raise your posterior if they do.
+On a substantive matter, we believe both kinds of research are common in psychology.
+It is, therefore, mostly a disagreement on terminology.
+This disagreement only highlights why using a mathematical framework to investigate such things is so fruitful because we can clearly see where and how we differ.
+Both frameworks are highly compatible mathematically, and it might be interesting to combine their idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.
+
 With the goal to facilitate rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].

### 2023-10-16 manuscript.Rmd
found smallish errors

@@
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers believe.
+Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers' believe.
 One may take such theoretical properties as a first starting point to form a substantive believe but surely researchers ought to take more factors into consideration.
@@
 In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
-If we know that no matter what, the results engineered to support the hypothesis, there is no reason to collect data.
+If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
 Case 2 has a similar problem.
@@
 In case 4, the theoretical risk is neither zero (unless the researcher plans to do everything in their power to obtain favourable results, then we have a particular instance case of 1) nor high (that is what makes it exploratory).
-However, we can use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
+However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
 This favorable position leads us to recommend preregistration of exploratory studies.

### 2023-10-16 manuscript.Rmd
remove psych methods submission

@@
   This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] and was submitted to [Psychological Methods](https://www.apa.org/pubs/journals/) but has not been peer reviewed.
+  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] but was not approved by peer-review.
 

### 2023-10-19 manuscript.Rmd
add declaration for submission
take same affiliation as in software evaluation

--- 2023-10-16 manuscript.Rmd
@@
   - name: Maximilian S. Ernst
-    affiliation: "1"
+    affiliation: "1, 4"
     role:
@@
   - name: Andreas M. Brandmaier
-    affiliation: "1, 2, 4"
+    affiliation: "1, 3, 5"
     role:
@@
 affiliation:
-  - id: 1
-    institution: Center for Lifespan Psychology, Max Planck Institute for Human Development
-  - id: 2
-    institution: Max Planck UCL Centre for Computational Psychiatry and Ageing Research
-  - id: 3
-    institution: Department of Psychology, Humboldt-Universität zu Berlin
-  - id: 4
-    institution: Department of Psychology, MSB Medical School Berlin
+  - id            : "1"
+    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
+  - id            : "2"
+    institution   : "Department of Imaging Neuroscience, University College London, London, UK"
+  - id            : "3"
+    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany"
+  - id            : "4"
+    institution   : "Max Planck School of Cognition, Leipzig, Germany"
+  - id            : "5"
+    institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
 keywords: |
@@
 <!--introduce confirmatory vs. exploratory-->
-The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018; @tukeyWeNeedBoth1980].
+The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis generation and hypothesis testing, or prediction and postdiction  [@hoyningen-hueneContextDiscoveryContext2006; @shmueliExplainPredict2010; @nosekPreregistrationRevolution2018].
 Despite the different names, it is fundamentally the same dichotomy that is at stake here.
@@
 However, two problems become apparent under closer inspection.
-First, many researchers do not subscribe to a purely confirmatory research agenda [@tukeyExploratoryDataAnalysis1972; @baumeisterChartingFutureSocial2016; @finkelReplicabilityOtherFeatures2017].
+First, many researchers do not subscribe to a purely confirmatory research agenda.
 Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.
@@
 This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
-The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
 Both strategies may be used for sensible scientific reasons or with the self-serving intent of generating desirable results.
@@
 <!---outline-structure--->
-To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold].
+To form such an objective for preregistration, we first introduce some tools of Bayesian philosophy of science and map the exploration/confirmation distinction onto a dimensional quantity we call "theoretical risk" [a term borrowed from @meehlTheoreticalRisksTabular1978, but formalized as the probability of proving a hypothesis wrong if it does not hold], which is inversely related to the type-I error rate in null hypothesis testing.
 
-We are interested in why preregistrations should change researchers' evaluation of evidence.
-Applying a Bayesian framework allows us to investigate our research question most straightforwardly. 
-Specifically, it allows us to model changes in subjective degrees of belief due to preregistration or, more simply, "persuasion".
-Please note that our decision to adopt a Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
-In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we dislike "failing to disprove" too much to give it up), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
-While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with the behavior of researchers.
-
-We outline two interpretations of preregistration.
+Further, we outline two interpretations of preregistration.
 The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
@@
 The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
-However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974] or as recent psychological literature come to call it "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019].
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974].
 That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
+
+Please note that our decision to adopt this aspect of the Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.
+In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
 
@@
 The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
-In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
-We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
-In reality, surely researchers differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
-The goal is therefore to reason about the space of possible measures researchers might apply.
-However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.
+In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ @carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
+As noted before, we concentrate on posterior probability as a measure of epistemic value since no measure shows universally better properties than others.
+However, it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, and our argument applies to those measures as well.
 
@@
 When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
-However, we latter discuss how issues detectability must be considered for the issue of preregistration.
+However, while detectability is of great importance for study design, it is not directly relevant to the objective of preregistration.
 Thus, $P(E)$ remains to be considered.
@@
 Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (an inversion of statistical relevancy).
+If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold.
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).
@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence beyond failing to disprove it).
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value.
 
@@
 Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
-This is similar to the well known tradeoff between type-I error rate and power.
+In other words, one cannot increase power while maintaining the same type-I error rate without increasing the invested resources.
 Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-As noted before, we want high detectability and high theoretical risk.
-Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
-One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.
 
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then "any value except zero").
-In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
+Second, the research hypothesis seldomly equals the statistical alternative-hypothesis.
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers believe.
-One may take such theoretical properties as a first starting point to form a substantive believe but surely researchers ought to take more factors into consideration.
-For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, that should decrease the theorethical risk; after all they might find an association purely on account of the confounders.
 
@@
 
-## Sources of uncertainty
+## Sources of Uncertainty
 
@@
 
-Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher (and even the measure of confirmation), but we can illustrate the effect of uncertainty with an example.
+Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher, but we can illustrate the effect of uncertainty with an example.
 Assuming $P(E|H) = 0.8$ (relective of the typically strived for power of 80%).
@@
 ```{r other-measures-fig, cache=TRUE, warning=FALSE}
-#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line).
-Measure taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p.
-51."
+#| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line)."
 
@@
 Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
-We find it helpful to calculate some example because of the nonlinear nature of the evidence functions.
 
@@
 They still might be uncertain about the actual theoretical risk the authors took, but to a much smaller extent than if the study would not have been preregistered.
-
 The remaining sources of uncertainty might be unfamiliarity with statistical methods or experimental paradigms used, the probability of an implementation error in the statistical analyses, a bug in the software used for analyses, etc.
-To further reduce the uncertainty about theoretical risk, researchers must therefore publish code and ideally data.
-After all, computational reproducibility is only possible if the data analytic procedure was communicated clearly enough to allow others to retrace the computational steps [@peikertReproducibleDataAnalysis2021].
-
 In any case, a well-written preregistration should aim to reduce the uncertainty about the theoretical risk and hence increase the persuasiveness of evidence.
@@
 Crucially, perfect means here that the theoretical risk can be judged with low uncertainty, not that the theoretical risk is necessarily high.
-
-# Hacking, harking, and other harms
-
-The importance of distinguishing between low and highly uncertain theoretical risk becomes perhaps clearer if we consider couple of hypothetical cases.
-
-1. We know with absolute certainty that the results are hacked.
-2. The hypothesis was picked to explain the results.
-3. We can not exclude the possibility of p-hacking.
-4. The results are obtained by planned exploration.
-5. The results are obtained by unplanned exploration.
-
-In case 1, there is no theoretical risk ($P(\neg{}E|\neg{}H) = 0$).
-If we know that no matter what, the results engineered to support the hypothesis, there is no reason to collect data.
-Case 2 has a similar problem.
-After all, the hypothesis that it had to happen the way it did happen is irrefutable.
-In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
-Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain and perhaps represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
-In case 4, the theoretical risk is neither zero (unless the researcher plans to do everything in their power to obtain favourable results, then we have a particular instance case of 1) nor high (that is what makes it exploratory).
-However, we can use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
-This favorable position leads us to recommend preregistration of exploratory studies.
-Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
-Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.
 
@@
 Funding agencies should demand an explicit statement about theoretical risk in relation to detectability and must be willing to provide the necessary resources to reach adequate levels of both [@kooleRewardingReplicationsSure2012].
-
-Theoretical risk may conceptually be related to the very fruitful conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
-However, there are crucial differences.
-First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
-This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
-Second, severity considerations are mainly post-data.
-Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
-However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
 
@@
 
-Our emphasis on transparency echoes other justifications of preregistration, especially @lakensValuePreregistrationPsychological2019's, although we come from a very different philosophical starting point.
-What we hope to contribute, however, is a rationale that more stringently deals with the issues of exploration and confirmation, post-hoc changes, as well as subjective evaluations of evidence.
-We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
-In our framework, researchers can input what they assume other researchers think and calculate what their actions, be it the preregistration itself or later deviations, may have on their readers.
-We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors), researchers who suspect confounding variables (they have lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
-We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a tool to navigate the complicated questions that arise in the practice of preregistration.
-
-However, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
-To our knowledge, it was not yet applied to the problem of preregistration.
-However, @oberauerAddressingTheoryCrisis2019 made use of the formalism to model the relation between theory, hypothesis, and evidence.
-In the wake of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
-Most interestingly, they are rather critical of the idea that preregistration has tangible benefits.
-Instead, they prefer multiverse analyses but contend that those could be preregistered if one fancies it.
-Their reasoning is based on two intuitions about what should *not* influence the evaluation of evidence: temporal order and the mental state of the originator.
-For one, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
-However, we agree that the information flow matters for preregistration, not the temporal order.
-For the other, we find that the mental state of the originator does matter.
-Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
-Where each $k$ has a $P(E_k | \neg H) > 0$.
-If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = \prod_{k=1}^K P(\neg E_k | \neg H)$?
-We think the latter.
-
-Whatever the difference in evaluating preregistration as a tool, maybe conceptually more profound is that @oberauerAddressingTheoryCrisis2019 conceptualizes "discovery-oriented research" differently than we do "exploratory".
-They assume the same theoretical risk ($P(\neg E| \neg H) = .05$) and detectability ($P(E| H)) = .8$) in their calculation example as we but assign different prior probabilities, namely $.06$ for discovery versus $.6$ for theory testing.
-They then conclude that discovery-oriented researcher requires a much lower type-I error rate to control false positive in light of the low prior probability.
-This runs counter to our definition of exploratory research having low theoretical risk.
-Of course, we agree that low priors require more persuasive evidence; our disagreement, therefore, lies mainly in terminology.
-They imagine discovery-oriented researchers to conduct experiments where they have low expectations that they obtain positive evidence ($.06 \cdot .8 + .94 \cdot .05 = 0.095$), but if they do, it raises the posterior significantly (from $.06$ to $.51$)
-In our view, researchers who set out to explore a data set often find "something" (due to low $P(\neg E| \neg H)$); therefore, it should only slightly raise your posterior if they do.
-On a substantive matter, we believe both kinds of research are common in psychology.
-It is, therefore, mostly a disagreement on terminology.
-This disagreement only highlights why using a mathematical framework to investigate such things is so fruitful because we can clearly see where and how we differ.
-Both frameworks are highly compatible mathematically, and it might be interesting to combine their idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.
-
-With the goal to facilitate rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
+We have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].
 In a PAC, researchers use computer code for the planned analysis as well as a verbal description of theory and methods for the preregistration.

### 2023-10-19 manuscript.Rmd
Update manuscript.Rmd
I don't think that we need to be explicit about the previous rejection, particularly because the current manuscript is not the version that was rejected anymore.

--- 2023-10-16 manuscript.Rmd
@@
   This version was created from git commit ``r repro::current_hash(backend = "gert")``.
-  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023] but was not approved by peer-review.
+  The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
 
@@
 
+# Declarations {-}
+
+All code and materials required to reproduce this article are available under https://github.com/aaronpeikert/bayes-prereg [@zenodo-release].
+The authors have no competing interests to declare that are relevant to the content of this article.
+
 \newpage

### 2023-10-20 manuscript.Rmd
Update manuscript.Rmd
Apply andreas' suggestions
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 
-We outline two interpretations of preregistration.
+Now, we outline two possible perspectives on the utility of preregistration.
 The first one corresponds to the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
@@
 The posterior probability $P(H|E)$ is of great relevance since it is often used directly or indirectly as a measure of confirmation of a hypothesis.
-In the tradition of Carnap, in its direct use, it is called confirmation as firmness; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
+In the tradition of Carnap, in its direct use, it is called _confirmation as firmness_; in its relation to the a priori probability $P(H)$, it is called _increase in firmness_ [@carnapLogicalFoundationsProbability1950, preface to the 1962 edition].
 We concentrate on the posterior probability because of its simplicity but take it only as one example of a possible measure.
-In reality, surely researchers differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
+In reality, researchers surely differ in what function they apply to judge evidence and it is often most fruitful to compare more than two competing hypotheses.
 The goal is therefore to reason about the space of possible measures researchers might apply.
@@
 However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
-However, we latter discuss how issues detectability must be considered for the issue of preregistration.
+We later discuss how issues of detectability must be considered in a preregistration.
 Thus, $P(E)$ remains to be considered.
@@
 Second, if the theoretical risk is smaller than detectability $P(E|H)$ it follows that the posterior probability must decrease when observing the evidence.
-If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (an inversion of statistical relevancy).
+If detectability exceeds theoretical risk, the evidence is less likely under the theory than it is when the theory does not hold (the inverse of statistical relevancy).
 Third, if the theoretical risk equals zero, then posterior probability is at best equal to prior probability but only if detectability is perfect ($P(H|E)$ = 1).

### 2023-10-23 manuscript.Rmd
commit obvious improvements
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
-However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we dislike "failing to disprove" too much to give it up), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
+However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we regard "failing to disprove" as too limited), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
 While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with the behavior of researchers.
@@
 Increasing theoretical risk $P(\neg{}E|\neg{}H)$ will usually decrease detectability $P(E|H)$ unless more resources are invested.
-This is similar to the well known tradeoff between type-I error rate and power.
+This is similar to the well known tradeoff between type-I error rate and statistical power.
 Tasking preregistration with an increase in theoretical risk makes it difficult to balance this trade-off.
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-As noted before, we want high detectability and high theoretical risk.
+As noted before, we strive for high detectability and high theoretical risk.
 Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis seldomly equals the statistical alternative-hypothesis (most research hypothesis are more specific then "any value except zero").
+Second, the research hypothesis rarely equals the statistical alternative hypothesis (most research hypothesis are more specific then "any value except zero").
 In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
 We argue that theoretical risk (and hence its complement, $P(E|\neg{}H)$) also encompasses factors outside the statistical realm, most notably the study design and broader analytical strategies.
-Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers' believe.
-One may take such theoretical properties as a first starting point to form a substantive believe but surely researchers ought to take more factors into consideration.
-For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, that should decrease the theorethical risk; after all they might find an association purely on account of the confounders.
+Type-I error rate is the property of a statistical test under some assumptions, whereas theoretical risk is a researchers' belief.
+One may take such theoretical properties as a first starting point to form a substantive belief but surely researchers ought to take other factors into consideration.
+For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, this should decrease theorethical risk; after all they might find an association purely on account of the confounders.
 
@@
 
-The importance of distinguishing between low and highly uncertain theoretical risk becomes perhaps clearer if we consider couple of hypothetical cases.
+The importance of distinguishing between low and highly uncertain theoretical risk becomes perhaps clearer if we consider a few hypothetical cases for illustration.
 
-1. We know with absolute certainty that the results are hacked.
+1. We know with absolute certainty that reported results came about by p-hacking.
 2. The hypothesis was picked to explain the results.
-3. We can not exclude the possibility of p-hacking.
-4. The results are obtained by planned exploration.
-5. The results are obtained by unplanned exploration.
+3. We cannot exclude the possibility of p-hacking having led to the reported results.
+4. Reported results were obtained by planned exploration.
+5. Reported results were obtained by unplanned exploration.
 
@@
 In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
-Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain and perhaps represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}.
-In case 4, the theoretical risk is neither zero (unless the researcher plans to do everything in their power to obtain favourable results, then we have a particular instance case of 1) nor high (that is what makes it exploratory).
+Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
+In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
 However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
@@
 
-Theoretical risk may conceptually be related to the very fruitful conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
-However, there are crucial differences.
-First, we are not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
-This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .9999.
-Second, severity considerations are mainly post-data.
-Only after the evidence was observed can the severity with which we can rule out alternatives be calculated.
-However, they also share properties, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
+Theoretical risk may conceptually be related to the conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
+However, there are crucial differences between the two.
+First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
+This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
+Second, severity considerations are mainly after the fact.
+Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
+However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
 
@@
 
-Our emphasis on transparency echoes other justifications of preregistration, especially @lakensValuePreregistrationPsychological2019's, although we come from a very different philosophical starting point.
-What we hope to contribute, however, is a rationale that more stringently deals with the issues of exploration and confirmation, post-hoc changes, as well as subjective evaluations of evidence.
+Our emphasis on transparency aligns with other justifications of preregistration, especially those put forth by @lakensValuePreregistrationPsychological2019's, although based on quite different philosophical foundations.
+Our goal is to contribute a rationale that more comprehensively captures the spectrum of exploration and confirmation in relation to preregistrations, post-hoc changes of preregistrations, and subjective evaluations of evidence.
 We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
 In our framework, researchers can input what they assume other researchers think and calculate what their actions, be it the preregistration itself or later deviations, may have on their readers.
-We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors), researchers who suspect confounding variables (they have lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
-We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a tool to navigate the complicated questions that arise in the practice of preregistration.
+We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
+We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a framework to navigate the complicated questions that arise in the practice of preregistration.
 
-However, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
+At the same time, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
 To our knowledge, it was not yet applied to the problem of preregistration.
 However, @oberauerAddressingTheoryCrisis2019 made use of the formalism to model the relation between theory, hypothesis, and evidence.
-In the wake of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
-Most interestingly, they are rather critical of the idea that preregistration has tangible benefits.
+In the context of this conceptualization, they discussed the usefulness of preregistration, though without applying the formalism there.
+Most importantly, they are rather critical of the idea that preregistration has tangible benefits.
 Instead, they prefer multiverse analyses but contend that those could be preregistered if one fancies it.
@@
 However, we agree that the information flow matters for preregistration, not the temporal order.
-For the other, we find that the mental state of the originator does matter.
+For the other, we argue that the mental state of the originator does matter.
 Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
@@
 Whatever the difference in evaluating preregistration as a tool, maybe conceptually more profound is that @oberauerAddressingTheoryCrisis2019 conceptualizes "discovery-oriented research" differently than we do "exploratory".
-They assume the same theoretical risk ($P(\neg E| \neg H) = .05$) and detectability ($P(E| H)) = .8$) in their calculation example as we but assign different prior probabilities, namely $.06$ for discovery versus $.6$ for theory testing.
-They then conclude that discovery-oriented researcher requires a much lower type-I error rate to control false positive in light of the low prior probability.
+They assume the same theoretical risk ($P(\neg E| \neg H) = .05$) and detectability ($P(E| H)) = .8$) in their calculation example as we do but assign different prior probabilities, namely $.06$ for discovery versus $.6$ for theory testing.
+Then, they conclude that discovery-oriented researcher requires a much lower type-I error rate to control false positive in light of the low prior probability.
 This runs counter to our definition of exploratory research having low theoretical risk.
@@
 It is, therefore, mostly a disagreement on terminology.
-This disagreement only highlights why using a mathematical framework to investigate such things is so fruitful because we can clearly see where and how we differ.
+This disagreement only highlights why using a mathematical framework to investigate such things is so useful and ultimately indispensable because we can clearly see where and how we differ in our reasoning.
 Both frameworks are highly compatible mathematically, and it might be interesting to combine their idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.

### 2023-10-23 manuscript.Rmd
Update manuscript.Rmd
cite streamlined p hacking

@@
 Mindlessly maximizing theoretical risk would either never produce evidence or require huge amounts of resources.
-As noted before, we strive for high detectability and high theoretical risk.
+As noted before, we strive for high detectability and high theoretical risk in planning, conducting, and analyzing studies.
 Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
@@
 If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
+A prime example of this case is the $p_\text{pointless}$ metric [@husseyMethodStreamlinePhacking2021].
 Case 2 has a similar problem.

### 2023-10-23 manuscript.Rmd
cite HARKing properly
Update manuscript.Rmd

@@
 1. We know with absolute certainty that reported results came about by p-hacking.
-2. The hypothesis was picked to explain the results.
+2. A hypothesis was picked to explain reported results after the fact [HARKing, @kerrHARKingHypothesizingResults1998].
 3. We cannot exclude the possibility of p-hacking having led to the reported results.
@@
 We find it difficult to content ourselves with vague terms like "control" or "transparency" if they ultimately remain unconnected to how much researchers believe in a theory.
-In our framework, researchers can input what they assume other researchers think and calculate what their actions, be it the preregistration itself or later deviations, may have on their readers.
+Within our framework, researchers have the ability to input their assumptions regarding the perspectives of other researchers and calculate the potential impact of their actions on their readership, whether these actions relate to study design, to the preregistration itself, or subsequent deviations from it.
 We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).

### 2023-10-23 manuscript.Rmd
put figs into manuscript

@@
 wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~7000')`"
-figsintext: false
+figsintext: true
 figurelist: false

### 2023-10-23 manuscript.Rmd
update wordcount

@@
   preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
-wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~7000')`"
+wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~8390')`"
 figsintext: true

### 2023-10-23 manuscript.Rmd
cite the defeatist perspective

@@
 We think the latter.
+However, this "Defeatist" intuition is not universally warranted and depends on what we take $H$ to mean specifically [@kotzenMultipleStudiesEvidential2013].
+Addressing, this problem might benefit from combining @oberauerAddressingTheoryCrisis2019's idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.
 
@@
 This disagreement only highlights why using a mathematical framework to investigate such things is so useful and ultimately indispensable because we can clearly see where and how we differ in our reasoning.
-Both frameworks are highly compatible mathematically, and it might be interesting to combine their idea of updating on two nested levels (theory-hypothesis layered on top of hypothesis-evidence) with our approach to modelling uncertainty.
 

### 2023-10-23 manuscript.Rmd
cite höfler

@@
 
+We believe that our reasoning is quite similar to @hoflerMeansValuableExploration2022, who call for transparent exploration using preregistration.
+We could be surer of our agreement, if they had formulated their arguments within a mathematical framework, which would also have helped to dissolve an apparent conflict in their definitions of confirmation, exploration, and transparency.
+On the one hand, they define "The principle difference between confirmation and exploration is that confirmation adheres to an evidential norm for the test of a hypothesis to pass.", but then suggest that transparent exploration can be conducted using inferences tests as a filtering mechanism.
+Their distinction between confirmation, intransparent and transparent exploration are otherwise just as well placed along the dimensions, theoretical risk and uncertainty about theoretical risk.
+
 With the goal to facilitate rigorous exploration, we have proposed a workflow for preregistration called _preregistration as code_ (PAC) elsewhere [@peikertReproducibleResearchTutorial2021].

### 2023-10-30 manuscript.Rmd
Apply suggestions from maximilian
Co-authored-by: Maximilian-Stefan-Ernst <34346372+Maximilian-Stefan-Ernst@users.noreply.github.com>

@@
 
-This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration, to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
+This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
 The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
@@
 The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
-However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974] or as recent psychological literature come to call it "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019].
-That is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974], that is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
+In particular, "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019], a concept highlighted in recent psychological literature, seems to adhere to the statistical relevancy condition.
 
@@
 The goal is therefore to reason about the space of possible measures researchers might apply.
-However, since it is reasonable that any measure of confirmation increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.
+However, since any measure fulfilling the statistical relevancy condition increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.
 
@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but Popper would disagree with attaching any form of believe in $H$ on the basis of evidence beyond failing to disprove it).
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but we of course differ from Popper on the central issue of induction in science).
 
@@
 First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis [@mayoStatisticalInferenceSevere2018, 5.3].
-Second, the research hypothesis rarely equals the statistical alternative hypothesis (most research hypothesis are more specific then "any value except zero").
+Second, the research hypothesis rarely equals the statistical alternative hypothesis (most research hypothesis are more specific than "any value except zero").
 In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
@@
 #| fig.cap: "\\label{fig:other-measures}Several measures for confirmation as an increase in firmness as a function of $\\tau$, where $\\tau$ is either certain (solid line) or maximally uncertain (dotted line).
-Measure taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p.
+Measures taken from @sprengerBayesianPhilosophyScience2019, Table 1.3, p.
 51."
@@
 Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
-We find it helpful to calculate some example because of the nonlinear nature of the evidence functions.
+We find it helpful to calculate an example because of the nonlinear nature of the evidence functions.
 
@@
 
-1. We know with absolute certainty that reported results came about by p-hacking.
+1. We know with absolute certainty that researchers will revert to p-hacking to create evidence that is favorable for the theory.
 2. A hypothesis was picked to explain reported results after the fact [HARKing, @kerrHARKingHypothesizingResults1998].
@@
 In fact, both cases should be problematic to anyone who subscribes to the statistical relevancy condition because if we choose the hypothesis in accordance with the data or vice versa, without restrictions, they are not related anymore (i.e., observing the data does not tell us anything about the hypothesis and the other way around).
-Case 3, is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
+Case 3 is different since here the theoretical risk is not necessarily low but simply uncertain (and perhaps best represented by the dotted line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
 In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
@@
 Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
-Its low theoretical risk combined with high uncertainty about it, make it difficult to produce compelling evidence.
+Its low *and uncertain* theoretical risk make it difficult to produce compelling evidence.
 
@@
 
-Theoretical risk may conceptually be related to the conceptual framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
+Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
 However, there are crucial differences between the two.
@@
 Within our framework, researchers have the ability to input their assumptions regarding the perspectives of other researchers and calculate the potential impact of their actions on their readership, whether these actions relate to study design, to the preregistration itself, or subsequent deviations from it.
-We put subjective evaluations at the centre of our consideration; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence then others (they differ in their confirmation function).
+We put subjective evaluations at the centre of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).
 We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a framework to navigate the complicated questions that arise in the practice of preregistration.
@@
 Their reasoning is based on two intuitions about what should *not* influence the evaluation of evidence: temporal order and the mental state of the originator.
-For one, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
-However, we agree that the information flow matters for preregistration, not the temporal order.
+In our opinion, they disregard the temporal order a bit too hastily, as it is a long-standing issue in Bayesian philosophy of science known as the "problem of old evidence" [@chiharaProblemsBayesianConfirmation1987].
+However, we agree that not the temporal order is decisive but if the researchers incorporated the information into the hypothesis the evidence is supposed to confirm.
 For the other, we argue that the mental state of the originator does matter.
@@
 Where each $k$ has a $P(E_k | \neg H) > 0$.
-If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = \prod_{k=1}^K P(\neg E_k | \neg H)$?
+If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = P(\neg E_1 \land \ldots \land \neg E_k | \neg H)$?
 We think the latter.

### 2023-10-30 manuscript.Rmd
cite deviation paper
cite fiedler again

@@
 One may take such theoretical properties as a first starting point to form a substantive belief but surely researchers ought to take other factors into consideration.
-For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, this should decrease theorethical risk; after all they might find an association purely on account of the confounders.
+For example, if a researcher believes that there might be confounding variables at play for the relation between two variables, this should decrease theoretical risk; after all they might find an association purely on account of the confounders [@fiedlerWhatConstitutesStrong2017].
 
@@
 Or, if there is no obvious candidate, the researchers could conduct a multiverse analysis of the available options to deal with missings to show the influence of the decision [@steegenIncreasingTransparencyMultiverse2016].
+In any case, deviations must be transparently reported and we applaud recent developments to standardize and normalize this process [@willrothBestLaidPlans2023].
 

### 2023-10-30 manuscript.Rmd
add sem trees

@@
 However, two problems become apparent under closer inspection.
-First, many researchers do not subscribe to a purely confirmatory research agenda [@tukeyExploratoryDataAnalysis1972; @baumeisterChartingFutureSocial2016; @finkelReplicabilityOtherFeatures2017].
+First, many researchers do not subscribe to a purely confirmatory research agenda [@tukeyExploratoryDataAnalysis1972; @baumeisterChartingFutureSocial2016; @finkelReplicabilityOtherFeatures2017; @brandmaierStructuralEquationModel2013].
 Second, there is no strict mapping of the categories preregistered vs. non-preregistered onto the categories confirmatory vs. exploratory research.

### 2023-10-31 manuscript.Rmd

@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but we of course differ from Popper on the central issue of induction in science).
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but we of course differ from Popper on the central role of induction in science).
 
@@
 For the other, we argue that the mental state of the originator does matter.
-Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data.
-Where each $k$ has a $P(E_k | \neg H) > 0$.
-If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(\neg E| \neg H) = P(\neg E_1 | \neg H)$ or $P(\neg E| \neg H) = P(\neg E_1 \land \ldots \land \neg E_k | \neg H)$?
+Suppose there are $k = 1, 2, \ldots, K$ ways to analyze data, where each $k$ has a $P(E_k | \neg H) > 0$.
+If they intend to try each way after another but happen to be "lucky" on the first try and stop, should we then apply $P(E| \neg H) = P(E_1 | \neg H)$ or $P(E| \neg H) = P(E_1 \lor \ldots \lor E_k | \neg H)$?
 We think the latter.
@@
 We believe that our reasoning is quite similar to @hoflerMeansValuableExploration2022, who call for transparent exploration using preregistration.
-We could be surer of our agreement, if they had formulated their arguments within a mathematical framework, which would also have helped to dissolve an apparent conflict in their definitions of confirmation, exploration, and transparency.
+We could be more sure of our agreement, if they had formulated their arguments within a mathematical framework, which would also have helped to dissolve an apparent conflict in their definitions of confirmation, exploration, and transparency.
 On the one hand, they define "The principle difference between confirmation and exploration is that confirmation adheres to an evidential norm for the test of a hypothesis to pass.", but then suggest that transparent exploration can be conducted using inferences tests as a filtering mechanism.

### 2023-12-19 manuscript.Rmd

@@
 header-includes: |
-  \raggedbottom
-  \makeatletter
-
-  \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
-    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-    {-1em}%
-    {\normalfont\normalsize\bfseries\typesectitle}}
-  \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
-    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
-    {-\z@\relax}%
-    {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
-  \makeatother
+  \usepackage{amsmath}
+  \usepackage{amssymb}
 repro:
@@
     - inkscape
+output: rticles::springer_article
+keywords:
+  - preregistration
+  - confirmation
+  - exploration
+  - hypothesis testing
+  - Bayesian
+  - Open Science
 note: |
@@
   The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
-
 ---

### 2023-12-19 manuscript.Rmd
fix failed merge

@@
 class: "man"
-documentclass: "apa7"
-csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
-output:
-  papaja::apa6_pdf:
-    latex_engine: xelatex
 header-includes: |

### 2023-12-19 manuscript.Rmd
fix failed merge 2

@@
     institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
-keywords: |
-  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
 wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~8390')`"

### 2023-12-19 manuscript.Rmd
add csquotes

@@
   \usepackage{amssymb}
+  \usepackage{csquotes}
 repro:

### 2024-01-19 manuscript.Rmd
add meta psychology submission statement

--- 2023-12-19 manuscript.Rmd
@@
     institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
+keywords: |
+  preregistration; confirmation; exploration; hypothesis testing; Bayesian; Open Science
 wordcount         : "`r tryCatch(wordcountaddin::word_count(), error = function(x)'~8390')`"
@@
 class: "man"
+documentclass: "apa7"
+csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
+output:
+  papaja::apa6_pdf:
+    latex_engine: xelatex
 header-includes: |
-  \usepackage{amsmath}
-  \usepackage{amssymb}
+  \raggedbottom
+  \makeatletter
+
+  \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+    {-1em}%
+    {\normalfont\normalsize\bfseries\typesectitle}}
+  \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
+    {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+    {-\z@\relax}%
+    {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
+  \makeatother
 repro:
@@
     - inkscape
-output: rticles::springer_article
-keywords:
-  - preregistration
-  - confirmation
-  - exploration
-  - hypothesis testing
-  - Bayesian
-  - Open Science
 note: |
@@
   The manuscript is available as [preprint](https://www.doi.org/10.31234/osf.io/cs8wb) [@peikertWhyDoesPreregistration2023].
+
+  Submitted to *Meta-Psychology*.
+Participate in open peer review by sending an email to open.peer.reviewer@gmail.com.
+The full editorial process of all articles under review at Meta-Psychology can be found following this link:
+  https://tinyurl.com/mp-submissions  
+  You will find this preprint by searching for the first author's name.
+
 ---

### 2024-03-20 manuscript.Rmd
add missing software to docker image

@@
     - crsh/papaja@1c488f7
+    - pacman
   tlmgr:
@@
     - preview
+    - scalerel
+    - threeparttable
+    - endfloat
+    - was
+    - multirow
+    - threeparttablex
+    - environ
+    - trimspaces
+    - xpatch
+    - csquotes
   apt:

### 2024-03-20 README.Rmd
add documentation / fix local reproduction

@@
 
-## References
+## How to reproduce?
+
+### GitHub Actions
+
+Reproduction is done automatically on GitHub: [![Render and deploy PDF](https://github.com/aaronpeikert/bayes-prereg/actions/workflows/render-n-deploy-pdf.yml/badge.svg)](https://github.com/aaronpeikert/bayes-prereg/actions/workflows/render-n-deploy-pdf.yml), so forking the repo and pushing changes should be sufficient for you to reproduce.
+
+### Using Docker
+
+#### Prebuild Image
+
+This workflow requires Docker, Git, and Make installed:
+
+1. Get docker image, and give it a local tag:
+
+```
+docker pull ghcr.io/aaronpeikert/bayes-prereg:main
+docker tag ghcr.io/aaronpeikert/bayes-prereg:main bayesprereg:latest
+```
+
+2. Reproduce:
+
+```
+git clone https://github.com/aaronpeikert/bayes-prereg.git
+cd bayes-prereg
+make DOCKER=TRUE
+```
+
+#### Self build Image
+
+1. Build the image locally:
+
+```
+make docker
+```
+
+2. Reproduce:
+
+```
+make DOCKER=TRUE
+```
+
+#### Without Make/Git
+
+Since Make and Git are installed in the image, the docker container is technically all you need.
+However, retrieving the end result requires you to [bind a volume to the container (modify first comand).](https://docs.docker.com/storage/volumes/)
+
+```
+docker run ghcr.io/aaronpeikert/bayes-prereg:main bash
+git clone https://github.com/aaronpeikert/bayes-prereg.git
+cd bayes-prereg
+make
+```
+
+An alternative with graphical user interface is:
+
+1. Get docker image:
+
+```
+docker run ghcr.io/aaronpeikert/bayes-prereg:main
+```
+
+2. Use the url and password that is shown to you in the terminal to open RStudio in your web browser.
+3. Create a new project, click on "version control", click on "git", paste url: `https://github.com/aaronpeikert/bayes-prereg.git`
+4. Reproduce by running `make` in the RStudio terminal (not console!).
+
+### Locally
+
+#### With Git and Make
+
+Make sure you have all installed software you need.
+
+```
+git clone https://github.com/aaronpeikert/bayes-prereg.git
+cd bayes-prereg
+make
+```
+
+#### Per pedes
+
+1. Open RStudio
+2. Create a new project, click on "version control", click on "git", paste url: `https://github.com/aaronpeikert/bayes-prereg.git`
+3. Open `manuscript.Rmd`.
+4. Install all the packages listed under `packages` in the yaml metadata.
+5. Open `R/simulation.R` and run it.
+6. Open `manuscript.Rmd` and click on knit.

### 2024-06-03 manuscript.Rmd
typo

@@
 We are interested in why preregistrations should change researchers' evaluation of evidence.
-Applying a Bayesian framework allows us to investigate our research question most straightforwardly. 
+Applying a Bayesian framework allows us to investigate our research question most straightforwardly because it directly deals with what we ought to believe, given the evidence presented.
 Specifically, it allows us to model changes in subjective degrees of belief due to preregistration or, more simply, "persuasion".
@@
 In fact, this conceptualization is intentionally as minimal as possible to be compatible with a wide range of philosophies of science and statistical methods researchers might subscribe to.
+One feature of the Bayesian framework, is the strong emphasis on subjective yet rational judgement.
+Therefore, we assume that researchers will differ significantly in how they value evidence but that by making assumptions about the general process, we can make general statements that apply to all these subjective evaluations.
 However, we should note that Popperians would be appalled that we are content with positive inductive inferences (but we regard "failing to disprove" as too limited), and Neopopperians would flinch that we assign probabilities to beliefs (we are fond of calculating things).
-While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with the behavior of researchers.
+While the latter move is not strictly necessary it allows us to connect the more abstract considerations more closely with what researchers believe.
 
@@
 Within our framework, researchers have the ability to input their assumptions regarding the perspectives of other researchers and calculate the potential impact of their actions on their readership, whether these actions relate to study design, to the preregistration itself, or subsequent deviations from it.
-We put subjective evaluations at the centre of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).
+We put subjective evaluations at the center of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).
 We, therefore, hope to not only provide a rationale for preregistration for those who subscribe to a Bayesian philosophy of science but also a framework to navigate the complicated questions that arise in the practice of preregistration.

### 2024-06-05 reviews/methods1.Rmd
draft responses to Lakens

---
output:
  pdf_document: default
  html_document: default
---

All changes of the manuscript since this review can be viewed on GitHub:

<https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main#diff-a9a4aad3fa8c9c10c5404b632bc3a01a25d2d8430eb932bc35c76769963e4b70>

Our responses to the review are formatted as follows.

Normal text = the reviewer's comment

> Indented text = our response

> *Indented italicized = quote from the revised paper*

Reviewer #1: The authors set out to explain what the primary aim of preregistration is.
I think they are correct that many scientists have hardly thought about this question.
This is not in itself bad - preregistration intuitively makes sense - but it is important to clearly articulate what the aim of preregistration is.

> We agree that any tool applied thoughtlessly will seldom be effective.
> A clear aim of preregistration could be helpfull to shape its application, especially given that the scientific comunity is still in an experimenting phase around central issues of preregistration, e.g., what level of detail a preregistration should have, under what circumstances to deviate from it, how to review it, or how to deal with exploratory research questions.

So I applaud the authors for believing this is important.
At the same time, their arguments were, I believe, not always equally convincing, and they do not adequately discuss the existing literature.

> To not adequately cite Laken's work was a grave and frankly embarrassing mistake that we regret and did without intention.
> Laken's (and Mayo's) work on severity was at the front of our mind while conceptualising the ideas presented here, so how we missed that we did not explicitly cite them is still baffling to us.
> This mistake can only be explained by us being (wrongly) convinced that we cite them.

The main argument of the authors is straightforward: preregistration reduces uncertainty about the Type 1 error rate, and under the assumption that it is possible to inflate the Type 1 error rate when a study is not preregistered, reducing the uncertainty through preregistration increases the posterior probability of the hypothesis after a test that yields a significant result.
The authors say they provide a 'Bayesian rationalization'.
This is not correct.

> We respectfully disagree.
> What makes the arguments "Bayesian" is what we take probabilities to mean.
> For a Bayesian, they represent degrees of belief.
> For a frequentist, they represent (relative) frequencies of events that result from a random process.
This is a subtle but most important distinction.
> The Bayesian position deals with what agents (here "researchers") ought to believe, assuming they are rational.
> However, even here, there is no universally agreed upon function that maps evidence onto belief (indeed, there can not be).
> We tried to make this even clearer and more explicit as to why we include several functions (and that our reasoning applies to all possible functions satisfying our assumptions).
> The frequentist understanding of probability refers strictly to statistical procedures.
> From this position, one must either argue that preregistration is changing the underlying random process/ statistical procedure or trace the impact of preregistration via indirect paths outside of the core philosophy of frequentism.
> Mayo's severity perspective is taking one more step in this direction by asserting that belief in a hypothesis is justified to the extent that it survives stringent scrutiny.
> However, even this "strong" severity principle is only a proportional statement.
> Faced with a specific tradeoff, e.g., is a particular post hoc change of a preregistration beneficial, we must revert to vague notions and "gut" instinct under this philosophy.
> Within the Bayesian framework, we may directly ask: What consequences should preregistration have on the beliefs of individual scientists or the research community as a whole?

Their rationalization is in essence based on the likelihood ratio.
All else equal, a lower Type 1 error rate means that observing a significant result is relatively more likely when H1 is true.
Of course a higher likelihood ratio also means a higher Bayes factor, which means a higher probability that H1 is true in a Bayesian framework.

> It is well known that the math for frequentists is similar or even identical to Bayesian under the right conditions.
> Still, a frequentist would not conclude anything about the probability of H1.
> However, our goal is to say something about how preregistration changes how believable researchers should find claims made in the literature.

But the workhorse in the rationalization that the authors provide is the likelihood ratio.
This does not change when we put a distribution on the Type 1 error rate (as the authors do).

> Yes, the Bayes formula can be conceptualized as a likelihood ratio scaled by priors.
> However, the mere use of the likelihood ratio in the Bayes theorem can not be grounds to reduce Bayesianism to likelihood ratios.

The distribution (or the move to a Bayesian rationalization) does not add anything (although of course it can be added).

> We find that it adds value by making an otherwise implicit process explicit to the point of calculatable.
> By conceptualising uncertainty with a specific distribution, we rid ourselves of vague notions of transparency, control, or exploration.
> This is of utmost importance because only this clarity enables us to effectively navigate the tradeoffs involved.

Importantly, the likelihood function of course also underlies the uniformly most powerful test in a Neyman-Pearson hypothesis testing framework.
So for a frequentist (and I mean a frequentist not in statistical terms, but also in Popperian anti-Bayesian view on confirmation) the same argument for preregistration plays a role.

> We agree that this argument could perhaps be made for a frequentist as well.
> However, it is not the math of the likelihood ratio that eludes us; it is how a significant result of a statistical test changes how a frequentist judges a hypothesis.
> There is a great, yet unresolved, debate in frequentism about how to do just this.
> In a Fisherian approach, no heat is paid to the power of a test; only the p-value is interpreted as the strength of evidence against the null.
> In the Neyman-Person paradigm, a decision is to be made about rejecting one or the other hypothesis.
> Mayo's severe testing returns to a gradual inference (instead of a decision) by evaluating the extent of evidence in favour of H as the extent to which a test was capable of showing that H is false.

The title should not read a Bayesian rationalization, but a likelihood rationalization.
A higher likelihood is not always better, as one can cherry pick a high likelihood ratio.
This is extensively discussed by Mayo (2018).
After p-hacking, we could get a very high likelihood (or Bayes factor) that supports H1 - but it was cherry-picked, and is highly likely to be a fluke.
Therefore, the authors limit themselves to relative likelihoods.

> We did not limit ourselves to relative liklihoods.
> We explicitly include both the "firmness" and "increase in firmness" perspective.
> That a firmness perspective has significant flaws is of no matter to our argument since we intend to show that someone who subscribes to either must be more persuaded by a preregistered study.

They point out that, assuming people only increase their alpha level when p-hacking, preregistration leads to tests with a lower alpha level.
As such, they basically restate Mayo's (2018) argument about the severity of a test.

> Let us cite <https://doi.org/10.3758/s13423-022-02069-1> "Nowhere in Mayo’s severity principle or in the specifications discussed below is the riskiness or specificity of a hypothesis, and the predictions it makes, incorporated as a requirement for a severe test (Meehl, 1978; 1986; 1990b, a; Roberts & Pashler, 2000)."

However, they never make this explicit, and I think this is to the detriment of the paper.

> We added a new section that deals with the concept of severity, making it explicit how our approach differs fundamentally.

The fact that they subsequently add a Bayesian sauce to the likelihood ratio (by adding a distribution to the Type 1 error rate when a study is not preregistered, instead of simply making some assumption about a point-value of the inflated Type 1 error rate) does not add anything to their argument.
So, I believe the authors have basically attempted to re-invent Mayo's severity argument.

> The whole second part of the paper deals with the problem of not being able to reduce the theoretical risk to a point estimate.
> The statistical procedure has an objective Type 1 error rate, for which we can come up with a point estimate if we agree on the conditions, which is a starting point for evaluating the theoretical risk of the scientific (!= statistical) procedure.
> Since, a reader does not have access to the whole scientific procedure, they must judge it with uncertainty.
> We find it important to distentengle the statistical procedure, the scientific procedure, and the outside perspective of the reader.

The fact that this is not immediately clear is mainly due to missing discussions of the existing literature.
Point 1: Preregistration can not objectively minimize uncertainty about the Type 1 error rate The main argument of the authors is that risky predictions create persuasive evidence if they turn out to be correct.
Preregistrations are ideal for communicating the theoretical risk researchers took when testing a claim, and communicating this removes uncertainty in the reader about the theoretical risk the authors took.
I think this argument is in the right direction, but it is not fully developed.
I was somewhat disappointed not to see a discussion of my own paper on a conceptual analysis of the value of preregistration, as I had a similar goal as the authors when I wrote my paper, and my arguments are in the same direction, but dare I say, slightly better developed.

> We find ourselves in a curios position, that perhaps may be summarised as "doing the right thing, for the wrong reasons", i.e., we agree with Laken's conclusions about preregistration but we disagree how to come to these conclusions.
> We take a Bayesian/inductionist view, whereas Lakens takes a frequentist/anti-inductionist view, agreeing on the end result but disagreeing on everything that comes before.
> While we find it important that readers are aware, how anti-inductionist deal with preregistration, this paper was not meant as an exercise to compare Bayesian justification for preregistration with its frequentist counterpart.

Maybe everyone thinks this of their own papers, so forgive me if I am too biased, but I will attempt to point out some weaknesses in the current paper, sometimes by citing my own paper.
I hope the authors see how strongly the ideas are related, and forgive me for citing myself.

> Of course, we see the relevance of Laken's work; we met in Eindhoven to discuss an early version of the paper.

The authors argue that one interpretation of preregistration is "the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
I would say it is better to argue preregistration aims to *control* the error rate, not maximize the true negative rate.
The authors then propose their own view: "the objective of preregistration is not the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk."
I think the authors are in the right direction, but miss some things.
What the authors get right is that preregistration is about communicating something about the test to others.

> This is why we deliberately chose to use the term "uncertainty" and not "control".
> Reducing uncertainty directly invokes the concept of communication.
> Control also seems to imply that it is a property of the procedure, as it seems curious to suggest that control differs between the reader and the original researcher (we argue that the uncertainty is lower for the latter as they have all the information).

And the authors are correct that preregistration can remove uncertainty if all readers of a preregistration 1) agree on the analyses plan, and 2) there are no deviations of the preregistration.

> We make no such claim.
> We explicitly address that readers may differ considerably in how highly they judge the theoretical risk and their confirmation function.
> *One feature of the Bayesian framework is the strong emphasis on subjective yet rational judgement.*
> *Therefore, we assume that researchers will differ significantly in how they value evidence but that by making assumptions about the general process, we can make general statements that apply to all these subjective evaluations.*
> They may, therefore, very well disagree on the analysis.
> We also discuss at length under which circumstances a post hoc change to the preregistration is justified (if the increase in confirmation does offset the negative effects of uncertainty).

However, the authors already say 2 is not always true, as people often need to deviate.
And 1 is also not true.
This is important, because if 1 and 2 are true, then we can not minimize uncertainty about theoretical risk (as the authors argue the goal of preregistration is), because this uncertainty is based on a subjective evaluation.
If you deviate from an analysis plan because you think after looking at the data that you should have excluded some outliers, but this was not specified in the preregistration, as a reader I am left with an evaluation of whether 1) this analysis is a *better* test, and it reduces the possibility of incorrectly claiming there was an effect, because these outliers might make a null effect significant, while they should have been removed, or 2) this analysis is a *worse* test, and the reader believes the choice for this test without the outliers is driven by the nicer results the researcher observed.
The preregistration does not minimize uncertainty about the Type 1 error rate.
For some readers, the deviation increased the Type 1 error rate, and for others it did not.

> This aligns with our reasoning: Any change should increase uncertainty about the theoretical risk.
> How much and how this ultimately changes the persuasiveness of the evidence will differ between readers.

So what *does* preregistration achieve?
In my own words "Preregistration has the goal to allow others to transparently evaluate the capacity of a test to falsify a prediction." The lower the Type 1 error rate, the higher the capacity of a test to falsify a prediction.
But whether a test, especially after a deviation, has a higher or lower Type 1 error rate, remains a subjective decision.

> Agreed. It remains a subjective decision.
> As subjective yet rational judgement is such a fundamental property of Bayesian reasoning, we may not have emphasized this enough.
> We now take up this subject in the discussion:
> *We put subjective evaluations at the centre of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).*

Preregistration does not give certainty about the theoretical risk, as the authors argue.
It gives more certainty in most cases at best, but it allows others to evaluate it.
This might sound like a subtle difference, but I think it matters.

> Agreed.

The authors might reply that the did not really mean 'minimize uncertainty about the theoretical risk' but mainly meant that readers can judge for themselves what the Type 1 error of a test is (based on their own assumptions about what the most appropriate test is).
In that case, their argument does not differ from mine in Lakens (2019), and the authors should explain what their contribution adds.

> We took considerable time to put what we meant on paper.
> In the revision, we tried hard to be clearer that uncertainty always refers to the uncertainty of the reader.
> However, being content with claiming that readers have to judge for themselves what the theoretical risk/severity of a test results in a weak argument for preregistration.
> We go further and show that for all individual assessments, preregistration has a benefit and that if one is willing to assume some parameters, this benefit is striking.

Point 2: It remains unconvincing why preregistration is useful for exploratory studies.
The authors seem to argue that even in exploratory studies preregistration will remove *some* uncertainty about the ability of a test to falsify a prediction.
That is true.
However, they do not argue why it would do so to an extent that makes preregistration useful in real life. If an action I do reduces my uncertainty from 98% to 94%, it has relatively reduced my uncertainty.
At the same time, it has done so in a manner that does not matter at all.
My point here is that the error rate in exploratory research is inflated above alpha% to an unknown extent.
A preregistration might reduce the inflation somewhat.
But the rate is quickly too high so that I wouldn't trust the claim anyway.

> Actually, we think this is one of the strengths of the paper.
> With the mathematical analysis, we show exactly how much uncertainty, theoretical risk, and detectability result in how much confirmation, as seen in Figures 1 and 2.
> The vague intuition of a "[Type 1 error] rate [that] is quickly too high so that I wouldn't trust the claim anyway" is replaced by a clear functional form that enables navigating tradeoffs involving exploration, post-hoc changes or detectability.

If a claim is made with 10% lower Type 1 error rate than it would have been made when it was not preregistered is not useful - because I do not know what the error rate associated with the claim is.

> Our claim is that uncertainty is reduced, not theoretical risk directly.
> We claim that there are exploratory methods with a well-known error rate.
> For example, a neural network has extreme flexibility in fitting the data closely, yet it is calculatable.

Thus, preregistration in exploratory studies reduces the error rate inflation, but it does not make the error rate known.
Therefore, I do not believe preregistration in exploratory studies is useful, and the authors do not make an argument for it.
We do not want to reduce the error rate inflation in preregistration - we want to control it.

> Somehow, our argument seems to be flipped here.
> An exploratory study has low theoretical risk.
> A preregistered study has low uncertainty about theoretical risk.
> A preregistered exploratory study is more persuasive than a study with the same exploration going on that is not preregistered.

If the error rate is not controlled, I do not know how severely a claim has been tested, and I can not use the claim to decide how to act as a scientist.
The authors seem happy with the fact that preregistration reduces the Type 1 error rate even in an exploratory study.

> No, we are happy with low uncertainty about theoretical risk.
> An exploratory study is in our framework, signified by low theoretical risk (high Type 1 error rate).

But I am not. Only claims with a known error rate are useful in science.

> Exactly. The error rate must be known (readers must be certain about them).
> It may be high however (=exploratory).

If the authors disagree, they should say how an unknown relative decrease in the Type 1 error rate to a final still unknown type 1 error rate is beneficial.

> We have a hard time following this reasoning.
> Exploratory does not imply that theoretical risk is unknown.

Point 3: The Type 2 error rate matters as much as the type 1 error rate I do not believe the relationship the authors draw between 'theoretical risk', 'risky predictions', 'evidence', and the 'probability of a hypothesis' is coherent.
I especially disagree with the authors that the posterior probability P(H\|E) is important in preregistration.
The authors say this probability is 'often used directly or indirectly as a measure of confirmation of a hypothesis'.
This is simply not true.
No one discusses posterior probabilities of hypotheses in psychological science.

> Neither is severity as an absolute measure of corroboration a common consideration in empirical psychological work.
> However, the posterior probability is certainly the most classic measure of absolute confirmation in the philosophy of science.

The authors then discuss an 'increase in firmness'.
Logically, this would be the Bayes factor, not the posterior probability of the hypothesis, that the authors are talking about.
The Bayes factor is a continuous measure of how much we update our belief.
If we want to argue for any function of preregistration from a Bayesian perspective, it must be focused on the Bayes factor.
The authors then say 'If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal'.

> We explicitly consider the whole space of possible measures. From the revised paper:
> *The goal is therefore to reason about the space of possible measures researchers might apply.*
> *However, since any measure fulfilling the statistical relevancy condition increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.*
> Later in the paper: *Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher (and even the measure of confirmation), but we can illustrate the effect of uncertainty with an example.*

This statement is not in line with the earlier statement about an 'increase in firmness'.
Of course, if a hypothesis is highly probable (the Stroop effect), any evidence will end up with a higher posterior probability that if the prior probability was low.
But the same evidence would lead to a greater 'increase in firmness' (or a higher Bayes factor) if the prior was lower.
The same data would lead to a much greater update in belief.
The authors dismiss the prior probability of the hypothesis as it 'is nothing our study design can change'.
However, this is not true.
Although the authors only discuss inflating the alpha level in their paper, one function of preregistration is to prevent HARKing, or hypothesizing after the results are true.
Of course, if one comes up with a hypothesis after looking at the data, the 'prior' is 1: We already know the hypothesis will be supported by the data.

> We included a section called "Hacking, harking, and other harms", which hopefully makes our thinking on these things clearer.

HARKing is a major reason the field has moved towards preregistration, but the term 'HARKing' is not even mentioned in the article.
The reason for this is that the arguments the authors provide can not deal with HARKing.
This is obviously a limitation, compared to existing arguments for preregistration, and it is incorrectly ignored in the article.
As Mayo (2018) explains, the problem with HARKing is that it leads to a non-severe test:
The data have no probability of falsifying the claim.
This error-statistical perspective seems to me to be much more fruitful than a Bayesian approach.

> As stated above, we come to the same conclusions despite being inductionists.

The authors then move on the P(E\|H), or power.
They argue 'while detectability is of great importance for study design, it is not directly relevant to the object of preregistration.' This is incorrect.
As I write in Lakens, 2019: "In line with the general tendency to weigh Type 1 error rates (the probability of obtaining a statistically significant result when there is no true effect) as more serious than Type 2 error rates (the probability of obtaining a non-significant result when there is a true effect), publications that discuss preregistration have often been more concerned with inflated Type 1 error rates than with low power.
However, in situations where researchers want to find a null effect low power is a bigger concern."
The idea of null-hacking is not crazy - sometimes a certain variable should *not* be significant to exclude alternative explanations. Therefore, preregistration certainly should have the function to prevent researchers from incorrectly claiming there is no effect, when there is an effect, by inflating the Type 2 error rate (or equivalently, by reducing the power).

> Absolutely. However, note that our H is always the research hypothesis, independent of the statistical H1 or H0.
> If a researcher wants to provide evidence for a null hypothesis, this is their H, theoretical risk, then is related to power.
> From the revised paper:
> *In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.*
> We reference the ideas of frequentist null hypothesis testing to provide readers who are more familiar with this line of reasoning with a better intuition, not to claim equivalent.

This aspect of Bayes' formula is therefore also incorrectly ignored in the article.
The authors then argue that P(E) is the only part of the formula that remains to be considered.
Decreasing P(E) can increase the posterior probability.
Of course, P(E) is a function both of the power and the Type 1 error rate.
But for some unclear reason, the authors here dismiss power, as they write "We have already noted that there is not much to be done about prior probability ((), and hence its counter probability (¬)), and that it is common sense to increase detectability (\|)."
But this latter statement dismisses the important role of power much too easily.
One could just as well say it is 'common sense' to reduce the Type 1 error rate.
Furthermore, once again, we do not 'increase' error rates, we *control* them.

> Perhaps, we are a bit daft but a clear definition of error rate control eludes us.
> We are under the impression that two competing definitions are used here from time to time.
> First, a definition in which control means the error rates are known (but then known by whom?).
> Second, control means ensuring certain acceptable bounds of error rates are not exceeded.

And while controlling them, we always have to consider resource constraints (Lakens, 2022).
Indeed, one might even say that we are always balancing Type 1 and Type 2 error rates, and that it would be sensible to move away from a fixed alpha level for all studies (Maier and Lakens, 2022).
In any case, the authors fail to adequately argue why power can be ignored, and the Type 1 error rate is important.

> We can not deliver a treaty on statistical inference as a whole.
> Nowhere have we stated that power can be ignored for statistical inference; we just do not think that power is the main reason why preregistration is a good thing.
> We agree that detectability is important; after all, it shows in every mathematical consideration we made that it is a vital consideration for a Bayesian.

According to the error-statistical justification of preregistration (Mayo, 2018; Lakens, 2019) both are equally important.
At the very least, increasing the alpha level would also increase the statistical power, so the two can not even be seen as separate.

> We wholeheartedly agree with the importance of being able to navigate these tradeoffs.
> That is why we find a mathematical treatment of these ideas to be so important.
> The Bayesian perspective allows us to ask, very concretely, how persuasive an idealized Bayesian thinker finds different configurations, including the ratio of theoretical risk and detectability.
> We clearly say multiple times that these tradeoffs can not be optimized separately.
> In the revised paper:
> *As noted before, we strive for high detectability and high theoretical risk in planning, conducting, and analyzing studies.
Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
One advantage of our framework is that researchers can employ it to balance the trade-off more effectively, assuming they are willing to make some simplifying assumptions.*

POINT 4: Terminology As terminology can always be fixed, this point is not as problematic to the reasoning as the previous 3.
The authors use the word 'theoretical risk' for the probability that a test of a hypothesis yields a true negative.
As the authors exclusively talk about the probability that statistical tests will yield a correct conclusion, 'theoretical risk' feels as a misnomer (statistical risk would be better, but severity might also be a good general term).

> We talk about the scientific procedures, of which a statistical test is a piece.

True, in page 13 they give an example about theoretical risk that is actually theoretical risk - but this example has nothing to do with the remainder of the paper, as all three studies can have exactly the same alpha level!

> We are unable to follow this argument without more elaboration.

They are talking about a low frequentist error rate in all their examples, and as 'theoretical risk' is the inverse of the false positive rate, it might be easier to just talk about the Type 1 error rate.

> The point of not using a Type 1 error rate is to differentiate the statistical test from the the researcher's scientific judgement.
> Type 1 error rate is a property of a test under certain conditions.
> Theoretical risk refers to the individual judgement of a researcher about the substantive hypothesis and the study design as a whole.

The authors say they borrowed the term 'theoretical risk' from Meehl, but that he did not assign it to (¬\|¬).
The authors cite Kukla (1990), but after reading it, I am not convinced they are correct in equating 'theoretical risk' with the Type 1 error rate.
Theoretical risk means making an a-priori unlikely prediction if the theory was not true.

> This sentence sounds exactly like P(¬E|¬H) to us.
> The judgement about P(¬E|¬H) is a-priori, is about "prediction" as a statement about evidence E, given that the theory is not true (¬H).
> One might argue, that it should be multiplied by ¬H perhaps? 
> But that would not change any of the arguments.
> In any case, we provide clear mathematical definitions for our terms.
> If a reader does not agree with our terminology, they are free to term them differently.
> In fact, this happens in reference to a finding we found in the meantime (Oberauer and Lewandowsky (2019)), which we added to the discussion.

This could be, as Meehl would like, a range prediction. Or, in the example of the authors on page 13, it is a test without confounds.
In Lakens, 2019, on page 224, I explain how a two-sided test, one-sided test, and a range prediction all differ in how severe they test the hypothesis, even if they all have the same 5% alpha level ("Note that the three tests differ in severity even when they are tested with the same Type 1 error rate.").

> Yes, indeed, this could all have low theoretical risk.
> We could include the disclaimer to the examples on p. 13 "Note that the three tests differ in theoretical risk even when they are tested with the same Type 1 error rate."

This is another reason why I think the term 'theoretical risk' is not a good choice.
Indeed, based on the example on page 13, it seems the authors themselves confuse real theoretical risk with Type 1 error control.
However, it is better to use terminology that encompasses Type 1 and Type 2 errors (such as 'the probability that a test will falsify a hypothesis', or even better: the *severity* of a test).

> Different philosophies differ in their terminology.
> Deborah Mayo has good reasons to differentiate Type 1 error rate from severity, as do we.

Using the term 'severity of the test' would also embed the work more strongly in philosophy of science (and specifically the work of Mayo, as she builds on Popper).

> This would embed us in a different philosophy of science.
> Popper is the most famous anti-inductionist; we deliver an inductionist account.

Mayo's work has been exactly on this topic, and it is a shame it has not been discussed, even though she discusses the problem of p-hacking and Type 1 error rate inflation extensively in her 2018 book.
The authors write "The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value".
This is an unfortunate term, as epistemic values are already used in philosophy of science to refer to qualities such as predictive accuracy, internal coherence, external consistency, unifying power, and fertility - see Elliott, K. C. (2022). Values in Science (1st ed.). Cambridge University Press. <https://doi.org/10.1017/9781009052597>.
I would pick another terminology.
In our own work published in Psychological Methods we called this "the marginal gain in expected utility (or usefulness) of scientific claims". (Isager, P. M., van Aert, R. C. M., Bahník, Š., Brandt, M. J., DeSoto, K. A., Giner-Sorolla, R., Krueger, J. I., Perugini, M., Ropovik, I., van 't Veer, A. E., Vranka, M., & Lakens, D. (2023).
Deciding what to replicate: A decision model for replication study selection under resource and knowledge constraints. Psychological Methods, 28(2), 438-451. [https://doi.org/10.1037/met0000438)/](https://doi.org/10.1037/met0000438)/){.uri}.
There are probably other existing terms for the same idea in the literature. The authors discuss P(E\|H) as 'detectability' or the probability of detecting evidence if a hypothesis holds.
This is thus simply statistical power, or the true positive rate.
I think the term detectability is not particularly useful, as it is more strongly associated with signal detection theory - but that is not what the authors mean with the term, as they admit the term is simply statistical power when applied to hypothesis testing (and I don't know what else it could be applied to).

> Hypothesis testing and scientific inference are different things.
> A test might be significant, but if a researcher suspects strongly enough that there might be a problem, they may decide not to believe the hypothesis.

Minor points: The authors say "We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value." when they relate a property of their Bayesian model with Popper's work.

> No, we relate our work to beliefs we think are commonly held.
> Researchers find the idea of a risky test appealing.
> Popper gives one explanation for this appeal, but we happen to have a different view.

However, this is misleading.
Popper would never agree with the approach in this paper, as he does not think subjective prior or posterior beliefs should play a role in scientific claims.
According to Popper, verisimilitude is not about belief. Popper, p. 434: "As to degree of corroboration, it is nothing but a measure of the degree to which a hypothesis h has been tested, and of the degree to which it has stood up to tests.
It must not be interpreted, therefore, as a degree of the rationality of our belief in the truth of h".
The author's proposal violates the philosophical work of Popper at it's very core, and is incommensurable.

> As stated above, if Popper would happen to agree, we indeed would have to rethink our arguments.

Just because the authors also need to rely on the likelihood ratio, does not mean it 'reconstructs' Popperian philosophy of science.

> Which we never claimed.

It is not clear what the researchers mean with 'Researchers judge the evidence for or against a hypothesis rationally'. Researchers often differ in whether a study is evidence for or against a hypothesis, often because they disagree on auxiliary assumptions (e.g., A thinks X is a good measure of Y, but B does not think X measures Y).
Are these two researchers irrational? Are the authors talking about some utopian world, in which case this assumption seems violated at least for now.

> No, these researchers are not irrational according to our definition.
> For a researcher to be rational, they have to be internally consistent.
> They must not change their process of weighing evidence just because they do not like the outcome.
> We take this assumption from the literature of the Bayesian philosophy of science and cite it appropriately.

The authors admit they are talking about an 'ideal' estimation process.

> We meant "ideal" in the sense of a physicist talking about ideal gas.

But in an ideal estimation process we would also have infinite data.
Maybe I am being too pragmatic, but shouldn't we discuss how research is actually done?

> We agree to assume infinite data would go a little too far into wishful thinking.
> Also, with infinite data, it stops being fun to apply Bayesian reasoning since there is no uncertainty to deal with.

The authors write "Indeed, limiting the type-I error rate is commonly stated as the central goal of preregistration (Nosek et al., 2018; Oberauer, 2019; Rubin, 2020)."
It might be commonly stated. But then, people also commonly state the incorrect definition of a p-value.
Instead, the authors could cite a better paper, like my own: "The goal of preregistration is not simply to control the Type 1 error rate in hypothesis tests, but to prevent researchers from non-transparently reducing the capacity of the test to falsify a prediction in general."
I worked quite hard on conceptually clarifying the role of preregistration, correcting exactly the kind of misconception the authors also respond against.

> We appreciate this hard work; we just do not think it is quite done yet.

The authors write "First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis (Mayo, 2018, p. 5.3)."
This is generally true, but of course null effects often *are* evidence for the null - as would be clear if an equivalence test was performed, or a Bayes factor or likelihood ratio was computed.

> With the following statement, we hoped to make clear that there are many statistical procedures that can be investigated through the lens we adopted here:
> *Please note that our decision to adopt a Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.*

The authors argue (page 15) theoretical risk also encompasses factors outside the statistical realm.
I agree, and say so in Lakens, 2019: "Making very narrow range predictions is a way to make it statistically likely to falsify your prediction if it is wrong.
But the severity of a test is determined by all characteristics of a study that increases the capability of a prediction to be wrong, if it is wrong.
If a researcher predicts that an effect will only be observed under a very specific set of experimental conditions that all follow from a single theory, it is possible to make theoretically risky predictions."
Line 425: There are 3 arguments, not 2.
Conclusion: Again, I appreciate the authors digging in to the justification of preregistration, but I think there are some important things missing in this version of the manuscript.
How can the article be improved? The authors should address harking.

> We included a section called "Hacking, harking, and other harms", which hopefully makes our thinking on these thinkgs clearer.

They should explain better why they think preregistration of exploratory studies is useful.

> We hope that the improvements that we made to better explain the topics of subjectivity and uncertainty make clearer that theoretical risk and uncertainty about theoretical risk are two dimensions.

They should improve terminology.

> We take the criticism of the terminology seriously.
> Although we have opted to stick with most of the terms as we defined them, we provided more references.
> It might also be the case that the differences in philosophical traditions lead to some criticism.

The should better integrate their ideas with existing ideas, especially Mayo's work on severity, but at least engage with the arguments for preregistration in the literature more than a somewhat strawman account of minimizing the type 1 error rate.

> We have discussed Mayo's Severity to show that these approaches fundamentally differ, and we have significantly extended the discussion to include literature that also uses a Bayesian rationale to investigate what methods researchers should employ.
> Additionally, we substantiated the "strawman" further because we believe that the minimizing vs communicating researchers' degrees of freedom confusion is at the centre of many debates around preregistration.

They should also address 'null-hacking'.

> We have merely mentioned this topic in passing since it is conceptually only difficult if the null hypothesis has a special position, as in frequentist reasoning.
> However, we significantly expanded on the topic of "hacking" in general, under which this topic falls for us.

After these changes, I am not sure what their argument will end up looking like, but I would hope the authors explore their ideas further.
Sorry for citing myself, no need to cite me - others often make the same points I make, so pointing you to my own papers is mainly just easier for myself. Lakens, D. (2019). The value of preregistration for psychological science: A conceptual analysis. Japanese Psychological Review, 62(3), 221-230. <https://doi.org/10.24602/sjpr.62.3_221> Lakens, D. (2022). Sample Size Justification. Collabra: Psychology, 8(1), 33267. <https://doi.org/10.1525/collabra.33267> Maier, M., & Lakens, D. (2022). Justify Your Alpha: A Primer on Two Practical Approaches. Advances in Methods and Practices in Psychological Science, 5(2), 25152459221080396. <https://doi.org/10.1177/25152459221080396>
Signed,
Daniel Lakens

> Thank you for your review and for allowing us to share it!

### 2024-06-07 reviews/PsychBulletinEditor.Rmd

---
output:
  pdf_document: default
  html_document: default
---

All changes of the manuscript since this review can be viewed on GitHub:

<https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main#diff-a9a4aad3fa8c9c10c5404b632bc3a01a25d2d8430eb932bc35c76769963e4b70>

Our responses to the review are formatted as follows.

Normal text = the reviewer's comment

> Indented text = our response

Dear Mr. Peikert:

I have now read Manuscript PBR-TR-23-119 entitled "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization" that you submitted to Psychonomic Bulletin & Review.
You offer a new perspective on an important issue, and I truly enjoyed reading through your arguments for the role of preregistration in reducing uncertainty about theoretical risk.

> We are thankful for the kind words.

That said, this is more of a think piece than a theoretical review, and the value of assessing theoretical risk remains hypothetical.
I think you would agree that your examples are highly schematic and serve to illustrate the principles that you are proposing as opposed to reporting specific examples of how these principles can improve psychological research.

> We fully agree.

Perhaps it would be possible to compare a range of studies with and without preregistrations to show how the act of preregistration helps clarify the reported results even when the preregistration does not report the level of detail required for a truly confirmatory study.

> We are sure this is possible and would be enlightening but we would not be sure how to relate the results to the paper.
> The issue is, that how close a theory is to truth, is very difficult to measure.
> That implies that measuring if our approach is successfull is similarly difficult to measure empierically.
> Our goal is more modest: Showing that if you subscribe to a certain class of philosphies of science, you should find preregistered evidence more persuasive.

Given your broad scope, the reviewed studies would ideally come from a variety of research domains.
In other words, perhaps it is possible to put your ideas through a few test runs to see how they can be applied when all the messy details of specific research endeavors are considered.

> Applying preregistration with "all the messy details of specific research endeavors" arguably already leads to quite some confusion.
> We hope that some conceptual clarity helps researcher in practice.
> This hope is grounded in Kurt Lewins Maxim: "Nothing as practical as a good theory".

In my view, theoretical reviews propose a new set of ideas *and* systematically apply the new ideas to existing findings to establish new discoveries or bolster conclusions with more rigorous methods.
I don’t see the second element here, and I don’t see how getting reviewer comments would change that impression.

> We agree. Empierically establishing our perspective on preregistration as "correct" would be very difficult.

ACTION: I am rejecting the manuscript.
Although a desk reject seems like a very negative outcome, I want to again emphasize that I found your analysis creative and valuable.
I just don’t think a PBR theoretical review is the appropriate outlet.

Thank you for considering Psychonomic Bulletin & Review for the publication of your research.

Sincerely, Jeff Starns Action Editor, Psychonomic Bulletin & Review

Editor: James Brockmole

### 2024-06-27 manuscript.Rmd
remove the term "epistemic value"

@@
 
-This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward.
+This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward [@lakensWhenHowDeviate2024].
 The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
@@
 
-# Epistemic value and the Bayesian rationale
+# Persuation and the Bayesian rationale
 
-Let us start by defining what we call expected epistemic value.
 If researchers plan to conduct a study, they usually hope that it will change their assessment of some theory's verisimilitude [@niiniluotoVerisimilitudeThirdPeriod1998].
-In other words, they hope to learn something from conducting the study.
-The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value.
-Researchers cannot know what exactly they will learn from a study before they run it.
-However, they can develop an expectation that helps them decide about the specifics of a planned study.
-This expectation is what we term expected epistemic value.
+Moreover, they hope to convince other researchers can be persuaded to change their believe in a theory as well.
+Beforehand, researchers cannot know what evidence a study will provide but still must form an expectation in order to decide about the specifics of a planned study, including if they should preregister it.
+If they can expect that preregistration helps them to persuade other researchers to change their believe, it is only rational to employ preregistration.
 To make our three arguments, we must assume three things about what an ideal estimation process  entails and how it relates to what studies (preregistered vs not preregistered) to conduct.
@@
 2. They expect other researchers to apply a similar rational process.
-2. Researchers try to maximize the expected epistemic value for other researchers.
+2. Researchers try to maximize the expected persuasiveness for *other* researchers.
 
@@
 Researchers who decide to conduct a certain study are actually choosing a study to bet on.
-They have to "place the bet" by conducting the study by investing resources and stand to gain epistemic value with some probability.
+They have to "place the bet" by conducting the study by investing resources and stand to gain evidence for or against a theory with some probability.
 This conceptualization of choosing a study as a betting problem allows us to apply a "Dutch book" argument [@christensenCleverBookiesCoherent1991].
@@
 Fully developing a Dutch book argument for this problem requires careful consideration of what kind of studies to include as possible bets, defining a conversion rate from the stakes to the reward, and modeling what liberties researchers have in what studies to conduct.
-Without deliberating these concepts further, we find it persuasive that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
+Without deliberating these concepts further, we find it reasonable that researchers should not violate the axioms of probability if they have some expectation about what they stand to gain with some likelihood from conducting a study.
 The axioms of probability are sufficient to derive the Bayes formula, on which we will heavily rely for our further arguments.
-The argument is not sufficient, however, to warrant conceptualizing the kind of epistemic value we reason about in terms of posterior probability; that remains a leap of faith.
-However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974], that is, evidence only increases epistemic value for a theory if the evidence is more likely to be observed under the theory than under the alternative.
+The argument is not sufficient, however, to warrant conceptualizing persuasiveness in terms of posterior probability; that remains a leap of faith.
+In fact, persuasiveness depends on how other researchers weigh evidence which differs between individuals.
+
+However, the argument applies to any reward function that satisfies the "statistical relevancy condition" [@wesleyc.salmonStatisticalExplanation1970; @fetzerStatisticalExplanations1974], that is, evidence only increases believe for a theory if the evidence is more likely to be observed under the theory than under the alternative.
 In particular, "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019], a concept highlighted in recent psychological literature, seems to adhere to the statistical relevancy condition.
 
-# Epistemic value and theoretical risk
+# Theoretical risk
 
-Our first argument is that theoretical risk is crucial for judging evidential support for theories.
+Our first argument is that theoretical risk is crucial for judging the persuasiveness of evidence.
 Put simply, risky predictions create persuasive evidence if they turn out to be correct.
@@
 In short, we want to increase posterior probability $P(H|E)$.
-Increases in posterior probability $P(H|E)$ are associated with increased epistemic value, of which we want to maximize the expectation.
+Increases in posterior probability $P(H|E)$ are associated with increases in persuasiveness, of which we want to maximize the expectation.
 So how can we increase posterior probability?
@@
 When applied strictly within the bounds of null hypothesis testing, detectability is equivalent to power (or the complement of type-II error rate).
-However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is comunicating to other researchers.
+However, while detectability is of great importance for study design, it is not directly relevant to what a preregistration is communicating to other researchers.
 We later discuss how issues of detectability must be considered in a preregistration.
@@
 18].
-We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value (but we of course differ from Popper on the central role of induction in science).
+We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on how "risky" predictions are valued (but we of course differ from Popper on the central role of induction in science).
 
@@
 
-Having discussed that increasing the theoretical risk will increase the epistemic value, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
+Having discussed that increasing the theoretical risk will increase the persuasiveness, it is intuitive to task preregistration with maximizing theoretical risk, i.e., a confirmatory research agenda.
 Indeed, limiting the type-I error rate is commonly stated as  *the* central goal of preregistration [@nosekPreregistrationRevolution2018; @oberauerPreregistrationForkingPath2019; @rubinDoesPreregistrationImprove2020].
@@
 We will later show that preregistration increases the degree of belief in the further specified theory, though it remains low till being substantiated by testing the theory again.
-This is because preregistration increases the expected epistemic value regardless of the theory being tested, as we will show.
+This is because preregistration increases the expected persuasiveness regardless of the theory being tested, as we will show.
 
@@
 As noted before, we strive for high detectability and high theoretical risk in planning, conducting, and analyzing studies.
-Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
+Maximizing one at the expense of the other is not necessarily beneficial for increasing persuasiveness but depends on the specific function they apply to judge evidence and their specific location on the curve.
 One advantage of our framework is that researchers can employ it to balance the trade-off more effectively assuming they are willing to make some simplifiying assumptions.
@@
 Figure \ref{fig:posterior} shows exemplary the effect of theoretical risk (x-axis) on the posterior probability (y-axis) being certain (solid line) or uncertain (dashed line) about the theoretical risk of a study.
-Our expectation of the gained epistemic value varies considerably depending on how uncertain we are about the theoretical risk a study took on.
+Our expectation of the persuasiveness varies considerably depending on how uncertain we are about the theoretical risk a study took on.
 Mathematically, uncertainty about theoretical risk is expressed through the variance (or rather entropy) of the distribution.
-The increase in uncertainty (expressed as more entropic distributions) leads to a decreased expected epistemic value.
+The increase in uncertainty (expressed as more entropic distributions) leads to a decreased expected persuasiveness.
 
-The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected epistemic value, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
+The argument for a confirmatory research agenda is that by increasing theoretical risk we increase expected persuasiveness, i.e., moving to the right on the x-axis in Figure \ref{fig:posterior} increases posterior probability (on the y-axis).
 However, if a hypothesis in a certain study has low theoretical risk, there is not much researchers can do about it.
@@
 A study that has a very high theoretical risk (e.g., `r (1-confirmatory_tau)*100`% chance that if the hypothesis is wrong, evidence in its favor will be observed,) but has also maximum uncertainty will result in a posterior probability of `r round(confirmatory_uncertain*100)`%, while the same study with maximum certainty will result in `r round(confirmatory_certain*100)`% posterior probability.
-The other factors (detectability, prior beliefs, measure of epistemic value) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
+The other factors (detectability, prior beliefs, measure of confirmation) and, therefore, the extent of the benefit varies, of course, with the specifics of the study.
 Crucially, even studies with some exploratory aspects benefit from preregistration, e.g., in this scenario with a $\tau$ = `r some_exploration_tau` (false positive rate of `r 1-some_exploration_tau`) moving from uncertain to certain increases the posterior from `r round(some_exploration_uncertain, 2)` to `r round(some_exploration_certain, 2)`.
@@
 2. They expect other researchers to apply a similar rational process.
-2. Researchers try to maximize the expected epistemic value for other researchers.
+2. Researchers try to maximize the expected persuasiveness for other researchers.
 
@@
 
-To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher expected epistemic value across a variety of measures.
-The former result that increasing theoretical risk leads to higher expected epistemic value reconstructs the appeal and central goal of preregistration of confirmatory research agendas.
+To summarize, we showed that both higher theoretical risk and lower uncertainty about theoretical risk lead to higher persuasiveness across a variety of measures.
+The former result that increasing theoretical risk leads to higher expected persuasiveness reconstructs the appeal and central goal of preregistration of confirmatory research agendas.
 However, theoretical risk is something researchers have only limited control over.
@@
 Second, even under less-than-ideal circumstances for confirmatory research, preregistration is beneficial.
-Preregistering exploratory studies increases the expected epistemic value by virtue of reducing uncertainty about theoretical risk.
-Nevertheless, exploratory studies will have a lower expected epistemic value than a more confirmatory study if both are preregistered and have equal detectability.
+Preregistering exploratory studies increases the expected persuasiveness by virtue of reducing uncertainty about theoretical risk.
+Nevertheless, exploratory studies will have a lower expected persuasiveness than a more confirmatory study if both are preregistered and have equal detectability.
 
@@
 Note that a reader cannot rule out that the researchers leveraged the decision to decrease theoretical risk, i.e., picking among all options the one that delivers the most beneficial results for the theory (in the previous example, chosing between various options of handling missing values).
-Whatever decision they make, increased uncertainty about the theoretical risk is inevitable and the expected epistemic value is decreased compared to a world where they anticipated the need to deal with missing data.
+Whatever decision they make, increased uncertainty about the theoretical risk is inevitable and the expected persuasiveness is decreased compared to a world where they anticipated the need to deal with missing data.
 However, it is still justified to deviate.
@@
 Confirmatory research would almost always be less persuasive without preregistration, but in our view, preregistration primarily communicates the extent of confirmation, i.e., theoretical risk, of a study.
-Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected epistemic value.
+Clearly communicating theoretical risk is important because it reduces the uncertainty and hence increases expected persuasiveness.
 

### 2024-06-27 reviews/methods1.Rmd
move some arguments to a later point

@@
 > We respectfully disagree.
-> What makes the arguments "Bayesian" is what we take probabilities to mean.
+> What makes the arguments "Bayesian" is our interpretation of probabilities.
 > For a Bayesian, they represent degrees of belief.
 > For a frequentist, they represent (relative) frequencies of events that result from a random process.
-This is a subtle but most important distinction.
-> The Bayesian position deals with what agents (here "researchers") ought to believe, assuming they are rational.
-> However, even here, there is no universally agreed upon function that maps evidence onto belief (indeed, there can not be).
-> We tried to make this even clearer and more explicit as to why we include several functions (and that our reasoning applies to all possible functions satisfying our assumptions).
-> The frequentist understanding of probability refers strictly to statistical procedures.
-> From this position, one must either argue that preregistration is changing the underlying random process/ statistical procedure or trace the impact of preregistration via indirect paths outside of the core philosophy of frequentism.
-> Mayo's severity perspective is taking one more step in this direction by asserting that belief in a hypothesis is justified to the extent that it survives stringent scrutiny.
-> However, even this "strong" severity principle is only a proportional statement.
-> Faced with a specific tradeoff, e.g., is a particular post hoc change of a preregistration beneficial, we must revert to vague notions and "gut" instinct under this philosophy.
-> Within the Bayesian framework, we may directly ask: What consequences should preregistration have on the beliefs of individual scientists or the research community as a whole?
+> This is a subtle but most important distinction.
 
@@
 > However, our goal is to say something about how preregistration changes how believable researchers should find claims made in the literature.
+> This necessitates a Bayesian perspective on the process of scientific inquiry (which we adopted) yet it applies to cases of researchers either using frequentist and Bayesian inference in their daily routine.
 
@@
 
-> We find that it adds value by making an otherwise implicit process explicit to the point of calculatable.
+> We find that it adds value by making an otherwise implicit process (that is, the process of updating beliefs in the theories) explicit to the point of making it calculable.
+To our knowledge, there is no other perspective in the literature that makes this part of the process explicit.
 > By conceptualising uncertainty with a specific distribution, we rid ourselves of vague notions of transparency, control, or exploration.
@@
 > Mayo's severe testing returns to a gradual inference (instead of a decision) by evaluating the extent of evidence in favour of H as the extent to which a test was capable of showing that H is false.
+> The Bayesian position deals with what agents (here "researchers") ought to believe, assuming they are rational.
+> However, even here, there is no universally agreed upon function that maps evidence onto belief (indeed, there can not be).
+> We tried to make this even clearer and more explicit as to why we include several functions (and that our reasoning applies to all possible functions satisfying our assumptions).
+> The frequentist understanding of probability refers strictly to statistical procedures.
+> From this position, one must either argue that preregistration is changing the underlying random process/ statistical procedure or trace the impact of preregistration via indirect paths outside of the core philosophy of frequentism.
+> Mayo's severity perspective is taking one more step in this direction by asserting that belief in a hypothesis is justified to the extent that it survives stringent scrutiny.
+> However, even this "strong" severity principle is only a proportional statement.
+> Faced with a specific tradeoff, e.g., is a particular post hoc change of a preregistration beneficial, we must revert to vague notions and "gut" instinct under this philosophy.
+> Within the Bayesian framework, we may directly ask: What consequences should preregistration have on the beliefs of individual scientists or the research community as a whole?
 
@@
 
-> We took considerable time to put what we meant on paper.
 > In the revision, we tried hard to be clearer that uncertainty always refers to the uncertainty of the reader.
@@
 > We claim that there are exploratory methods with a well-known error rate.
-> For example, a neural network has extreme flexibility in fitting the data closely, yet it is calculatable.
+> For example, a neural network has extreme flexibility in fitting the data closely, yet it is calculable.
 
@@
 
-> We explicitly consider the whole space of possible measures. From the revised paper:
+> We explicitly consider the whole space of possible measures.
+> We tried to clarify this.
+> Quoting From the revised paper:
 > *The goal is therefore to reason about the space of possible measures researchers might apply.*
@@
 
-> We included a section called "Hacking, harking, and other harms", which hopefully makes our thinking on these things clearer.
+> We included a section called "Hacking, harking, and other harms", which hopefully clarifies our position.
 
@@
 > If a researcher wants to provide evidence for a null hypothesis, this is their H, theoretical risk, then is related to power.
-> From the revised paper:
+> We tried to clarify this in our revision:
 > *In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.*

### 2024-07-04 reviews.Rmd
change format

---
title          : "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization"
authors        : "Aaron Peikert, Maximilian S. Ernst & Andreas M. Brandmaier"
journal        : "Meta-Psychology"
manuscript-id  : "4139"
manuscript-src : "headings"

cutout-quotes  : true
class          : "draft"
output         : papaja::revision_letter_pdf
---

Dear Dr. Rickard Carlsson,

thank you very much for taking the time to consider our manuscript for publication at _`r rmarkdown::metadata$journal`_.
In the following we address your and each reviewers'/editors' concerns point-by-point.
For Daniel Laken's review we took the liberty to additionally include a high level summary, since we found it, at times, hard to follow the different interwoven discussions.

All changes of the manuscript since this review can be viewed on GitHub:

[https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main](https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main#diff-a9a4aad3fa8c9c10c5404b632bc3a01a25d2d8430eb932bc35c76769963e4b70)

# Daniel Lakens

:::reviewer
Reviewer #1: The authors set out to explain what the primary aim of preregistration is.
I think they are correct that many scientists have hardly thought about this question.
This is not in itself bad - preregistration intuitively makes sense - but it is important to clearly articulate what the aim of preregistration is.
:::

We agree.

:::reviewer
So I applaud the authors for believing this is important.
At the same time, their arguments were, I believe, not always equally convincing, and they do not adequately discuss the existing literature.
:::

We thank the reviewer for pointing out that we failed to include opposing philosophical perspectives.
We are now citing Laken's and Mayo's positions and relate them to our perspective.

:::reviewer
The main argument of the authors is straightforward: preregistration reduces uncertainty about the Type 1 error rate, and under the assumption that it is possible to inflate the Type 1 error rate when a study is not preregistered, reducing the uncertainty through preregistration increases the posterior probability of the hypothesis after a test that yields a significant result.
The authors say they provide a 'Bayesian rationalization'.
This is not correct.
:::

We respectfully disagree.
What makes the arguments "Bayesian" is our interpretation of probabilities.
For a Bayesian, they represent degrees of belief.
For a frequentist, they represent (relative) frequencies of events that result from a random process.
This is a subtle but most important distinction.

:::reviewer
Their rationalization is in essence based on the likelihood ratio.
All else equal, a lower Type 1 error rate means that observing a significant result is relatively more likely when H1 is true.
Of course a higher likelihood ratio also means a higher Bayes factor, which means a higher probability that H1 is true in a Bayesian framework.
:::

It is well known that the math for frequentists is similar or even identical to Bayesian under the right conditions.
Still, a frequentist would not conclude anything about the probability of H1.
However, our goal is to say something about how preregistration changes how believable researchers should find claims made in the literature.
This necessitates a Bayesian perspective on the process of scientific inquiry (which we adopted) yet it applies to cases of researchers either using frequentist and Bayesian inference in their daily routine.

:::reviewer
But the workhorse in the rationalization that the authors provide is the likelihood ratio.
This does not change when we put a distribution on the Type 1 error rate (as the authors do).
:::

Yes, the Bayes formula can be conceptualized as a likelihood ratio scaled by priors.
However, the mere use of the likelihood ratio in the Bayes theorem can not be grounds to reduce Bayesianism to likelihood ratios.

:::reviewer
The distribution (or the move to a Bayesian rationalization) does not add anything (although of course it can be added).
:::

We find that it adds value by making an otherwise implicit process (that is, the process of updating beliefs in the theories) explicit to the point of making it calculable.
To our knowledge, there is no other perspective in the literature that makes this part of the process explicit.
By conceptualising uncertainty with a specific distribution, we rid ourselves of vague notions of transparency, control, or exploration.
This is of utmost importance because only this clarity enables us to effectively navigate the tradeoffs involved.

:::reviewer
Importantly, the likelihood function of course also underlies the uniformly most powerful test in a Neyman-Pearson hypothesis testing framework.
So for a frequentist (and I mean a frequentist not in statistical terms, but also in Popperian anti-Bayesian view on confirmation) the same argument for preregistration plays a role.
:::

We agree that this argument could perhaps be made for a frequentist as well.
However, it is not the math of the likelihood ratio that eludes us; it is how a significant result of a statistical test changes how a frequentist judges a hypothesis.
There is a great, yet unresolved, debate in frequentism about how to do just this.
In a Fisherian approach, no heat is paid to the power of a test; only the p-value is interpreted as the strength of evidence against the null.
In the Neyman-Person paradigm, a decision is to be made about rejecting one or the other hypothesis.
Mayo's severe testing returns to a gradual inference (instead of a decision) by evaluating the extent of evidence in favour of H as the extent to which a test was capable of showing that H is false.
The Bayesian position deals with what agents (here "researchers") ought to believe, assuming they are rational.
However, even here, there is no universally agreed upon function that maps evidence onto belief (indeed, there can not be).
We tried to make this even clearer and more explicit as to why we include several functions (and that our reasoning applies to all possible functions satisfying our assumptions).
The frequentist understanding of probability refers strictly to statistical procedures.
From this position, one must either argue that preregistration is changing the underlying random process/ statistical procedure or trace the impact of preregistration via indirect paths outside of the core philosophy of frequentism.
Mayo's severity perspective is taking one more step in this direction by asserting that belief in a hypothesis is justified to the extent that it survives stringent scrutiny.
However, even this "strong" severity principle is only a proportional statement.
Faced with a specific tradeoff, e.g., is a particular post hoc change of a preregistration beneficial, we must revert to vague notions and "gut" instinct under this philosophy.
Within the Bayesian framework, we may directly ask: What consequences should preregistration have on the beliefs of individual scientists or the research community as a whole?

:::reviewer
The title should not read a Bayesian rationalization, but a likelihood rationalization.
A higher likelihood is not always better, as one can cherry pick a high likelihood ratio.
This is extensively discussed by Mayo (2018).
After p-hacking, we could get a very high likelihood (or Bayes factor) that supports H1 - but it was cherry-picked, and is highly likely to be a fluke.
Therefore, the authors limit themselves to relative likelihoods.
:::

We did not limit ourselves to relative liklihoods.
We explicitly include both the "firmness" and "increase in firmness" perspective.
That a firmness perspective has significant flaws is of no matter to our argument since we intend to show that someone who subscribes to either must be more persuaded by a preregistered study.

:::reviewer
They point out that, assuming people only increase their alpha level when p-hacking, preregistration leads to tests with a lower alpha level.
As such, they basically restate Mayo's (2018) argument about the severity of a test.
:::

Let us cite <https://doi.org/10.3758/s13423-022-02069-1"Nowhere in Mayo’s severity principle or in the specifications discussed below is the riskiness or specificity of a hypothesis, and the predictions it makes, incorporated as a requirement for a severe test (Meehl, 1978; 1986; 1990b, a; Roberts & Pashler, 2000)."

:::reviewer
However, they never make this explicit, and I think this is to the detriment of the paper.
:::

We added a new section that deals with the concept of severity, making it explicit how our approach differs fundamentally.

:::reviewer
The fact that they subsequently add a Bayesian sauce to the likelihood ratio (by adding a distribution to the Type 1 error rate when a study is not preregistered, instead of simply making some assumption about a point-value of the inflated Type 1 error rate) does not add anything to their argument.
So, I believe the authors have basically attempted to re-invent Mayo's severity argument.
:::

The whole second part of the paper deals with the problem of not being able to reduce the theoretical risk to a point estimate.
The statistical procedure has an objective Type 1 error rate, for which we can come up with a point estimate if we agree on the conditions, which is a starting point for evaluating the theoretical risk of the scientific (!= statistical) procedure.
Since, a reader does not have access to the whole scientific procedure, they must judge it with uncertainty.
We find it important to distentengle the statistical procedure, the scientific procedure, and the outside perspective of the reader.

:::reviewer
The fact that this is not immediately clear is mainly due to missing discussions of the existing literature.
Point 1: Preregistration can not objectively minimize uncertainty about the Type 1 error rate The main argument of the authors is that risky predictions create persuasive evidence if they turn out to be correct.
Preregistrations are ideal for communicating the theoretical risk researchers took when testing a claim, and communicating this removes uncertainty in the reader about the theoretical risk the authors took.
I think this argument is in the right direction, but it is not fully developed.
I was somewhat disappointed not to see a discussion of my own paper on a conceptual analysis of the value of preregistration, as I had a similar goal as the authors when I wrote my paper, and my arguments are in the same direction, but dare I say, slightly better developed.
:::

We find ourselves in a curios position, that perhaps may be summarised as "doing the right thing, for the wrong reasons", i.e., we agree with Laken's conclusions about preregistration but we disagree how to come to these conclusions.
We take a Bayesian/inductionist view, whereas Lakens takes a frequentist/anti-inductionist view, agreeing on the end result but disagreeing on everything that comes before.
While we find it important that readers are aware, how anti-inductionist deal with preregistration, this paper was not meant as an exercise to compare Bayesian justification for preregistration with its frequentist counterpart.

:::reviewer
Maybe everyone thinks this of their own papers, so forgive me if I am too biased, but I will attempt to point out some weaknesses in the current paper, sometimes by citing my own paper.
I hope the authors see how strongly the ideas are related, and forgive me for citing myself.
:::

Of course, we see the relevance of Laken's work; we met in Eindhoven to discuss an early version of the paper.

:::reviewer
The authors argue that one interpretation of preregistration is "the traditional application of preregistration to research paradigms that focus on confirmation by maximizing the theoretical risk or, equivalently, by limiting type-I error (when dichotomous decisions about theories are an inferential goal).
I would say it is better to argue preregistration aims to *control* the error rate, not maximize the true negative rate.
The authors then propose their own view: "the objective of preregistration is not the maximization of theoretical risk but rather the minimization of uncertainty about the theoretical risk."
I think the authors are in the right direction, but miss some things.
What the authors get right is that preregistration is about communicating something about the test to others.
:::

This is why we deliberately chose to use the term "uncertainty" and not "control".
Reducing uncertainty directly invokes the concept of communication.
Control also seems to imply that it is a property of the procedure, as it seems curious to suggest that control differs between the reader and the original researcher (we argue that the uncertainty is lower for the latter as they have all the information).

:::reviewer
And the authors are correct that preregistration can remove uncertainty if all readers of a preregistration 1) agree on the analyses plan, and 2) there are no deviations of the preregistration.
:::

We make no such claim.
We explicitly address that readers may differ considerably in how highly they judge the theoretical risk and their confirmation function.

> One feature of the Bayesian framework is the strong emphasis on subjective yet rational judgement.
> Therefore, we assume that researchers will differ significantly in how they value evidence but that by making assumptions about the general process, we can make general statements that apply to all these subjective evaluations.

They may, therefore, very well disagree on the analysis.
We also discuss at length under which circumstances a post hoc change to the preregistration is justified (if the increase in confirmation does offset the negative effects of uncertainty).

:::reviewer
However, the authors already say 2 is not always true, as people often need to deviate.
And 1 is also not true.
This is important, because if 1 and 2 are true, then we can not minimize uncertainty about theoretical risk (as the authors argue the goal of preregistration is), because this uncertainty is based on a subjective evaluation.
If you deviate from an analysis plan because you think after looking at the data that you should have excluded some outliers, but this was not specified in the preregistration, as a reader I am left with an evaluation of whether 1) this analysis is a *better* test, and it reduces the possibility of incorrectly claiming there was an effect, because these outliers might make a null effect significant, while they should have been removed, or 2) this analysis is a *worse* test, and the reader believes the choice for this test without the outliers is driven by the nicer results the researcher observed.
The preregistration does not minimize uncertainty about the Type 1 error rate.
For some readers, the deviation increased the Type 1 error rate, and for others it did not.
:::

This aligns with our reasoning: Any change should increase uncertainty about the theoretical risk.
How much and how this ultimately changes the persuasiveness of the evidence will differ between readers.

:::reviewer
So what *does* preregistration achieve?
In my own words "Preregistration has the goal to allow others to transparently evaluate the capacity of a test to falsify a prediction." The lower the Type 1 error rate, the higher the capacity of a test to falsify a prediction.
But whether a test, especially after a deviation, has a higher or lower Type 1 error rate, remains a subjective decision.
:::

Agreed. It remains a subjective decision.
As subjective yet rational judgement is such a fundamental property of Bayesian reasoning, we may not have emphasized this enough.
We now take up this subject in the discussion:

> We put subjective evaluations at the centre of our considerations; we deal explicitly with researchers who are proponents of some theory (they have higher priors for the theory being true), researchers who suspect confounding variables (they assume lower theoretical risk), or those who remain doubtful if everything relevant was reported (they have higher uncertainty about theoretical risk) or even those who place greater value on incongruent evidence than others (they differ in their confirmation function).

:::reviewer
Preregistration does not give certainty about the theoretical risk, as the authors argue.
It gives more certainty in most cases at best, but it allows others to evaluate it.
This might sound like a subtle difference, but I think it matters.
:::

Agreed.

:::reviewer
The authors might reply that the did not really mean 'minimize uncertainty about the theoretical risk' but mainly meant that readers can judge for themselves what the Type 1 error of a test is (based on their own assumptions about what the most appropriate test is).
In that case, their argument does not differ from mine in Lakens (2019), and the authors should explain what their contribution adds.
:::

In the revision, we tried hard to be clearer that uncertainty always refers to the uncertainty of the reader.
However, being content with claiming that readers have to judge for themselves what the theoretical risk/severity of a test results in a weak argument for preregistration.
We go further and show that for all individual assessments, preregistration has a benefit and that if one is willing to assume some parameters, this benefit is striking.

:::reviewer
Point 2: It remains unconvincing why preregistration is useful for exploratory studies.
The authors seem to argue that even in exploratory studies preregistration will remove *some* uncertainty about the ability of a test to falsify a prediction.
That is true.
However, they do not argue why it would do so to an extent that makes preregistration useful in real life. If an action I do reduces my uncertainty from 98% to 94%, it has relatively reduced my uncertainty.
At the same time, it has done so in a manner that does not matter at all.
My point here is that the error rate in exploratory research is inflated above alpha% to an unknown extent.
A preregistration might reduce the inflation somewhat.
But the rate is quickly too high so that I wouldn't trust the claim anyway.
:::

Actually, we think this is one of the strengths of the paper.
With the mathematical analysis, we show exactly how much uncertainty, theoretical risk, and detectability result in how much confirmation, as seen in Figures 1 and 2.
The vague intuition of a "[Type 1 error] rate [that] is quickly too high so that I wouldn't trust the claim anyway" is replaced by a clear functional form that enables navigating tradeoffs involving exploration, post-hoc changes or detectability.

:::reviewer
If a claim is made with 10% lower Type 1 error rate than it would have been made when it was not preregistered is not useful - because I do not know what the error rate associated with the claim is.
:::

Our claim is that uncertainty is reduced, not theoretical risk directly.
We claim that there are exploratory methods with a well-known error rate.
For example, a neural network has extreme flexibility in fitting the data closely, yet it is calculable.

:::reviewer
Thus, preregistration in exploratory studies reduces the error rate inflation, but it does not make the error rate known.
Therefore, I do not believe preregistration in exploratory studies is useful, and the authors do not make an argument for it.
We do not want to reduce the error rate inflation in preregistration - we want to control it.
:::

Somehow, our argument seems to be flipped here.
An exploratory study has low theoretical risk.
A preregistered study has low uncertainty about theoretical risk.
A preregistered exploratory study is more persuasive than a study with the same exploration going on that is not preregistered.

:::reviewer
If the error rate is not controlled, I do not know how severely a claim has been tested, and I can not use the claim to decide how to act as a scientist.
The authors seem happy with the fact that preregistration reduces the Type 1 error rate even in an exploratory study.
:::

No, we are happy with low uncertainty about theoretical risk.
An exploratory study is in our framework, signified by low theoretical risk (high Type 1 error rate).

:::reviewer
But I am not. Only claims with a known error rate are useful in science.
:::

Exactly. The error rate must be known (readers must be certain about them).
It may be high however (=exploratory).

:::reviewer
If the authors disagree, they should say how an unknown relative decrease in the Type 1 error rate to a final still unknown type 1 error rate is beneficial.
:::

We have a hard time following this reasoning.
Exploratory does not imply that theoretical risk is unknown.

:::reviewer
Point 3: The Type 2 error rate matters as much as the type 1 error rate I do not believe the relationship the authors draw between 'theoretical risk', 'risky predictions', 'evidence', and the 'probability of a hypothesis' is coherent.
I especially disagree with the authors that the posterior probability P(H\|E) is important in preregistration.
The authors say this probability is 'often used directly or indirectly as a measure of confirmation of a hypothesis'.
This is simply not true.
No one discusses posterior probabilities of hypotheses in psychological science.
:::

Neither is severity as an absolute measure of corroboration a common consideration in empirical psychological work.
However, the posterior probability is certainly the most classic measure of absolute confirmation in the philosophy of science.

:::reviewer
The authors then discuss an 'increase in firmness'.
Logically, this would be the Bayes factor, not the posterior probability of the hypothesis, that the authors are talking about.
The Bayes factor is a continuous measure of how much we update our belief.
If we want to argue for any function of preregistration from a Bayesian perspective, it must be focused on the Bayes factor.
The authors then say 'If a hypothesis is more probable to begin with, observing evidence in its favor will result in a hypothesis that is more strongly confirmed, all else being equal'.
:::

We explicitly consider the whole space of possible measures.
We tried to clarify this.
Quoting from the revised paper:

> The goal is therefore to reason about the space of possible measures researchers might apply.
> However, since any measure fulfilling the statistical relevancy condition increases monotonically with an increase in posterior probability $P(H|E)$, we might well take it to illustrate our reasoning.
> Later in the paper: Of course, the assigned probabilities and the distribution $Q$ vary from study to study and researcher to researcher (and even the measure of confirmation), but we can illustrate the effect of uncertainty with an example.

:::reviewer
This statement is not in line with the earlier statement about an 'increase in firmness'.
Of course, if a hypothesis is highly probable (the Stroop effect), any evidence will end up with a higher posterior probability that if the prior probability was low.
But the same evidence would lead to a greater 'increase in firmness' (or a higher Bayes factor) if the prior was lower.
The same data would lead to a much greater update in belief.
The authors dismiss the prior probability of the hypothesis as it 'is nothing our study design can change'.
However, this is not true.
Although the authors only discuss inflating the alpha level in their paper, one function of preregistration is to prevent HARKing, or hypothesizing after the results are true.
Of course, if one comes up with a hypothesis after looking at the data, the 'prior' is 1: We already know the hypothesis will be supported by the data.
:::

We included a section called "Hacking, harking, and other harms", which hopefully clarifies our position.

:::reviewer
HARKing is a major reason the field has moved towards preregistration, but the term 'HARKing' is not even mentioned in the article.
The reason for this is that the arguments the authors provide can not deal with HARKing.
This is obviously a limitation, compared to existing arguments for preregistration, and it is incorrectly ignored in the article.
As Mayo (2018) explains, the problem with HARKing is that it leads to a non-severe test:
The data have no probability of falsifying the claim.
This error-statistical perspective seems to me to be much more fruitful than a Bayesian approach.
:::

As stated above, we come to the same conclusions despite being inductionists.

:::reviewer
The authors then move on the P(E\|H), or power.
They argue 'while detectability is of great importance for study design, it is not directly relevant to the object of preregistration.' This is incorrect.
As I write in Lakens, 2019: "In line with the general tendency to weigh Type 1 error rates (the probability of obtaining a statistically significant result when there is no true effect) as more serious than Type 2 error rates (the probability of obtaining a non-significant result when there is a true effect), publications that discuss preregistration have often been more concerned with inflated Type 1 error rates than with low power.
However, in situations where researchers want to find a null effect low power is a bigger concern."
The idea of null-hacking is not crazy - sometimes a certain variable should *not* be significant to exclude alternative explanations. Therefore, preregistration certainly should have the function to prevent researchers from incorrectly claiming there is no effect, when there is an effect, by inflating the Type 2 error rate (or equivalently, by reducing the power).
:::

Absolutely. However, note that our H is always the research hypothesis, independent of the statistical H1 or H0.
If a researcher wants to provide evidence for a null hypothesis, this is their H, theoretical risk, then is related to power.
We tried to clarify this in our revision:

> In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.

We reference the ideas of frequentist null hypothesis testing to provide readers who are more familiar with this line of reasoning with a better intuition, not to claim equivalence.

:::reviewer
This aspect of Bayes' formula is therefore also incorrectly ignored in the article.
The authors then argue that P(E) is the only part of the formula that remains to be considered.
Decreasing P(E) can increase the posterior probability.
Of course, P(E) is a function both of the power and the Type 1 error rate.
But for some unclear reason, the authors here dismiss power, as they write "We have already noted that there is not much to be done about prior probability ((), and hence its counter probability (¬)), and that it is common sense to increase detectability (\|)."
But this latter statement dismisses the important role of power much too easily.
One could just as well say it is 'common sense' to reduce the Type 1 error rate.
Furthermore, once again, we do not 'increase' error rates, we *control* them.
:::

Perhaps, we are a bit daft but a clear definition of error rate control eludes us.
We are under the impression that two competing definitions are used here from time to time.
First, a definition in which control means the error rates are known (but then known by whom?).
Second, control means ensuring certain acceptable bounds of error rates are not exceeded.

:::reviewer
And while controlling them, we always have to consider resource constraints (Lakens, 2022).
Indeed, one might even say that we are always balancing Type 1 and Type 2 error rates, and that it would be sensible to move away from a fixed alpha level for all studies (Maier and Lakens, 2022).
In any case, the authors fail to adequately argue why power can be ignored, and the Type 1 error rate is important.
:::

We can not deliver a treaty on statistical inference as a whole.
Nowhere have we stated that power can be ignored for statistical inference; we just do not think that power is the main reason why preregistration is a good thing.
We agree that detectability is important; after all, it shows in every mathematical consideration we made that it is a vital consideration for a Bayesian.

:::reviewer
According to the error-statistical justification of preregistration (Mayo, 2018; Lakens, 2019) both are equally important.
At the very least, increasing the alpha level would also increase the statistical power, so the two can not even be seen as separate. 
:::

We wholeheartedly agree with the importance of being able to navigate these tradeoffs.
That is why we find a mathematical treatment of these ideas to be so important.
The Bayesian perspective allows us to ask, very concretely, how persuasive an idealized Bayesian thinker finds different configurations, including the ratio of theoretical risk and detectability.
We clearly say multiple times that these tradeoffs can not be optimized separately.
In the revised paper:

> As noted before, we strive for high detectability and high theoretical risk in planning, conducting, and analyzing studies.
> Maximizing one at the expense of the other is not necessarily beneficial for increasing epistemic value but depends on the specific function they apply to judge evidence and their specific location on the curve.
> One advantage of our framework is that researchers can employ it to balance the trade-off more effectively, assuming they are willing to make some simplifying assumptions.

:::reviewer
POINT 4: Terminology As terminology can always be fixed, this point is not as problematic to the reasoning as the previous 3.
The authors use the word 'theoretical risk' for the probability that a test of a hypothesis yields a true negative.
As the authors exclusively talk about the probability that statistical tests will yield a correct conclusion, 'theoretical risk' feels as a misnomer (statistical risk would be better, but severity might also be a good general term).
:::

We talk about the scientific procedures, of which a statistical test is a piece.

:::reviewer
True, in page 13 they give an example about theoretical risk that is actually theoretical risk - but this example has nothing to do with the remainder of the paper, as all three studies can have exactly the same alpha level!
:::

We are unable to follow this argument without more elaboration.

:::reviewer
They are talking about a low frequentist error rate in all their examples, and as 'theoretical risk' is the inverse of the false positive rate, it might be easier to just talk about the Type 1 error rate.
:::

The point of not using a Type 1 error rate is to differentiate the statistical test from the the researcher's scientific judgement.
Type 1 error rate is a property of a test under certain conditions.
Theoretical risk refers to the individual judgement of a researcher about the substantive hypothesis and the study design as a whole.

:::reviewer
The authors say they borrowed the term 'theoretical risk' from Meehl, but that he did not assign it to (¬\|¬).
The authors cite Kukla (1990), but after reading it, I am not convinced they are correct in equating 'theoretical risk' with the Type 1 error rate.
Theoretical risk means making an a-priori unlikely prediction if the theory was not true.
:::

This sentence sounds exactly like P(¬E|¬H) to us.
The judgement about P(¬E|¬H) is a-priori, is about "prediction" as a statement about evidence E, given that the theory is not true (¬H).
One might argue, that it should be multiplied by ¬H perhaps? 
But that would not change any of the arguments.
In any case, we provide clear mathematical definitions for our terms.
If a reader does not agree with our terminology, they are free to term them differently.
In fact, this happens in reference to a finding we found in the meantime (Oberauer and Lewandowsky (2019)), which we added to the discussion.

:::reviewer
This could be, as Meehl would like, a range prediction. Or, in the example of the authors on page 13, it is a test without confounds.
In Lakens, 2019, on page 224, I explain how a two-sided test, one-sided test, and a range prediction all differ in how severe they test the hypothesis, even if they all have the same 5% alpha level ("Note that the three tests differ in severity even when they are tested with the same Type 1 error rate.").
:::

Yes, indeed, this could all have low theoretical risk.
We could include the disclaimer to the examples on p. 13 "Note that the three tests differ in theoretical risk even when they are tested with the same Type 1 error rate."

:::reviewer
This is another reason why I think the term 'theoretical risk' is not a good choice.
Indeed, based on the example on page 13, it seems the authors themselves confuse real theoretical risk with Type 1 error control.
However, it is better to use terminology that encompasses Type 1 and Type 2 errors (such as 'the probability that a test will falsify a hypothesis', or even better: the *severity* of a test).
:::

Different philosophies differ in their terminology.
Deborah Mayo has good reasons to differentiate Type 1 error rate from severity, as do we.

:::reviewer
Using the term 'severity of the test' would also embed the work more strongly in philosophy of science (and specifically the work of Mayo, as she builds on Popper).
:::

This would embed us in a different philosophy of science.
Popper is the most famous anti-inductionist; we deliver an inductionist account.

:::reviewer
Mayo's work has been exactly on this topic, and it is a shame it has not been discussed, even though she discusses the problem of p-hacking and Type 1 error rate inflation extensively in her 2018 book.
The authors write "The amount of knowledge researchers gain from a particular study concerning the verisimilitude of a specific theory is what we call epistemic value".
This is an unfortunate term, as epistemic values are already used in philosophy of science to refer to qualities such as predictive accuracy, internal coherence, external consistency, unifying power, and fertility - see Elliott, K. C. (2022). Values in Science (1st ed.). Cambridge University Press. <https://doi.org/10.1017/9781009052597>.
I would pick another terminology.
In our own work published in Psychological Methods we called this "the marginal gain in expected utility (or usefulness) of scientific claims". (Isager, P. M., van Aert, R. C. M., Bahník, Š., Brandt, M. J., DeSoto, K. A., Giner-Sorolla, R., Krueger, J. I., Perugini, M., Ropovik, I., van 't Veer, A. E., Vranka, M., & Lakens, D. (2023).
Deciding what to replicate: A decision model for replication study selection under resource and knowledge constraints. Psychological Methods, 28(2), 438-451. [https://doi.org/10.1037/met0000438)/](https://doi.org/10.1037/met0000438)/){.uri}.
There are probably other existing terms for the same idea in the literature. The authors discuss P(E\|H) as 'detectability' or the probability of detecting evidence if a hypothesis holds.
This is thus simply statistical power, or the true positive rate.
I think the term detectability is not particularly useful, as it is more strongly associated with signal detection theory - but that is not what the authors mean with the term, as they admit the term is simply statistical power when applied to hypothesis testing (and I don't know what else it could be applied to).
:::

Hypothesis testing and scientific inference are different things.
A test might be significant, but if a researcher suspects strongly enough that there might be a problem, they may decide not to believe the hypothesis.

:::reviewer
Minor points: The authors say "We note these relations to underline that the Bayesian rationale we apply here is able to reconstruct many commonly held views on riskiness and epistemic value." when they relate a property of their Bayesian model with Popper's work.
:::

No, we relate our work to beliefs we think are commonly held.
Researchers find the idea of a risky test appealing.
Popper gives one explanation for this appeal, but we happen to have a different view.

:::reviewer
However, this is misleading.
Popper would never agree with the approach in this paper, as he does not think subjective prior or posterior beliefs should play a role in scientific claims.
According to Popper, verisimilitude is not about belief. Popper, p. 434: "As to degree of corroboration, it is nothing but a measure of the degree to which a hypothesis h has been tested, and of the degree to which it has stood up to tests.
It must not be interpreted, therefore, as a degree of the rationality of our belief in the truth of h".
The author's proposal violates the philosophical work of Popper at it's very core, and is incommensurable.
:::

As stated above, if Popper would happen to agree, we indeed would have to rethink our arguments.

:::reviewer
Just because the authors also need to rely on the likelihood ratio, does not mean it 'reconstructs' Popperian philosophy of science.
:::

Which we never claimed.

:::reviewer
It is not clear what the researchers mean with 'Researchers judge the evidence for or against a hypothesis rationally'. Researchers often differ in whether a study is evidence for or against a hypothesis, often because they disagree on auxiliary assumptions (e.g., A thinks X is a good measure of Y, but B does not think X measures Y).
Are these two researchers irrational? Are the authors talking about some utopian world, in which case this assumption seems violated at least for now.
:::

No, these researchers are not irrational according to our definition.
For a researcher to be rational, they have to be internally consistent.
They must not change their process of weighing evidence just because they do not like the outcome.
We take this assumption from the literature of the Bayesian philosophy of science and cite it appropriately.

:::reviewer´
The authors admit they are talking about an 'ideal' estimation process.
:::

We meant "ideal" similar to how a physicist would use the term in the context of "ideal gas".

:::reviewer
But in an ideal estimation process we would also have infinite data.
Maybe I am being too pragmatic, but shouldn't we discuss how research is actually done?
:::

We agree to assume infinite data would go a little too far into wishful thinking.
Also, with infinite data, it stops being fun to apply Bayesian reasoning since there is no uncertainty to deal with.

:::reviewer
The authors write "Indeed, limiting the type-I error rate is commonly stated as the central goal of preregistration (Nosek et al., 2018; Oberauer, 2019; Rubin, 2020)."
It might be commonly stated. But then, people also commonly state the incorrect definition of a p-value.
Instead, the authors could cite a better paper, like my own: "The goal of preregistration is not simply to control the Type 1 error rate in hypothesis tests, but to prevent researchers from non-transparently reducing the capacity of the test to falsify a prediction in general."
I worked quite hard on conceptually clarifying the role of preregistration, correcting exactly the kind of misconception the authors also respond against.
:::

We appreciate this hard work; we just do not think it is quite done yet.

:::reviewer
The authors write "First, according to most interpretations of null hypothesis testing, the absence of a significant result should not generally be interpreted as evidence against the hypothesis (Mayo, 2018, p. 5.3)."
This is generally true, but of course null effects often *are* evidence for the null - as would be clear if an equivalence test was performed, or a Bayes factor or likelihood ratio was computed.
:::

With the following statement, we hoped to make clear that there are many statistical procedures that can be investigated through the lens we adopted here:

> Please note that our decision to adopt a Bayesian philosophy of science does not make assumptions about the statistical methods researchers use.

:::reviewer
The authors argue (page 15) theoretical risk also encompasses factors outside the statistical realm.
I agree, and say so in Lakens, 2019: "Making very narrow range predictions is a way to make it statistically likely to falsify your prediction if it is wrong.
But the severity of a test is determined by all characteristics of a study that increases the capability of a prediction to be wrong, if it is wrong.
If a researcher predicts that an effect will only be observed under a very specific set of experimental conditions that all follow from a single theory, it is possible to make theoretically risky predictions."
Line 425: There are 3 arguments, not 2.
:::

:::reviewer
\label{pt:harking}
Conclusion: Again, I appreciate the authors digging in to the justification of preregistration, but I think there are some important things missing in this version of the manuscript.
How can the article be improved? The authors should address harking.
:::

We included a section called "Hacking, harking, and other harms", which hopefully makes our thinking on these thinkgs clearer.

:::reviewer
\label{pt:exploratory}
They should explain better why they think preregistration of exploratory studies is useful.
:::

We hope that the improvements that we made to better explain the topics of subjectivity and uncertainty make clearer that theoretical risk and uncertainty about theoretical risk are two dimensions.

:::reviewer
\label{pt:terminology}
They should improve terminology.
:::

We take the criticism of the terminology seriously.
Although we have opted to stick with most of the terms as we defined them, we provided more references.
It might also be the case that the differences in philosophical traditions lead to some criticism.

:::reviewer
\label{pt:severity}
The should better integrate their ideas with existing ideas, especially Mayo's work on severity, but at least engage with the arguments for preregistration in the literature more than a somewhat strawman account of minimizing the type 1 error rate.
:::

We have discussed Mayo's Severity to show that these approaches fundamentally differ, and we have significantly extended the discussion to include literature that also uses a Bayesian rationale to investigate what methods researchers should employ.
Additionally, we substantiated the "strawman" further because we believe that the minimizing vs communicating researchers' degrees of freedom confusion is at the centre of many debates around preregistration.

:::reviewer
\label{pt:nullhacking}
They should also address 'null-hacking'.
:::

We have merely mentioned this topic in passing since it is conceptually only difficult if the null hypothesis has a special position, as in frequentist reasoning.
However, we significantly expanded on the topic of "hacking" in general, under which this topic falls for us.

:::reviewer
After these changes, I am not sure what their argument will end up looking like, but I would hope the authors explore their ideas further.
Sorry for citing myself, no need to cite me - others often make the same points I make, so pointing you to my own papers is mainly just easier for myself. Lakens, D. (2019). The value of preregistration for psychological science: A conceptual analysis. Japanese Psychological Review, 62(3), 221-230. <https://doi.org/10.24602/sjpr.62.3_221Lakens, D. (2022). Sample Size Justification. Collabra: Psychology, 8(1), 33267. <https://doi.org/10.1525/collabra.33267Maier, M., & Lakens, D. (2022). Justify Your Alpha: A Primer on Two Practical Approaches. Advances in Methods and Practices in Psychological Science, 5(2), 25152459221080396. <https://doi.org/10.1177/25152459221080396>
Signed,
Daniel Lakens
:::

We thank Daniel Lakens for the review and for allowing us to share it!

### 2024-08-22 reviews.Rmd
add summary

@@
 # Daniel Lakens
+
+Lakens and our own positions on preregistration built upon two long standing traditions in philosophy of science, namely inductionist and anti-inductionist, that differ fundamentally in their core assumptions.
+
+Integration with Existing Literature:
+
+* Reviewer Concern: The manuscript needs better engagement with existing ideas, particularly those from Mayo’s work on severity.
+* Our Response: We expanded our discussion to include more literature that employs a Bayesian rationale and highlighted how our approach differs fundamentally from the concept of severity, situating our work within the broader philosophical landscape that is not limited to (Neo-)Popperian approaches.
+
+Bayesian vs. Frequentist Interpretation:
+
+* Reviewer Concern: The reviewer challenges our Bayesian approach, suggesting it merely re-frames a frequentist argument using the likelihood ratio.
+* Our Response: We clarified that our Bayesian approach, while indeed utilizing the likelihood ratio, offers a more explicit and calculable framework for updating beliefs based on evidence---a hallmark of Bayesian reasoning.
+
+Objective vs. Subjective Judgement:
+
+* Reviewer Concern: The reviewer is skeptical about the extent to which preregistration can objectively minimize uncertainty, particularly in exploratory studies, suggesting that our focus on subjective evaluation might downplay the need for more objective control of error rates.
+* Our Response: We argued that preregistration reduces uncertainty about theoretical risk, which is inherently a subjective evaluation made by the reader.
+Bayesian reasoning accommodates the subjectivity in scientific judgment, allowing for rational, individual interpretations of evidence.
+We acknowledged the reviewer’s concerns but maintained that the subjective aspect of Bayesian analysis is a strength, not a weakness.
+
+Control vs Transparency:
+
+* Reviewer Concern: The reviewer agrees that preregistration enhances transparency but believes we overstate its ability to minimize uncertainty, suggesting that error control should be the primary focus.
+* Our response: We contend that error control, as traditionally understood in the frequentist framework, assumes fully known random processes, making it difficult to analyse situations with incomplete or absent preregistration.
+As a consequence, a severe tester must differentiate error control and transparency, the former supported by formal reasoning, and the latter a necessary addition outside of the core philosophy.
+
+HARKing and Error-Statistical Perspective:
+
+* Reviewer Concern: The manuscript lacked a discussion on HARKing, a practice detrimental to the integrity of hypothesis testing, and the error-statistical perspective emphasizing the importance of severe tests.
+* Our Response: We acknowledged this oversight and added a dedicated section titled “Hacking, harking, and other harms” to address how our Bayesian framework addresses these concerns.
+
+Type I and Type II error rates:
+
+* Reviewer Concern: The manuscript was perceived as downplaying the importance of statistical power (P(E|H)) in favor of focusing on Type 1 error rates.
+* Our Response: The Bayesian rational considers theoretical risk and detectability simultaneously a priori.
+From a severity perspective, either Type I or Type II, never both, is considered post hoc, depending on the test results and if the research hypothesis is H1 or H2.
+
+Terminology and Conceptual Clarity:
+* Reviewer Concern: The terminology used, particularly “theoretical risk,” was seen as potentially confusing and misaligned with existing literature, such as the concept of severity.
+* Our Response: We provided additional references and explanations to support our terminology.
+We believe our terms accurately reflect the concepts within our Bayesian framework, but we differentiated our terminology further from frequentist reasoning.
 
@@
 
-:::reviewer´
+:::reviewer
 The authors admit they are talking about an 'ideal' estimation process.

### 2024-08-29 reviews.Rmd
point out our strenghts

@@
 * Our response: We contend that error control, as traditionally understood in the frequentist framework, assumes fully known random processes, making it difficult to analyse situations with incomplete or absent preregistration.
-As a consequence, a severe tester must differentiate error control and transparency, the former supported by formal reasoning, and the latter a necessary addition outside of the core philosophy.
+As a consequence, a severe tester must differentiate error control and transparency, the former supported by formal reasoning, and the latter a necessary addition outside of the core philosophy of severity.
+Not so in our Bayesian framework, where transparency maps cleanly onto uncertainty which can be mathematically modeled.
 

### 2024-09-02 manuscript.Rmd
add definition of severity and another critizism

--- 2024-06-03 manuscript.Rmd
@@
 Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
+Severity, is a Neopopperian view which asserts that there is evidence for a hypothesis just to the extent that it survives stringent scrutiny.
 However, there are crucial differences between the two.
@@
 Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
+This makes it difficult to guide a priori decisions in planning a study, after all severity disregards power, if we observe evidence, and disregards Type I error rate when we do not.
+This implies that for a priori balancing Type I and Type II error rate, a researcher must assign a priori probabilities to, for example, the size of an effect.
+Since such a move not in line with frequentist rationale there is no guidelines available on how to do this.
+Third, we would argue that severity considerations assume full information about how the evidence came about and hence imply axiomatically the need for perfect preregistration.
+This comes down to frequentist understanding of probability as the outcome of a well defined random experiments.
+When judging a particular study, a frequentist, and hence a severe tester, may not assign probability to the event that the researchers did, for example, p-hack.
+The lack of knowledge on the readers side does not turn the p hacking into a random event of which we can calculate the long run frequency aka frequentist probability.
+A severe test, hence, must assume that they know the Type I and Type II error rate precisely.
+Full transparency, is hence assumed, and we can not imagine many ways except preregistration that get close to this ideal.
+This assumptions also makes it difficult to deal with less than perfect preregistrations and post-hoc changes without appealing to principles outside the core philosophy of severity.
+
 However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

### 2024-09-02 reviews/PsychBulletinEditor.Rmd

@@
 I think you would agree that your examples are highly schematic and serve to illustrate the principles that you are proposing as opposed to reporting specific examples of how these principles can improve psychological research.
-
-> We fully agree.
-
 Perhaps it would be possible to compare a range of studies with and without preregistrations to show how the act of preregistration helps clarify the reported results even when the preregistration does not report the level of detail required for a truly confirmatory study.
-
-> We are sure this is possible and would be enlightening but we would not be sure how to relate the results to the paper.
-> The issue is, that how close a theory is to truth, is very difficult to measure.
-> That implies that measuring if our approach is successfull is similarly difficult to measure empierically.
-> Our goal is more modest: Showing that if you subscribe to a certain class of philosphies of science, you should find preregistered evidence more persuasive.
-
 Given your broad scope, the reviewed studies would ideally come from a variety of research domains.
@@
 
-> Applying preregistration with "all the messy details of specific research endeavors" arguably already leads to quite some confusion.
-> We hope that some conceptual clarity helps researcher in practice.
-> This hope is grounded in Kurt Lewins Maxim: "Nothing as practical as a good theory".
+> We are sure this would be possible and enlightening, but also way beyond the scope of the current paper.
+> Out goal is more modest: we want to provide a bayesian justification of pre-registration that 1.
+clearly articulates the aims of a preregistration and 2.
+describes how preregistration achieves these goals.
+> As a result, it eliminates common missconceptions about preregistrations, for example, that you are not allowed to deviate under any circumstances.
+> To achieve these goals, our examples are purposely schematic, as we want them to apply to all empirical sciences.
 

### 2024-09-02 reviews.Rmd

@@
 This is a subtle but most important distinction.
+A Bayesian may assign a probability to a hypothetical world where a paper was p hacked, since probabilities represent their subjective believes.
+A frequentist (and hence severe tester), can not possibly assign a probability here because there is no random process but a single event that happened in the past, so the paper was either p hacked or not.
+This is similar to the mistake, in frequentist philosophy, to say that an effect is with 95% probability in the confidence interval.
 
@@
 
-Yes, the Bayes formula can be conceptualized as a likelihood ratio scaled by priors.
 However, the mere use of the likelihood ratio in the Bayes theorem can not be grounds to reduce Bayesianism to likelihood ratios.
@@
 
-Let us cite <https://doi.org/10.3758/s13423-022-02069-1"Nowhere in Mayo’s severity principle or in the specifications discussed below is the riskiness or specificity of a hypothesis, and the predictions it makes, incorporated as a requirement for a severe test (Meehl, 1978; 1986; 1990b, a; Roberts & Pashler, 2000)."
+Let us cite <https://doi.org/10.3758/s13423-022-02069-1> "Nowhere in Mayo’s severity principle or in the specifications discussed below is the riskiness or specificity of a hypothesis, and the predictions it makes, incorporated as a requirement for a severe test (Meehl, 1978; 1986; 1990b, a; Roberts & Pashler, 2000)."
 
@@
 We claim that there are exploratory methods with a well-known error rate.
-For example, a neural network has extreme flexibility in fitting the data closely, yet it is calculable.
+In our argument, the difference between a t-test and a random forest is a matter of degree and not of kind.
+A t-test may maintain its 5%-alpha error and a random forest may find a variable important 30% of the time.
+For both, however, their behavior is mathematically analyzable, yielding a concrete expectation for their theoretical risk.
 
@@
 A preregistered study has low uncertainty about theoretical risk.
-A preregistered exploratory study is more persuasive than a study with the same exploration going on that is not preregistered.
+A preregistered exploratory study is more persuasive than a study with the same degree of exploration that is not preregistered.
 
@@
 
-Yes, indeed, this could all have low theoretical risk.
+Yes, indeed, these could all have low theoretical risk.
 We could include the disclaimer to the examples on p. 13 "Note that the three tests differ in theoretical risk even when they are tested with the same Type 1 error rate."
@@
 We have discussed Mayo's Severity to show that these approaches fundamentally differ, and we have significantly extended the discussion to include literature that also uses a Bayesian rationale to investigate what methods researchers should employ.
-Additionally, we substantiated the "strawman" further because we believe that the minimizing vs communicating researchers' degrees of freedom confusion is at the centre of many debates around preregistration.
+Additionally, we substantiated the "strawman" further because we believe that the minimizing vs communicating researchers' degrees of freedom confusion is at the center of many debates around preregistration.
 

### 2024-09-02 reviews.Rmd
merge both reviews in one file

@@
 We thank Daniel Lakens for the review and for allowing us to share it!
+
+# James Brockmole
+
+:::reviewer
+Dear Mr. Peikert:
+
+I have now read Manuscript PBR-TR-23-119 entitled "Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization" that you submitted to Psychonomic Bulletin & Review.
+You offer a new perspective on an important issue, and I truly enjoyed reading through your arguments for the role of preregistration in reducing uncertainty about theoretical risk.
+:::
+
+We are thankful for the kind words.
+
+:::reviewer
+That said, this is more of a think piece than a theoretical review, and the value of assessing theoretical risk remains hypothetical.
+I think you would agree that your examples are highly schematic and serve to illustrate the principles that you are proposing as opposed to reporting specific examples of how these principles can improve psychological research.
+Perhaps it would be possible to compare a range of studies with and without preregistrations to show how the act of preregistration helps clarify the reported results even when the preregistration does not report the level of detail required for a truly confirmatory study.
+Given your broad scope, the reviewed studies would ideally come from a variety of research domains.
+In other words, perhaps it is possible to put your ideas through a few test runs to see how they can be applied when all the messy details of specific research endeavors are considered.
+:::
+
+We are sure this would be possible and enlightening, but also way beyond the scope of the current paper.
+Out goal is more modest: we want to provide a bayesian justification of pre-registration that 1.
+clearly articulates the aims of a preregistration and 2.
+describes how preregistration achieves these goals.
+As a result, it eliminates common missconceptions about preregistrations, for example, that you are not allowed to deviate under any circumstances.
+To achieve these goals, our examples are purposely schematic, as we want them to apply to all empirical sciences.
+
+:::reviewer
+In my view, theoretical reviews propose a new set of ideas *and* systematically apply the new ideas to existing findings to establish new discoveries or bolster conclusions with more rigorous methods.
+I don’t see the second element here, and I don’t see how getting reviewer comments would change that impression.
+:::
+
+We agree. Empirically establishing our perspective on preregistration as "correct" would be very difficult.
+
+:::reviewer
+ACTION: I am rejecting the manuscript.
+Although a desk reject seems like a very negative outcome, I want to again emphasize that I found your analysis creative and valuable.
+I just don’t think a PBR theoretical review is the appropriate outlet.
+
+Thank you for considering Psychonomic Bulletin & Review for the publication of your research.
+
+Sincerely, Jeff Starns Action Editor, Psychonomic Bulletin & Review
+
+Editor: James Brockmole
+:::

### 2024-09-02 manuscript.Rmd
small language edits

@@
 In case 4, the theoretical risk is neither zero (unless the researcher plans to do run variations of analyses until a favourable outcome is obtained, then we have a particular instance case of 1) nor high (as this is the nature of exploratory approaches).
-However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk and hence are in a somewhat favourable position (i.e., close to the solid line in n Figure \ref{fig:posterior}/\ref{fig:other-measures}).
+However, we can take advantage of computational reproducibility, use statistical properties, simulation or resampling methods, together with scientific reasoning, to get a reasonably certain evaluation of the theoretical risk.
+Low uncertainty about high theoretical risk is a somewhat favourable position (i.e., close to the solid line in Figure \ref{fig:posterior}/\ref{fig:other-measures}).
 This favorable position leads us to recommend preregistration of exploratory studies.
-Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard did researcher try to come up with favourable results).
+Case 5 shares the neither zero nor high theoretical risk of case 4 but has additional uncertainty about how much exploration was going on (how hard exactly did the researchers try to come up with favourable results).
 Its low *and uncertain* theoretical risk make it difficult to produce compelling evidence.

### 2024-09-04 reviews.Rmd
accept some of maximilians and andreas changes

@@
 
-It is well known that the math for frequentists is similar or even identical to Bayesian under the right conditions.
+Indeed, the math for frequentist inference is identical to Bayesian under the right conditions.
 Still, a frequentist would not conclude anything about the probability of H1.
@@
 We explicitly include both the "firmness" and "increase in firmness" perspective.
-That a firmness perspective has significant flaws is of no matter to our argument since we intend to show that someone who subscribes to either must be more persuaded by a preregistered study.
+Since we showed that our justification of preregistration holds for both perspectives, the discussion about specific flaws of the firmness perspective does not concern our argument.
 
@@
 
+> Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
+Severity, is a Neopopperian view which asserts that there is evidence for a hypothesis just to the extent that it survives stringent scrutiny.
+However, there are crucial differences between the two.
+First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
+This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
+Second, severity considerations are mainly after the fact.
+Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
+This makes it difficult to guide a priori decisions in planning a study, after all severity disregards power, if we observe evidence, and disregards Type I error rate when we do not.
+This implies that for a priori balancing Type I and Type II error rate, a researcher must assign a priori probabilities to, for example, the size of an effect.
+Since such a move not in line with frequentist rationale there is no guidelines available on how to do this.
+Third, we would argue that severity considerations assume full information about how the evidence came about and hence imply axiomatically the need for perfect preregistration.
+This comes down to frequentist understanding of probability as the outcome of a well defined random experiments.
+When judging a particular study, a frequentist, and hence a severe tester, may not assign probability to the event that the researchers did, for example, p-hack.
+The lack of knowledge on the readers side does not turn the p hacking into a random event of which we can calculate the long run frequency aka frequentist probability.
+A severe test, hence, must assume that they know the Type I and Type II error rate precisely.
+Full transparency, is hence assumed, and we can not imagine many ways except preregistration that get close to this ideal.
+This assumptions also makes it difficult to deal with less than perfect preregistrations and post-hoc changes without appealing to principles outside the core philosophy of severity.
+
+> However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
+
 :::reviewer
@@
 The statistical procedure has an objective Type 1 error rate, for which we can come up with a point estimate if we agree on the conditions, which is a starting point for evaluating the theoretical risk of the scientific (!= statistical) procedure.
-Since, a reader does not have access to the whole scientific procedure, they must judge it with uncertainty.
-We find it important to distentengle the statistical procedure, the scientific procedure, and the outside perspective of the reader.
+Since readers do not have access to the whole scientific procedure, they must judge it with uncertainty.
+We find it important to disentangle the statistical procedure, the scientific procedure, and the outside perspective of the reader.
 
@@
 Of course, we see the relevance of Laken's work; we met in Eindhoven to discuss an early version of the paper.
+With respect to how well developed our respective arguments are, we would like to point out that we think that a key part of our argument is the mathematical formulation that enables research to actually compute tradeoffs they face during the research process instead of juggling with vague notions of unquantifiable philosophical principles.
 
@@
 
-We can not deliver a treaty on statistical inference as a whole.
 Nowhere have we stated that power can be ignored for statistical inference; we just do not think that power is the main reason why preregistration is a good thing.
@@
 
-We appreciate this hard work; we just do not think it is quite done yet.
+We appreciate this hard work.
+Yet we believe that this is not the only useful perspective on the utility of preregistration.
 

### 2024-09-04 reviews.Rmd
fix references

@@
 manuscript-src : "headings"
-
+bibliography   : references.bib
 cutout-quotes  : true
@@
 Terminology and Conceptual Clarity:
+
 * Reviewer Concern: The terminology used, particularly “theoretical risk,” was seen as potentially confusing and misaligned with existing literature, such as the concept of severity.
@@
 
-However, the mere use of the likelihood ratio in the Bayes theorem can not be grounds to reduce Bayesianism to likelihood ratios.
+The mere use of the likelihood ratio in the Bayes theorem can not be grounds to reduce Bayesianism to likelihood ratios.
 

### 2024-09-04 manuscript.Rmd
critize Lakens definition in his recent paper

@@
 figurelist: false
-tablelist : false
+tablelist: false
 footnotelist: false
@@
 
-This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward [@lakensWhenHowDeviate2024].
+This is the result of researchers applying one of two strategies to evade the self-imposed restrictions of preregistrations: writing a loose preregistration to begin with [@stefanBigLittleLies2022] or deviating from the preregistration afterward [@lakensDeviate2024].
 The latter is a frequent occurrence and, perhaps more worryingly, often remains undisclosed [@akkerEffectivenessPreregistrationPsychology2023; @claesenComparingDreamReality2021].
@@
 This implies that for a priori balancing Type I and Type II error rate, a researcher must assign a priori probabilities to, for example, the size of an effect.
-Since such a move not in line with frequentist rationale there is no guidelines available on how to do this.
+Since such a move is not in line with frequentist rationale there is no guidelines available on how to do this.
 Third, we would argue that severity considerations assume full information about how the evidence came about and hence imply axiomatically the need for perfect preregistration.
@@
 This assumptions also makes it difficult to deal with less than perfect preregistrations and post-hoc changes without appealing to principles outside the core philosophy of severity.
+One such approach is @lakensDeviate2024' introduction of validity as an additional consideration to severity when evaluating deviations from preregistrations.
+Interestingly, in this work he unconventionally defines high severity as high $P(E|H)$ and high $P(\neg{}E|\neg{}H)$, which is closer to definitions of "diagnosticity" [@fiedlerWhatConstitutesStrong2017; @oberauerAddressingTheoryCrisis2019] and falls under the broad class of measures for evaluating evidence we consider here.
+Notably, the original definition of severity does not satisfy the statistical relevancy condition and is not such a measure; @mayoStatisticalInferenceSevere2018, p. 14:
 
-However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
+> Severity Principle (strong): We have evidence for a claim C just to the extent it survives a stringent scrutiny.
+If C passes a test that was highly capable of findings flaws or discrepancies from C, and yet none or few are found, the passing result, x, is evidence for C.
+
+However, there also are communalities between our approach and severity, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
 

### 2024-09-04 reviews.Rmd
adjust intro to letter

@@
 manuscript-id  : "4139"
-manuscript-src : "headings"
-bibliography   : references.bib
 cutout-quotes  : true
@@
 output         : papaja::revision_letter_pdf
+manuscript-src : manuscript.Rmd
 ---
@@
 In the following we address your and each reviewers'/editors' concerns point-by-point.
-For Daniel Laken's review we took the liberty to additionally include a high level summary, since we found it, at times, hard to follow the different interwoven discussions.
 
@@
 [https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main](https://github.com/aaronpeikert/bayes-prereg/compare/v0.2.0...main#diff-a9a4aad3fa8c9c10c5404b632bc3a01a25d2d8430eb932bc35c76769963e4b70)
+
+For Daniel Laken's review we took the liberty to additionally include a high level summary, since we found it, at times, hard to follow the different interwoven discussions.
+We took special care to change the manuscript in response to his summary of suggestions (RC \@ref(pt:change-start) to RC \@ref(pt:change-end)).
+
+Due to a technical issue, the citations in the manuscript excerpts in this letter do not work correctly (they are correct in the manuscript, however).
 
@@
 
-Control vs Transparency:
+Control vs. Transparency:
 
@@
 
-* Reviewer Concern: The terminology used, particularly “theoretical risk,” was seen as potentially confusing and misaligned with existing literature, such as the concept of severity.
+* Reviewer Concern: The terminology used, particularly "theoretical risk," was seen as potentially confusing and misaligned with existing literature, such as the concept of severity.
 * Our Response: We provided additional references and explanations to support our terminology.
@@
 :::reviewer
+\label{pt:change-start}
+
 \label{pt:harking}
@@
 :::reviewer
+\label{pt:change-end}
 \label{pt:nullhacking}

### 2024-09-30 reviews.Rmd
Update reviews.Rmd
I don't think it was per se an oversight but it's a good thing to add indeed.

@@
 * Reviewer Concern: The manuscript lacked a discussion on HARKing, a practice detrimental to the integrity of hypothesis testing, and the error-statistical perspective emphasizing the importance of severe tests.
-* Our Response: We acknowledged this oversight and added a dedicated section titled “Hacking, harking, and other harms” to address how our Bayesian framework addresses these concerns.
+* Our Response: We thank the reviewer for raising this issue. We added a dedicated section titled “Hacking, harking, and other harms” to address how our Bayesian framework addresses these concerns.
 

### 2024-09-30 reviews.Rmd
Update reviews.Rmd
Update reviews.Rmd
some more changes

@@
 
-We thank the reviewer for pointing out that we failed to include opposing philosophical perspectives.
+We thank the reviewer for pointing out that we failed to better acknowledge opposing philosophical perspectives.
 We are now citing Laken's and Mayo's positions and relate them to our perspective.
@@
 A Bayesian may assign a probability to a hypothetical world where a paper was p hacked, since probabilities represent their subjective believes.
-A frequentist (and hence severe tester), can not possibly assign a probability here because there is no random process but a single event that happened in the past, so the paper was either p hacked or not.
-This is similar to the mistake, in frequentist philosophy, to say that an effect is with 95% probability in the confidence interval.
+A frequentist (and thus a severe tester, too), cannot possibly assign a probability here because there is no random process but a single event that happened in the past, so the paper was either p hacked or not.
+This is similar to the mistake, in frequentist philosophy, to claim that an effect size found in a given study is with 95% probability inside the confidence interval.
+However, a frequentist can only make a statement about an expected values but not about a single case.
 
@@
 
-We did not limit ourselves to relative liklihoods.
+We did not limit ourselves to relative likelihoods.
 We explicitly include both the "firmness" and "increase in firmness" perspective.
@@
 
-We added a new section that deals with the concept of severity, making it explicit how our approach differs fundamentally.
+We added a new section that puts the concept of severity into perspective and explicates how our approach differs fundamentally from it:
 

### 2024-09-30 reviews.Rmd
Update reviews.Rmd

@@
 Third, we would argue that severity considerations assume full information about how the evidence came about and hence imply axiomatically the need for perfect preregistration.
-This comes down to frequentist understanding of probability as the outcome of a well defined random experiments.
+This comes down to frequentist understanding of probability as the outcome of a well defined random experiment.
 When judging a particular study, a frequentist, and hence a severe tester, may not assign probability to the event that the researchers did, for example, p-hack.
@@
 Full transparency, is hence assumed, and we can not imagine many ways except preregistration that get close to this ideal.
-This assumptions also makes it difficult to deal with less than perfect preregistrations and post-hoc changes without appealing to principles outside the core philosophy of severity.
+This assumptions also makes it hard to deal with less-than-perfect preregistrations and post-hoc changes without relying to principles outside the core philosophy of severity.
 
@@
 
-We find ourselves in a curios position, that perhaps may be summarised as "doing the right thing, for the wrong reasons", i.e., we agree with Laken's conclusions about preregistration but we disagree how to come to these conclusions.
+Importantly, we agree with Laken's conclusions about preregistration but we fundamentally disagree with how to arrive at these conclusions.
 We take a Bayesian/inductionist view, whereas Lakens takes a frequentist/anti-inductionist view, agreeing on the end result but disagreeing on everything that comes before.
@@
 
-No, we are happy with low uncertainty about theoretical risk.
-An exploratory study is in our framework, signified by low theoretical risk (high Type 1 error rate).
+No. 
+As we argue in our manuscript, we are happy when uncertainty about theoretical risk is low.
+A typical exploratory study is in our framework characterized by low theoretical risk (high Type 1 error rate).
+Reducing uncertainty about its theoretical risk is still useful for scientific progress, in order for us to be better able to judge its theoretical risk (for example, the theoretical risk may low but higher than another exploratory study).
 

### 2024-09-30 reviews.Rmd
Update reviews.Rmd

@@
 
-Neither is severity as an absolute measure of corroboration a common consideration in empirical psychological work.
+Neither is severity -- as an absolute measure of corroboration -- a common consideration in empirical psychological work.
 However, the posterior probability is certainly the most classic measure of absolute confirmation in the philosophy of science.
@@
 
-> In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling, where the roles of detectability, theoretical risk and type-I/II error rate switch.
+> In fact, it is entirely possible to assume the null hypothesis as a research hypothesis, as is commonly done in e.g., structural equation modelling.
+Then, the roles of detectability, theoretical risk and type-I/II error rate switch, yet our line of reasoning stays the same.
 
@@
 
-Perhaps, we are a bit daft but a clear definition of error rate control eludes us.
-We are under the impression that two competing definitions are used here from time to time.
+We are under the impression that two inconsistent definitions of *control* of error rates are used by the reviewer.
 First, a definition in which control means the error rates are known (but then known by whom?).

### 2024-09-30 reviews.Rmd
Update reviews.Rmd

@@
 
-We talk about the scientific procedures, of which a statistical test is a piece.
+Please note that our framework adressess the entire scientific procedure, of which a statistical test is only one piece.
+It would cost our framework its generality if we reduced 'theoretical risk' to only a property of the statistical test (such as severity).
 
@@
 
+This concern is conceptually very similar to the second-to-last concern.
 The point of not using a Type 1 error rate is to differentiate the statistical test from the the researcher's scientific judgement.
@@
 Researchers find the idea of a risky test appealing.
-Popper gives one explanation for this appeal, but we happen to have a different view.
+Popper gives one explanation for this appeal, but we happen to adopt a different philosophical view here.
 
@@
 
-As stated above, if Popper would happen to agree, we indeed would have to rethink our arguments.
+As stated above, if Popper happened to agree, we indeed would have to rethink our arguments.
 
@@
 
-We appreciate this hard work.
+We appreciate this important work.
 Yet we believe that this is not the only useful perspective on the utility of preregistration.
@@
 We take the criticism of the terminology seriously.
-Although we have opted to stick with most of the terms as we defined them, we provided more references.
-It might also be the case that the differences in philosophical traditions lead to some criticism.
+We have opted to stick with most of the terms as we defined them initially, as we did with good reasons, we now back them up with additional references.
 

### 2024-10-21 reviews.Rmd

@@
 A typical exploratory study is in our framework characterized by low theoretical risk (high Type 1 error rate).
-Reducing uncertainty about its theoretical risk is still useful for scientific progress, in order for us to be better able to judge its theoretical risk (for example, the theoretical risk may low but higher than another exploratory study).
+Reducing uncertainty about its theoretical risk is still useful for scientific progress, in order to be better able to judge its theoretical risk.
 

### 2024-10-21 reviews.Rmd
Update reviews.Rmd
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 
-For Daniel Laken's review we took the liberty to additionally include a high level summary, since we found it, at times, hard to follow the different interwoven discussions.
+Concerning Daniel Laken's review, we took the liberty to provide a high level summary of his main points because they are repeated across different subdiscussions throughout his letter.
 We took special care to change the manuscript in response to his summary of suggestions (RC \@ref(pt:change-start) to RC \@ref(pt:change-end)).

### 2024-10-21 reviews.Rmd
Update reviews.Rmd
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 Concerning Daniel Laken's review, we took the liberty to provide a high level summary of his main points because they are repeated across different subdiscussions throughout his letter.
-We took special care to change the manuscript in response to his summary of suggestions (RC \@ref(pt:change-start) to RC \@ref(pt:change-end)).
+We are grateful for his comments and revised the manuscript in response to his summary of suggestions (RC \@ref(pt:change-start) to RC \@ref(pt:change-end)).
 

### 2024-10-21 reviews.Rmd
Update reviews.Rmd
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>
Apply suggestions from code review
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>

@@
 
-# Daniel Lakens
-
-Lakens and our own positions on preregistration built upon two long standing traditions in philosophy of science, namely inductionist and anti-inductionist, that differ fundamentally in their core assumptions.
+# Reviewer 1: Daniel Lakens
+We begin with a summarizing our understanding of Daniel Lakens philosophical position, his main points of criticisim and our general take on how to address these issues.
+Thereafter, we will provide point-by-point responses in the usual style of response letters.
+Lakens and our own positions on preregistration fundamentally differ.
+They each build upon two long standing traditions in philosophy of science, namely _inductionist_ and _anti-inductionist_, which differ in their core epistemological assumptions.
+We believe that a lot of seemingly critical points raised by Daniel Lakens can simply be traced back to differences in these philosophical stances and thus cannot be resolved in the various sub points he raised.
 
@@
 * Our Response: The Bayesian rational considers theoretical risk and detectability simultaneously a priori.
-From a severity perspective, either Type I or Type II, never both, is considered post hoc, depending on the test results and if the research hypothesis is H1 or H2.
+From a severity perspective, either Type I or Type II, never both, is considered post hoc (i.e., after a test result was obtained from data) depending on the test results and if the research hypothesis is H0 or H1.
 

### 2024-10-21 manuscript.Rmd
allow cross referencing to review letter
Apply suggestions from code review
Co-authored-by: Andreas Brandmaier <brandmaier@mpib-berlin.mpg.de>
its not P_pointless

@@
 )
+options(tinytex.clean = FALSE)
 knitr::read_chunk("R/plots.R")
@@
 
-# Persuation and the Bayesian rationale
+# Persuasion and the Bayesian rationale
 
@@
 If we know that the results will be engineered to support the hypothesis no matter what, there is no reason to collect data.
-A prime example of this case is the $p_\text{pointless}$ metric [@husseyMethodStreamlinePhacking2021].
+A prime example of this case is the $p_\text{ointless}$ metric [@husseyMethodStreamlinePhacking2021].
 Case 2 has a similar problem.

### 2024-10-21 manuscript.Rmd
missing semicolon messed this up badly

@@
 
-At the same time, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990 @fiedlerWhatConstitutesStrong2017;@sprengerBayesianPhilosophyScience2019].
+At the same time, approaching the evaluation of evidence using a Bayesian formalism is far from novel [e.g., @kuklaClinicalStatisticalTheory1990; @fiedlerWhatConstitutesStrong2017; @sprengerBayesianPhilosophyScience2019].
 To our knowledge, it was not yet applied to the problem of preregistration.

### 2024-10-21 reviews.Rmd

@@
 manuscript-src : manuscript.Rmd
+bibliography   : references.bib
 ---
@@
 We are sure this would be possible and enlightening, but also way beyond the scope of the current paper.
-Out goal is more modest: we want to provide a bayesian justification of pre-registration that 1.
+Out goal is more modest: we want to provide a Bayesian justification of pre-registration that 1.
 clearly articulates the aims of a preregistration and 2.
@@
 :::
+
+# References

### 2024-10-21 reviews.Rmd
add excerpt

@@
 
-This sentence sounds exactly like P(¬E|¬H) to us.
-The judgement about P(¬E|¬H) is a-priori, is about "prediction" as a statement about evidence E, given that the theory is not true (¬H).
-One might argue, that it should be multiplied by ¬H perhaps? 
+This sentence sounds exactly like $P(\neg{}E|\neg{}H)$ to us.
+The judgement about $P(\neg{}E|\neg{}H)$ is a-priori, is about "prediction" as a statement about evidence E, given that the theory is not true ($\neg{}H$).
+One might argue, that it should be multiplied by $\neg{}$ perhaps? 
 But that would not change any of the arguments.
@@
 In fact, this happens in reference to a finding we found in the meantime (Oberauer and Lewandowsky (2019)), which we added to the discussion.
+
+> Whatever the difference in evaluating preregistration as a tool, maybe conceptually more profound is that @oberauerAddressingTheoryCrisis2019 conceptualizes "discovery-oriented research" differently than we do "exploratory".
+They assume the same theoretical risk ($P(\neg E| \neg H) = .05$) and detectability ($P(E| H)) = .8$) in their calculation example as we do but assign different prior probabilities, namely $.06$ for discovery versus $.6$ for theory testing.
+Then, they conclude that discovery-oriented researcher requires a much lower type-I error rate to control false positive in light of the low prior probability.
+This runs counter to our definition of exploratory research having low theoretical risk.
+Of course, we agree that low priors require more persuasive evidence; our disagreement, therefore, lies mainly in terminology.
+They imagine discovery-oriented researchers to conduct experiments where they have low expectations that they obtain positive evidence ($.06 \cdot .8 + .94 \cdot .05 = 0.095$), but if they do, it raises the posterior significantly (from $.06$ to $.51$)
+In our view, researchers who set out to explore a data set often find "something" (due to low $P(\neg E| \neg H)$); therefore, it should only slightly raise your posterior if they do.
+On a substantive matter, we believe both kinds of research are common in psychology.
+It is, therefore, mostly a disagreement on terminology.
+This disagreement only highlights why using a mathematical framework to investigate such things is so useful and ultimately indispensable because we can clearly see where and how we differ in our reasoning.
 

### 2024-10-21 reviews.Rmd
references still does not work

@@
 manuscript-src : manuscript.Rmd
-bibliography   : references.bib
 ---

### 2024-10-22 reviews.Rmd
add link to preprint and osf

@@
 
+We have updated the preprint:
+
+<https://osf.io/preprints/psyarxiv/cs8wb>
+
+ OSF (<https://osf.io/nc5g8/>) and Zenodo (<https://doi.org/10.5281/zenodo.7648471>) are mirroring our GitHub repo (<https://github.com/aaronpeikert/bayes-prereg>).
+
 Concerning Daniel Laken's review, we took the liberty to provide a high level summary of his main points because they are repeated across different subdiscussions throughout his letter.
@@
 # Reviewer 1: Daniel Lakens
+
 We begin with a summarizing our understanding of Daniel Lakens philosophical position, his main points of criticisim and our general take on how to address these issues.

### 2025-06-24 manuscript.Rmd
accepted in Metapsychology

@@
 
-  Submitted to *Meta-Psychology*.
-Participate in open peer review by sending an email to open.peer.reviewer@gmail.com.
-The full editorial process of all articles under review at Meta-Psychology can be found following this link:
-  https://tinyurl.com/mp-submissions  
-  You will find this preprint by searching for the first author's name.
-
+  This preprint has now been peer reviewed and accepted for publication in Meta-Psychology.
 ---
